### 2013-09-18 tex/ms.tex
streams: start on document

\documentclass[12pt,preprint]{aastex}
%\documentclass{emulateapj}
\usepackage{amssymb,amsmath}
%\usepackage{caption}
%\DeclareCaptionLabelSeparator{dotemdash}{.--- }
%\captionsetup[figure]{labelformat=simple, labelsep=dotemdash}
%\usepackage{color,hyperref}
% hypertex insanity
%\definecolor{linkcolor}{rgb}{0,0,0.25}
%\hypersetup{
%  colorlinks=true,        % false: boxed links; true: colored links
%  linkcolor=linkcolor,    % color of internal links
%  citecolor=linkcolor,    % color of links to bibliography
%  filecolor=linkcolor,    % color of file links
%  urlcolor=linkcolor      % color of external links
%}
\newcounter{address}
\setcounter{address}{1}
%\usepackage{sidecap}
\setlength{\emergencystretch}{2em}%No overflowing references
\newcommand{\ie}{i.e.}
\newcommand{\etal}{et al.}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\eg}{e.g.}
\newcommand{\eqnname}{equation}
\newcommand{\equationname}{\eqnname}
\renewcommand{\tablename}{Table}
\renewcommand{\figurename}{Figure}
\newcommand{\figurenames}{\figurename~s}
\newcommand{\sectionname}{$\mathsection$}
\newcommand{\normal}{\ensuremath{\mathcal{N}}}
\newcommand{\flag}[1]{\texttt{\lowercase{#1}}}
\newcommand{\feh}{\ensuremath{[\mathrm{Fe/H}]}}
\newcommand{\afe}{\ensuremath{[\alpha\mathrm{/Fe}]}}
\newcommand{\logg}{log g}
\newcommand{\Ro}{\ensuremath{R_0}}

\renewcommand{\vec}[1]{\ensuremath{\mathbf{#1}}}
\newcommand{\vecx}{\ensuremath{\vec{x}}}
\newcommand{\vecv}{\ensuremath{\vec{v}}}
\newcommand{\vecj}{\ensuremath{\vec{J}}}
\newcommand{\veco}{\ensuremath{\vec{\Omega}}}
\newcommand{\veca}{\ensuremath{\vec{\theta}}}
\newcommand{\df}{\ensuremath{f}}
\newcommand{\tdf}{\ensuremath{\mathrm{tDF}}}
\newcommand{\paramsdf}{\ensuremath{\vec{p}_{\mathrm{DF}}}}
\newcommand{\paramspot}{\ensuremath{\vec{p}_\Phi}}

\newcommand{\DF}{\df}
\newcommand{\dff}{\ensuremath{\mathbf{f}}}
\newcommand{\dex}{\ensuremath{\,\mathrm{dex}}}
\newcommand{\Gyr}{\ensuremath{\,\mathrm{Gyr}}}
\newcommand{\kpc}{\ensuremath{\,\mathrm{kpc}}}
\newcommand{\pc}{\ensuremath{\,\mathrm{pc}}}
\newcommand{\kms}{\ensuremath{\,\mathrm{km\ s}^{-1}}}
\newcommand{\msun}{\ensuremath{\,\mathrm{M}_{\odot}}}
\newcommand{\inv}{\ensuremath{^{-1}}}

\newcommand{\jr}{\ensuremath{J_R}}
\newcommand{\jphi}{\ensuremath{J_\phi}}
\newcommand{\jz}{\ensuremath{J_Z}}
\newcommand{\lz}{\ensuremath{L_Z}}
\newcommand{\Or}{\ensuremath{\Omega_R}}
\newcommand{\Ophi}{\ensuremath{\Omega_\phi}}
\newcommand{\Oz}{\ensuremath{\Omega_Z}}
\newcommand{\ar}{\ensuremath{\theta_R}}
\newcommand{\aphi}{\ensuremath{\theta_\phi}}
\newcommand{\az}{\ensuremath{\theta_Z}}

%\submitted{}

\begin{document}

\title{Dynamical modeling of tidal streams}
\author{Jo~Bovy\altaffilmark{1,2,3} \etal}
%  Hans-Walter~Rix\altaffilmark{4}}
\altaffiltext{\theaddress}{\label{IAS}\stepcounter{address} Institute
  for Advanced Study, Einstein Drive, Princeton, NJ 08540, USA}
\altaffiltext{\theaddress}{\label{Hubble}\stepcounter{address} Hubble
  fellow}
\altaffiltext{\theaddress}{\label{email}\stepcounter{address}
  Correspondence should be addressed to bovy@ias.edu~.}
%\altaffiltext{\theaddress}{\label{MPIA}\stepcounter{address}
%  Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117
%  Heidelberg, Germany}

\begin{abstract} 
  We present a generative, dynamical model for a tidal stream.
\end{abstract}

\keywords{
	Galaxy: fundamental parameters
	---
        Galaxy: kinematics and dynamics
        ---
	Galaxy: structure
}

\section{Introduction}

%BOVY: reference point

\section{Two equivalent generative models of tidal streams}\label{sec:method}

%BOVY: reference point

\acknowledgements It is a pleasure to thank \dots for helpful comments
and assistance. J.B. was supported by NASA through Hubble Fellowship
grant HST-HF-51285.01 from the Space Telescope Science Institute,
which is operated by the Association of Universities for Research in
Astronomy, Incorporated, under NASA contract NAS5-26555. J.B. 
%and H.-W.R  (BOVY: ALSO ACKNOWLEDGE WITHOUT S)
acknowledges support from SFB 881 (A3) funded by the German
Research Foundation DFG.

\appendix

\section{An efficient, general method for calculating action-angle coordinates using orbit integration}\label{sec:aa}

\section{Dynamical modeling of two-star streams}\label{sec:twostar}

%BOVY: reference point

\begin{thebibliography}{}

\bibitem[{{Binney} \& {Tremaine}(2008)}]{binneytremaine}
  Binney,~J. \& Tremaine,~S. 2008, Galactic Dynamics: Second Edition
\end{thebibliography}

\end{document}

### 2013-09-18 tex/ms.tex
streams: start on document

@@
 \newcommand{\vecv}{\ensuremath{\vec{v}}}
+\newcommand{\vech}{\ensuremath{\vec{H}}}
 \newcommand{\vecj}{\ensuremath{\vec{J}}}
 \newcommand{\veco}{\ensuremath{\vec{\Omega}}}
-\newcommand{\veca}{\ensuremath{\vec{\theta}}}
+\newcommand{\veca}{\ensuremath{\boldsymbol\theta}}
 \newcommand{\df}{\ensuremath{f}}
 \newcommand{\tdf}{\ensuremath{\mathrm{tDF}}}
+\newcommand{\paramsdiff}{\ensuremath{\vec{p}_\Delta}}
 \newcommand{\paramsdf}{\ensuremath{\vec{p}_{\mathrm{DF}}}}
@@
 
+In this Section, we describe two mathematically equivalent generative
+models for a tidal stream. The first is based on orbit integrations in
+real (position-velocity) space: a tidal stream is generated from small
+(phase-space) offsets between a star and a source\footnote{The word
+  `progenitor' is typically used to indicate the actual parent body of
+  a star in a tidal stream. For the dynamical model only a reference
+  orbit from which the tidal stream can be thought of to originate is
+  necessary; this reference orbit does \emph{not} have to coincide
+  with the actual progenitor (this point is most clear in the
+  description of the dynamics in action-angle coordinates). To avoid
+  confusion between the hypothetical progenitor and the actual parent
+  body, we refer to the reference orbit as the `source' rather than
+  the `progenitor'.} that are evolved in time using orbit
+integrations. The second method relies on the equivalent description
+of this process in action-angle coordinates. While these two methods
+are equivalent, the latter provides a conceptually simpler description
+of the dynamical model and, when fitting a tidal stream, it allows for
+fast marginalization over nuisance parameters. In particular, the
+marginalization over the time at which a star was removed from the
+source is analytic and the integration over possible source orbits can
+be performed much faster (by avoiding orbit integrations of all
+possible source orbits).
+
+\subsection{Real-space model}\label{sec:realmethod}
+
+The description of the stripping of a star from a satellite in real
+(\vecx,\vecv) space assumes that at a time $\Delta t$ ago the star was
+offset by a small amount ($\Delta \vecx,\Delta \vecv$) from the
+position of the source $(\vecx^s,\vecv^s)(-\Delta t)$ at that
+time. Then the star orbited under the influence of the same
+gravitational potential $\Phi$ as the source until the present day
+$(t=0)$, when it was observed: $(\vecx,\vecv)(t=0)$ = $\vech_{\Delta
+  t}(\vecx^s(-\Delta t)+\Delta \vecx,\vecv^s(-\Delta t)+\Delta
+\vecv)$, where $\vech$ denotes the Hamiltonian flow from $t=-\Delta t$
+to $t=0$ (see \citealt{binneytremaine}); this is simply the orbit of
+the star between these times. In phase-space this orbit can be
+calculated by solving Hamilton's equations: $\dot{\vecx} = \vecv;
+\dot{\vecv} = - \dd \Phi / \dd \vecx$.
+
+A generative model for the positions and velocities
+($\vecx_i,\vecv_i$) of stars $i$ in a tidal stream is then fully
+specified by: (a) the gravitational potential $\Phi$, (b) the current
+phase-space position of the source $(\vecx_s,\vecv_s)(t=0)$, which can
+be integrated backwards in time by reversing the Hamiltonian flow
+defined by $\Phi$, (c) the time $\Delta t_i$ at which each star was
+removed from the source, and (d) the distribution $f(\Delta
+\vecx,\Delta \vecv|\paramsdiff)$, characterized by some parameters
+$\paramsdiff$, of initial phase-space offsets from which the actual
+offsets $(\Delta \vecx_i,\Delta \vecv_i)$ are drawn. In practice, $f$
+can be taken to be a Gaussian, characterized by a width in position
+$\sigma_x$ and in velocity $\sigma_v$, corresponding to the tidal
+radius and escape velocity of the progenitor, respectively. However,
+more complicated models for the initial phase-space distribution of
+the debris can be used here as well (see discussion below). In the
+case of a Gaussian $f$, the source can be thought of as the mean of
+this Gaussian, with the additional requirement that this mean is an
+orbit in the potential $\Phi$.
+
+The probability for observing a star today at $(\vecx,\vecv)$ given
+items (a) through (d) is then
+\begin{align}\label{eq:realdft}
+  p(\vecx_i,\vecv_i | \Phi,\vecx_s,\vecv_s,\Delta t_i,\paramsdiff) 
+  & = p(\vecx_i(-\Delta t),\vecv_i(-\Delta t) | \Phi,\vecx_s,\vecv_s,\Delta t_i,\paramsdiff) \,,\nonumber\\
+  & = p([\vecx_i-\vecx_s](-\Delta t),[\vecv_i-\vecv_s](-\Delta t) | \Phi,\vecx_s,\vecv_s,\Delta t_i,\paramsdiff) \,\nonumber\\
+  & = p(\Delta\vecx_i(-\Delta t),\Delta\vecv_i(-\Delta t) | \Phi,\vecx_s,\vecv_s,\Delta t_i,\paramsdiff)\,,\\
+  & = f(\Delta\vecx_i(-\Delta t),\Delta\vecv_i(-\Delta t) | \Phi,\vecx_s,\vecv_s,\Delta t_i,\paramsdiff)\, ,\nonumber
+\end{align}
+where the first equality holds because the Hamiltonian flow is unique
+and conserves phase-space volume by Liousville's theorem and the
+second one because the Jacobian of the transformation
+$(\vecx_i,\vecv_i)(-\Delta t) \rightarrow (\vecx_i,\vecv_i)(-\Delta t)
+- (\vecx_s,\vecv_s)(-\Delta t)$ is one. The final probability can be
+evaluated using the parameters $\paramsdiff$ following (d).
+
+The time $\Delta t_i$ at which a star was removed is unobservable and
+the probability in \eqnname~(\ref{eq:realdft}) should therefore be
+integrated over this time using a prior $p(\Delta t_i|\ldots)$, where
+$\ldots$ indicates that this prior could depend on other factors such
+as the potential and the progenitor orbit\footnote{For example, one
+  could insist that stars are removed around pericenter passages. Note
+  that this assumption would require the source orbit to be the actual
+  progenitor orbit.}. For simplicity, we will assume a uniform prior
+for the times $\Delta t$. The probability for a single star then becomes
+\begin{align}\label{eq:realdf}
+  p(\vecx_i,\vecv_i | \Phi,\vecx_s,\vecv_s,\paramsdiff) & = \int \dd
+  \Delta t_i \,p(\vecx_i,\vecv_i | \Phi,\vecx_s,\vecv_s,\Delta t_i,
+  \paramsdiff)\,.
+\end{align}
+If the progenitor's current phase-space position is unknown, then this
+probability should be further marginalized over $(\vecx_s,\vecv_s)$;
+for multiple stars, this marginalization is done after multiplying the
+individual-star likelihoods in \eqnname~(\ref{eq:realdf}).
+
+In practice, the likelihood in \eqnname~(\ref{eq:realdf}) can be
+calculated as follows: for a given potential $\Phi$ and source
+$(\vecx_s,\vecv_s)$, we (a) integrate the orbits of all stars
+$(\vecx_i,\vecv_i)$ and (b) that of the source backward in time for
+many Gyr and record the phase-space positions over time on a grid $j$
+in $\Delta^j_i$; (c) we evaluate $f(\Delta \vecx_i,\Delta
+\vecv_i)(\Delta t^j_i)$ for all stars $i$ and times $j$; (d) $f(\Delta
+\vecx_i,\Delta \vecv_i)(\Delta t^j_i)$ is then summed over $j$ and the
+logarithm of that is summed over $i$. The most computationally
+expensive part of this calculation is the integration of the various
+orbits. The evaluation of the likelihood for different parameters
+\paramsdiff\ of $f$ can be efficiently computed by caching the orbit
+integrations of (a) and (b); different source orbits
+$(\vecx_s,\vecv_s)$ can also make use of the stored orbits in (a).
+
+\subsection{Action-angle-space model}\label{sec:aamethod}
+
+An alternative description of orbits in a gravitational potential
+$\Phi$ is provided by the action-angle formalism (see
+\citealt{binneytremaine} for an extensive description). The
+Hamiltonian flow in the action-angle formalism is trivial: actions are
+conserved and angles increase linearly with time. In
+\appendixname~\ref{sec:aa} we summarize a method to calculate
+action-angle coordinates from an orbit integration with little extra
+work in any (static) potential for any kind of orbit. Therefore, it is
+practical to use action-angle coordinates to describe the dynamical
+model for a tidal stream in these coordinates and make use of the many
+conceptual advantages provided by this formalism.
+
+In the action-angle formalism, a star is removed from the source with
+a small offset $(\Delta \vecj,\Delta \veca)(-\Delta t)$ from the
+source's actions $\vecj_s$ and angles $\veca_s$ a time $\Delta t$ ago,
+similar to the small offset $(\Delta \vecx,\Delta \vecv)(-\Delta t)$
+in the real-space model. The offset in the actions corresponds to an
+offset in the frequencies $\Delta \veco$; ultimately, this offset is
+responsible for the spreading of the stream over long stretches on the
+sky. Because both the source's angles and the star's angles increase
+linearly with time, albeit at different frequencies, the difference in
+the angles also increases linearly in time. Therefore at time $t=0$
+the difference in angles is
+\begin{equation}
+\Delta \veca(t=0) = \Delta \veca(-\Delta t) + \Delta \veco \Delta t\,,
+\end{equation}
+while the difference in actions is constant
+\begin{equation}
+\Delta \vecj(t=0) = \Delta \vecj(-\Delta t)\,.
+\end{equation}
+
+With these preliminaries, we can write down a generative model for the
+positions and velocities ($\vecx_i,\vecv_i$) of stars $i$ in a tidal
+stream in the action-angle formalism by specifying: (a) the
+gravitational potential $\Phi$, (b) the current action-angle
+coordinates of the source $(\vecj_s,\veca_s)(t=0)$, (c) the time
+$\Delta t_i$ at which each star was removed from the source, and (d)
+the distribution $f(\Delta \vecj,\Delta \veca|\paramsdiff)$,
+characterized by some parameters $\paramsdiff$, of initial phase-space
+offsets from which the actual offsets $(\Delta \vecj_i,\Delta
+\veca_i)$ are drawn. In practice, $f$ can again be taken to be a
+Gaussian, although more general models could take the expected
+`bowtie' structure of the action distribution into account
+\citep{Eyre11a}.
+
+The probability for observing a star today at $(\vecx,\vecv)$ given
+items (a) through (d) is then
+\begin{align}\label{eq:aadft}
+  p(\vecx_i,\vecv_i | \Phi,\vecj_s,\veca_s,\Delta t_i,\paramsdiff) 
+  & = p(\vecj_i,\veca_i | \Phi,\vecj_s,\veca_s,\Delta t_i,\paramsdiff) \,,\nonumber\\
+  & = p(\vecj_i-\vecj_s,\veca_i-\veca_s-\Delta \veco_i\Delta t_i | \Phi,\vecj_s,\veca_s,\Delta t_i,\paramsdiff) \,,\\
+  & = f(\Delta \vecj_i,\Delta \veca_i(-\Delta t_i) | \Phi,\vecj_s,\veca_s,\Delta t_i,\paramsdiff) \,,\nonumber
+\end{align}
+where the first equality holds because the Jacobian of the
+$(\vecx,\vecv)\rightarrow(\vecj,\veca)$ transformation is unity and
+the second equality is a simple linear transformation. The final
+probability can again be evaluated using the parameters
+\paramsdiff\ following (d).
+
+As in the real-space formalism, we should integrate
+\eqnname~(\ref{eq:aadft}) over the time $\Delta t_i$ at which the star
+is removed from the source (cf.~\eqnname~[\ref{eq:realdf}]):
+\begin{equation}\label{eq:aadf}
+  p(\vecx_i,\vecv_i | \Phi,\vecj_s,\veca_s,\paramsdiff) = \int \dd
+  \Delta t_i\, f(\Delta \vecj_i,\Delta \veca_i(-\Delta t_i) |
+  \Phi,\vecj_s,\veca_s,\Delta t_i,\paramsdiff) \,.
+\end{equation}
+Without further assumptions the action-angle formalism proceeds in the
+same way as the real-space formalism, evaluating
+\eqnname~(\ref{eq:aadft}) for a grid $j$ of times $\Delta t_i^j$ and
+summing $f(\Delta \vecj_i,\Delta \veca_i(-\Delta t^j_i) |
+\Phi,\vecj_s,\veca_s,\Delta t_i,\paramsdiff)$ over $j$.
+
+Even without further assumptions about the shape of $f(\Delta
+\vecj,\Delta \veca|\paramsdiff)$, the action-angle formalism already
+allows for a simplification over the real-space method. Because the
+source orbit is specified in action-angle coordinates, it can be
+changed \emph{without} expensive orbit integrations of all possible
+source orbits. As the source's current phase-space position is a
+six-dimensional parameter, this is a major simplification of the
+dynamical model when the source orbit is unknown. Note that this
+simplification does require us to be able to compute $\veco(\vecj)$,
+to calculate the new source frequencies $\veco_s$ when changing the
+source actions $\vecj_s$. This mapping can be constructed using a
+small number of (action,frequency) calculations covering the range of
+possible source actions.
+
+We can simplify \eqnname~(\ref{eq:aadf}) by assuming (a) that
+$f(\Delta \vecj,\Delta \veca|\paramsdiff)$ is separable into
+$f_{\veca}(\Delta \veca|\paramsdiff)\,f_{\vecj}(\Delta
+\vecj|\paramsdiff)$ and (b) that the distribution $f_{\veca}(\Delta
+\veca|\paramsdiff)$ is Gaussian. In what follows, we will assume that
+this Gaussian is isotropic with dispersion $\sigma_\theta$, but all of
+the results are the same after appropriate diagonalization and
+rescaling of the angles for more general covariance matrices. In this
+case, \eqnname~(\ref{eq:aadf}) can be written as
+\begin{equation}
+\begin{split}
+  p(\vecx_i,\vecv_i & | \Phi,\vecj_s,\veca_s,\paramsdiff) 
+   = \\
+   & f_{\vecj}(\Delta \vecj_i|\paramsdiff) \int \dd \Delta t_i\,
+  \frac{1}{(2\,\pi)^{3/2}\,\sigma_\theta^3}\,\exp\left(-\frac{1}{2\,\sigma_\theta^2}\left[\sum_{R,\phi,Z}\left(\Delta \veca_i-\Delta \veco_i\Delta t_i\right)^2\right]\right)\,,
+\end{split}
+\end{equation}
+where the sum in the argument of the exponential is over the three
+components of $\Delta \veca_i$ and $\Delta \veco_i$. This integral can
+be done analytically, resulting in
+\begin{equation}\label{eq:aadfsimple}
+\begin{split}
+  p(\vecx_i,\vecv_i & | \Phi,\vecj_s,\veca_s,\paramsdiff) 
+   = \\
+   & f_{\vecj}(\Delta \vecj_i|\paramsdiff) \,  \frac{1}{2\,\pi\,\sigma_\theta^2\,\sqrt{\sum_{R,\phi,Z}\Delta \veco_i^2}}\,\exp\left(-\frac{1}{2\,\sigma_\theta^2}\left[\sum_{R,\phi,Z}\Delta \veca^2_i-\frac{\left(\sum_{R,\phi,Z} \Delta \veco_i\Delta \veca_i\right)^2}{\sum_{R,\phi,Z}\Delta \veco_i^2}\right]\right)\,.
+\end{split}
+\end{equation}
+In this expression, $\Delta X^2 \equiv (X_1-X_2)^2$. The expression in
+square brackets is positive by using the Cauchy-Schwarz inequality.
+
+The quantity in the square brackets in \eqnname~(\ref{eq:aadfsimple})
+expresses the fact that the present-day difference in angles between a
+star in a tidal stream and the source must be linearly related by a
+scalar (the time since the removal of the star) to the difference in
+frequencies, within a tolerance set by the width of the distribution
+of initial angle differences ($\sigma_\theta$). The ratio
+$\sum_{R,\phi,Z} \Delta \veco_i\Delta \veca_i/\sum_{R,\phi,Z}\Delta
+\veco_i^2$ is the best-guess for $\Delta t_i$ when performing a linear
+regression between $\Delta \veca_i$ and $\Delta \veco_i$. The
+expression in square brackets is minimized when the correlation
+between $\Delta \veca_i$ and $\Delta \veco_i$ is the largest, \ie,
+when a single time difference $\Delta t_i$ explains all three of the
+angle differences.
+
+This view of the dynamics of the stream in action-angle coordinates
+also shows that the orbit of the source does not have to coincide with
+that of the actual progenitor.
+
+The probability in \eqnname~(\ref{eq:aadfsimple}) incorporates and
+elucidates commonly used methods for fitting tidal streams. On the one
+hand are approaches that search to minimize the spread in energy,
+integrals of the motion, or actions (\eg,
+\citealt{Binney08a,Penarrubia12a}, Sanderson \etal, 2013, in
+preparation). These approaches ignore the correlations between actions
+and angles in the stream (the exponential in
+\eqnname~(\ref{eq:aadfsimple}), which describes this correlation
+through the frequencies) and only use $f_{\vecj}(\Delta
+\vecj_i|\paramsdiff)$; finding the gravitational potential that
+minimizes the spread in actions optimizes $f_{\vecj}(\Delta
+\vecj_i|\paramsdiff)$. Recently, a method was proposed that only uses
+the correlation between the actions and the angles, expressed by the
+exponential in \eqnname~(\ref{eq:aadfsimple}) \citep{Sanders12a}; this
+method does not use the fact that the spread in actions in a tidal
+stream is small (expressed by $f_{\vecj}(\Delta
+\vecj_i|\paramsdiff)$).
+
 %BOVY: reference point
@@
   Binney,~J. \& Tremaine,~S. 2008, Galactic Dynamics: Second Edition
+\bibitem[Binney(2008)]{Binney08a}
+  Binney,~J. 2008, \mnras, 386, L47
+\bibitem[Eyre \& Binney(2011)]{Eyre11a}
+  Eyre,~A. \& Binney,~J. 2011, \mnras, 413, 1852
+\bibitem[Pe{\~n}arrubia \etal(2012)]{Penarrubia12a}
+  Pe{\~n}arrubia, J., Koposov,~S.~E., \& Walker,~M.~G. 2012, \apj, 760, 2
+\bibitem[Sanders \& Binney(2013)]{Sanders13a}
+  Sanders,~J. \& Binney,~J. 2013, \mnras, 433, 1826
 \end{thebibliography}

### 2013-09-19 py/make_action_movie.py
streams: make move illustrating action-angle coordinates

@@ -1,0 +1,74 @@
+import sys
+import os, os.path
+import numpy
+import subprocess
+from galpy import potential
+from galpy.orbit import Orbit
+from galpy.util import bovy_plot
+from galpy.actionAngle_src.actionAngleIsochroneApprox import actionAngleIsochroneApprox
+def make_action_movie(aa=False):
+    skip= 1
+    if aa:
+        savedir= '../movies/aa/'
+        savedirpng= '../movies/aa/pngs/'
+        moviefilename= os.path.join(savedir,'aa_orbit_skip%i.mpg' % skip)
+        basefilename= 'aa_'
+        xrange=[-0.7,7.]
+        yrange=[-0.7,7.]
+        xlabel=r'$\theta_R$'
+        ylabel=r'$\theta_Z$'    
+    else:
+        savedir= '../movies/actions/'
+        savedirpng= '../movies/actions/pngs/'
+        basefilename= 'real_'
+        moviefilename= os.path.join(savedir,'real_orbit_skip%i.mpg' % skip)
+        xrange=[0.,7.]
+        yrange=[-4.25,4.25]
+        xlabel=r'$R$'
+        ylabel=r'$Z$'    
+    pot= potential.MWPotential
+    o= Orbit([1.,0.8,1.5,0.8,0.5,0.])
+    ts= numpy.linspace(0.,300.,1000)
+    o.integrate(ts,pot)
+    Rs= numpy.zeros(len(ts)+23)
+    zs= numpy.zeros(len(ts)+23)
+    Rs[:23]= numpy.nan
+    zs[:23]= numpy.nan
+    if aa:
+        aAIA= actionAngleIsochroneApprox(b=5.,pot=pot)
+        acfs= aAIA.actionsFreqsAngles(o,maxn=3)
+        Rs[23:]= (acfs[6]+acfs[3]*ts) % (2.*numpy.pi)
+        zs[23:]= (acfs[8]+acfs[5]*ts) % (2.*numpy.pi)
+    else:
+        Rs[23:]= o.R(ts)
+        zs[23:]= o.z(ts)
+        print numpy.nanmax(Rs)
+        print numpy.nanmax(numpy.fabs(zs))
+    if True:
+        for ii in range(len(ts)-1):       
+            bovy_plot.bovy_print()
+            bovy_plot.bovy_plot(Rs[ii:ii+24:skip],zs[ii:ii+24:skip],
+                                scatter=True,color='k',
+                                s=numpy.arange(24/skip)*skip,
[…]

### 2013-09-20 tex/ms.tex
streams: further work on action-angle framework

@@
 \newcommand{\eqnname}{equation}
+\newcommand{\Eqnname}{Equation}
 \newcommand{\equationname}{\eqnname}
@@
 \renewcommand{\vec}[1]{\ensuremath{\mathbf{#1}}}
+\newcommand{\unitvec}[1]{\ensuremath{\mathbf{\hat{#1}}}}
 \newcommand{\vecx}{\ensuremath{\vec{x}}}
@@
 
+The discussion so far has been entirely general and applies to any
+sort of stellar debris that originates from a single orbit, without
+necessarily generating a one-dimensional structure. In most
+observations of tidal streams, however, they appear as one-dimensional
+objects rather than the six-dimensional objects that we have
+considered so far. The reason for this is both physical and
+observational: (a) narrow tidal streams form because a star separates
+from the progenitor over time mostly along a single direction in
+angle-space: the Hessian matrix of the Hamiltonian, $\partial^2 H /
+\partial \vecj \partial \vecj$, has a single large eigenvalue, which
+determines the direction along which the stream is disrupted in
+angle-space (because $\partial \veco / \partial \vecj$ is equal to the
+Hessian); (b) because many tidal streams are dynamically cold with
+velocity dispersions $\lesssim10\kms$, observations are unable to
+resolve the angle differences along the two eigenvectors of the
+Hessian corresponding to the two smallest eigenvalues or the spread in
+actions among stream members. If the latter can be resolved, the
+stream is four-dimensional. In what follows we will be mainly
+concerned with the stream only spreading along a single direction in
+angle space.
+
+If the stream is one-dimensional in angle space then the time since
+the removal from the progenitor becomes unobservable if the progenitor
+orbit is unknown: above, determining the time depended on explaining
+the three-dimensional angle difference $\Delta \veca$ with the
+three-dimensional frequency difference $\Delta \veco$ using a single
+time difference; if two of the angle and frequency differences are
+unable to be resolved, then the frequency difference and the time
+difference become degenerate (aka, you can draw a straight line
+connecting any two points). To make use of this fact to simplify the
+dynamical model, we write the frequency difference as
+\begin{equation}
+  \Delta \veco = \Delta \Omega_1\,\unitvec{e}_1\,,
+\end{equation}
+where $\unitvec{e}_1$ is a unit vector. Then we write the angle
+difference $\Delta \veca$ 
+\begin{equation}
+  \Delta \veca = \Delta \theta_1\,\unitvec{e}_1 + \Delta \veca_\perp\,\unitvec{e}_1^\perp\,,
+\end{equation}
+where$\unitvec{e}_1^\perp$ is the orthogonal complement of
+$\unitvec{e}_1$ (the two directions perpendicular to $\unitvec{e}_1$)
+and $\Delta \veca_\perp$ are the components of $\Delta \veca$ in that
+direction. We can then write \eqnname~(\ref{eq:aadfsimple}) as
+\begin{equation}\label{eq:aadfsimpler}
+\begin{split}
+  p(\vecx_i,\vecv_i & | \Phi,\vecj_s,\veca_s,\paramsdiff) 
+   = \\
+   & f_{\vecj}(\Delta \vecj_i|\paramsdiff) \,  \frac{1}{2\,\pi\,\sigma_\theta^2\,\sqrt{\sum_{R,\phi,Z}\Delta \veco_i^2}}\,\exp\left(-\frac{|\Delta\veca_{i,\perp}|^2}{2\,\sigma_\theta^2}\right)\,.
+\end{split}
+\end{equation}
+\Eqnname~(\ref{eq:aadfsimpler}) is equivalent to
+\eqnname~(\ref{eq:aadfsimple}), but it expresses the dynamical model
+in a slightly different way. Whereas the angle part of
+\eqnname~(\ref{eq:aadfsimple}) allows the correct potential and
+progenitor to be inferred by asking whether a single time difference
+explains the three angle differences given the three frequencies,
+\eqnname~(\ref{eq:aadfsimpler}) `rewinds' the star's orbit along the
+direction of $\Delta \veco$ and asks whether the angle difference
+perpendicular to $\Delta \veco$ is consistent with the initial angle
+spread (because it has not evolved since the removal of the star). In
+re-writing this, we have not yet used the approximation that the
+stream only forms along a single direction in $\Delta \veco$.
+
 This view of the dynamics of the stream in action-angle coordinates
@@
 the correlation between the actions and the angles, expressed by the
-exponential in \eqnname~(\ref{eq:aadfsimple}) \citep{Sanders12a}; this
+exponential in \eqnname~(\ref{eq:aadfsimple}) \citep{Sanders13a}; this
 method does not use the fact that the spread in actions in a tidal

### 2013-09-30 py/pal5GC.py
stream: start on plotting the stream

@@ -103,7 +103,7 @@ def fitDirect():
                                     numpy.log(1.),21))
     #Load progenitor orbit
     progo= progenitorOrbit(V0=_REFV0,R0=_REFR0)
-    ts= numpy.linspace(0.,300.,10000)
+    ts= numpy.linspace(0.,300.,1000)
     outfilename= 'pal5GC_100_direct.sav'
     if not os.path.exists(outfilename):
         out= numpy.zeros((len(vos),len(qs),len(sigvs),len(sigxs)))
@@ -113,6 +113,7 @@ def fitDirect():
         out= pickle.load(outfile)
         ii= pickle.load(outfile)
         jj= pickle.load(outfile)
+        ii, jj= 14,0
         outfile.close()
     while ii < len(vos):
         progo= progenitorOrbit(V0=_REFV0*vos[ii],R0=_REFR0)
@@ -133,11 +134,12 @@ def fitDirect():
                                                                          q=qs[jj]),
                                   progenitor=progo,ts=ts)
                     out[ii,jj,kk,ll]= sdf(sX,sY,sZ,svX,svY,svZ,
-                                          rect=True,log=True)
+                                          rect=True,log=True)-3.*len(sX)*numpy.log(vos[ii])
             jj+= 1
-            if jj == len(qs):
-                jj= 0
-                ii+= 1
             save_pickles(outfilename,out,ii,jj)
+        if ii != (len(vos)-1):
+            jj= 0
+        ii+= 1
+        save_pickles(outfilename,out,ii,jj)
     save_pickles(outfilename,out,ii,jj)
     return None

### 2013-09-30 py/plot_stream.py

@@ -1,0 +1,152 @@
+import sys
+import os, os.path
+import numpy
+from galpy.util import bovy_plot, bovy_coords
+from galpy import potential
+from galpy.orbit import Orbit
+_STREAMSNAPDIR= '../sim/snaps'
+_STREAMSNAPAADIR= '../sim/snaps_aai'
+def plot_stream_xz(plotfilename):
+    #Read stream
+    data= numpy.loadtxt(os.path.join(_STREAMSNAPDIR,'gd1_evol_hitres_01312.dat'),
+                        delimiter=',')
+    includeorbit= True
+    if includeorbit:
+        npts= 201
+        pot= potential.LogarithmicHaloPotential(normalize=1.,q=0.9)
+        pts= numpy.linspace(0.,4.,npts)
+        #Calculate progenitor orbit around this point
+        pox= numpy.median(data[:,1])
+        poy= numpy.median(data[:,3])
+        poz= numpy.median(data[:,2])
+        povx= numpy.median(data[:,4])
+        povy= numpy.median(data[:,6])
+        povz= numpy.median(data[:,5])
+        pR,pphi,pZ= bovy_coords.rect_to_cyl(pox,poy,poz)
+        pvR,pvT,pvZ= bovy_coords.rect_to_cyl_vec(povx,povy,povz,pR,
+                                                 pphi,pZ,cyl=True)
+        ppo= Orbit([pR/8.,pvR/220.,pvT/220.,pZ/8.,pvZ/220.,pphi])
+        pno= Orbit([pR/8.,-pvR/220.,-pvT/220.,pZ/8.,-pvZ/220.,pphi])
+        ppo.integrate(pts,pot)
+        pno.integrate(pts,pot)
+        pvec= numpy.zeros((3,npts*2-1))
+        pvec[0,:npts-1]= pno.x(pts)[::-1][:-1]
+        pvec[1,:npts-1]= pno.z(pts)[::-1][:-1]
+        pvec[2,:npts-1]= pno.y(pts)[::-1][:-1]
+        pvec[0,npts-1:]= ppo.x(pts)
+        pvec[1,npts-1:]= ppo.z(pts)
+        pvec[2,npts-1:]= ppo.y(pts)
+        pvec*= 8.
+    #Plot
+    bovy_plot.bovy_print()
+    bovy_plot.bovy_plot(data[:,1],data[:,2],'k,',
+                        xlabel=r'$X\,(\mathrm{kpc})$',
[…]

### 2013-09-30 tex/ms.tex
stream: some plots of the stream
streams: action-angle structure of simulation

@@
 
+\begin{figure}[t!]
+  \includegraphics[width=0.5\textwidth,clip=]{gd1_evol_xz.ps}
+  \caption{}\label{fig:gd1_xy}
+%python plot_stream.py ../tex/gd1_evol_xz.ps
+\end{figure}
+
+\begin{figure}[t!]
+  \includegraphics[width=\textwidth,clip=]{gd1_evol_lb.ps}
+  \includegraphics[width=\textwidth,clip=]{gd1_evol_ld.ps}
+  \caption{}\label{fig:gd1_xy}
+%python plot_stream.py ../tex/gd1_evol_lb.ps
+%python plot_stream.py ../tex/gd1_evol_ld.ps
+\end{figure}
+
+\begin{figure}[t!]
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aajrjz.ps}
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaoroz.ps}
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaaraz.ps}\\
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aajrjp.ps}
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaorop.ps}
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaarap.ps}
+  \caption{}\label{fig:gd1_xy}
+%python plot_stream.py ../tex/gd1_evol_aajrjz.ps
+%python plot_stream.py ../tex/gd1_evol_aaoroz.ps
+%python plot_stream.py ../tex/gd1_evol_aaaraz.ps
+%python plot_stream.py ../tex/gd1_evol_aajrjp.ps
+%python plot_stream.py ../tex/gd1_evol_aaorop.ps
+%python plot_stream.py ../tex/gd1_evol_aaarap.ps
+\end{figure}
+
 \end{document}

### 2013-09-30 py/plot_stream.py
streams: tweak to plots

@@ -1,7 +1,7 @@
 import sys
 import os, os.path
 import numpy
-from galpy.util import bovy_plot, bovy_coords
+from galpy.util import bovy_plot, bovy_coords, bovy_conversion
 from galpy import potential
 from galpy.orbit import Orbit
 _STREAMSNAPDIR= '../sim/snaps'
@@ -118,28 +118,110 @@ def plot_stream_aa(plotfilename):
         progfile= '../sim/gd1_evol_hitres_progaai.dat'
         progaa= numpy.loadtxt(progfile,delimiter=',')
     if 'araz' in plotfilename:
-        plotx= data[:,6]
-        ploty= data[:,8]
-        plotz= data[:,7]
-        plotx= (numpy.pi+(plotx-numpy.median(plotx))) % (2.*numpy.pi)
-        ploty= (numpy.pi+(ploty-numpy.median(ploty))) % (2.*numpy.pi)
-        plotz= (numpy.pi+(plotz-numpy.median(plotz))) % (2.*numpy.pi)
+        thetar= data[:,6]
+        thetar= (numpy.pi+(thetar-numpy.median(thetar))) % (2.*numpy.pi)
+        indx= numpy.fabs(thetar-numpy.pi) > (5.*numpy.median(numpy.fabs(thetar-numpy.median(thetar))))
+        plotx= data[indx,6]
+        ploty= data[indx,8]
+        plotx= (numpy.pi+(plotx-numpy.median(data[:,6]))) % (2.*numpy.pi)
+        ploty= (numpy.pi+(ploty-numpy.median(data[:,8]))) % (2.*numpy.pi)
         xrange=[numpy.pi-1.,numpy.pi+1.]
         yrange=[numpy.pi-1.,numpy.pi+1.]
-        zrange=[numpy.pi-.5,numpy.pi+.5]
         xlabel=r'$\theta_R$'
         ylabel=r'$\theta_Z$'
+    elif 'arap' in plotfilename:
+        thetar= data[:,6]
+        thetar= (numpy.pi+(thetar-numpy.median(thetar))) % (2.*numpy.pi)
+        indx= numpy.fabs(thetar-numpy.pi) > (5.*numpy.median(numpy.fabs(thetar-numpy.median(thetar))))
+        plotx= data[indx,6]
+        ploty= data[indx,7]
+        plotx= (numpy.pi+(plotx-numpy.median(data[:,6]))) % (2.*numpy.pi)
[…]

### 2013-11-20 tex/ms.tex
streams: new section 2

@@
 %BOVY: reference point
+
+\section{The dynamics of tidal streams}\label{sec:dynamics}
+
+The dynamics of tidal streams both in real (position--velocity) and
+action--angle space has been described by various previous authors
+\citep[\eg,][]{Helmi99a,Tremaine99a,SomeJohnstonPaper,Sanders13a}. I
+provide here a brief description of the dynamics of tidal streams that
+is relevant for what follows.
+
+In real space, a tidal stream forms as stars are stripped from a
+progenitor some time $\Delta t$ in the past and then evolve (largely)
+independently in a galaxy's host potential, moving away both in
+position and velocity from the progenitor object as time goes
+on. Thus, a star was offset by a small amount $(\Delta \vecx,\Delta
+\vecv)$ from the position of the progenitor $(\vecx^p,\vecv^p)(-\Delta
+t)$ at that time. Then the star orbited under the influence of the
+same gravitational potential $\Phi$ as the progenitor until the
+present day $(t=0)$, when it was observed: $(\vecx,\vecv)(t=0)$ =
+$\vech_{\Delta t}(\vecx^p(-\Delta t)+\Delta \vecx,\vecv^p(-\Delta
+t)+\Delta \vecv)$, where $\vech$ denotes the Hamiltonian flow from
+$t=-\Delta t$ to $t=0$ (see \citealt{binneytremaine}); this is simply
+the orbit of the star between these times. In phase-space this orbit
+can be calculated by solving Hamilton's equations: $\dot{\vecx} =
+\vecv; \dot{\vecv} = - \dd \Phi / \dd \vecx$.
+
+In action--angle space the dynamics of stream formation is much
+simpler. A star received an offset ($\Delta \vecj,\Delta \veca$) at
+time $-\Delta t$ from the action--angle coordinates of the progenitor
+$(\vecj^p,\veca^p)$, similar to the small offset ($\Delta \vecx,\Delta
+\vecv)$ in the real-space description. The offset in the actions
+corresponds to an offset in the frequencies $\Delta \veco$;
+ultimately, this offset is responsible for the spreading of the stream
+over long stretches on the sky. Because both the source's angles and
+the star's angles increase linearly with time, albeit at different
+frequencies, the difference in the angles also increases linearly in
+time. Therefore at time $t=0$ the difference in angles is
+\begin{equation}
+\Delta \veca(t=0) = \Delta \veca(-\Delta t) + \Delta \veco \Delta t\,,
+\end{equation}
+while the difference in actions is constant
+\begin{equation}
+\Delta \vecj(t=0) = \Delta \vecj(-\Delta t)\,.
+\end{equation}
+As the initial offset $\Delta \veca(-\Delta t$ in angles is small
+compared to the growth $\Delta\veco\Delta t$, $\Delta \veca(t=0)
+\approx \Delta \veco \Delta t$.
+
+Thus, in action--angle coordinates, the dynamics of tidal-stream
+formation is simplified to linear growth of small initial angle
+differences due to small action/frequency differences. The angle
+direction $\unitvec{\veco} \equiv \Delta \veco / | \Delta \veco|$
+along which a star moves away from the progenitor is related to the
+initial action offset by $\Delta \veco \approx (\partial^2 H
+/ \partial \vecj \partial \vecj) \Delta \vecj$, in the limit of small
+action offsets, as $\veco = \partial H / \partial \vecj$. For an
+approximately one-dimensional stream to form, the Hessian matrix
+$(\partial^2 H / \partial \vecj \partial \vecj)\large|_{\vecj^p}$ has
+to be dominated by a single large eigenvalue; in this case the angle
+difference of stars stripped from the stream at any time will fall
+along approximately the same direction $\unitvec\veco$, and the stream
+will essentially be one dimensional. In detail, because the
+distribution of initial offsets $\Delta \vecj$ is not isotropic, the
+direction $\unitvec \veco$ is not just determined by the potential
+(through $H$) and the progenitor orbit, but also by the distribution
+of $\Delta \vecj$, which is set through a combination of the internal
+dynamics of the progenitor and the stripping process (see discussion
+in \citealt{Sanders13a}). For dynamical modeling we can be agnostic
+about the relative importance of all of these contributors to
+$\unitvec \veco$ by flexible modeling of the distribution of $\Delta
+\vecj$.
 

### 2013-11-20 tex/ms.tex
streams: more discussion

@@
 \newcommand{\vecj}{\ensuremath{\vec{J}}}
+\newcommand{\vecp}{\ensuremath{\vec{p}}}
 \newcommand{\veco}{\ensuremath{\vec{\Omega}}}
@@
 $(\vecj^p,\veca^p)$, similar to the small offset ($\Delta \vecx,\Delta
-\vecv)$ in the real-space description. The offset in the actions
+\vecv)$ in the real--space description. The offset in the actions
 corresponds to an offset in the frequencies $\Delta \veco$;
 ultimately, this offset is responsible for the spreading of the stream
-over long stretches on the sky. Because both the source's angles and
+over long stretches on the sky. Because both the progenitor's angles and
 the star's angles increase linearly with time, albeit at different
@@
 
+\section{Generative models of tidal streams}\label{sec:method}
+
+The discussions of the dynamics of stream formation in the previous
+section directly give rise of a generative model of a tidal stream. A
+star in a tidal stream is \emph{generated} through being stripped a
+time $\Delta t$ ago from a progenitor $(\vecx^p,\vecv^p) \equiv
+(\vecj^p,\veca^p)$ by receiving a small offset $(\Delta \vecx,\Delta
+\vecv) \equiv (\Delta \vecj,\Delta \veca)$. The offset then evolves
+using the dynamical equations given in the previous section and the
+star is observed at time $t=0$ (\ie, today). Because phase-space
+density is conserved by Liousville's theorem, a probabilistic model
+results by relating the likelihood $p(\vecx,\vecv|\vecp)$ (where
+$\vecp$ denotes the set of all model parameters; see below) for the
+observed phase-space position today to the distribution of initial
+offsets. This probabilistic model can be marginalized over missing
+observational data or extended to include a model for outliers.
+
+\subsection{Real--space modeling}
+
+In more detail, in the real--space approach the model can be specified
+by: (a) the gravitational potential $\Phi$, (b) the current
+phase-space position of the progenitor $(\vecx_p,\vecv_p)(t=0)$, which
+can be integrated backwards in time by reversing the Hamiltonian flow
+defined by $\Phi$, (c) the time $\Delta t_i$ at which each star was
+removed from the progenitor, and (d) the distribution $f(\Delta
+\vecx,\Delta \vecv|\paramsdiff)$, characterized by some parameters
+$\paramsdiff$, of initial phase-space offsets from which the actual
+offsets $(\Delta \vecx_i,\Delta \vecv_i)$ are drawn. In practice, $f$
+can be taken to be a Gaussian, characterized by a width in position
+$\sigma_x$ and in velocity $\sigma_v$, related to the tidal radius and
+escape velocity of the progenitor, respectively. However, more
+complicated models for the initial phase-space distribution of the
+debris can be used here as well.
+
+The probability for observing a star today at $(\vecx,\vecv)$ given
+items (a) through (d) is then
+\begin{align}\label{eq:realdft}
+  p(\vecx_i,\vecv_i | \Phi,\vecx_p,\vecv_p,\Delta t_i,\paramsdiff) 
+  & = p(\vecx_i(-\Delta t),\vecv_i(-\Delta t) | \Phi,\vecx_p,\vecv_p,\Delta t_i,\paramsdiff) \,,\nonumber\\
+  & = p([\vecx_i-\vecx_p](-\Delta t),[\vecv_i-\vecv_p](-\Delta t) | \Phi,\vecx_p,\vecv_p,\Delta t_i,\paramsdiff) \,\nonumber\\
+  & = p(\Delta\vecx_i(-\Delta t),\Delta\vecv_i(-\Delta t) | \Phi,\vecx_p,\vecv_p,\Delta t_i,\paramsdiff)\,,\\
+  & = f(\Delta\vecx_i(-\Delta t),\Delta\vecv_i(-\Delta t) | \Phi,\vecx_p,\vecv_p,\Delta t_i,\paramsdiff)\, ,\nonumber
+\end{align}
+where the first equality holds because the Hamiltonian flow is unique
+and conserves phase-space volume by Liousville's theorem and the
+second one because the Jacobian of the transformation
+$(\vecx_i,\vecv_i)(-\Delta t) \rightarrow (\vecx_i,\vecv_i)(-\Delta t)
+- (\vecx_p,\vecv_p)(-\Delta t)$ is one. The final probability can be
+evaluated using the parameters $\paramsdiff$ following (d).
+
+The time $\Delta t_i$ at which a star was removed is unobservable and
+the probability in \eqnname~(\ref{eq:realdft}) should therefore be
+integrated over this time using a prior $p(\Delta t_i|\ldots)$, where
+$\ldots$ indicates that this prior could depend on other factors such
+as the potential and the progenitor orbit\footnote{For example, one
+  could insist that stars are removed around pericenter
+  passages.}. For simplicity, we can assume a uniform prior for the
+times $\Delta t$. The probability for a single star then becomes
+\begin{align}\label{eq:realdf}
+  p(\vecx_i,\vecv_i | \Phi,\vecx_p,\vecv_p,\paramsdiff) & = \int \dd
+  \Delta t_i \,p(\vecx_i,\vecv_i | \Phi,\vecx_p,\vecv_p,\Delta t_i,
+  \paramsdiff)\,.
+\end{align}
+If the progenitor's current phase-space position is unknown, then this
+probability should be further marginalized over $(\vecx_p,\vecv_p)$;
+for multiple stars, this marginalization is done after multiplying the
+individual-star likelihoods in \eqnname~(\ref{eq:realdf}).
+
+In practice, the likelihood in \eqnname~(\ref{eq:realdf}) can be
+calculated as follows: for a given potential $\Phi$ and progenitor
+$(\vecx_p,\vecv_p)$, we (a) integrate the orbits of all stars
+$(\vecx_i,\vecv_i)$ and (b) that of the progenitor backward in time for
+many Gyr and record the phase-space positions over time on a grid $j$
+in $\Delta^j_i$; (c) we evaluate $f(\Delta \vecx_i,\Delta
+\vecv_i)(\Delta t^j_i)$ for all stars $i$ and times $j$; (d) $f(\Delta
+\vecx_i,\Delta \vecv_i)(\Delta t^j_i)$ is then summed over $j$ and the
+logarithm of that is summed over $i$. The most computationally
+expensive part of this calculation is the integration of the various
+orbits. The evaluation of the likelihood for different parameters
+\paramsdiff\ of $f$ can be efficiently computed by caching the orbit
+integrations of (a) and (b); different progenitor orbits
+$(\vecx_p,\vecv_p)$ can also make use of the stored orbits in (a).
+
+The approach outlined in this section is straightforward, but very
+computationally expensive, as it requires large numbers of orbit
+integrations to explore all plausible removal times $\Delta t$,
+progenitor orbits, and potentials. Moreover, when the data have large
+uncertainties, more orbit integrations are necessary to explore each
+data points uncertainty range. While it might be possible to speed up
+the computation through linear approximations for nearby orbits, the
+action--angle formalism allows for significant and elegant
+simplifications and I consider the real--space approach further.
+
+\subsection{Action--angle stream modeling}
+
+Similar as for the real--space approach, we can write down a
+generative model for the positions and velocities ($\vecx_i,\vecv_i$)
+of stars $i$ in a tidal stream in the action--angle formalism by
+specifying: (a) the gravitational potential $\Phi$, (b) the current
+action-angle coordinates of the progenitor $(\vecj_p,\veca_p)(t=0)$,
+(c) the time $\Delta t_i$ at which each star was removed from the
+progenitor, and (d) the distribution $f(\Delta \vecj,\Delta
+\veca|\paramsdiff)$, characterized by some parameters $\paramsdiff$,
+of initial phase-space offsets from which the actual offsets $(\Delta
+\vecj_i,\Delta \veca_i)$ are drawn. In practice, $f$ can again be
+taken to be a Gaussian, although more general models could take the
+expected `bowtie' structure of the action distribution into account
+(\citealt{Eyre11a}; see also \figurename~\ref{fig:gd1_jt}).
+
+The probability for observing a star today at $(\vecx,\vecv)$ given
+items (a) through (d) is then
+\begin{align}\label{eq:aadft}
+  p(\vecx_i,\vecv_i | \Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff) 
+  & = p(\vecj_i,\veca_i | \Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff) \,,\nonumber\\
+  & = p(\vecj_i-\vecj_p,\veca_i-\veca_p-\Delta \veco_i\Delta t_i | \Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff) \,,\\
+  & = f(\Delta \vecj_i,\Delta \veca_i(-\Delta t_i) | \Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff) \,,\nonumber
+\end{align}
+where the first equality holds because the Jacobian of the
+$(\vecx,\vecv)\rightarrow(\vecj,\veca)$ transformation is unity and
+the second equality is a simple linear transformation. The final
+probability can again be evaluated using the parameters
+\paramsdiff\ following (d).
+
+As in the real--space formalism, we should integrate
+\eqnname~(\ref{eq:aadft}) over the time $\Delta t_i$ at which the star
+is removed from the progenitor (cf.~\eqnname~[\ref{eq:realdf}]):
+\begin{equation}\label{eq:aadf}
+  p(\vecx_i,\vecv_i | \Phi,\vecj_p,\veca_p,\paramsdiff) = \int \dd
+  \Delta t_i\, f(\Delta \vecj_i,\Delta \veca_i(-\Delta t_i) |
+  \Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff) \,.
+\end{equation}
+Without further assumptions the action-angle formalism proceeds in the
+same way as the real--space formalism, evaluating
+\eqnname~(\ref{eq:aadft}) for a grid $j$ of times $\Delta t_i^j$ and
+summing $f(\Delta \vecj_i,\Delta \veca_i(-\Delta t^j_i) |
+\Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff)$ over $j$.
+
+Even without further assumptions about the shape of $f(\Delta
+\vecj,\Delta \veca|\paramsdiff)$, the action-angle formalism already
+allows for a simplification over the real--space method. Because the
+progenitor orbit is specified in action-angle coordinates, it can be
+changed \emph{without} expensive orbit integrations of all possible
+source orbits. As the progenitor's current phase-space position is a
+six-dimensional parameter, this is a major simplification of the
+dynamical model when the source orbit is unknown. Note that this
+simplification does require us to be able to compute $\veco(\vecj)$,
+to calculate the new source frequencies $\veco_p$ when changing the
+source actions $\vecj_p$. This mapping can be constructed using a
+small number of (action,frequency) calculations covering the range of
+possible source actions and is relatively easy to construct due to the
+approximately one-dimensional structure of a stream (see discussion at
+the end of \sectionname~\ref{sec:dynamics}).
+
+We can simplify \eqnname~(\ref{eq:aadf}) by assuming (a) that
+$f(\Delta \vecj,\Delta \veca|\paramsdiff)$ is separable into
+$f_{\veca}(\Delta \veca|\paramsdiff)\,f_{\vecj}(\Delta
+\vecj|\paramsdiff)$ and (b) that the distribution $f_{\veca}(\Delta
+\veca|\paramsdiff)$ is Gaussian. In what follows, we will assume that
+this Gaussian is isotropic with dispersion $\sigma_\theta$, but all of
+the results are the same after appropriate diagonalization and
+rescaling of the angles for more general covariance matrices. In this
+case, \eqnname~(\ref{eq:aadf}) can be written as
+\begin{equation}
+\begin{split}
+  p(\vecx_i,\vecv_i & | \Phi,\vecj_p,\veca_p,\paramsdiff) 
+   = \\
+   & f_{\vecj}(\Delta \vecj_i|\paramsdiff) \int \dd \Delta t_i\,
+  \frac{1}{(2\,\pi)^{3/2}\,\sigma_\theta^3}\,\exp\left(-\frac{1}{2\,\sigma_\theta^2}\left[\sum_{R,\phi,Z}\left(\Delta \veca_i-\Delta \veco_i\Delta t_i\right)^2\right]\right)\,,
+\end{split}
+\end{equation}
+where the sum in the argument of the exponential is over the three
+components of $\Delta \veca_i$ and $\Delta \veco_i$. This integral can
+be done analytically, resulting in
+\begin{equation}\label{eq:aadfsimple}
+\begin{split}
+  p(\vecx_i,\vecv_i & | \Phi,\vecj_p,\veca_p,\paramsdiff) 
+   = \\
+   & f_{\vecj}(\Delta \vecj_i|\paramsdiff) \,  \frac{1}{2\,\pi\,\sigma_\theta^2\,\sqrt{\sum_{R,\phi,Z}\Delta \veco_i^2}}\,\exp\left(-\frac{1}{2\,\sigma_\theta^2}\left[\sum_{R,\phi,Z}\Delta \veca^2_i-\frac{\left(\sum_{R,\phi,Z} \Delta \veco_i\Delta \veca_i\right)^2}{\sum_{R,\phi,Z}\Delta \veco_i^2}\right]\right)\,.
+\end{split}
+\end{equation}
+In this expression, $\Delta X^2 \equiv (X_1-X_2)^2$. The expression in
+square brackets is positive by using the Cauchy-Schwarz inequality.
+
+The quantity in the square brackets in \eqnname~(\ref{eq:aadfsimple})
+expresses the fact that the present-day difference in angles between a
+star in a tidal stream and the source must be linearly related by a
+scalar (the time since the removal of the star) to the difference in
+frequencies, within a tolerance set by the width of the distribution
+of initial angle differences ($\sigma_\theta$). The ratio
+$\sum_{R,\phi,Z} \Delta \veco_i\Delta \veca_i/\sum_{R,\phi,Z}\Delta
+\veco_i^2$ is the best-guess for $\Delta t_i$ when performing a linear
+regression between $\Delta \veca_i$ and $\Delta \veco_i$. The
+expression in square brackets is minimized when the correlation
+between $\Delta \veca_i$ and $\Delta \veco_i$ is the largest, \ie,
+when a single time difference $\Delta t_i$ explains all three of the
+angle differences.
+
+A detail that is ignored in \equationname~(\ref{eq:aadfsimple}) is
+that the uniform prior on $\Delta t$ should only have support for
+$\Delta t > 0$, \ie, for times in the past, which introduces an
+additional factor of $[1+\mathrm{erf}(t_0)]$, with
+\[ t_0 = \frac{\sqrt{\sum_{R,\phi,Z}\Delta
+  \veco_i^2}}{\sqrt{2}\,\sigma_\theta}\,\frac{\sum_{R,\phi,Z} \Delta
+  \veco_i\Delta \veca_i}{\sum_{R,\phi,Z}\Delta \veco_i^2}\,.\] Because
+the second factor on the right-hand side is the best guess for $\Delta
+t_i$, if this is large and negative (\ie, the star was stripped
+\emph{in the future}), the probability of the star being stripped from
+the progenitor is strongly suppressed. Only when it appears that the
+star was only recently stripped with $\Delta t_i$ negative, but
+$\approx \sigma_\theta/\sqrt{\sum_{R,\phi,Z}\Delta \veco_i^2}$, is the
+suppression for negative stripping times less.
+
+The discussion so far has been entirely general and applies to any
+sort of stellar debris that originates from a single orbit, without
+necessarily generating a one-dimensional structure. In most
+observations of tidal streams, however, they appear as one-dimensional
+objects rather than the six-dimensional objects that we have
+considered so far. The reason for this is both physical and
+observational: (a) narrow tidal streams form because a star separates
+from the progenitor over time mostly along a single direction in
+angle-space: the Hessian matrix of the Hamiltonian, $\partial^2 H /
+\partial \vecj \partial \vecj$, has a single large eigenvalue, which
+determines the direction along which the stream is disrupted in
+angle-space (because $\partial \veco / \partial \vecj$ is equal to the
+Hessian); (b) because many tidal streams are dynamically cold with
+velocity dispersions $\lesssim10\kms$, observations are unable to
+resolve the angle differences along the two eigenvectors of the
+Hessian corresponding to the two smallest eigenvalues or the spread in
+actions among stream members. If the latter can be resolved, the
+stream is four-dimensional. In what follows we will be mainly
+concerned with the stream only spreading along a single direction in
+angle space.
+
+If the stream is one-dimensional in angle space then the time since
+the removal from the progenitor becomes unobservable if the progenitor
+orbit is unknown: above, determining the time depended on explaining
+the three-dimensional angle difference $\Delta \veca$ with the
+three-dimensional frequency difference $\Delta \veco$ using a single
+time difference; if two of the angle and frequency differences are
+unable to be resolved, then the frequency difference and the time
+difference become degenerate (aka, you can draw a straight line
+connecting any two points). To make use of this fact to simplify the
+dynamical model, we write the frequency difference as
+\begin{equation}
+  \Delta \veco = \Delta \Omega_1\,\unitvec{e}_1\,,
+\end{equation}
+where $\unitvec{e}_1$ is a unit vector. Then we write the angle
+difference $\Delta \veca$ 
+\begin{equation}
+  \Delta \veca = \Delta \theta_1\,\unitvec{e}_1 + \Delta \veca_\perp\,\unitvec{e}_1^\perp\,,
+\end{equation}
+where$\unitvec{e}_1^\perp$ is the orthogonal complement of
+$\unitvec{e}_1$ (the two directions perpendicular to $\unitvec{e}_1$)
+and $\Delta \veca_\perp$ are the components of $\Delta \veca$ in that
+direction. We can then write \eqnname~(\ref{eq:aadfsimple}) as
+\begin{equation}\label{eq:aadfsimpler}
+\begin{split}
+  p(\vecx_i,\vecv_i & | \Phi,\vecj_p,\veca_p,\paramsdiff) 
+   = \\
+   & f_{\vecj}(\Delta \vecj_i|\paramsdiff) \,  \frac{1}{2\,\pi\,\sigma_\theta^2\,\sqrt{\sum_{R,\phi,Z}\Delta \veco_i^2}}\,\exp\left(-\frac{|\Delta\veca_{i,\perp}|^2}{2\,\sigma_\theta^2}\right)\,.
+\end{split}
+\end{equation}
+\Eqnname~(\ref{eq:aadfsimpler}) is equivalent to
+\eqnname~(\ref{eq:aadfsimple}), but it expresses the dynamical model
+in a slightly different way. Whereas the angle part of
+\eqnname~(\ref{eq:aadfsimple}) allows the correct potential and
+progenitor to be inferred by asking whether a single time difference
+explains the three angle differences given the three frequencies,
+\eqnname~(\ref{eq:aadfsimpler}) `rewinds' the star's orbit along the
+direction of $\Delta \veco$ and asks whether the angle difference
+perpendicular to $\Delta \veco$ is consistent with the initial angle
+spread (because it has not evolved since the removal of the star). In
+re-writing this, we have not yet used the approximation that the
+stream only forms along a single direction in $\Delta \veco$.
+
 \section{Two equivalent generative models of tidal streams}\label{sec:method}
@@
 
-\subsection{Real-space model}\label{sec:realmethod}
+\subsection{Real--space model}\label{sec:realmethod}
 
@@
 specified by: (a) the gravitational potential $\Phi$, (b) the current
-phase-space position of the source $(\vecx_s,\vecv_s)(t=0)$, which can
+phase-space position of the source $(\vecx_p,\vecv_p)(t=0)$, which can
 be integrated backwards in time by reversing the Hamiltonian flow
@@
 \begin{align}\label{eq:realdft}
-  p(\vecx_i,\vecv_i | \Phi,\vecx_s,\vecv_s,\Delta t_i,\paramsdiff) 
-  & = p(\vecx_i(-\Delta t),\vecv_i(-\Delta t) | \Phi,\vecx_s,\vecv_s,\Delta t_i,\paramsdiff) \,,\nonumber\\
-  & = p([\vecx_i-\vecx_s](-\Delta t),[\vecv_i-\vecv_s](-\Delta t) | \Phi,\vecx_s,\vecv_s,\Delta t_i,\paramsdiff) \,\nonumber\\
-  & = p(\Delta\vecx_i(-\Delta t),\Delta\vecv_i(-\Delta t) | \Phi,\vecx_s,\vecv_s,\Delta t_i,\paramsdiff)\,,\\
-  & = f(\Delta\vecx_i(-\Delta t),\Delta\vecv_i(-\Delta t) | \Phi,\vecx_s,\vecv_s,\Delta t_i,\paramsdiff)\, ,\nonumber
+  p(\vecx_i,\vecv_i | \Phi,\vecx_p,\vecv_p,\Delta t_i,\paramsdiff) 
+  & = p(\vecx_i(-\Delta t),\vecv_i(-\Delta t) | \Phi,\vecx_p,\vecv_p,\Delta t_i,\paramsdiff) \,,\nonumber\\
+  & = p([\vecx_i-\vecx_p](-\Delta t),[\vecv_i-\vecv_p](-\Delta t) | \Phi,\vecx_p,\vecv_p,\Delta t_i,\paramsdiff) \,\nonumber\\
+  & = p(\Delta\vecx_i(-\Delta t),\Delta\vecv_i(-\Delta t) | \Phi,\vecx_p,\vecv_p,\Delta t_i,\paramsdiff)\,,\\
+  & = f(\Delta\vecx_i(-\Delta t),\Delta\vecv_i(-\Delta t) | \Phi,\vecx_p,\vecv_p,\Delta t_i,\paramsdiff)\, ,\nonumber
 \end{align}
@@
 $(\vecx_i,\vecv_i)(-\Delta t) \rightarrow (\vecx_i,\vecv_i)(-\Delta t)
-- (\vecx_s,\vecv_s)(-\Delta t)$ is one. The final probability can be
+- (\vecx_p,\vecv_p)(-\Delta t)$ is one. The final probability can be
 evaluated using the parameters $\paramsdiff$ following (d).
@@
 \begin{align}\label{eq:realdf}
-  p(\vecx_i,\vecv_i | \Phi,\vecx_s,\vecv_s,\paramsdiff) & = \int \dd
-  \Delta t_i \,p(\vecx_i,\vecv_i | \Phi,\vecx_s,\vecv_s,\Delta t_i,
+  p(\vecx_i,\vecv_i | \Phi,\vecx_p,\vecv_p,\paramsdiff) & = \int \dd
+  \Delta t_i \,p(\vecx_i,\vecv_i | \Phi,\vecx_p,\vecv_p,\Delta t_i,
   \paramsdiff)\,.
@@
 If the progenitor's current phase-space position is unknown, then this
-probability should be further marginalized over $(\vecx_s,\vecv_s)$;
+probability should be further marginalized over $(\vecx_p,\vecv_p)$;
 for multiple stars, this marginalization is done after multiplying the
@@
 calculated as follows: for a given potential $\Phi$ and source
-$(\vecx_s,\vecv_s)$, we (a) integrate the orbits of all stars
+$(\vecx_p,\vecv_p)$, we (a) integrate the orbits of all stars
 $(\vecx_i,\vecv_i)$ and (b) that of the source backward in time for
@@
 integrations of (a) and (b); different source orbits
-$(\vecx_s,\vecv_s)$ can also make use of the stored orbits in (a).
+$(\vecx_p,\vecv_p)$ can also make use of the stored orbits in (a).
 
@@
 a small offset $(\Delta \vecj,\Delta \veca)(-\Delta t)$ from the
-source's actions $\vecj_s$ and angles $\veca_s$ a time $\Delta t$ ago,
+source's actions $\vecj_p$ and angles $\veca_p$ a time $\Delta t$ ago,
 similar to the small offset $(\Delta \vecx,\Delta \vecv)(-\Delta t)$
-in the real-space model. The offset in the actions corresponds to an
+in the real--space model. The offset in the actions corresponds to an
 offset in the frequencies $\Delta \veco$; ultimately, this offset is
@@
 gravitational potential $\Phi$, (b) the current action-angle
-coordinates of the source $(\vecj_s,\veca_s)(t=0)$, (c) the time
+coordinates of the source $(\vecj_p,\veca_p)(t=0)$, (c) the time
 $\Delta t_i$ at which each star was removed from the source, and (d)
@@
 \begin{align}\label{eq:aadft}
-  p(\vecx_i,\vecv_i | \Phi,\vecj_s,\veca_s,\Delta t_i,\paramsdiff) 
-  & = p(\vecj_i,\veca_i | \Phi,\vecj_s,\veca_s,\Delta t_i,\paramsdiff) \,,\nonumber\\
-  & = p(\vecj_i-\vecj_s,\veca_i-\veca_s-\Delta \veco_i\Delta t_i | \Phi,\vecj_s,\veca_s,\Delta t_i,\paramsdiff) \,,\\
-  & = f(\Delta \vecj_i,\Delta \veca_i(-\Delta t_i) | \Phi,\vecj_s,\veca_s,\Delta t_i,\paramsdiff) \,,\nonumber
+  p(\vecx_i,\vecv_i | \Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff) 
+  & = p(\vecj_i,\veca_i | \Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff) \,,\nonumber\\
+  & = p(\vecj_i-\vecj_p,\veca_i-\veca_p-\Delta \veco_i\Delta t_i | \Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff) \,,\\
+  & = f(\Delta \vecj_i,\Delta \veca_i(-\Delta t_i) | \Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff) \,,\nonumber
 \end{align}
@@
 
-As in the real-space formalism, we should integrate
+As in the real--space formalism, we should integrate
 \eqnname~(\ref{eq:aadft}) over the time $\Delta t_i$ at which the star
@@
 \begin{equation}\label{eq:aadf}
-  p(\vecx_i,\vecv_i | \Phi,\vecj_s,\veca_s,\paramsdiff) = \int \dd
+  p(\vecx_i,\vecv_i | \Phi,\vecj_p,\veca_p,\paramsdiff) = \int \dd
   \Delta t_i\, f(\Delta \vecj_i,\Delta \veca_i(-\Delta t_i) |
-  \Phi,\vecj_s,\veca_s,\Delta t_i,\paramsdiff) \,.
+  \Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff) \,.
 \end{equation}
 Without further assumptions the action-angle formalism proceeds in the
-same way as the real-space formalism, evaluating
+same way as the real--space formalism, evaluating
 \eqnname~(\ref{eq:aadft}) for a grid $j$ of times $\Delta t_i^j$ and
 summing $f(\Delta \vecj_i,\Delta \veca_i(-\Delta t^j_i) |
-\Phi,\vecj_s,\veca_s,\Delta t_i,\paramsdiff)$ over $j$.
+\Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff)$ over $j$.
 
@@
 \vecj,\Delta \veca|\paramsdiff)$, the action-angle formalism already
-allows for a simplification over the real-space method. Because the
+allows for a simplification over the real--space method. Because the
 source orbit is specified in action-angle coordinates, it can be
@@
 simplification does require us to be able to compute $\veco(\vecj)$,
-to calculate the new source frequencies $\veco_s$ when changing the
-source actions $\vecj_s$. This mapping can be constructed using a
+to calculate the new source frequencies $\veco_p$ when changing the
+source actions $\vecj_p$. This mapping can be constructed using a
 small number of (action,frequency) calculations covering the range of
@@
 \begin{split}
-  p(\vecx_i,\vecv_i & | \Phi,\vecj_s,\veca_s,\paramsdiff) 
+  p(\vecx_i,\vecv_i & | \Phi,\vecj_p,\veca_p,\paramsdiff) 
    = \\
@@
   \includegraphics[width=\textwidth,clip=]{gd1_evol_ld.ps}
-  \caption{}\label{fig:gd1_xy}
+  \caption{}\label{fig:gd1_lbd}
 %python plot_stream.py ../tex/gd1_evol_lb.ps
@@
   \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaarap.ps}
-  \caption{}\label{fig:gd1_xy}
+  \caption{}\label{fig:gd1_jt}
 %python plot_stream.py ../tex/gd1_evol_aajrjz.ps

### 2013-11-20 tex/ms.tex
streams: change t_0

@@
 $\Delta t > 0$, \ie, for times in the past, which introduces an
-additional factor of $[1+\mathrm{erf}(t_0)]$, with
-\[ t_0 = \frac{\sqrt{\sum_{R,\phi,Z}\Delta
+additional factor of $[1+\mathrm{erf}(a_0)]$, with
+\[ a_0 = \frac{\sqrt{\sum_{R,\phi,Z}\Delta
   \veco_i^2}}{\sqrt{2}\,\sigma_\theta}\,\frac{\sum_{R,\phi,Z} \Delta

### 2013-11-30 py/plot_stream.py
streams: plot times

@@ -225,10 +225,98 @@ def plot_stream_aa(plotfilename):
                             zorder=0)
     bovy_plot.bovy_end_print(plotfilename)
 
+def plot_stream_times(plotfilename):
+    #Read stream
+    data= numpy.loadtxt(os.path.join(_STREAMSNAPAADIR,
+                                     'gd1_evol_hitres_aa_01312.dat'),
+                        delimiter=',')
+    #Calculate times at which stars were stripped, angles
+    thetar= data[:,6]
+    thetar= (numpy.pi+(thetar-numpy.median(thetar))) % (2.*numpy.pi)
+    indx= numpy.fabs(thetar-numpy.pi) > (5.*numpy.median(numpy.fabs(thetar-numpy.median(thetar))))
+    thetar= thetar[indx]
+    thetap= data[:,7]
+    thetap= (numpy.pi+(thetap-numpy.median(thetap))) % (2.*numpy.pi)
+    thetap= thetap[indx]
+    thetaz= data[:,8]
+    thetaz= (numpy.pi+(thetaz-numpy.median(thetaz))) % (2.*numpy.pi)
+    thetaz= thetaz[indx]
+    #center around 0 (instead of pi)
+    thetar-= numpy.pi
+    thetap-= numpy.pi
+    thetaz-= numpy.pi
+    #Frequencies
+    Or= data[:,3]
+    Op= data[:,4]
+    Oz= data[:,5]
+    dOr= Or[indx]-numpy.median(Or)
+    dOp= Op[indx]-numpy.median(Op)
+    dOz= Oz[indx]-numpy.median(Oz)
+    #Times
+    dangle= numpy.vstack((thetar,thetap,thetaz))
+    dO= numpy.vstack((dOr,dOp,dOz))*bovy_conversion.freq_in_Gyr(220.,8.)
+    dts= numpy.sum(dO*dangle,axis=0)/numpy.sum(dO**2.,axis=0)
+    if 'hist' in plotfilename:
+        bovy_plot.bovy_print()
+        bovy_plot.bovy_hist(dts,range=[0.,6.],bins=13,
+                            xlabel=r'$\Delta t\ (\mathrm{Gyr})$',
+                            histtype='step',color='k',normed=True,lw=2.)
+        bovy_plot.bovy_hist(dts,range=[0.,6.],bins=61,normed=True,
+                            overplot=True,ls='dotted',histtype='step',
+                            color='0.5',lw=2.)
[…]

### 2013-12-02 py/plot_stream.py
streams: plot vlos

@@ -56,6 +56,11 @@ def plot_stream_lb(plotfilename):
     #Transform to (l,b)
     XYZ= bovy_coords.galcenrect_to_XYZ(data[:,1],data[:,3],data[:,2],Xsun=8.)
     lbd= bovy_coords.XYZ_to_lbd(XYZ[0],XYZ[1],XYZ[2],degree=True)
+    vXYZ= bovy_coords.galcenrect_to_vxvyvz(data[:,4],data[:,6],data[:,5],
+                                           vsun=[0.,30.24*8.,0.])
+    vlbd= bovy_coords.vxvyvz_to_vrpmllpmbb(vXYZ[0],vXYZ[1],vXYZ[2],
+                                           lbd[:,0],lbd[:,1],lbd[:,2],
+                                           degree=True)
     includeorbit= True
     if includeorbit:
         npts= 201
@@ -75,36 +80,64 @@ def plot_stream_lb(plotfilename):
         pno= Orbit([pR/8.,-pvR/220.,-pvT/220.,pZ/8.,-pvZ/220.,pphi])
         ppo.integrate(pts,pot)
         pno.integrate(pts,pot)
-        pvec= numpy.zeros((3,npts*2-1))
+        pvec= numpy.zeros((6,npts*2-1))
         pvec[0,:npts-1]= pno.x(pts)[::-1][:-1]
         pvec[1,:npts-1]= pno.z(pts)[::-1][:-1]
         pvec[2,:npts-1]= pno.y(pts)[::-1][:-1]
         pvec[0,npts-1:]= ppo.x(pts)
         pvec[1,npts-1:]= ppo.z(pts)
         pvec[2,npts-1:]= ppo.y(pts)
-        pvec*= 8.
+        pvec[3,:npts-1]= -pno.vx(pts)[::-1][:-1]
+        pvec[4,:npts-1]= -pno.vz(pts)[::-1][:-1]
+        pvec[5,:npts-1]= -pno.vy(pts)[::-1][:-1]
+        pvec[3,npts-1:]= ppo.vx(pts)
+        pvec[4,npts-1:]= ppo.vz(pts)
+        pvec[5,npts-1:]= ppo.vy(pts)
+        pvec[:3,:]*= 8.
+        pvec[3:,:]*= 220.
         pXYZ= bovy_coords.galcenrect_to_XYZ(pvec[0,:],pvec[2,:],pvec[1,:],
[…]

### 2013-12-02 py/plot_stream.py
streams: plot dO histogram

@@ -1,6 +1,8 @@
 import sys
 import os, os.path
+import copy
 import numpy
+from scipy import optimize, special
 from galpy.util import bovy_plot, bovy_coords, bovy_conversion
 from galpy import potential
 from galpy.orbit import Orbit
@@ -214,6 +216,42 @@ def plot_stream_aa(plotfilename):
         yrange=[-14.64,-14.23]
         xlabel=r'$J_R\,(220\,\mathrm{km\,s}^{-1}\,\mathrm{kpc})$'
         ylabel=r'$L_Z\,(220\,\mathrm{km\,s}^{-1}\,\mathrm{kpc})$'
+    elif 'dohist' in plotfilename:
+        thetar= data[:,6]
+        thetar= (numpy.pi+(thetar-numpy.median(thetar))) % (2.*numpy.pi)
+        indx= numpy.fabs(thetar-numpy.pi) > (5.*numpy.median(numpy.fabs(thetar-numpy.median(thetar))))
+        #Frequencies
+        Or= data[:,3]
+        Op= data[:,4]
+        Oz= data[:,5]
+        dOr= Or[indx]-numpy.median(Or)
+        dOp= Op[indx]-numpy.median(Op)
+        dOz= Oz[indx]-numpy.median(Oz)
+        dO= numpy.vstack((dOr,dOp,dOz))*bovy_conversion.freq_in_Gyr(220.,8.)
+        dO4dir= copy.copy(dO)
+        dO4dir[:,dO4dir[:,0] < 0.]*= -1.
+        dOdir= numpy.median(dO4dir,axis=1)
+        dOdir/= numpy.sqrt(numpy.sum(dOdir**2.))
+        dO1d= numpy.dot(dOdir,dO)
+        dO1d[dO1d < 0.]*= -1.
+        bovy_plot.bovy_print()
+        bovy_plot.bovy_hist(dO1d,range=[0.,0.4],bins=61,
+                           normed=True,
+                           xlabel=r'$|\Delta \mathbf{\Omega}|\,(\mathrm{Gyr}^{-1})$',
+                            histtype='step',color='k',zorder=10)
+        #Overplot best-fit Gaussian
+        xs= numpy.linspace(0.,0.4,1001)
+        bovy_plot.bovy_plot(xs,1./numpy.sqrt(2.*numpy.pi)/numpy.std(dO1d)\
+                                *numpy.exp(-(xs-numpy.mean(dO1d))**2./2./numpy.var(dO1d)),
[…]

### 2013-12-03 py/plot_stream.py
streams: add delta angle histogram

@@ -252,6 +252,66 @@ def plot_stream_aa(plotfilename):
                             '-',color='0.4',overplot=True,lw=2.,zorder=0)
         bovy_plot.bovy_end_print(plotfilename)
         return None
+    elif 'dahist' in plotfilename:
+        thetar= data[:,6]
+        thetar= (numpy.pi+(thetar-numpy.median(thetar))) % (2.*numpy.pi)
+        indx= numpy.fabs(thetar-numpy.pi) > (5.*numpy.median(numpy.fabs(thetar-numpy.median(thetar))))
+        thetar= thetar[indx]
+        thetap= data[:,7]
+        thetap= (numpy.pi+(thetap-numpy.median(thetap))) % (2.*numpy.pi)
+        thetap= thetap[indx]
+        thetaz= data[:,8]
+        thetaz= (numpy.pi+(thetaz-numpy.median(thetaz))) % (2.*numpy.pi)
+        thetaz= thetaz[indx]
+        #center around 0 (instead of pi)
+        thetar-= numpy.pi
+        thetap-= numpy.pi
+        thetaz-= numpy.pi
+        #Frequencies
+        Or= data[:,3]
+        Op= data[:,4]
+        Oz= data[:,5]
+        dOr= Or[indx]-numpy.median(Or)
+        dOp= Op[indx]-numpy.median(Op)
+        dOz= Oz[indx]-numpy.median(Oz)
+        dO= numpy.vstack((dOr,dOp,dOz))*bovy_conversion.freq_in_Gyr(220.,8.)
+        #Direction in which the stream spreads
+        dO4dir= copy.copy(dO)
+        dO4dir[:,dO4dir[:,0] < 0.]*= -1.
+        dOdir= numpy.median(dO4dir,axis=1)
+        dOdir/= numpy.sqrt(numpy.sum(dOdir**2.))
+        #Gram-Schmidt to get the perpendicular directions
+        v2= numpy.array([1.,0.,0.])
+        v3= numpy.array([0.,1.,0.])
+        u2= v2-numpy.sum(dOdir*v2)*dOdir
+        u2/= numpy.sqrt(numpy.sum(u2**2.))
+        u3= v3-numpy.sum(dOdir*v3)*dOdir-numpy.sum(u2*v3)*u2
+        #Times
+        dangle= numpy.vstack((thetar,thetap,thetaz))
[…]

### 2013-12-06 py/plot_stream.py
streams: plot quantities as a function of stream position

@@ -148,6 +148,7 @@ def plot_stream_aa(plotfilename):
                                      'gd1_evol_hitres_aa_01312.dat'),
                         delimiter=',')
     includeorbit= True
+    fmt= 'k,'
     if includeorbit:
         #Read progenitor actions
         progfile= '../sim/gd1_evol_hitres_progaai.dat'
@@ -164,7 +165,7 @@ def plot_stream_aa(plotfilename):
         yrange=[numpy.pi-1.,numpy.pi+1.]
         xlabel=r'$\theta_R$'
         ylabel=r'$\theta_Z$'
-    elif 'arap' in plotfilename:
+    elif 'arap' in plotfilename and not 'aparaperp' in plotfilename:
         thetar= data[:,6]
         thetar= (numpy.pi+(thetar-numpy.median(thetar))) % (2.*numpy.pi)
         indx= numpy.fabs(thetar-numpy.pi) > (5.*numpy.median(numpy.fabs(thetar-numpy.median(thetar))))
@@ -312,8 +313,159 @@ def plot_stream_aa(plotfilename):
         print "perpendicular 2", numpy.mean(newdangle[2,:]), numpy.std(newdangle[2,:])
         bovy_plot.bovy_end_print(plotfilename)
         return None
+    elif 'aparopar' in plotfilename:
+        thetar= data[:,6]
+        thetar= (numpy.pi+(thetar-numpy.median(thetar))) % (2.*numpy.pi)
+        indx= numpy.fabs(thetar-numpy.pi) > (5.*numpy.median(numpy.fabs(thetar-numpy.median(thetar))))
+        thetar= thetar[indx]
+        thetap= data[:,7]
+        thetap= (numpy.pi+(thetap-numpy.median(thetap))) % (2.*numpy.pi)
+        thetap= thetap[indx]
+        thetaz= data[:,8]
+        thetaz= (numpy.pi+(thetaz-numpy.median(thetaz))) % (2.*numpy.pi)
+        thetaz= thetaz[indx]
+        #center around 0 (instead of pi)
+        thetar-= numpy.pi
+        thetap-= numpy.pi
+        thetaz-= numpy.pi
+        #Frequencies
+        Or= data[:,3]
+        Op= data[:,4]
+        Oz= data[:,5]
[…]

### 2013-12-06 tex/ms.tex
streams: add times figures to ms
streams: add vlos
streams: add plots as a function of angle along the stream

@@
   \includegraphics[width=\textwidth,clip=]{gd1_evol_ld.ps}
+  \includegraphics[width=\textwidth,clip=]{gd1_evol_lvlos.ps}
   \caption{}\label{fig:gd1_lbd}
@@
 %python plot_stream.py ../tex/gd1_evol_ld.ps
+%python plot_stream.py ../tex/gd1_evol_lvlos.ps
 \end{figure}
@@
 
+\begin{figure}[t!]
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_timeshist.ps}
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_timesdO.ps}
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_timesda.ps}
+  \caption{}\label{fig:gd1_times}
+%python plot_stream.py ../tex/gd1_evol_timeshist.ps
+%python plot_stream.py ../tex/gd1_evol_timesdO.ps
+%python plot_stream.py ../tex/gd1_evol_timesda.ps
+\end{figure}
+
+\begin{figure}[t!]
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aadohist.ps}
+  \caption{}\label{fig:gd1_dohist}
+%python plot_stream.py ../tex/gd1_evol_aadohist.ps
+\end{figure}
+
+\begin{figure}[t!]
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaaparopar.ps}
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaaparoperp.ps}\\
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaaparaperp.ps}
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaapartime.ps}
+  \caption{}\label{fig:gd1_apar}
+%python plot_stream.py ../tex/gd1_evol_aaaparopar.ps
+%python plot_stream.py ../tex/gd1_evol_aaaparoperp.ps
+%python plot_stream.py ../tex/gd1_evol_aaaparaperp.ps
+%python plot_stream.py ../tex/gd1_evol_aaapartime.ps
+\end{figure}
+
 \end{document}

### 2013-12-06 py/plot_stream.py
streams: fix label

@@ -424,7 +424,7 @@ def plot_stream_aa(plotfilename):
         xrange=[0.,1.3]
         yrange=[0.,0.03]
         xlabel= r'$\Large|\Delta \mathbf{\theta}_\parallel\Large|$'
-        ylabel= r'$\Large|\Delta \mathbf{\theta}_\perp\Large|\,(\mathrm{Gyr}^{-1})$'
+        ylabel= r'$\Large|\Delta \mathbf{\theta}_\perp\Large|$'
         fmt= 'k.'
     elif 'apartime' in plotfilename:
         thetar= data[:,6]

### 2013-12-20 py/plot_stream.py
streams: xz plot with inset

@@ -6,8 +6,14 @@ from scipy import optimize, special
 from galpy.util import bovy_plot, bovy_coords, bovy_conversion
 from galpy import potential
 from galpy.orbit import Orbit
+from galpy.actionAngle_src.actionAngleIsochroneApprox\
+    import actionAngleIsochroneApprox
+from galpy.df_src.streamdf import streamdf
+from matplotlib import pyplot
+from matplotlib.ticker import NullFormatter
 _STREAMSNAPDIR= '../sim/snaps'
 _STREAMSNAPAADIR= '../sim/snaps_aai'
+_NTRACKCHUNKS= 4
 def plot_stream_xz(plotfilename):
     #Read stream
     data= numpy.loadtxt(os.path.join(_STREAMSNAPDIR,'gd1_evol_hitres_01312.dat'),
@@ -39,16 +45,60 @@ def plot_stream_xz(plotfilename):
         pvec[1,npts-1:]= ppo.z(pts)
         pvec[2,npts-1:]= ppo.y(pts)
         pvec*= 8.
+    includetrack= True
+    if includetrack:
+        #Setup stream model
+        lp= potential.LogarithmicHaloPotential(q=0.9,normalize=1.)
+        aAI= actionAngleIsochroneApprox(b=0.8,pot=lp)
+        obs= numpy.array([1.56148083,0.35081535,-1.15481504,
+                          0.88719443,-0.47713334,0.12019596])
+        sdf= streamdf(0.3/220.,progenitor=Orbit(obs),pot=lp,aA=aAI,
+                      leading=True,nTrackChunks=_NTRACKCHUNKS)
+        sdft= streamdf(0.3/220.,progenitor=Orbit(obs),pot=lp,aA=aAI,
+                       leading=False,nTrackChunks=_NTRACKCHUNKS)
     #Plot
     bovy_plot.bovy_print()
     bovy_plot.bovy_plot(data[:,1],data[:,2],'k,',
                         xlabel=r'$X\,(\mathrm{kpc})$',
                         ylabel=r'$Z\,(\mathrm{kpc})$',
                         xrange=[0.,16.],
-                        yrange=[-0.5,12.])
+                        yrange=[-0.5,11.])
     if includeorbit:
         bovy_plot.bovy_plot(pox,poz,'o',color='0.5',mec='none',overplot=True,ms=8)
[…]

### 2013-12-20 py/plot_stream.py
streams: add inset for ll bb figure

@@ -155,6 +155,17 @@ def plot_stream_lb(plotfilename):
         pvlbd= bovy_coords.vxvyvz_to_vrpmllpmbb(pvXYZ[0],pvXYZ[1],pvXYZ[2],
                                                 plbd[:,0],plbd[:,1],plbd[:,2],
                                                 degree=True)
+    includetrack= True
+    if includetrack:
+        #Setup stream model
+        lp= potential.LogarithmicHaloPotential(q=0.9,normalize=1.)
+        aAI= actionAngleIsochroneApprox(b=0.8,pot=lp)
+        obs= numpy.array([1.56148083,0.35081535,-1.15481504,
+                          0.88719443,-0.47713334,0.12019596])
+        sdf= streamdf(0.3/220.,progenitor=Orbit(obs),pot=lp,aA=aAI,
+                      leading=True,nTrackChunks=_NTRACKCHUNKS)
+        sdft= streamdf(0.3/220.,progenitor=Orbit(obs),pot=lp,aA=aAI,
+                       leading=False,nTrackChunks=_NTRACKCHUNKS)
     #Plot
     bovy_plot.bovy_print(fig_width=8.25,fig_height=3.5)
     if 'ld' in plotfilename:
@@ -190,6 +201,52 @@ def plot_stream_lb(plotfilename):
             bovy_plot.bovy_plot(plbd[npts,0],plbd[npts,lbindx],
                                 'o',color='0.5',mec='none',overplot=True,ms=8)
             bovy_plot.bovy_plot(plbd[:,0],plbd[:,lbindx],'k--',overplot=True)
+    if includetrack:
+        d1= 'll'
+        if 'vlos' in plotfilename:
+            d2= 'vlos'
+        elif 'ld'  in plotfilename:
+            d2= 'dist'
+        else:
+            d2= 'bb'
+        sdf.plotTrack(d1=d1,d2=d2,interp=True,color='k',spread=0,
+                      overplot=True,lw=1.)
+        sdft.plotTrack(d1=d1,d2=d2,interp=True,color='k',spread=0,
+                       overplot=True,lw=1.)
+        #Insets
+        if 'vlos' in plotfilename:
+            pass
+        elif 'ld' in plotfilename:
+            pass
+        else:
+            xmin, xmax= 90., 165.
+            ymin, ymax= 47., 59.
+            pyplot.plot([xmin,xmin],[ymin,ymax],'k-')
[…]

### 2013-12-20 py/plot_stream.py
streams: also plot distance insets

@@ -217,7 +217,57 @@ def plot_stream_lb(plotfilename):
         if 'vlos' in plotfilename:
             pass
         elif 'ld' in plotfilename:
-            pass
+            xmin, xmax= 158., 227.
+            ymin, ymax= 7.7,9.5
+            pyplot.plot([xmin,xmin],[ymin,ymax],'k-')
+            pyplot.plot([xmax,xmax],[ymin,ymax],'k-')
+            pyplot.plot([xmin,xmax],[ymin,ymin],'k-')
+            pyplot.plot([xmin,xmax],[ymax,ymax],'k-')
+            pyplot.plot([xmin,70.],[ymax,18.5],'k:')
+            pyplot.plot([xmax,248.],[ymax,18.5],'k:')
+            #2nd inset
+            xmin2, xmax2= 72.,100.
+            ymin2, ymax2= 11.5, 16.1
+            pyplot.plot([xmin2,xmin2],[ymin2,ymax2],'k-')
+            pyplot.plot([xmax2,xmax2],[ymin2,ymax2],'k-')
+            pyplot.plot([xmin2,xmax2],[ymin2,ymin2],'k-')
+            pyplot.plot([xmin2,xmax2],[ymax2,ymax2],'k-')
+            pyplot.plot([xmin2,66.5],[ymax2,15.85],'k:')
+            pyplot.plot([xmin2,66.5],[ymin2,0.5],'k:')
+            insetAxes= pyplot.axes([0.31,0.6,0.48,0.27])
+            pyplot.sca(insetAxes)
+            bovy_plot.bovy_plot(lbd[:,0],lbd[:,lbindx],'k,',
+                                overplot=True)
+            sdf.plotProgenitor(d1=d1,d2=d2,color='k',ls='--',
+                                overplot=True)
+            sdf.plotTrack(d1=d1,d2=d2,interp=True,color='k',spread=0,
+                           overplot=True,lw=1.)
+            nullfmt   = NullFormatter()         # no labels
+            insetAxes.xaxis.set_major_formatter(nullfmt)
+            insetAxes.yaxis.set_major_formatter(nullfmt)
+            insetAxes.set_xlim(xmin,xmax)
+            insetAxes.set_ylim(ymin,ymax)
+            pyplot.tick_params(\
+                axis='both',          # changes apply to the x-axis
+                which='both',      # both major and minor ticks are affected
+                bottom='off',      # ticks along the bottom edge are off
+                top='off',         # ticks along the top edge are off
[…]

### 2013-12-20 py/plot_stream.py
streams: change number of track points

@@ -13,7 +13,7 @@ from matplotlib import pyplot
 from matplotlib.ticker import NullFormatter
 _STREAMSNAPDIR= '../sim/snaps'
 _STREAMSNAPAADIR= '../sim/snaps_aai'
-_NTRACKCHUNKS= 4
+_NTRACKCHUNKS= 11
 def plot_stream_xz(plotfilename):
     #Read stream
     data= numpy.loadtxt(os.path.join(_STREAMSNAPDIR,'gd1_evol_hitres_01312.dat'),

### 2013-12-20 py/plot_stream.py
streams: plot vlos and track

@@ -13,7 +13,7 @@ from matplotlib import pyplot
 from matplotlib.ticker import NullFormatter
 _STREAMSNAPDIR= '../sim/snaps'
 _STREAMSNAPAADIR= '../sim/snaps_aai'
-_NTRACKCHUNKS= 11
+_NTRACKCHUNKS= 4
 def plot_stream_xz(plotfilename):
     #Read stream
     data= numpy.loadtxt(os.path.join(_STREAMSNAPDIR,'gd1_evol_hitres_01312.dat'),
@@ -163,9 +163,11 @@ def plot_stream_lb(plotfilename):
         obs= numpy.array([1.56148083,0.35081535,-1.15481504,
                           0.88719443,-0.47713334,0.12019596])
         sdf= streamdf(0.3/220.,progenitor=Orbit(obs),pot=lp,aA=aAI,
-                      leading=True,nTrackChunks=_NTRACKCHUNKS)
+                      leading=True,nTrackChunks=_NTRACKCHUNKS,
+                      vsun=[0.,30.24*8.,0.])
         sdft= streamdf(0.3/220.,progenitor=Orbit(obs),pot=lp,aA=aAI,
-                       leading=False,nTrackChunks=_NTRACKCHUNKS)
+                       leading=False,nTrackChunks=_NTRACKCHUNKS,
+                       vsun=[0.,30.24*8.,0.])
     #Plot
     bovy_plot.bovy_print(fig_width=8.25,fig_height=3.5)
     if 'ld' in plotfilename:
@@ -215,7 +217,27 @@ def plot_stream_lb(plotfilename):
                        overplot=True,lw=1.)
         #Insets
         if 'vlos' in plotfilename:
-            pass
+            xmin, xmax= 220., 250.
+            ymin, ymax= 230., 390.
+            pyplot.plot([xmin,xmin],[ymin,ymax],'k-')
+            pyplot.plot([xmax,xmax],[ymin,ymax],'k-')
+            pyplot.plot([xmin,xmax],[ymin,ymin],'k-')
+            pyplot.plot([xmin,xmax],[ymax,ymax],'k-')
+            pyplot.plot([xmin,152.],[ymin,-100.],'k:')
+            pyplot.plot([xmin,152.],[ymax,460.],'k:')
+            insetAxes= pyplot.axes([0.15,0.42,0.38,0.45])
+            pyplot.sca(insetAxes)
+            bovy_plot.bovy_plot(lbd[:,0],vlbd[:,lbindx],'k,',
+                                overplot=True)
[…]

### 2013-12-20 py/plot_stream.py
streams: also plot pmll and pmbb

@@ -13,7 +13,7 @@ from matplotlib import pyplot
 from matplotlib.ticker import NullFormatter
 _STREAMSNAPDIR= '../sim/snaps'
 _STREAMSNAPAADIR= '../sim/snaps_aai'
-_NTRACKCHUNKS= 4
+_NTRACKCHUNKS= 11
 def plot_stream_xz(plotfilename):
     #Read stream
     data= numpy.loadtxt(os.path.join(_STREAMSNAPDIR,'gd1_evol_hitres_01312.dat'),
@@ -178,11 +178,19 @@ def plot_stream_lb(plotfilename):
         lbindx= 0
         ylabel=r'$V_\mathrm{los}\,(\mathrm{km\,s}^{-1})$'
         yrange=[-500.,500.]
+    elif 'lpmll' in plotfilename:
+        lbindx= 1
+        ylabel=r'$\mu_{l}\cos b\,(\mathrm{mas\,yr}^{-1})$'
+        yrange=[-2.,13.5]
+    elif 'lpmbb' in plotfilename:
+        lbindx= 2
+        ylabel=r'$\mu_{b}\,(\mathrm{mas\,yr}^{-1})$'
+        yrange=[-8.,7.]
     else:
         lbindx= 1 
         yrange=[-10.,60.]
         ylabel=r'$\mathrm{Galactic\ latitude}\,(\mathrm{deg})$'
-    if 'vlos' in plotfilename:
+    if 'vlos' in plotfilename or 'pm' in plotfilename:
         bovy_plot.bovy_plot(lbd[:,0],vlbd[:,lbindx],'k,',
                             xlabel=r'$\mathrm{Galactic\ longitude}\,(\mathrm{deg})$',
                             ylabel=ylabel,
@@ -195,7 +203,7 @@ def plot_stream_lb(plotfilename):
                             xrange=[0.,290.],
                             yrange=yrange)
     if includeorbit:
-        if 'vlos' in plotfilename:
+        if 'vlos' in plotfilename or 'pm' in plotfilename:
             bovy_plot.bovy_plot(plbd[npts,0],pvlbd[npts,lbindx],
                                 'o',color='0.5',mec='none',overplot=True,ms=8)
             bovy_plot.bovy_plot(plbd[:,0],pvlbd[:,lbindx],'k--',overplot=True)
@@ -207,6 +215,10 @@ def plot_stream_lb(plotfilename):
         d1= 'll'
         if 'vlos' in plotfilename:
[…]

### 2013-12-20 py/plot_stream.py
streams: add predictions to plots of stream properties as a function of angle

@@ -14,6 +14,7 @@ from matplotlib.ticker import NullFormatter
 _STREAMSNAPDIR= '../sim/snaps'
 _STREAMSNAPAADIR= '../sim/snaps_aai'
 _NTRACKCHUNKS= 11
+_SIGV=0.365
 def plot_stream_xz(plotfilename):
     #Read stream
     data= numpy.loadtxt(os.path.join(_STREAMSNAPDIR,'gd1_evol_hitres_01312.dat'),
@@ -52,10 +53,12 @@ def plot_stream_xz(plotfilename):
         aAI= actionAngleIsochroneApprox(b=0.8,pot=lp)
         obs= numpy.array([1.56148083,0.35081535,-1.15481504,
                           0.88719443,-0.47713334,0.12019596])
-        sdf= streamdf(0.3/220.,progenitor=Orbit(obs),pot=lp,aA=aAI,
-                      leading=True,nTrackChunks=_NTRACKCHUNKS)
-        sdft= streamdf(0.3/220.,progenitor=Orbit(obs),pot=lp,aA=aAI,
-                       leading=False,nTrackChunks=_NTRACKCHUNKS)
+        sdf= streamdf(_SIGV/220.,progenitor=Orbit(obs),pot=lp,aA=aAI,
+                      leading=True,nTrackChunks=_NTRACKCHUNKS,
+                      tdisrupt=4.5/bovy_conversion.time_in_Gyr(220.,8.))
+        sdft= streamdf(_SIGV/220.,progenitor=Orbit(obs),pot=lp,aA=aAI,
+                       leading=False,nTrackChunks=_NTRACKCHUNKS,
+                       tdisrupt=4.5/bovy_conversion.time_in_Gyr(220.,8.))
     #Plot
     bovy_plot.bovy_print()
     bovy_plot.bovy_plot(data[:,1],data[:,2],'k,',
@@ -162,12 +165,14 @@ def plot_stream_lb(plotfilename):
         aAI= actionAngleIsochroneApprox(b=0.8,pot=lp)
         obs= numpy.array([1.56148083,0.35081535,-1.15481504,
                           0.88719443,-0.47713334,0.12019596])
-        sdf= streamdf(0.3/220.,progenitor=Orbit(obs),pot=lp,aA=aAI,
+        sdf= streamdf(_SIGV/220.,progenitor=Orbit(obs),pot=lp,aA=aAI,
[…]

### 2013-12-20 py/plot_stream.py
streams: update plots

@@ -795,7 +795,10 @@ def plot_stream_aa(plotfilename):
                             '-',color='0.75',lw=2.,overplot=True,zorder=10)
         pyplot.fill_between(das,(dOs+sOs)*bovy_conversion.freq_in_Gyr(220.,8),
                             (dOs-sOs)*bovy_conversion.freq_in_Gyr(220.,8),
-                            color='0.6',zorder=0)
+                            color='0.6',zorder=1)
+        pyplot.fill_between(das,(dOs+2*sOs)*bovy_conversion.freq_in_Gyr(220.,8),
+                            (dOs-2*sOs)*bovy_conversion.freq_in_Gyr(220.,8),
+                            color='0.8',zorder=0)
     elif includetrack and 'apartime' in plotfilename:
         das= numpy.linspace(0.01,1.3,101)
         mts= numpy.array([sdf.meantdAngle(da) for da in das])
@@ -809,7 +812,7 @@ def plot_stream_aa(plotfilename):
                             (mts-sts)*bovy_conversion.time_in_Gyr(220.,8),
                             color='0.6',zorder=1)
     elif includetrack and 'aparaperp' in plotfilename:
-        das= numpy.linspace(0.01,1.3,11)
+        das= numpy.linspace(0.01,1.3,101)
         sas= numpy.array([sdf.sigangledAngle(da) for da in das])
         sass= numpy.array([sdf.sigangledAngle(da,simple=True) for da in das])
         pyplot.fill_between(das,0.,

### 2013-12-23 py/plot_stream_2Gya.py
streams: add plot of stream about 2 Gya

@@ -1,0 +1,151 @@
+import sys
+import os, os.path
+import numpy
+from galpy.util import bovy_plot, bovy_coords, bovy_conversion
+from galpy import potential
+from galpy.orbit import Orbit
+from galpy.actionAngle_src.actionAngleIsochroneApprox\
+    import actionAngleIsochroneApprox
+from galpy.df_src.streamdf import streamdf
+from matplotlib import pyplot
+from matplotlib.ticker import NullFormatter
+_STREAMSNAPDIR= '../sim/snaps'
+_STREAMSNAPAADIR= '../sim/snaps_aai'
+_NTRACKCHUNKS= 4
+_SIGV=0.365
+def plot_stream_xz(plotfilename):
+    #Read stream
+    data= numpy.loadtxt(os.path.join(_STREAMSNAPDIR,'gd1_evol_hitres_00800.dat'),
+                        delimiter=',')
+    aadata= numpy.loadtxt(os.path.join(_STREAMSNAPAADIR,
+                                       'gd1_evol_hitres_aa_00800.dat'),
+                        delimiter=',')
+    thetar= aadata[:,6]
+    thetar= (numpy.pi+(thetar-numpy.median(thetar))) % (2.*numpy.pi)
+    sindx= numpy.fabs(thetar-numpy.pi) > (1.5*numpy.median(numpy.fabs(thetar-numpy.median(thetar))))
+    includeorbit= True
+    if includeorbit:
+        npts= 201
+        pot= potential.LogarithmicHaloPotential(normalize=1.,q=0.9)
+        pts= numpy.linspace(0.,4.,npts)
+        #Calculate progenitor orbit around this point
+        pox= numpy.median(data[:,1])
+        poy= numpy.median(data[:,3])
+        poz= numpy.median(data[:,2])
+        povx= numpy.median(data[:,4])
+        povy= numpy.median(data[:,6])
+        povz= numpy.median(data[:,5])
+        pR,pphi,pZ= bovy_coords.rect_to_cyl(pox,poy,poz)
+        pvR,pvT,pvZ= bovy_coords.rect_to_cyl_vec(povx,povy,povz,pR,
+                                                 pphi,pZ,cyl=True)
+        ppo= Orbit([pR/8.,pvR/220.,pvT/220.,pZ/8.,pvZ/220.,pphi])
+        pno= Orbit([pR/8.,-pvR/220.,-pvT/220.,pZ/8.,-pvZ/220.,pphi])
[…]

### 2013-12-23 py/plot_stream_2Gya.py
streams: tweaks to 2Gya plot

@@ -11,7 +11,7 @@ from matplotlib import pyplot
 from matplotlib.ticker import NullFormatter
 _STREAMSNAPDIR= '../sim/snaps'
 _STREAMSNAPAADIR= '../sim/snaps_aai'
-_NTRACKCHUNKS= 4
+_NTRACKCHUNKS= 11
 _SIGV=0.365
 def plot_stream_xz(plotfilename):
     #Read stream
@@ -22,7 +22,10 @@ def plot_stream_xz(plotfilename):
                         delimiter=',')
     thetar= aadata[:,6]
     thetar= (numpy.pi+(thetar-numpy.median(thetar))) % (2.*numpy.pi)
-    sindx= numpy.fabs(thetar-numpy.pi) > (1.5*numpy.median(numpy.fabs(thetar-numpy.median(thetar))))
+    if 'sim' in plotfilename:
+        sindx= numpy.fabs(thetar-numpy.pi) > (4.*numpy.median(numpy.fabs(thetar-numpy.median(thetar))))
+    else:
+        sindx= numpy.fabs(thetar-numpy.pi) > (1.5*numpy.median(numpy.fabs(thetar-numpy.median(thetar))))
     includeorbit= True
     if includeorbit:
         npts= 201
@@ -107,6 +110,12 @@ def plot_stream_xz(plotfilename):
     if includeorbit:
         bovy_plot.bovy_plot(pox,poz,'o',color='0.5',mec='none',overplot=True,ms=8)
         bovy_plot.bovy_plot(pvec[0,:],pvec[1,:],'k--',overplot=True,lw=1.)
+    if 'sim' in plotfilename:
+        bovy_plot.bovy_text(r'$\mathrm{simulated\ stream}$',
+                            bottom_left=True,size=16.)
+    else:
+        bovy_plot.bovy_text(r'$N\!\!-\!\!\mathrm{body\ stream}$',
+                            bottom_left=True,size=16.)
     if includetrack:
         d1= 'x'
         d2= 'z'

### 2013-12-23 tex/ms.tex
streams: add pmbb figure, add to document
streams: add simulated stream figure
streams: abstract

@@
 \title{Dynamical modeling of tidal streams}
-\author{Jo~Bovy\altaffilmark{1,2,3} \etal}
-%  Hans-Walter~Rix\altaffilmark{4}}
-\altaffiltext{\theaddress}{\label{IAS}\stepcounter{address} Institute
-  for Advanced Study, Einstein Drive, Princeton, NJ 08540, USA}
-\altaffiltext{\theaddress}{\label{Hubble}\stepcounter{address} Hubble
-  fellow}
-\altaffiltext{\theaddress}{\label{email}\stepcounter{address}
-  Correspondence should be addressed to bovy@ias.edu~.}
-%\altaffiltext{\theaddress}{\label{MPIA}\stepcounter{address}
-%  Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117
-%  Heidelberg, Germany}
+\author{Jo~Bovy\altaffilmark{\ref{Hubble}}}
+\affil{Institute for Advanced Study, Einstein Drive, Princeton, NJ 08540, USA}
+\email{bovy@ias.edu}
+\altaffiltext{\theaddress}{\label{Hubble}\stepcounter{address} Hubble fellow}
 
 \begin{abstract} 
-  We present a generative, dynamical model for a tidal stream.
+  I present a new framework for modeling the dynamics of cold tidal
+  streams that is useful for constraining the gravitational host
+  potential and progenitor properties from observed tidal-stream data
+  and for quickly generating mock tidal streams. The framework
+  consists of simple models for the initial action--angle distribution
+  of tidal debris as a function of time, which can be
+  straightforwardly evolved forward in time. Taking advantage of the
+  essentially one-dimensional nature of cold tidal streams, the
+  transformation to position--velocity coordinates can be linearized
+  and interpolated near a small number of points along the progenitor
+  orbit, thus allowing for efficient computations of a stream's
+  properties in observable quantities. I illustrate how to calculate
+  the stream's track in different coordinate systems, how to quickly
+  estimate the dispersion perpendicular to the track, and how to draw
+  mock stream data. As a generative model, this framework allows one
+  to compute the full probability distribution function and
+  marginalize over or condition it on certain phase--space dimensions
+  as well as convolve it with observational uncertainties. This will
+  be instrumental in proper data analysis of stream data. In addition
+  to providing a computationally-efficient practical tool for modeling
+  the dynamics of tidal streams, the action--angle nature of the
+  framework helps elucidate exactly in what manner streams do not
+  follow single orbits, how the observed width of the stream relates
+  to the velocity dispersion or mass of the progenitor, and how the
+  progenitor of ``orphan'' streams could be located.
+
+  The practical usefulness of the proposed framework crucially depends
+  on the ability to calculate action--angle variables for any orbit in
+  any gravitational potential. A novel method for calculating actions,
+  frequencies, and angles in any static potential using a single orbit
+  integration is described in an Appendix.
 \end{abstract}
@@
 \keywords{
-	Galaxy: fundamental parameters
+        dark matter
+        ---
+	Galaxy: halo
 	---
@@
 	Galaxy: structure
+        ---
+        galaxies: interactions
+        ---
+        stellar dynamics
 }
@@
   \includegraphics[width=\textwidth,clip=]{gd1_evol_ld.ps}
-  \includegraphics[width=\textwidth,clip=]{gd1_evol_lvlos.ps}
   \caption{}\label{fig:gd1_lbd}
@@
 %python plot_stream.py ../tex/gd1_evol_ld.ps
+\end{figure}
+\begin{figure}[t!]
+  \includegraphics[width=\textwidth,clip=]{gd1_evol_lvlos.ps}
+  \includegraphics[width=\textwidth,clip=]{gd1_evol_lpmll.ps}
+  \includegraphics[width=\textwidth,clip=]{gd1_evol_lpmbb.ps}
+  \caption{}\label{fig:gd1_lbv}
 %python plot_stream.py ../tex/gd1_evol_lvlos.ps
+%python plot_stream.py ../tex/gd1_evol_lpmll.ps
+%python plot_stream.py ../tex/gd1_evol_lpmbb.ps
 \end{figure}
@@
 
+\begin{figure}[t!]
+ \includegraphics[width=0.48\textwidth,clip=]{gd1_evol_2Gya_xz.ps}
+ \includegraphics[width=0.48\textwidth,clip=]{gd1_evol_2Gya_xz_sim.ps}
+  \caption{}\label{fig:gd1_sim}
+%python plot_stream_2Gya.py ../tex/gd1_evol_2Gya_xz.ps
+%python plot_stream_2Gya.py ../tex/gd1_evol_2Gya_xz_sim.ps
+\end{figure}
+
 \end{document}

### 2013-12-24 tex/ms.tex
streams: intro and re-worded abstract

@@
   I present a new framework for modeling the dynamics of cold tidal
-  streams that is useful for constraining the gravitational host
-  potential and progenitor properties from observed tidal-stream data
-  and for quickly generating mock tidal streams. The framework
-  consists of simple models for the initial action--angle distribution
-  of tidal debris as a function of time, which can be
-  straightforwardly evolved forward in time. Taking advantage of the
-  essentially one-dimensional nature of cold tidal streams, the
-  transformation to position--velocity coordinates can be linearized
-  and interpolated near a small number of points along the progenitor
-  orbit, thus allowing for efficient computations of a stream's
-  properties in observable quantities. I illustrate how to calculate
-  the stream's track in different coordinate systems, how to quickly
-  estimate the dispersion perpendicular to the track, and how to draw
-  mock stream data. As a generative model, this framework allows one
-  to compute the full probability distribution function and
+  streams. The framework consists of simple models for the initial
+  action--angle distribution of tidal debris as a function of time,
+  which can be straightforwardly evolved forward in time. Taking
+  advantage of the essentially one-dimensional nature of cold tidal
+  streams, the transformation to position--velocity coordinates can be
+  linearized and interpolated near a small number of points along the
+  progenitor orbit, thus allowing for efficient computations of a
+  stream's properties in observable quantities. I illustrate how to
+  calculate the stream's track in different coordinate systems, how to
+  quickly estimate the dispersion perpendicular to the track, and how
+  to draw mock stream data. As a generative model, this framework
+  allows one to compute the full probability distribution function and
   marginalize over or condition it on certain phase--space dimensions
@@
   the dynamics of tidal streams, the action--angle nature of the
-  framework helps elucidate exactly in what manner streams do not
+  framework helps elucidate in exactly what manner streams do not
   follow single orbits, how the observed width of the stream relates
@@
 
+Tidal streams hold enormous promise as probes of both the large-scale
+structure of the Milky Way halo's density distribution
+\citep[\eg,][]{Johnston99a,Koposov10a} and its small-scale
+fluctuations \citep{Carlberg09a,Carlberg12a}. However, two factors
+have hampered the practical use of tidal streams as constraints on the
+gravitational potential. First, streams do not trace a single orbit,
+which has led to confusion over how problematic the single-orbit
+approximation really is \citep{Eyre11a,Sanders13a}. Second, approaches
+that go beyond the single-orbit assumption encounter practical
+difficulties of computational cost and observational data quality
+\citep{Sanders13b,PriceWhelan13a}. Similarly, spurious underdensities
+created by the dynamics of stream stars hinder our ability to use
+underdensities in star counts along a tidal stream in order to
+constrain the number of encounters with dark satellites and their
+masses \citep{Ngan14a}. Fast, generative models of tidal streams would
+greatly benefit both of these applications.
+
+It has long been clear that the dynamics of tidal streams is most
+simply described in terms of action--angle coordinates
+\citep{Tremaine99a,Helmi99a}. Once a star has been tidally stripped
+from the progenitor, the self-gravity of the stream can be neglected
+and the orbital actions of a stream member are conserved while the
+angles increase linearly with time. The action--angle structure of a
+stream is therefore characterized by strong but straightforward
+correlations between the actions and angles of stream members, the
+exploitation of which is crucial for using streams to measure the host
+gravitational potential \citep{Sanders13b}. The description of a
+stream in action--angle coordinates also makes the connection between
+the orbit and velocity dispersion (or mass) of the progenitor and the
+action distribution of the tidal debris clear \citep{Eyre11a},
+allowing simple physical models of the stream to be specified.
+
+However, streams are observed in position--velocity coordinates and
+for action--angle descriptions to be useful, we must be able to
+calculate the transformation between these coordinate systems
+efficiently. Until now this has required accurate phase--space data
+\citep{Sanders13b} and specialized algorithms to calculate
+action--angle coordinates that break down for the eccentric orbits on
+which tidal-stream progenitors are typically found
+\citep{Sanders12a}. However, even with Gaia, stream data in the future
+will typically have large uncertainties compared to the intrinsic
+dispersion of the stream and in particular the line-of-sight
+velocities of the faintest stream members will not be observed for
+many stars. Additionally, streams are superimposed on a non-negligible
+background of field stars and background contamination cannot be
+easily taken into account in any of the current stream-fitting
+methods.
+
+In this paper, I present a new method for modeling cold tidal streams
+that fundamentally lives in frequency--angle space, because after
+specifying the time at which a star was stripped, the stream
+distribution function is essentially one-dimensional in this
+space. While many of the results of this paper are more generally
+valid, the fiducial stream model in this paper consists of a
+three-dimensional (close to) Gaussian distribution of frequencies in
+the stream, a three-dimensional Gaussian distribution of initial angle
+offsets between stream members and the progenitor, and a uniform
+distribution of stripping times; the frequency and angle distributions
+are independent of stripping time. While this model oversimplifies the
+real stripping process, it is argued that at least for constraining
+the gravitational potential with stream data, this simple model is
+likely unbiased. For detailed mock stream data, more elaborate models
+of the stripping process as a function of time might be necessary and
+I discuss how these could easily be incorporated into the proposed
+framework.
+
+To evaluate the model in position--velocity coordinates, the
+transformation to and from action--angle coordinates is computed using
+a novel method for calculating action--angle coordinates presented in
+Appendix~\ref{sec:aa} and this transformation is linearized at a small
+number of points along the progenitor orbit that span the width of the
+stream. For cold streams, the separation between the stream track and
+the progenitor orbit as well as the internal differences within the
+stream are small, such that this linear approximation is accurate
+enough. The linear approximation allows for fast evaluation of the
+average stream position in position--velocity coordinates and
+efficient mock data generation. For a uniform distribution of
+stripping times, an observed stream member's phase--space probability
+distribution function (PDF) can be analytically marginalized over
+stripping time, such that using the linear approximation any
+evaluation of the PDF is extremely rapid. This clears the way for
+proper probabilistic inference of the Milky Way halo's gravitational
+potential and progenitor properties using stream data for individual
+stars.
+
+This paper is outlined as follows. In \sectionname~\ref{sec:dynamics}
+I briefly summarize the dynamics of tidal streams in terms of
+action--angle coordinates. A generative model of a tidal stream in
+frequency--angle coordinates is given in
+\sectionname~\ref{sec:generative}. \sectionname~\ref{sec:track}
+describes how to compute the stream's properties as a function of
+angle along the stream, both in action--angle coordinates and in
+poition--velocity coordinates. \sectionname~\ref{sec:mock} discusses
+how to generate mock stream data using the generative model of a tidal
+stream and in \sectionname~\ref{sec:pdf} I show how to calculate the
+stream PDF for individual stream members as well as its
+marginalization and convolution over missing and noisy directions. A
+discussion and outlook is presented in
+\sectionname~\ref{sec:discussion} and I conclude in
+\sectionname~\ref{sec:conclusion}. In Appendix
+\sectionname~\ref{sec:aa} a novel method for calculating actions,
+frequencies, and angles for any orbit in any static gravitational
+potential is presented.
+
 %BOVY: reference point
@@
 action--angle space has been described by various previous authors
-\citep[\eg,][]{Helmi99a,Tremaine99a,SomeJohnstonPaper,Sanders13a}. I
+\citep[\eg,][]{Helmi99a,Tremaine99a,SomeJohnstonPaper,Sanders13b}. I
 provide here a brief description of the dynamics of tidal streams that
@@
 dynamics of the progenitor and the stripping process (see discussion
-in \citealt{Sanders13a}). For dynamical modeling we can be agnostic
+in \citealt{Sanders13b}). For dynamical modeling we can be agnostic
 about the relative importance of all of these contributors to
@@
 the correlation between the actions and the angles, expressed by the
-exponential in \eqnname~(\ref{eq:aadfsimple}) \citep{Sanders13a}; this
+exponential in \eqnname~(\ref{eq:aadfsimple}) \citep{Sanders13b}; this
 method does not use the fact that the spread in actions in a tidal
@@
   Binney,~J. 2008, \mnras, 386, L47
+\bibitem[Carlberg(2009)]{Carlberg09a}
+  Carlberg,~R.~G. 2009, \apj, 705, L223
+\bibitem[Carlberg(2012)]{Carlberg12a}
+Carlberg,~R.~G. 2012, \apj, 748, 20
 \bibitem[Eyre \& Binney(2011)]{Eyre11a}
   Eyre,~A. \& Binney,~J. 2011, \mnras, 413, 1852
+\bibitem[Helmi \& White(1999)]{Helmi99a}
+  Helmi,~A. \& White,~S.~D.~M. 1999, \mnras, 307, 495
+\bibitem[Johnston \etal(1999)]{Johnston99a}
+  Johnston,~K.~V., Zhao,~H., Spergel,~D.~N., \& Hernquist,~L. 1999, \apj, 512, L109
+\bibitem[Koposov \etal(2010)]{Koposov10a}
+  Koposov,~S.~E., Rix,~H.-W., \& Hogg,~D.~W. 2010, \apj, 712, 260
+\bibitem[Ngan \& Carlberg(2014)]{Ngan14a}
+  Ngan,~W.~H.~W. \& Carlberg,~R.~G. 2014, \apj, submitted
 \bibitem[Pe{\~n}arrubia \etal(2012)]{Penarrubia12a}
   Pe{\~n}arrubia, J., Koposov,~S.~E., \& Walker,~M.~G. 2012, \apj, 760, 2
-\bibitem[Sanders \& Binney(2013)]{Sanders13a}
-  Sanders,~J. \& Binney,~J. 2013, \mnras, 433, 1826
+\bibitem[Price-Whelan \& Johnston(2013)]{PriceWhelan13a}
+  Price-Whelan,~A.~M. \& Johnston,~K.~V.~2013, \apj, 778, L12
+\bibitem[Sanders(2012)]{Sanders12a}
+  Sanders,~J. 2012, \mnras, 426, 128
+\bibitem[Sanders \& Binney(2013a)]{Sanders13a}
+  Sanders,~J. \& Binney,~J. 2013a, \mnras, 433, 1813
+\bibitem[Sanders \& Binney(2013b)]{Sanders13b}
+  Sanders,~J. \& Binney,~J. 2013b, \mnras, 433, 1826
+\bibitem[Tremaine(1999)]{Tremaine99a}
+  Tremaine,~S. 1999, \mnras, 307, 877
 \end{thebibliography}

### 2013-12-24 tex/ms.tex
streams: tweaks to the intro

@@
 Tidal streams hold enormous promise as probes of both the large-scale
-structure of the Milky Way halo's density distribution
+structure of the Milky Way (MW) halo's density distribution
 \citep[\eg,][]{Johnston99a,Koposov10a} and its small-scale
 fluctuations \citep{Carlberg09a,Carlberg12a}. However, two factors
-have hampered the practical use of tidal streams as constraints on the
-gravitational potential. First, streams do not trace a single orbit,
-which has led to confusion over how problematic the single-orbit
-approximation really is \citep{Eyre11a,Sanders13a}. Second, approaches
-that go beyond the single-orbit assumption encounter practical
-difficulties of computational cost and observational data quality
-\citep{Sanders13b,PriceWhelan13a}. Similarly, spurious underdensities
-created by the dynamics of stream stars hinder our ability to use
-underdensities in star counts along a tidal stream in order to
-constrain the number of encounters with dark satellites and their
-masses \citep{Ngan14a}. Fast, generative models of tidal streams would
-greatly benefit both of these applications.
+have hampered the practical use of tidal streams in obtaining
+constraints on the MW gravitational potential. First, streams do not
+trace a single orbit, which has led to confusion over how to best fit
+streams and over how problematic the single-orbit approximation really
+is \citep{Eyre11a,Sanders13a}. Second, approaches that go beyond the
+single-orbit assumption encounter practical difficulties of
+computational cost and observational data quality making many of these
+approaches impractical for real, noisy data
+\citep{Sanders13b,PriceWhelan13a}. Observed gaps in tidal streams may
+be due to interactions with dark--matter subhalos \citep{Yoon11a}, but
+spurious underdensities could be created by the dynamics of stream
+stars. This hinders our ability to use underdensities in star counts
+along a tidal stream in order to constrain the number of encounters
+with dark satellites and their masses \citep{Ngan14a}. Fast,
+generative models of tidal streams would greatly benefit both of these
+applications.
 
@@
 gravitational potential \citep{Sanders13b}. The description of a
-stream in action--angle coordinates also makes the connection between
-the orbit and velocity dispersion (or mass) of the progenitor and the
-action distribution of the tidal debris clear \citep{Eyre11a},
-allowing simple physical models of the stream to be specified.
+stream in action--angle coordinates also elucidates the connection
+between the orbit and velocity dispersion (or mass) of the progenitor
+and the action distribution of the tidal debris \citep{Eyre11a},
+allowing simple physical models of the stream to be used and
+constrained by observational data.
 
@@
 action--angle coordinates that break down for the eccentric orbits on
-which tidal-stream progenitors are typically found
-\citep{Sanders12a}. However, even with Gaia, stream data in the future
+which tidal-stream progenitors are typically found (that is, radial
+and/or vertical actions of similar size as the angular momentum;
+\citealt{Sanders12a}). However, even with Gaia, future stream data
 will typically have large uncertainties compared to the intrinsic
@@
 a novel method for calculating action--angle coordinates presented in
-Appendix~\ref{sec:aa} and this transformation is linearized at a small
+Appendix~\ref{sec:aa}. This transformation is linearized at a small
 number of points along the progenitor orbit that span the width of the
@@
 enough. The linear approximation allows for fast evaluation of the
-average stream position in position--velocity coordinates and
-efficient mock data generation. For a uniform distribution of
-stripping times, an observed stream member's phase--space probability
-distribution function (PDF) can be analytically marginalized over
-stripping time, such that using the linear approximation any
-evaluation of the PDF is extremely rapid. This clears the way for
-proper probabilistic inference of the Milky Way halo's gravitational
-potential and progenitor properties using stream data for individual
-stars.
+average stream location (the stream ``track'') in position--velocity
+coordinates and efficient mock data generation. For a uniform
+distribution of stripping times, an observed stream member's
+phase--space probability distribution function (PDF) can be
+analytically marginalized over stripping time, such that using the
+linear approximation any evaluation---including marginalization and
+uncertainty convolution---of the PDF is extremely rapid. This clears
+the way for proper probabilistic inference of the Milky Way halo's
+gravitational potential and progenitor properties using stream data
+for individual stars.
 
@@
   Tremaine,~S. 1999, \mnras, 307, 877
+\bibitem[Yoon \etal(2011)]{Yoon11a}
+  Yoon,~J.~H., Johnston,~K.~V., \& Hogg,~D.~W. 2011, \apj, 731, 58
 \end{thebibliography}

### 2013-12-24 py/plot_aAI.py
streams: make some plots illustrating the action-angle calculation

@@ -1,0 +1,30 @@
+#Make plots showing how the actionAngleIsochroneApprox works
+import sys
+import numpy
+from galpy.util import bovy_plot
+from galpy import potential
+from galpy.actionAngle_src.actionAngleIsochroneApprox import actionAngleIsochroneApprox 
+def plot_jr(plotfilename):
+    lp= potential.LogarithmicHaloPotential(q=0.9,normalize=1.)
+    aAI= actionAngleIsochroneApprox(b=0.8,pot=lp,tintJ=200,ntintJ=20000)
+    obs= numpy.array([1.56148083,0.35081535,-1.15481504,0.88719443,
+                      -0.47713334,0.12019596])
+    bovy_plot.bovy_print(fig_width=6.)
+    aAI.plot(*obs,type='jr',downsample=True)
+    bovy_plot.bovy_end_print(plotfilename)
+
+def plot_araz(plotfilename):
+    lp= potential.LogarithmicHaloPotential(q=0.9,normalize=1.)
+    aAI= actionAngleIsochroneApprox(b=0.8,pot=lp,tintJ=50)
+    obs= numpy.array([1.56148083,0.35081535,-1.15481504,0.88719443,
+                      -0.47713334,0.12019596])
+    bovy_plot.bovy_print(fig_width=6.)
+    aAI.plot(*obs,type='araz',downsample=True,deperiod=True)
+    bovy_plot.bovy_end_print(plotfilename)
+
+if __name__ == '__main__':
+    plotfilename= sys.argv[1]
+    if 'jr' in plotfilename:
+        plot_jr(plotfilename)
+    elif 'araz' in plotfilename:
+        plot_araz(plotfilename)

### 2013-12-24 tex/ms.tex
streams: re-arrange figures
streams: first pass of the appendix

@@
 \newcommand{\vecp}{\ensuremath{\vec{p}}}
+\newcommand{\vecn}{\ensuremath{\vec{n}}}
 \newcommand{\veco}{\ensuremath{\vec{\Omega}}}
@@
 \newcommand{\dex}{\ensuremath{\,\mathrm{dex}}}
+\newcommand{\Myr}{\ensuremath{\,\mathrm{Myr}}}
 \newcommand{\Gyr}{\ensuremath{\,\mathrm{Gyr}}}
@@
 Astronomy, Incorporated, under NASA contract NAS5-26555. J.B.
-%and H.-W.R  (BOVY: ALSO ACKNOWLEDGE WITHOUT S)
-acknowledges support from SFB 881 (A3) funded by the German
-Research Foundation DFG.
+acknowledges support from SFB 881 (A3) funded by the German Research
+Foundation DFG.
 
@@
 
-\section{Dynamical modeling of two-star streams}\label{sec:twostar}
+The wide-spread use of action--angle coordinates, in many ways the
+natural coordinates for studying the orbital structure of galaxies, in
+dynamical modeling has been frustrated by the difficulty in
+calculating the transformation to and from these coordinates,
+$(\vecx,\vecv) \leftrightarrow (\vecj,\veca)$. The most general method
+for calculating the transformation $(\vecj,\veca) \rightarrow
+(\vecx,\vecv)$ is provided by torus modeling \citep{McGill90a}. In the
+context of modeling observational data, computing the transformation
+$(\vecx,\vecv) \rightarrow (\vecj,\veca)$ is more important and the
+most general method consists of iteratively inverting the torus
+modeling \citep{McMillan08a}, which is computationally
+expensive. Recently, progress has been made in computing
+$(\vecx,\vecv) \rightarrow (\vecj,\veca)$ for Milky-Way-like
+potentials by either implicitly or explicitly approximating the
+gravitational potential as a St\"{a}ckel potential
+\citep{Binney12a,Sanders12a}, which allows action--angle coordinates
+to be computed with percent-level errors for moderately eccentric
+orbits. However, these methods break down for orbits with radial
+and/or vertical actions of a similar magnitude as the angular
+momentum---that is, orbits that feel the gravitational potential over
+a volume large enough that the St\"{a}ckel approximation fails. This
+is especially problematic for the orbits of stars in tidal streams, as
+the progenitors of tidal streams are typically on quite eccentric
+orbits \citep[\eg,][]{Sanders13a}.
+
+Recently, Fox \& Binney, 2014, in preparation, have suggested a new
+method for calculating $\vecj(\vecx,\vecv)$ inspired by torus
+modeling. This method requires only a single orbit integration and
+allows the actions to be calculated accurately and as precisely as
+desired (that is, a convergence criterion can be applied and
+convergence can be attained). I describe this method briefly here and
+then show how to extend it to also calculate the frequencies and
+angles: $(\vecx,\vecv) \rightarrow (\veco,\veca)$.
+
+As in torus modeling, the Fox \& Binney method uses a auxiliary
+isochrone potential $\Phi^A$ that should be close to the target
+potential $\Phi$ for which the action--angle coordinates should be
+calculated (a precise definition of ``close'' will be given
+below). Actions and angles $(\vecj^A,\veca^A)$ calculated in the
+auxiliary potential can be related to those in the target potential
+$(\vecj,\veca)$ by a generating function that can be written as
+\citep{McGill90a}
+\begin{equation}
+  S(\veca^A,\vecj) = \veca^A\cdot\vecj+2\sum_{\vecn > 0} S_{\vecn}(\vecj)\sin(\vecn\cdot\veca^A)\,,
+\end{equation}
+where the $\vecn > 0$ sum is over the integer two-dimensional
+half-plane that excludes the origin (two- rather than
+three-dimensional, as the azimuthal angle is perfectly correlated with
+the vertical angle in the auxiliary isochrone potential;
+\citealt{binneytremaine}), and the $S_{\vecn}$ are functions of the
+target actions that are to be determined. The canonical transformation
+corresponding to this generating function is
+\begin{equation}\label{eq:jjt}
+  \vecj^A = \frac{\partial S(\veca^A,\vecj)}{\partial \veca^A} = \vecj + 2\sum_{\vecn > 0} \vecn\,S_{\vecn}(\vecj)\cos(\vecn\cdot\veca^A)\,,
+\end{equation}
+and
+\begin{equation}\label{eq:aat}
+  \veca = \frac{\partial S(\veca^A,\vecj)}{\partial \vecj} = \veca^A +2\sum_{\vecn > 0} \frac{\partial S_{\vecn}(\vecj)}{\partial \vecj}\,\sin(\vecn\cdot\veca^A)\,.
+\end{equation}
+
+The Fox \& Binney method consists of averaging \eqnname~(\ref{eq:jjt})
+over the auxiliary angle along the path of the orbit, such that the
+oscillating cosine behavior cancels and the target action is obtained
+as the auxiliary-angle-averaged auxilary action. That is,
+\begin{equation}\label{eq:jjtint}
+  \int \dd \theta_i^A J_i^A = \int \dd \theta_i^A J_i + 2 \sum_{\vecn > 0} \vecn\,S_{\vecn}(J_i)\int \dd \theta_i^A\cos(\vecn\cdot\veca^A)\,,
+\end{equation}
+where the integral is over auxiliary angles calculated on the path of
+the orbit in the target potential. The factor $\vecn\,S_{\vecn}(J_i)$
+comes out of the integral as the target actions are conserved along
+the orbit. In this Equation, we integrate over the angle coordinate
+$\theta_i^A$ associated with the action $J_i$ that we want to compute
+(that is, $\theta_R$ for $J_R$ and so on). This is because of the
+factor $\vecn$: the condition $\vecn > 0$ means that when calculating
+$J_R$ there are terms with $\vecn = (n_R,0)$, which would not cancel
+according to the argument below when integrating over $\theta_Z$. For
+non-resonant orbits and after a long-enough orbit integration, the
+angles can be considered to be independent (that is, the orbit ``fills
+its torus''), such that we can write (for example, for $\theta_R$)
+\begin{equation}
+  \int \dd \theta_R^A\cos(\vecn\cdot\veca^A) = \mathrm{Re}\left[\int \dd \theta_R^A\,\exp(-i\vecn\cdot\veca^A)\right] = \mathrm{Re}\left[\exp(-i\,n_Z\,\theta^A_Z)\int \dd \theta_R^A\,\exp(-i\,n_R\,\theta^A_R)\right]\,.
+\end{equation}
+As long as the auxiliary angle $\theta_R$ goes through the full range
+$0$ to $2\pi$ along the orbit, the oscillatory behavior integrates to
+zero and \eqnname~(\ref{eq:jjtint}) simplifies to
+\begin{equation}\label{eq:aAIJ}
+  J_i = \frac{\int \dd \theta_i^A J_i^A}{\int \dd \theta_i^A}\,,
+\end{equation}
+that is, the target action is obtained by averaging the auxiliary
+action over the auxiliary angle.
+
+The left panel of \figurename~\ref{fig:aAI} shows a running average of
+the radial action calculated using \eqnname~(\ref{eq:aAIJ}) along the
+orbit of the progenitor in the simulation used in the main body of
+this paper. The time in this figure is in units of $35.56\Myr$, so the
+orbit is integrated for a total of $7.11\Gyr$. After a few radial
+periods, the radial action quickly converges with a remaining
+oscillatory behavior that is primarily a function of the radial angle,
+but that becomes smaller over time. With longer integrations and
+shorter time-steps, the radial action can converge to any desired
+tolerance. The duration of the orbit integration shown in this figure
+is what is typically used in this paper and gives a precision of
+$\approx2\,\%$ in the actions.
+
+Given the orbit integration already performed for calculating the
+actions, we can use \equationname~(\ref{eq:aat}) to calculate the
+frequencies and angles corresponding to the point $(\vecx,\vecv)$
+(this procedure is similar to that used in \citealt{McMillan08a}). We
+first re-write \equationname~(\ref{eq:aat}) in terms of the angle at
+time zero and the frequencies
+\begin{equation}\label{eq:aat2}
+  \veca = \veca(t=0) + \veco(\vecj) \,t = \veca^A +2\sum_{\vecn > 0} \frac{\partial S_{\vecn}(\vecj)}{\partial \vecj}\,\sin(\vecn\cdot\veca^A)\,.
+\end{equation}
+This a linear system of equations for each point along the orbit with
+unknowns $\veca(t=0)$, $\veco(\vecj)$, and the functions
+$\frac{\partial S_{\vecn}(\vecj)}{\partial \vecj}$. For a finely
+integrated orbit, this system is highly over-constrained (at least
+when the number of expansion terms is small) and its solution can be
+found by simple linear algebra for any desired number of expansion
+terms $\sin(\vecn\cdot\veca^A)$. For extra stability, we integrate the
+orbit both forward and backward in time, such that the desired angle
+$\veca(t=0)$ is at the center of the dependent variable $t$.
+
+The right panel of \figurename~\ref{fig:aAI} shows the auxiliary
+radial angle versus the auxiliary vertical angle along the orbit for
+the progenitor used in the main body of this paper. The periodic
+behavior in the angles is removed before the linear fit described in
+the previous paragraph, which is shown in this figure. It is clear
+that the contribution of the oscillatory terms
+$\sin(\vecn\cdot\veca^A)$ is small, such that a small number of
+expansion terms is sufficient (we limit the terms to $n_R < 4$ and
+$n_Z < 4$).
+
+The method described here is completely general, but the only obstacle
+to using it is that a sufficiently close isochrone potential has to be
+supplied. Sufficiently close means specifically that both the
+auxiliary radial and vertical angle have to go through the full range
+of $0$ to $2\pi$ along the orbit in the target potential. Whether they
+do or not is easy to check and I have found that with limited trial
+and error a good enough isochrone potential can be determined quickly
+for Milky-Way-like axisymmetric potentials, although it remains to be
+checked whether this holds for triaxial potentials as well. For
+analyzing stream data, the necessity of this trial-and-error
+determinition of the auxiliary potential does not pose a practical
+problem, as all of the stream stars are on very similar
+orbits. Therefore, the same auxiliary potential can be used for all
+action--angle calculations for a given potential and for different
+potentials as well, as long as they are not too different. More
+widespread use of this technique may require automated methods for
+determining a good auxiliary potential. In the unlikely circumstance
+that no good auxiliary isochrone potential can be found, the method
+discussed here can also be used with other auxiliary potentials for
+which the actions and angles can be calculated, such as other
+spherical potentials or the family of St\"{a}ckel potentials. This
+will, however, add a non-negligible amount of computation for the
+necessary numerical integrations.
 
@@
   Binney,~J. 2008, \mnras, 386, L47
+\bibitem[Binney(2012)]{Binney12a}
+  Binney,~J. 2012, \mnras, 426, 1324
 \bibitem[Carlberg(2009)]{Carlberg09a}
@@
   Koposov,~S.~E., Rix,~H.-W., \& Hogg,~D.~W. 2010, \apj, 712, 260
+\bibitem[McGill \& Binney(1990)]{McGill90a}
+  McGill,~C. \& Binney,~J. 1990, \mnras, 244, 634
+\bibitem[McMillan \& Binney(2008)]{McMillan08a}
+  McMillan,~P.~J. \& Binney,~J.~J. 2008, \mnras, 390, 429
 \bibitem[Ngan \& Carlberg(2014)]{Ngan14a}
@@
 \begin{figure}[t!]
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aajrjz.ps}
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaoroz.ps}
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaaraz.ps}\\
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aajrjp.ps}
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaorop.ps}
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaarap.ps}
+  \caption{}\label{fig:gd1_jt}
+%python plot_stream.py ../tex/gd1_evol_aajrjz.ps
+%python plot_stream.py ../tex/gd1_evol_aaoroz.ps
+%python plot_stream.py ../tex/gd1_evol_aaaraz.ps
+%python plot_stream.py ../tex/gd1_evol_aajrjp.ps
+%python plot_stream.py ../tex/gd1_evol_aaorop.ps
+%python plot_stream.py ../tex/gd1_evol_aaarap.ps
+\end{figure}
+
+\begin{figure}[t!]
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_timeshist.ps}
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_timesdO.ps}
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_timesda.ps}
+  \caption{}\label{fig:gd1_times}
+%python plot_stream.py ../tex/gd1_evol_timeshist.ps
+%python plot_stream.py ../tex/gd1_evol_timesdO.ps
+%python plot_stream.py ../tex/gd1_evol_timesda.ps
+\end{figure}
+
+\begin{figure}[t!]
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aadohist.ps}
+  \caption{}\label{fig:gd1_dohist}
+%python plot_stream.py ../tex/gd1_evol_aadohist.ps
+\end{figure}
+
+\begin{figure}[t!]
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaaparopar.ps}
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaaparoperp.ps}\\
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaaparaperp.ps}
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaapartime.ps}
+  \caption{}\label{fig:gd1_apar}
+%python plot_stream.py ../tex/gd1_evol_aaaparopar.ps
+%python plot_stream.py ../tex/gd1_evol_aaaparoperp.ps
+%python plot_stream.py ../tex/gd1_evol_aaaparaperp.ps
+%python plot_stream.py ../tex/gd1_evol_aaapartime.ps
+\end{figure}
+
+\begin{figure}[t!]
   \includegraphics[width=\textwidth,clip=]{gd1_evol_lb.ps}
@@
 \begin{figure}[t!]
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aajrjz.ps}
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaoroz.ps}
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaaraz.ps}\\
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aajrjp.ps}
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaorop.ps}
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaarap.ps}
-  \caption{}\label{fig:gd1_jt}
-%python plot_stream.py ../tex/gd1_evol_aajrjz.ps
-%python plot_stream.py ../tex/gd1_evol_aaoroz.ps
-%python plot_stream.py ../tex/gd1_evol_aaaraz.ps
-%python plot_stream.py ../tex/gd1_evol_aajrjp.ps
-%python plot_stream.py ../tex/gd1_evol_aaorop.ps
-%python plot_stream.py ../tex/gd1_evol_aaarap.ps
-\end{figure}
-
-\begin{figure}[t!]
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_timeshist.ps}
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_timesdO.ps}
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_timesda.ps}
-  \caption{}\label{fig:gd1_times}
-%python plot_stream.py ../tex/gd1_evol_timeshist.ps
-%python plot_stream.py ../tex/gd1_evol_timesdO.ps
-%python plot_stream.py ../tex/gd1_evol_timesda.ps
-\end{figure}
-
-\begin{figure}[t!]
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aadohist.ps}
-  \caption{}\label{fig:gd1_dohist}
-%python plot_stream.py ../tex/gd1_evol_aadohist.ps
-\end{figure}
-
-\begin{figure}[t!]
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaaparopar.ps}
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaaparoperp.ps}\\
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaaparaperp.ps}
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaapartime.ps}
-  \caption{}\label{fig:gd1_apar}
-%python plot_stream.py ../tex/gd1_evol_aaaparopar.ps
-%python plot_stream.py ../tex/gd1_evol_aaaparoperp.ps
-%python plot_stream.py ../tex/gd1_evol_aaaparaperp.ps
-%python plot_stream.py ../tex/gd1_evol_aaapartime.ps
-\end{figure}
-
-\begin{figure}[t!]
  \includegraphics[width=0.48\textwidth,clip=]{gd1_evol_2Gya_xz.ps}
@@
 
+\begin{figure}[t!]
+ \includegraphics[width=0.48\textwidth,clip=]{aAI_jr.ps}
+ \includegraphics[width=0.48\textwidth,clip=]{aAI_araz.ps}
+  \caption{}\label{fig:aAI}
+%python plot_aAI.py ../tex/aAI_jr.ps
+%python plot_aAI.py ../tex/aAI_araz.ps
+\end{figure}
+
 \end{document}

### 2013-12-26 tex/ms.tex
streams: mention that this is implemented in galpy
streams: description of the simulation

@@
 \sectionname~\ref{sec:discussion} and I conclude in
-\sectionname~\ref{sec:conclusion}. In Appendix
-\sectionname~\ref{sec:aa} a novel method for calculating actions,
-frequencies, and angles for any orbit in any static gravitational
-potential is presented.
+\sectionname~\ref{sec:conclusion}. In \appendixname~\ref{sec:aa} a
+novel method for calculating actions, frequencies, and angles for any
+orbit in any static gravitational potential is presented.
 
@@
 action--angle space has been described by various previous authors
-\citep[\eg,][]{Helmi99a,Tremaine99a,SomeJohnstonPaper,Sanders13b}. I
-provide here a brief description of the dynamics of tidal streams that
-is relevant for what follows.
+\citep[\eg,][]{Helmi99a,Tremaine99a,Johnston99a,Sanders13b}. I provide
+here a brief description of the dynamics of tidal streams that is
+relevant for what follows.
 
@@
 corresponds to an offset in the frequencies $\Delta \veco$;
-ultimately, this offset is responsible for the spreading of the stream
-over long stretches on the sky. Because both the progenitor's angles and
-the star's angles increase linearly with time, albeit at different
-frequencies, the difference in the angles also increases linearly in
-time. Therefore at time $t=0$ the difference in angles is
+ultimately, this frequency offset is responsible for the spreading of
+the stream over long stretches on the sky. Because both the
+progenitor's angles and the star's angles increase linearly with time,
+albeit at different frequencies, the difference in the angles also
+increases linearly in time. Therefore at time $t=0$ the difference in
+angles is
 \begin{equation}
@@
 \end{equation}
-As the initial offset $\Delta \veca(-\Delta t$ in angles is small
-compared to the growth $\Delta\veco\Delta t$, $\Delta \veca(t=0)
-\approx \Delta \veco \Delta t$.
+As the initial offset $\Delta \veca(-\Delta t)$ in angles is small
+compared to the growth, we have that $\Delta\veco\Delta t$, $\Delta
+\veca(t=0) \approx \Delta \veco \Delta t$.
 
@@
 along which a star moves away from the progenitor is related to the
-initial action offset by $\Delta \veco \approx (\partial^2 H
-/ \partial \vecj \partial \vecj) \Delta \vecj$, in the limit of small
-action offsets, as $\veco = \partial H / \partial \vecj$. For an
+initial action offset by $\Delta \veco \approx (\partial^2 H /
+\partial \vecj \partial \vecj) \Delta \vecj$, in the limit of small
+action offsets, because $\veco = \partial H / \partial \vecj$. For an
 approximately one-dimensional stream to form, the Hessian matrix
@@
 $\unitvec \veco$ by flexible modeling of the distribution of $\Delta
-\vecj$.
-
-\section{Generative models of tidal streams}\label{sec:method}
+\vecj$ or $\Delta \veco$.
+
+\section{Generative models of tidal streams}\label{sec:generative}
+
+\subsection{A simulation}\label{eq:sim}
+
+I illustrate the theoretical considerations set forth in this paper by
+using a simulated tidal stream. This stream is set up similarly to
+that used in \citep{Sanders13b}. The stream is constructed by running
+an $N$-body simulation of a King cluster on an orbit similar to that
+of the GD-1 stream \citep{Koposov10a}. Specifically, a King cluster
+with a mass of $2\times10^4\msun$, a tidal radius of $0.07\kpc$, and a
+ratio of the central potential to the velocity dispersion squared of
+$2.0$ is sampled using $10^4$ particles. The cluster is evolved in an
+external logarithmic host potential with $V_c = 220\kms$ and a
+flattening $q=0.9$ ($\Phi = V_c^2 \ln (R^2 + z^2/q^2$) using the
+gyrfalcon code \citep{Dehnen00a,Dehnen02a} with a softening of
+$1.5\pc$ in the NEMO toolkit \citep{Teuben95a}. The cluster is evolved
+for $5.011\Gyr$ ($=5.125\kpc/[\kms]$) from the initial condition
+$(X,Y,Z) = (-11.63337239,-20.76235661,-10.631736273934635)\kpc$ and
+$(V_X,V_Y,V_Z) = (-128.8281653,42.88727925,79.172383882274971)\kms$,
+chosen such that the cluster ends up at roughly the observed location
+of the GD-1 stream at the end of the simulation
+\citep{Koposov10a}. The progenitor's orbit has an eccentricity of
+$0.31$, a pericenter of $13.7\kpc$, a current $r=14.4\kpc$, such that
+the cluster is close to pericenter at the end of the simulation. The
+actions of the progenitor are $(J_R,L_Z,J_Z) =
+(288.5,-3173.7,897.6)\kpc\kms$ and its orbital frequencies are
+$(\Omega_R,\Omega_\phi,\Omega_Z) = (15.7,-10.8,11.9)\Gyr\inv$.
+
+The final position of the stream in $(X,Z)$ is shown in
+\figurename~\ref{fig:gd1_xy}. \figurename~\ref{fig:gd1_jt} shows the
+final position of the stream in action--frequency--angle coordinates,
+calculated using the algorithm described in
+\appendixname~\ref{sec:aa}. As described in
+\sectionname~\ref{sec:dynamics}, a one-dimensional tidal stream forms
+because even though the action distribution of the tidal debris is not
+isotropic or one-dimensional, the Hessian $(\partial^2 H / \partial
+\vecj \partial \vecj)\large|_{\vecj^p}$ is so strongly dominated by a
+single large eigenvalue (the ratio of the largest to the second
+largest eigenvalue is about $30$), that the frequency distribution of
+the debris is essentially one dimensional. The angle distribution of
+the debris at the end of the simulation is therefore essentially one
+dimensional as well.
+
+\subsection{A model in frequency--action space}\label{sec:modeloa}
+
+\section{The track of a tidal stream}\label{sec:track}
+
+\section{Mock stream data}\label{sec:mock}
+
+\section{The full stream PDF}\label{sec:pdf}
+
+\section{Discussion}\label{sec:discussion}
+
+\section{Conclusion}\label{sec:conclusion}
 
@@
 
+The action--angle method described in this Appendix is implemented in
+the \emph{galpy} Galactic dynamics code\footnote{Available at
+  \protect{\url{http://github.com/jobovy/galpy}}~.}.
+
 %BOVY: reference point
@@
 Carlberg,~R.~G. 2012, \apj, 748, 20
+\bibitem[Dehnen(2000)]{Dehnen00a}
+  Dehnen,~W. 2000, \apj, 536, L39
+\bibitem[Dehnen(2002)]{Dehnen02a}
+  Dehnen,~W. 2002, J.~Comput.~Phys., 179, 27
 \bibitem[Eyre \& Binney(2011)]{Eyre11a}
@@
   Sanders,~J. \& Binney,~J. 2013b, \mnras, 433, 1826
+\bibitem[Teuben(1995)]{Teuben95a}
+  Teuben,~P.~J. 1995, in Astronomical Data Analysis Software and Systems IV, 
+  ed.~R.~Shaw, H.~E.~Payne and J.~J.~E.~Hayes., PASP Conf Series 77, 398
 \bibitem[Tremaine(1999)]{Tremaine99a}

### 2013-12-26 py/plot_stream.py
streams: print misalignment

@@ -514,6 +514,7 @@ def plot_stream_aa(plotfilename):
         dOdir= numpy.median(dO4dir,axis=1)
         dOdir/= numpy.sqrt(numpy.sum(dOdir**2.))
         dO1d= numpy.dot(dOdir,dO)
+        print "Misalignment:", numpy.arccos(numpy.sum(dOdir*progaa[-1,3:6])/numpy.sqrt(numpy.sum(dOdir**2.)*numpy.sum(progaa[-1,3:6]**2.)))/numpy.pi*180.-180.
         dO1d[dO1d < 0.]*= -1.
         bovy_plot.bovy_print()
         bovy_plot.bovy_hist(dO1d,range=[0.,0.4],bins=61,

### 2013-12-26 tex/ms.tex
streams: some figure captions

@@
 
-\subsection{A simulation}\label{eq:sim}
+\subsection{A simulation}\label{sec:sim}
 
@@
   \includegraphics[width=0.5\textwidth,clip=]{gd1_evol_xz.ps}
-  \caption{}\label{fig:gd1_xy}
+  \caption{Mock tidal stream simulated as described in
+    \sectionname~\ref{sec:sim} after $5.01\Gyr$ of evolution in
+    Galactocentric $X$ and $Z$ coordinates. The grey dot shows the
+    current position of the progenitor cluster and the dashed line
+    shows its orbit. The solid line gives the stream track, calculated
+    using the method described in \sectionname~\ref{sec:track}. The
+    inset shows a zoomed-in version of part of the leading stream that
+    more clearly shows the offset between the stream track and the
+    progenitor orbit.}\label{fig:gd1_xy}
 %python plot_stream.py ../tex/gd1_evol_xz.ps
@@
   \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaarap.ps}
-  \caption{}\label{fig:gd1_jt}
+  \caption{Simulated tidal stream from \sectionname~\ref{sec:sim}
+    after $5.01\Gyr$ of evolution in action--frequency--angle
+    coordinates. The top row displays projections of the actions
+    (left), frequencies (middle), and angles (right) in the $R$ and
+    $Z$ directions; the bottom panels shows the same for $R$ and
+    $\phi$. While the action distribution has a quite complicated
+    form, the frequency distribution is close to one-dimensional and
+    the leading and trailing parts of the stream clearly separate. In
+    the right panels, the angles of the progenitor have each been
+    shifted to $\pi$ and the dashed lines show the direction of the
+    progenitor's frequency vector. The misalignment between the
+    progenitor's frequency vector and the mean frequency vector of the
+    debris is $0.18^\circ$.}\label{fig:gd1_jt}
 %python plot_stream.py ../tex/gd1_evol_aajrjz.ps
@@
   \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_timesda.ps}
-  \caption{}\label{fig:gd1_times}
+  \caption{Properties of the simulated tidal stream of
+    \sectionname~\ref{sec:sim} as a function of time. Stripping times
+    are calculated by linear regression of the angle difference
+    between a stream particle and the progenitor versus the equivalent
+    frequency difference. The distribution of stripping times is shown
+    in the left panel.  While the stripping process happens in bursts
+    with a period of roughly $40\Myr$ (the radial period of the
+    progenitor), it can be approximated as constant (more
+    coarsely-binned solid histogram). The magnitude of the frequency
+    offset between stream particles and the progenitor is shown in the
+    middle panel. While particles stripped at pericenter have larger
+    frequency offsets than those between pericenter passages, there is
+    no secular trend and the typical frequency offset (averaged over a
+    radial period) can be considered constant with time. The right
+    panel shows the magnitude of the angle offset between stream
+    particles and the progenitor as a function of stripping
+    time. There is a spread in the angle offset at all stripping times
+    because of the dispersion in frequency offsets, but typically
+    stars that have been stripped earlier reach larger angle
+    offsets. That is, self-sorting happens only to a small
+    degree.}\label{fig:gd1_times}
 %python plot_stream.py ../tex/gd1_evol_timeshist.ps

### 2013-12-26 py/plot_stream.py
streams: tweaks

@@ -518,11 +518,12 @@ def plot_stream_aa(plotfilename):
         dO1d[dO1d < 0.]*= -1.
         bovy_plot.bovy_print()
         bovy_plot.bovy_hist(dO1d,range=[0.,0.4],bins=61,
-                           normed=True,
-                           xlabel=r'$|\Delta \mathbf{\Omega}|\,(\mathrm{Gyr}^{-1})$',
+                            normed=True,
+                            xlabel=r'$\Large|\Delta \mathbf{\Omega}_\parallel\Large|\,(\mathrm{Gyr}^{-1})$',
                             histtype='step',color='k',zorder=10)
         #Overplot best-fit Gaussian
         xs= numpy.linspace(0.,0.4,1001)
+        print numpy.mean(dO1d), numpy.std(dO1d)
         bovy_plot.bovy_plot(xs,1./numpy.sqrt(2.*numpy.pi)/numpy.std(dO1d)\
                                 *numpy.exp(-(xs-numpy.mean(dO1d))**2./2./numpy.var(dO1d)),
                             '--',color='k',overplot=True,lw=2.,zorder=0)
@@ -530,6 +531,7 @@ def plot_stream_aa(plotfilename):
                                       numpy.array([numpy.log(numpy.mean(dO1d)*2.),
                                                    numpy.log(numpy.std(dO1d))]),
                                       args=(dO1d,))
+        print numpy.exp(bestfit)
         bovy_plot.bovy_plot(xs,gausstimesvalue(bestfit,xs,nologsum=True),
                             '-',color='0.4',overplot=True,lw=2.,zorder=0)
         bovy_plot.bovy_end_print(plotfilename)
@@ -664,6 +666,7 @@ def plot_stream_aa(plotfilename):
         plotx= numpy.fabs(numpy.dot(dangle.T,dOdir))
         ploty= numpy.sqrt(numpy.sum(dO**2.,axis=0)\
                               -(numpy.dot(dO.T,dOdir))**2.)
+        print numpy.std(ploty)
         xrange=[0.,1.3]
         yrange=[0.,0.005]
         xlabel= r'$\Large|\Delta \mathbf{\theta}_\parallel\Large|$'

### 2013-12-26 py/plot_stream_2Gya.py
streamsL change label

@@ -111,7 +111,7 @@ def plot_stream_xz(plotfilename):
         bovy_plot.bovy_plot(pox,poz,'o',color='0.5',mec='none',overplot=True,ms=8)
         bovy_plot.bovy_plot(pvec[0,:],pvec[1,:],'k--',overplot=True,lw=1.)
     if 'sim' in plotfilename:
-        bovy_plot.bovy_text(r'$\mathrm{simulated\ stream}$',
+        bovy_plot.bovy_text(r'$\mathrm{mock\ stream}$',
                             bottom_left=True,size=16.)
     else:
         bovy_plot.bovy_text(r'$N\!\!-\!\!\mathrm{body\ stream}$',

### 2013-12-26 tex/ms.tex
streams: more figure captions and rest of model description

@@
 \newcommand{\paramspot}{\ensuremath{\vec{p}_\Phi}}
+\newcommand{\sigv}{\ensuremath{\sigma_v}}
 
@@
 external logarithmic host potential with $V_c = 220\kms$ and a
-flattening $q=0.9$ ($\Phi = V_c^2 \ln (R^2 + z^2/q^2$) using the
+flattening $q=0.9$ ($\Phi = V_c^2 \ln (R^2 + z^2/q^2)$) using the
 gyrfalcon code \citep{Dehnen00a,Dehnen02a} with a softening of
@@
 The final position of the stream in $(X,Z)$ is shown in
-\figurename~\ref{fig:gd1_xy}. \figurename~\ref{fig:gd1_jt} shows the
+\figurename~\ref{fig:gd1_xz}. \figurename~\ref{fig:gd1_jt} shows the
 final position of the stream in action--frequency--angle coordinates,
@@
 \subsection{A model in frequency--action space}\label{sec:modeloa}
+
+The action and frequency distributions shown in
+\figurename~\ref{fig:gd1_jt} combined with the knowledge that for
+\emph{any} tidal stream the Hessian $(\partial^2 H / \partial \vecj
+\partial \vecj)\large|_{\vecj^p}$ must be strongly dominated by a
+single large eigenvalue leads us to propose that frequency--angle
+space is a better coordinate system than action--angle space for
+modeling tidal streams. That is, the distribution of actions in a
+tidal stream has a complicated structure that is hard to capture using
+a simple, analytic form (see also the discussion in
+\citealt{Eyre11a}). Especially the ``bowtie'' structure in $(J_R,L_Z)$
+and the overlapping leading- and trailing-arm distributions in
+($J_R,J_Z$) are hard to model with a simple distribution function. The
+distribution of frequencies, however, is close-to one dimensional, its
+direction of largest variance can be well-modeled as a Gaussian
+(although below we will model it slightly differently), and its
+parameters can all be easily estimated from the velocity dispersion of
+the progenitor, the orbit of the progenitor, and the gravitational
+potential, as discussed below. Therefore, we will model tidal streams
+in frequency--angle space.
+
+A generative model of a tidal stream in frequency--angle space
+requires three ingredients: (a) a model for the distribution of times
+at which stars are stripped from the progenitor, (b) a prescription
+for the distribution of frequency offsets from the progenitor at every
+given stripping time, and (c) a description of the angle offsets for
+any given frequency offset and stripping time. Having specified these
+ingredients, we can then generate tidal debris at any given stripping
+time and evolve it forward using the simple linear dynamics in
+frequency--angle space discussed in
+\sectionname~\ref{sec:dynamics}. The initial angle offsets are small
+enough that they are likely unobservable even with futuristic data
+(especially in the direction of the frequency offset, where the
+initial angle offset is quickly overwhelmed by the subsequent
+dynamical evolution; for the simulated stream used here this happens
+after approximately $20\Myr$). For that reason, we will assume that
+the initial angle offsets are independent of both stripping time and
+frequency offset. In what follows, we will model them using a simple
+isotropic Gaussian distribution.
+
+Ingredients (a) and (b) require more careful modeling. The right panel
+of \figurename~\ref{fig:gd1_times} shows the distribution of times at
+which stream particles are stripped from the progenitor. These
+stripping times are calculated from the final snapshot by linear
+regression of a particle's angle offset $\Delta \veca$ from the
+progenitor, versus the frequency offset $\Delta \veco$ as 
+\begin{equation}
+  \Delta t = \frac{\Delta \veco \cdot \Delta \veca}{|\Delta \veco|^2}\,.
+\end{equation}
+The dotted line shows a finely-binned histogram of these stripping
+times, which indicates that stripping happens in bursts with a radial
+period of about $400\Myr$, the radial period of the progenitor's
+orbit. Thus, as expected, stripping happens primarily at pericenter
+passages, with a smaller number of particles lost between pericenter
+passages. Realistic models of a stream, especially those that require
+good models of the surface-density structure of the stream as a
+function of stream angle, need to model this non-uniform stripping
+process.
+
+The middle panel of \figurename~\ref{fig:gd1_times} shows the
+frequency offsets between stream particles and the progenitor as a
+function of strippng time, corresponding to ingredient (b) above. The
+same burstiness that is apparent in the distribution of stripping
+times shows up here. Particles stripped at pericenter typically have
+larger frequency offsets than those lost at larger radii. This can be
+described as saying that particles are \emph{removed} at pericenter,
+but only \emph{peeled off} between pericenter passages. Again,
+realistic models of tidal streams need to take this detailed structure
+into account.
+
+The right panel of \figurename~\ref{fig:gd1_times} shows the angle
+offset (at the end of the simulation) versus the stripping time. There
+is a spread in the angle offset reach for any given stripping time,
+but it is nevertheless typically the case that particles that have
+been stripped earlier are found at larger angle offsets
+today. Self-sorting---the tendency of particles stripped later with
+larger frequency offsets to overtake particles stripped earlier at
+smaller frequency differences and thus to erase correlations between
+particle position along the stream and stripping time---happens to a
+small extent but the narrow distribution of frequency offsets limits
+the ability of particles stripped later to reach large distances from
+the progenitor.
+
+\subsection{The fiducial stream model}
+
+The discussion of the simulated stream's properties as a function of
+stripping time in the previous section indicates that we can build
+simple models of the structure of tidal streams when considering them
+in frequency--angle space. In this section, we propose a very simple
+fiducial model that we will use in the remainder of this paper. This
+model combines simple analytic forms for ease of computation that are
+nevertheless realistic enough to provide adequate models for tidal
+streams in many applications.
+
+First, we approximate the bursty, non-uniform distribution of
+stripping times in \figurename~\ref{fig:gd1_times} with a uniform
+distribution
+\begin{align}\label{eq:pt}
+  \begin{split}
+    p(\Delta t) & \propto 1\,,\qquad 0 < \Delta t < t_d\\
+    & = 0\,,\qquad \mathrm{otherwise}\,.
+  \end{split}
+\end{align}
+The thick, solid histogram in the right panel of
+\figurename~\ref{fig:gd1_times} shows a coarser-binned histogram of
+stripping times, indicating that the uniform distribution is a good
+approximation over intervals larger than a radial period ($\approx
+400\Myr$). The \emph{disruption time} $t_d$ is a free parameter of the
+model that needs to be determined for each stream. For the model of
+the simulated stream used here we set $t_d = 4.5\Gyr$.
+
+Similarly, we model the distribution of frequency offsets between
+stream particles and the progenitor to be independent of stripping
+time, even though the middle panel of \figurename~\ref{fig:gd1_times}
+clearly shows that particles are stripped at larger frequency
+differences at pericentric passages than between them. The fact that
+there is no long-term trend in the typical frequency offset means that
+this is a good model on time-scales larger than the radial
+period. While this means that the small-scale structure of the stream
+will not be perfectly represented by the model, the large-scale
+structure of the stream will be fine. For using tidal streams to
+constrain the gravitational potential, large arcs are most important,
+such that this simple model will be adequate for such inferences.
+
+The distribution of frequencies is three dimensional. A general
+description in terms of a Gaussian distribution would have nine free
+parameters. However, we can model it with fewer free parameters as
+follows. First, following \citet{Eyre11a}, we approximate the
+\emph{action} distribution as a Gaussian with standard deviations
+\begin{equation}
+  \sigma_{J_R} = \frac{1}{\pi}\,\sigv\,\left(r_{\mathrm{apo}}-r_{\mathrm{peri}}\right)\,,\quad 
+  \sigma_{L_Z} = \sigv\,r_{\mathrm{peri}}\,,\quad 
+  \sigma_{J_Z} = \frac{2}{\pi}\,\sigv\,Z_{\mathrm{max}}\,,
+\end{equation}
+where $\sigv$ is the velocity dispersion of the progenitor,
+$r_{\mathrm{apo}}$ and $r_{\mathrm{peri}}$ are the apo- and pericenter
+radii of the progenitor orbit, respectively, and $Z_{\mathrm{max}}$ is
+the maximum $Z$ reached on this orbit. We assume that the all
+correlations between the actions are zero (see
+\figurename~\ref{fig:gd1_jt}). Then we propagate this Gaussian
+variance to frequency space using the Hessian $(\partial^2 H /
+\partial \vecj \partial \vecj)\large|_{\vecj^p}$\footnote{The Hessian
+  $(\partial^2 H / \partial \vecj \partial \vecj)\large|_{\vecj^p}$
+  can be computed using the method in \appendixname~\ref{sec:aa} by
+  computing $\partial (\vecj,\veca) / \partial (\vecx,\vecv)$ and
+  $\partial (\veco,\veca) / \partial (\vecx,\vecv)$ and forming
+  $\partial \veco / \partial \vecj = \left(\partial (\vecj,\veca) /
+  \partial (\vecx,\vecv)\right)^{-1}\,\left(\partial (\veco,\veca) /
+  \partial (\vecx,\vecv)\right)$.}, which gives the variance matrix
+$V_\veco$. The principal eigenvector of this matrix is the model
+direction along which the stream spreads. While the model I propose
+here is a simple one, it works well without having been tweaked for
+the simulation under scrutiny. The angle between the principal
+eigenvector of $V_\veco$ and the progenitor's frequency vector is
+$0.50^\circ$ while the value measured from the simulation is
+$0.18^\circ$. For a model with an isotropic action distribution, the
+misalignment would be $1.28^circ$.
+
+We then further constrain the mean frequency offset to lie along the
+principal eigenvector of $V_\veco$. This mean frequency offset can be
+described in terms of a parameter $\mu_\Omega = \Delta \Omega^m /
+\sigma_{\Omega,1}$, where $\sigma_{\Omega,1}$ is the square root of the
+largest eigenvalue of $V_\veco$ and $\Delta \Omega^m$ is the mean
+frequency offset. \figurename~\ref{fig:gd1_dohist} shows the
+distribution frequency offsets in the direction of the median
+frequency offset in the stream. The dashed line is a Gaussian fit to
+this distribution, while the solid line is a fit of the form
+\begin{equation}\label{eq:do}
+  p(|\Delta \veco|_\parallel) = |\Delta \veco|_\parallel\,\mathcal{N}\left(|\Delta \veco|_\parallel|\Delta \Omega^m,\sigma_{\Omega,1}^2\right)\,,
+\end{equation}
+where $\mathcal{N}(\cdot|\cdot,\cdot)$ represents a Gaussian
+distribution. The fit is equally good (because the dispersion is much
+smaller than the mean) and we choose the second form because it
+simplifies some of the stream-track calculations below. The best-fit
+$(\Delta \Omega^m,\sigma_{\Omega,1})$ is $(0.19,0.033)\Gyr\inv$ in
+both cases. Based on this information, a model with $\sigv =
+0.365\kms$ and $\mu_\Omega = 6$ provides a good fit to the simulation
+data. The distribution of frequency offsets in the two-dimensional
+space perpendicular to the principal eigenvector of $V_\veco$ is
+modeled as a zero-mean Gaussian with variance matrix given by the
+projection of $V_\veco$ onto this space.
+
+Because we expect the typical $\Delta\veco$ to scale with the
+velocity dispersion of the progenitor (see also \citealt{Sanders13a}),
+I conjecture that a constant $\mu_\Omega$ can be used for modeling
+cold tidal streams of any (small-ish) velocity dispersion. Further
+modeling of simulated streams, however, is necessary to check this and
+to find the optimizal value of $\mu_\Omega$. Fixing $\mu_\Omega$
+leaves a single free parameter \sigv\ for modeling the frequency
+offset distribution of a tidal stream. In modeling to observed data,
+this parameter needs to be fit and we expect it to be proportional to
+the velocity dispersion of the progenitor, although this
+proportionality needs to be checked more carefully (but see Figure 3
+in \citealt{Sanders13a}, which shows that the size of the frequency
+distribution scales as mass$^{1/3}$, or approximately as the velocity
+dispersion). Similarly, the initial angle distribution, modeled here
+as an isotropic Gaussian, has a charactaristic spread $\sigma_{\theta}$
+which we set here to $0.00365$ ($=\sigv/[100\kms]$), and we likewise
+expect this spread to scale with the velocity dispersion of the
+progenitor.
+
+The fiducial model for a tidal stream in frequency--angle coordinates
+he proposed here is therefore described by essentially two free
+parameters in addition to the progenitor's phase-space position: the
+progenitor's velocity dispersion \sigv\ and the disruption time
+$t_d$. As shown below, these two parameters are the most important in
+determining the track of the stream. More sophisticated models can
+leave $\mu_\Omega$ and $\sigma_{\theta}$ as free paramters as well. In
+what follows, we will fix these to the values given in the previous
+paragraph. It must be stressed, however, that all of these parameters
+here have been fixed ``by-eye'', and better fits might be possible
+using more quantitative fit procedures. But such procedures are not
+pursued here further.
 
@@
     more clearly shows the offset between the stream track and the
-    progenitor orbit.}\label{fig:gd1_xy}
+    progenitor orbit.}\label{fig:gd1_xz}
 %python plot_stream.py ../tex/gd1_evol_xz.ps
@@
     in the left panel.  While the stripping process happens in bursts
-    with a period of roughly $40\Myr$ (the radial period of the
+    with a period of roughly $400\Myr$ (the radial period of the
     progenitor), it can be approximated as constant (more
@@
   \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aadohist.ps}
-  \caption{}\label{fig:gd1_dohist}
+  \caption{Distribution of the magnitude of frequency offsets between
+    stream particles and the progenitor along the direction of the
+    mean frequency offset. The dashed line is a Gaussian fit to this
+    distribution, while the solid line is a fit of our preferred form
+    for this distribution, given in \equationname~(\ref{eq:do}). Both
+    forms provide adequate fits to the
+    distribution. }\label{fig:gd1_dohist}
 %python plot_stream.py ../tex/gd1_evol_aadohist.ps
@@
   \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaapartime.ps}
-  \caption{}\label{fig:gd1_apar}
+  \caption{Stream properties as a function of angle along the
+    stream. The top, left panel shows the magnitude of frequency
+    offsets along the stream. The gray line shows the mean predicted
+    frequency offset calculated using the framework of
+    \sectionname\sectionname~\ref{sec:modeloa} and \ref{sec:track},
+    while the grayscale bands show the $68\,\%$ and $95\,\%$ limits of
+    the predicted distribution. The top, right panel shows the
+    magnitude of frequency offsets perpendicular to those in the left
+    panel. The predicted distribution is constant as a function of
+    angle and therefore not shown. The bottom, left panel shows the
+    angle offset perpendicular to the stream and the grayscale bands
+    show $68\,\%$ and $95\,\%$ limits of the predicted distribution of
+    angle offsets; the white dashed line shows the $68\,\%$ limit
+    calculated using a simplified method. The bottom, right panel
+    shows the stripping time. The gray line and grayscale bands again
+    show the mean and $68\,\%$ and $95\,\%$ confidence limits of the
+    predicted distribution. While the ``burstiness'' related to the
+    energetic stripping at pericenter that is apparent in the
+    simulated data's distributions is (by design) absent in the model,
+    the overall scale and the mean trends with angle are well
+    represented by the model.}\label{fig:gd1_apar}
 %python plot_stream.py ../tex/gd1_evol_aaaparopar.ps
@@
   \includegraphics[width=\textwidth,clip=]{gd1_evol_ld.ps}
-  \caption{}\label{fig:gd1_lbd}
+  \caption{Simulated stream after $5.01\Gyr$ as a function of Galactic
+    longitude. The top panel shows the stream in Galactic latitude and
+    the bottom panel shows the distance from the Sun. The progenitor
+    is shown as the large gray dot and its orbit as the dashed
+    lines. The predicted track of the stream is shown as the solid
+    line. Insets show various zoomed-in parts of the stream to show
+    the stream, progenitor orbit, and predicted stream track up
+    close.}\label{fig:gd1_lbd}
 %python plot_stream.py ../tex/gd1_evol_lb.ps
@@
   \includegraphics[width=\textwidth,clip=]{gd1_evol_lpmbb.ps}
-  \caption{}\label{fig:gd1_lbv}
+  \caption{Simulated stream velocities as a function of Galactic
+    longitude. Same as \figurename~\ref{fig:gd1_lbd}, but for the
+    stream in heliocentric line-of-sight velocity $V_{\mathrm{los}}$
+    (top panel), proper motion in the direction of Galactic longitude
+    (middle panel), and in the direction of Galactic latitude (bottom
+    panel). Insets again show various interesting parts of the stream
+    up close.}\label{fig:gd1_lbv}
 %python plot_stream.py ../tex/gd1_evol_lvlos.ps
@@
  \includegraphics[width=0.48\textwidth,clip=]{gd1_evol_2Gya_xz_sim.ps}
-  \caption{}\label{fig:gd1_sim}
+  \caption{Mock stream data generated from the model. The left panel
+    shows the same $N$-body stream as shown in previous figures, but
+    $1.96\Gyr$ earlier in its evolution. Symbols, lines, and the
+    stream model are the same as those in
+    \figurename~\ref{fig:gd1_xz}, except that the model progenitor
+    phase-space location from \figurename~\ref{fig:gd1_xz} has been
+    integrated backward in time for $1.96\Gyr$ (no other changes were
+    made to the model's parameters). The right panel shows mock stream
+    data drawn from the model using the procedures discussed in
+    \sectionname~\ref{sec:mock}. Overall, the spatial distribution of
+    the mock data is very similar to that of the $N$-body
+    stream.}\label{fig:gd1_sim}
 %python plot_stream_2Gya.py ../tex/gd1_evol_2Gya_xz.ps

### 2013-12-26 tex/ms.tex
streams: more captions and tweaks to aa description

@@
 this paper. The time in this figure is in units of $35.56\Myr$, so the
-orbit is integrated for a total of $7.11\Gyr$. After a few radial
+orbit is integrated for a total of $7.11\Gyr$. The auxiliary isochrone
+potential has a scale parameter of $6.4\kpc$. After a few radial
 periods, the radial action quickly converges with a remaining
@@
  \includegraphics[width=0.48\textwidth,clip=]{aAI_araz.ps}
-  \caption{}\label{fig:aAI}
+  \caption{Action--angle calculations using an orbit-integration-based
+    method. The right panel shows the cumulative mean of the radial
+    actions calculated in an auxiliary isochrone potential calculated
+    along the orbit of the progenitor used in the main body of this
+    paper. The color-coding corresponds to the radial angle
+    (calculated in the isochrone potential) along the orbit. The time
+    is in units of $35.56\Myr$ and the radial action is normalized by
+    its value at $t=200$. The left panel shows the radial and vertical
+    angles calculated in the auxiliary isochrone potential, with
+    color-coding corresponding to the azimuthal angle. The wiggles in
+    the angles are due to the isochrone potential not being the
+    correct potential, but these wiggles are fit and removed to
+    calculate the true angles.}\label{fig:aAI}
 %python plot_aAI.py ../tex/aAI_jr.ps

### 2013-12-26 tex/ms.tex
streams: mention that we model leading and trailing separately, sign for leading/trailing, some more tweaks

@@
 using a simulated tidal stream. This stream is set up similarly to
-that used in \citep{Sanders13b}. The stream is constructed by running
+that used in \citet{Sanders13b}. The stream is constructed by running
 an $N$-body simulation of a King cluster on an orbit similar to that
@@
 external logarithmic host potential with $V_c = 220\kms$ and a
-flattening $q=0.9$ ($\Phi = V_c^2 \ln (R^2 + z^2/q^2)$) using the
+flattening $q=0.9$ ($\Phi = V_c^2 \ln [R^2 + z^2/q^2]$) using the
 gyrfalcon code \citep{Dehnen00a,Dehnen02a} with a softening of
@@
 potential, as discussed below. Therefore, we will model tidal streams
-in frequency--angle space.
+in frequency--angle space. Because the leading and trailing arm are
+well-separated in frequency space, we model each arm individually.
 
@@
 
-The distribution of frequencies is three dimensional. A general
-description in terms of a Gaussian distribution would have nine free
-parameters. However, we can model it with fewer free parameters as
-follows. First, following \citet{Eyre11a}, we approximate the
-\emph{action} distribution as a Gaussian with standard deviations
+The distribution of frequencies is three dimensional. Because we model
+the leading and the trailing arm separately, a single-peaked
+distribution suffices to describe the frequency distribution. A
+general description in terms of a Gaussian distribution would have
+nine free parameters. However, we can model it with fewer free
+parameters as follows. First, following \citet{Eyre11a}, we
+approximate the \emph{action} distribution as a Gaussian with standard
+deviations
 \begin{equation}
@@
 described in terms of a parameter $\mu_\Omega = \Delta \Omega^m /
-\sigma_{\Omega,1}$, where $\sigma_{\Omega,1}$ is the square root of the
-largest eigenvalue of $V_\veco$ and $\Delta \Omega^m$ is the mean
-frequency offset. \figurename~\ref{fig:gd1_dohist} shows the
-distribution frequency offsets in the direction of the median
-frequency offset in the stream. The dashed line is a Gaussian fit to
-this distribution, while the solid line is a fit of the form
+\sigma_{\Omega,1}$, where $\sigma_{\Omega,1}$ is the square root of
+the largest eigenvalue of $V_\veco$ and $\Delta \Omega^m$ is the mean
+frequency offset. The sign of this parameter sets whether we are
+modeling a leading or a trailing arm. \figurename~\ref{fig:gd1_dohist}
+shows the distribution frequency offsets in the direction of the
+median frequency offset in the stream. The dashed line is a Gaussian
+fit to this distribution, while the solid line is a fit of the form
 \begin{equation}\label{eq:do}
@@
 \section{Conclusion}\label{sec:conclusion}
-
-The discussions of the dynamics of stream formation in the previous
-section directly give rise of a generative model of a tidal stream. A
-star in a tidal stream is \emph{generated} through being stripped a
-time $\Delta t$ ago from a progenitor $(\vecx^p,\vecv^p) \equiv
-(\vecj^p,\veca^p)$ by receiving a small offset $(\Delta \vecx,\Delta
-\vecv) \equiv (\Delta \vecj,\Delta \veca)$. The offset then evolves
-using the dynamical equations given in the previous section and the
-star is observed at time $t=0$ (\ie, today). Because phase-space
-density is conserved by Liousville's theorem, a probabilistic model
-results by relating the likelihood $p(\vecx,\vecv|\vecp)$ (where
-$\vecp$ denotes the set of all model parameters; see below) for the
-observed phase-space position today to the distribution of initial
-offsets. This probabilistic model can be marginalized over missing
-observational data or extended to include a model for outliers.
-
-\subsection{Real--space modeling}
-
-In more detail, in the real--space approach the model can be specified
-by: (a) the gravitational potential $\Phi$, (b) the current
-phase-space position of the progenitor $(\vecx_p,\vecv_p)(t=0)$, which
-can be integrated backwards in time by reversing the Hamiltonian flow
-defined by $\Phi$, (c) the time $\Delta t_i$ at which each star was
-removed from the progenitor, and (d) the distribution $f(\Delta
-\vecx,\Delta \vecv|\paramsdiff)$, characterized by some parameters
-$\paramsdiff$, of initial phase-space offsets from which the actual
-offsets $(\Delta \vecx_i,\Delta \vecv_i)$ are drawn. In practice, $f$
-can be taken to be a Gaussian, characterized by a width in position
-$\sigma_x$ and in velocity $\sigma_v$, related to the tidal radius and
-escape velocity of the progenitor, respectively. However, more
-complicated models for the initial phase-space distribution of the
-debris can be used here as well.
-
-The probability for observing a star today at $(\vecx,\vecv)$ given
-items (a) through (d) is then
-\begin{align}\label{eq:realdft}
-  p(\vecx_i,\vecv_i | \Phi,\vecx_p,\vecv_p,\Delta t_i,\paramsdiff) 
-  & = p(\vecx_i(-\Delta t),\vecv_i(-\Delta t) | \Phi,\vecx_p,\vecv_p,\Delta t_i,\paramsdiff) \,,\nonumber\\
-  & = p([\vecx_i-\vecx_p](-\Delta t),[\vecv_i-\vecv_p](-\Delta t) | \Phi,\vecx_p,\vecv_p,\Delta t_i,\paramsdiff) \,\nonumber\\
-  & = p(\Delta\vecx_i(-\Delta t),\Delta\vecv_i(-\Delta t) | \Phi,\vecx_p,\vecv_p,\Delta t_i,\paramsdiff)\,,\\
-  & = f(\Delta\vecx_i(-\Delta t),\Delta\vecv_i(-\Delta t) | \Phi,\vecx_p,\vecv_p,\Delta t_i,\paramsdiff)\, ,\nonumber
-\end{align}
-where the first equality holds because the Hamiltonian flow is unique
-and conserves phase-space volume by Liousville's theorem and the
-second one because the Jacobian of the transformation
-$(\vecx_i,\vecv_i)(-\Delta t) \rightarrow (\vecx_i,\vecv_i)(-\Delta t)
-- (\vecx_p,\vecv_p)(-\Delta t)$ is one. The final probability can be
-evaluated using the parameters $\paramsdiff$ following (d).
-
-The time $\Delta t_i$ at which a star was removed is unobservable and
-the probability in \eqnname~(\ref{eq:realdft}) should therefore be
-integrated over this time using a prior $p(\Delta t_i|\ldots)$, where
-$\ldots$ indicates that this prior could depend on other factors such
-as the potential and the progenitor orbit\footnote{For example, one
-  could insist that stars are removed around pericenter
-  passages.}. For simplicity, we can assume a uniform prior for the
-times $\Delta t$. The probability for a single star then becomes
-\begin{align}\label{eq:realdf}
-  p(\vecx_i,\vecv_i | \Phi,\vecx_p,\vecv_p,\paramsdiff) & = \int \dd
-  \Delta t_i \,p(\vecx_i,\vecv_i | \Phi,\vecx_p,\vecv_p,\Delta t_i,
-  \paramsdiff)\,.
-\end{align}
-If the progenitor's current phase-space position is unknown, then this
-probability should be further marginalized over $(\vecx_p,\vecv_p)$;
-for multiple stars, this marginalization is done after multiplying the
-individual-star likelihoods in \eqnname~(\ref{eq:realdf}).
-
-In practice, the likelihood in \eqnname~(\ref{eq:realdf}) can be
-calculated as follows: for a given potential $\Phi$ and progenitor
-$(\vecx_p,\vecv_p)$, we (a) integrate the orbits of all stars
-$(\vecx_i,\vecv_i)$ and (b) that of the progenitor backward in time for
-many Gyr and record the phase-space positions over time on a grid $j$
-in $\Delta^j_i$; (c) we evaluate $f(\Delta \vecx_i,\Delta
-\vecv_i)(\Delta t^j_i)$ for all stars $i$ and times $j$; (d) $f(\Delta
-\vecx_i,\Delta \vecv_i)(\Delta t^j_i)$ is then summed over $j$ and the
-logarithm of that is summed over $i$. The most computationally
-expensive part of this calculation is the integration of the various
-orbits. The evaluation of the likelihood for different parameters
-\paramsdiff\ of $f$ can be efficiently computed by caching the orbit
-integrations of (a) and (b); different progenitor orbits
-$(\vecx_p,\vecv_p)$ can also make use of the stored orbits in (a).
-
-The approach outlined in this section is straightforward, but very
-computationally expensive, as it requires large numbers of orbit
-integrations to explore all plausible removal times $\Delta t$,
-progenitor orbits, and potentials. Moreover, when the data have large
-uncertainties, more orbit integrations are necessary to explore each
-data points uncertainty range. While it might be possible to speed up
-the computation through linear approximations for nearby orbits, the
-action--angle formalism allows for significant and elegant
-simplifications and I consider the real--space approach further.
 
@@
   \caption{Mock tidal stream simulated as described in
-    \sectionname~\ref{sec:sim} after $5.01\Gyr$ of evolution in
+    \sectionname~\ref{sec:sim} after $5.011\Gyr$ of evolution in
     Galactocentric $X$ and $Z$ coordinates. The grey dot shows the
@@
   \caption{Simulated tidal stream from \sectionname~\ref{sec:sim}
-    after $5.01\Gyr$ of evolution in action--frequency--angle
+    after $5.011\Gyr$ of evolution in action--frequency--angle
     coordinates. The top row displays projections of the actions
@@
   \includegraphics[width=\textwidth,clip=]{gd1_evol_ld.ps}
-  \caption{Simulated stream after $5.01\Gyr$ as a function of Galactic
+  \caption{Simulated stream after $5.011\Gyr$ as a function of Galactic
     longitude. The top panel shows the stream in Galactic latitude and

### 2013-12-27 py/illustrate_track.py
streams: llustrative figure of track determination

@@ -1,0 +1,129 @@
+import sys
+import numpy
+from galpy.util import bovy_plot, bovy_conversion
+from galpy import potential
+from galpy.orbit import Orbit
+from galpy.actionAngle_src.actionAngleIsochroneApprox\
+    import actionAngleIsochroneApprox
+from galpy.df_src.streamdf import streamdf
+from matplotlib import pyplot
+_NTRACKCHUNKS= 4
+_SIGV=0.365
+def illustrate_track(plotfilename1,plotfilename2,plotfilename3):
+    #Setup stream model
+    lp= potential.LogarithmicHaloPotential(q=0.9,normalize=1.)
+    aAI= actionAngleIsochroneApprox(b=0.8,pot=lp)
+    obs= numpy.array([1.56148083,0.35081535,-1.15481504,
+                      0.88719443,-0.47713334,0.12019596])
+    sdf= streamdf(_SIGV/220.,progenitor=Orbit(obs),pot=lp,aA=aAI,
+                  leading=True,nTrackChunks=_NTRACKCHUNKS,
+                  tdisrupt=4.5/bovy_conversion.time_in_Gyr(220.,8.))
+    #First calculate meanOmega and sigOmega
+    mOs= numpy.array([sdf.meanOmega(t,oned=True) for t in sdf._thetasTrack])
+    sOs= numpy.array([sdf.sigOmega(t) for t in sdf._thetasTrack])
+    mOs-= sdf._progenitor_Omega_along_dOmega
+    mOs*= -bovy_conversion.freq_in_Gyr(220.,8.)
+    sOs*= bovy_conversion.freq_in_Gyr(220.,8.)
+    progAngle= numpy.dot(sdf._progenitor_angle,sdf._dsigomeanProgDirection)
+    bovy_plot.bovy_print(fig_width=8.25,fig_height=3.5)
+    bovy_plot.bovy_plot(sdf._thetasTrack+progAngle,mOs,'ko',ms=8.,
+                        xlabel=r'$\Large|\Delta \mathbf{\theta}_\parallel\Large|$',
+                        ylabel=r'$\Large|\mathbf{\Omega}_\parallel\Large|\,(\mathrm{Gyr}^{-1})$',
+                        xrange=[-0.2-1.14,1.6-1.14],
+                        yrange=[22.05,22.55])
[…]

### 2013-12-27 tex/ms.tex
streams: tweak

@@
 along which a star moves away from the progenitor is related to the
-initial action offset by $\Delta \veco \approx (\partial^2 H /
-\partial \vecj \partial \vecj) \Delta \vecj$, in the limit of small
-action offsets, because $\veco = \partial H / \partial \vecj$. For an
-approximately one-dimensional stream to form, the Hessian matrix
-$(\partial^2 H / \partial \vecj \partial \vecj)\large|_{\vecj^p}$ has
-to be dominated by a single large eigenvalue; in this case the angle
-difference of stars stripped from the stream at any time will fall
-along approximately the same direction $\unitvec\veco$, and the stream
-will essentially be one dimensional. In detail, because the
-distribution of initial offsets $\Delta \vecj$ is not isotropic, the
-direction $\unitvec \veco$ is not just determined by the potential
-(through $H$) and the progenitor orbit, but also by the distribution
-of $\Delta \vecj$, which is set through a combination of the internal
-dynamics of the progenitor and the stripping process (see discussion
-in \citealt{Sanders13b}). For dynamical modeling we can be agnostic
-about the relative importance of all of these contributors to
-$\unitvec \veco$ by flexible modeling of the distribution of $\Delta
-\vecj$ or $\Delta \veco$.
+initial action offset by $\Delta \veco \approx (\partial \veco /
+\partial \vecj) \Delta \vecj = (\partial^2 H / \partial \vecj \partial
+\vecj) \Delta \vecj$, in the limit of small action offsets, because
+$\veco = \partial H / \partial \vecj$. For an approximately
+one-dimensional stream to form, the Hessian matrix $(\partial^2 H /
+\partial \vecj \partial \vecj)\large|_{\vecj^p}$ has to be dominated
+by a single large eigenvalue; in this case the angle difference of
+stars stripped from the stream at any time will fall along
+approximately the same direction $\unitvec\veco$, and the stream will
+essentially be one dimensional. In detail, because the distribution of
+initial offsets $\Delta \vecj$ is not isotropic, the direction
+$\unitvec \veco$ is not just determined by the potential (through $H$)
+and the progenitor orbit, but also by the distribution of $\Delta
+\vecj$, which is set through a combination of the internal dynamics of
+the progenitor and the stripping process (see discussion in
+\citealt{Sanders13b}). For dynamical modeling we can be agnostic about
+the relative importance of all of these contributors to $\unitvec
+\veco$ by flexible modeling of the distribution of $\Delta \vecj$ or
+$\Delta \veco$.
 
@@
 \figurename~\ref{fig:gd1_jt}). Then we propagate this Gaussian
-variance to frequency space using the Hessian $(\partial^2 H /
-\partial \vecj \partial \vecj)\large|_{\vecj^p}$\footnote{The Hessian
-  $(\partial^2 H / \partial \vecj \partial \vecj)\large|_{\vecj^p}$
-  can be computed using the method in \appendixname~\ref{sec:aa} by
-  computing $\partial (\vecj,\veca) / \partial (\vecx,\vecv)$ and
-  $\partial (\veco,\veca) / \partial (\vecx,\vecv)$ and forming
-  $\partial \veco / \partial \vecj = \left(\partial (\vecj,\veca) /
-  \partial (\vecx,\vecv)\right)^{-1}\,\left(\partial (\veco,\veca) /
-  \partial (\vecx,\vecv)\right)$.}, which gives the variance matrix
+variance to frequency space using the Hessian $(\partial \veco /
+\partial \vecj)\large|_{\vecj^p}$\footnote{The Hessian $(\partial
+  \veco / \partial \vecj)\large|_{\vecj^p}$ can be computed using the
+  method in \appendixname~\ref{sec:aa} by computing $\partial
+  (\vecj,\veca) / \partial (\vecx,\vecv)$ and $\partial (\veco,\veca)
+  / \partial (\vecx,\vecv)$ and forming $\partial \veco / \partial
+  \vecj = \left(\partial (\vecj,\veca) / \partial
+  (\vecx,\vecv)\right)^{-1}\,\left(\partial (\veco,\veca) / \partial
+  (\vecx,\vecv)\right)$.}, which gives the variance matrix
 $V_\veco$. The principal eigenvector of this matrix is the model

### 2013-12-27 py/illustrate_track.py
streams: tweak to latex

@@ -112,7 +112,7 @@ size=16.)
     bovy_plot._add_ticks()
     bovy_plot._add_axislabels(r'$\mathrm{Galactic\ longitude\, (deg)}$',
                               r'$\mathrm{distance\, (kpc)}$')
-    bovy_plot.bovy_text(165.,13.5,r"$\mathrm{finally, the\ interpolated\ track\ in}\ (\mathbf{x},\mathbf{v})\ \mathrm{is}"+'\n'+r"$\mathrm{converted\ to\ observable\ quantities\ (here}, l\ \mathrm{and}\ D)$",
+    bovy_plot.bovy_text(165.,13.5,r"$\mathrm{finally, the\ interpolated\ track\ in}\ (\mathbf{x},\mathbf{v})\ \mathrm{is}$"+'\n'+r"$\mathrm{converted\ to\ observable\ quantities\ (here}, l\ \mathrm{and}\ D)$",
                         size=16.)
     bovy_plot.bovy_plot([230.,sdf._interpolatedObsTrackLB[850,0]],
                         [13.25,sdf._interpolatedObsTrackLB[850,2]],

### 2013-12-27 py/illustrate_track.py
streams: plot tweaks

@@ -27,15 +27,19 @@ def illustrate_track(plotfilename1,plotfilename2,plotfilename3):
     progAngle= numpy.dot(sdf._progenitor_angle,sdf._dsigomeanProgDirection)
     bovy_plot.bovy_print(fig_width=8.25,fig_height=3.5)
     bovy_plot.bovy_plot(sdf._thetasTrack+progAngle,mOs,'ko',ms=8.,
-                        xlabel=r'$\Large|\Delta \mathbf{\theta}_\parallel\Large|$',
-                        ylabel=r'$\Large|\mathbf{\Omega}_\parallel\Large|\,(\mathrm{Gyr}^{-1})$',
+                        xlabel=r'$\theta_\parallel$',
+                        ylabel=r'$\Omega_\parallel\,(\mathrm{Gyr}^{-1})$',
                         xrange=[-0.2-1.14,1.6-1.14],
                         yrange=[22.05,22.55])
     bovy_plot.bovy_plot(sdf._thetasTrack+progAngle,mOs,'k-',lw=1.5,overplot=True)
     bovy_plot.bovy_plot(sdf._thetasTrack+progAngle,mOs+2*sOs,'ko',ms=6.,mfc='none',
-                        overplot=True)
+                        zorder=1,overplot=True)
     bovy_plot.bovy_plot(sdf._thetasTrack+progAngle,mOs-2*sOs,'ko',ms=6.,mfc='none',
-                        overplot=True)
+                        zorder=1,overplot=True)
+    bovy_plot.bovy_plot(sdf._thetasTrack+progAngle,mOs+2*sOs,'k-.',lw=1.5,
+                        zorder=0,overplot=True)
+    bovy_plot.bovy_plot(sdf._thetasTrack+progAngle,mOs-2*sOs,'k-.',lw=1.5,
+                        zorder=0,overplot=True)
     bovy_plot.bovy_plot(sdf._thetasTrack+progAngle,sdf._progenitor_Omega_along_dOmega*bovy_conversion.freq_in_Gyr(220.,8.)*numpy.ones(len(sdf._thetasTrack)),
                         'k--',lw=1.5,overplot=True)
[…]

### 2013-12-27 py/plot_stream.py
streams: also plot do*td

@@ -803,6 +803,11 @@ def plot_stream_aa(plotfilename):
         pyplot.fill_between(das,(dOs+2*sOs)*bovy_conversion.freq_in_Gyr(220.,8),
                             (dOs-2*sOs)*bovy_conversion.freq_in_Gyr(220.,8),
                             color='0.8',zorder=0)
+        #Also plot the apar at which t_d becomes important
+        pyplot.plot([sdf.meanOmega(0.01,oned=True)*sdf._tdisrupt,
+                     sdf.meanOmega(0.01,oned=True)*sdf._tdisrupt],
+                    [0.,0.3],
+                    'k--')                     
     elif includetrack and 'apartime' in plotfilename:
         das= numpy.linspace(0.01,1.3,101)
         mts= numpy.array([sdf.meantdAngle(da) for da in das])

### 2013-12-27 py/plot_stream.py
sreams: new estimate

@@ -822,8 +822,8 @@ def plot_stream_aa(plotfilename):
                             color='0.6',zorder=1)
     elif includetrack and 'aparaperp' in plotfilename:
         das= numpy.linspace(0.01,1.3,101)
-        sas= numpy.array([sdf.sigangledAngle(da) for da in das])
-        sass= numpy.array([sdf.sigangledAngle(da,simple=True) for da in das])
+        sas= numpy.array([sdf.sigangledAngle(da) for da in das])*numpy.sqrt(2.)
+        sass= numpy.array([sdf.sigangledAngle(da,simple=True) for da in das])*numpy.sqrt(2.)
         pyplot.fill_between(das,0.,
                             (2*sas),
                             color='0.8',zorder=0)

### 2013-12-27 tex/ms.tex
streams: add illustrative track figure
streams: section on stream track in aa

@@
 \newcommand{\az}{\ensuremath{\theta_Z}}
+\newcommand{\apar}{\ensuremath{\theta_\parallel}}
+\newcommand{\opar}{\ensuremath{\Omega_\parallel}}
+\newcommand{\aperp}{\ensuremath{\veca_\perp}}
+\newcommand{\aperpii}{\ensuremath{\theta_{\perp,ii}}}
+\newcommand{\operp}{\ensuremath{\veco_\perp}}
+\newcommand{\operpii}{\ensuremath{\Omega_{\perp,ii}}}
+
+\newcommand{\vlos}{\ensuremath{V_{\mathrm{los}}}}
+\newcommand{\pmll}{\ensuremath{\mu_l}}
+\newcommand{\pmbb}{\ensuremath{\mu_b}}
 
@@
 orbit in any static gravitational potential is presented.
-
-%BOVY: reference point
 
@@
 $0.18^\circ$. For a model with an isotropic action distribution, the
-misalignment would be $1.28^circ$.
+misalignment would be $1.28^\circ$.
 
@@
 \section{The track of a tidal stream}\label{sec:track}
+
+In this section, I discuss how to calculate the track of a model tidal
+stream in the generative model described in
+\sectionname~\ref{sec:generative}. The track consists of the mean
+location of the stream as a function of angle along the stream. I also
+describe how to estimate the dispersion of the stream along this
+track. In \sectionname~\ref{sec:trackaa}, I explain how to calculate
+the track of the stream in frequency--angle coordinates as well as how
+to estimate the spread around this track. In
+\sectionname~\ref{sec:trackxv} I discuss how to efficiently project
+the stream track and dispersion into position--velocity coordinates
+$(\vecx,\vecv)$ or observable coordinates $(l,b,D,\vlos,\pmll,\pmbb)$.
+
+\subsection{In frequency--angle coordinates}\label{sec:trackaa}
+
+Here, I discuss how to calculate the stream's properties as a function
+of the angle along the stream. In the generative model described
+above, the direction of the angle along the stream lies along the
+principal eigenvector of $V_\veco$, the variance matrix of the model
+$\Delta \veco$ distribution. We denote the angle in this direction as
+$\apar$ and the frequency in this direction $\opar$; the
+two-dimensional angle- and frequency-space perpendicular to this
+direction is denoted as $\aperp$ and $\operp$.
+
+First, we compute the distribution of $\Delta \opar$, the frequency
+offset between stream members and the progenitor, given the angle
+offset $\Delta \apar$ as follows
+\begin{equation}\label{eq:poparapargeneral}
+\begin{split}
+  p(\Delta \opar|\Delta \apar) & = \int \dd \Delta t\,p(\Delta \opar,\Delta t|\Delta \apar)\, \\
+  & = \int \dd \Delta t \,p(\Delta \apar | \Delta \opar,\Delta t)\,p(\Delta \opar,\Delta t)\,,\\
+  & \approx \int \dd \Delta t \,\delta(\Delta \apar - \Delta \opar\,\Delta t)\,p(\Delta \opar,\Delta t)\,,\\
+  & = \frac{1}{\left|\Delta \opar\right|}\,p(\Delta \opar,\frac{\Delta \apar}{\Delta \opar})\,,
+\end{split}
+\end{equation}
+where in the penultimate step we have approximated the initial angle
+distribution as a delta function (as the initial angle offset is small
+with respect to the final offset). This is a fully general
+expression. For the fiducial stream model, this can be simplified to
+\begin{equation}\label{eq:poparapar}
+\begin{split}
+  p(\Delta \opar|\Delta \apar) \propto &\ \frac{1}{\left|\Delta \opar\right|}\,p(\Delta \opar)\,,\qquad 0 < \frac{\Delta \apar}{\Delta \opar} < t_d\,.\\
+  & 0\,,\qquad \mathrm{otherwise}\,.
+\end{split}
+\end{equation}
+\equationname~(\ref{eq:poparapar}) shows why we chose to model
+$p(\Delta \opar)$ as a Gaussian multiplied with $|\Delta
+\opar|$. $p(\Delta \opar|\Delta \apar)$ is then a Gaussian with mean
+$\Delta \Omega^m$ and variance $\sigma^2_{\Omega,1}$, for $\Delta
+\opar > \Delta \apar / t_d$, and zero otherwise. The mean and variance
+of such a Gaussian is straightforward to calculate in terms of error
+functions. The top left panel of \figurename~\ref{fig:gd1_apar} shows
+the distribution of $|\Delta \opar|$ versus $|\Delta \apar|$ (absolute
+value in order to show the trailing and leading stream together) for
+the simulated stream, as well as the mean and dispersion calculated
+from the fiducial model. The dashed line shows $\Delta \apar = \Delta
+\Omega^m \,t_d$, approximately the angle where the finite disruption
+time starts to influence the mean $\Delta \opar$ of the stream. The
+finite disruption time, $\approx 4.5\Gyr$ in this case, of a tidal
+stream means that for stream members to have reached very large angle
+differences with respect to the progenitor, they must have been
+removed at large frequency differences. Around $\Delta \Omega^m
+\,t_d$, average stream stars, \ie, those removed with the average
+frequency offset, do not have a large enough offset to reach large
+angle offsets. Therefore, even though the average frequency offset
+does not change over time (see \figurename~\ref{fig:gd1_times}), the
+average frequency offset, and therefore the average orbit, does change
+with distance from the progenitor. It is in this sense that streams do
+not follow a single orbit. This is discussed further in BOVY:REF.
+
+In the fiducial model, the average $\Delta \operp$, and $\Delta
+\aperp$ as a function of $\Delta \apar$ are zero. Therefore, the
+average stream track as a function of $\Delta \apar$ is entirely
+specified by the mean $\Delta \opar(\Delta \apar)$ and zero offsets in
+the perpendicular direction. This track can be rotated into the
+$(R,\phi,Z)$ coordinate system using the eigenvectors of $V_\veco$.
+
+To estimate the spread around the stream track, we need to calculate
+the distributions $p(\Delta \operp|\Delta \apar)$ and $p(\Delta
+\aperp|\Delta \apar)$. The former is just the zero mean Gaussian with
+variance given by the projection of $V_\veco$ onto the direction
+perpendicular to the stream. The distribution $p(\Delta \aperp|\Delta
+\apar)$, however, does depend on $\Delta \apar$, because $\Delta
+\aperp = \Delta \aperp^{\mathrm{init}} + \Delta \operp \Delta t$ and
+the distribution of $\Delta t$ depends on $\Delta \apar$. Therefore,
+we first calculate $p(\Delta t | \Delta \apar)$.
+
+We can calculate $p(\Delta t | \Delta \apar)$ as follows
+\begin{equation}\label{eq:pdtapar}
+\begin{split}
+  p(\Delta t | \Delta \apar) & = \int \dd \Delta \opar\,p(\Delta t,\Delta \opar|\Delta \apar)\,,\\
+  & = \int \dd \Delta \opar\,p(\Delta t|\Delta \opar,\Delta \apar)\,p(\Delta \opar|\Delta \apar)\,,\\
+  & = p\left(\Delta \opar = \frac{\Delta \apar}{\Delta t}\right)\,\left|\frac{\Delta \apar}{(\Delta t)^2}\right|\,,
+\end{split}
+\end{equation}
+where in the last step we have again approximated the initial
+distribution of parallel angle offsets as a delta function. The first
+factor in this equation is given by the expression given in
+\equationname~(\ref{eq:poparapargeneral}) (in general) or
+\equationname~(\ref{eq:poparapar}) for the fiducial model. The lower
+right panel of \figurename~\ref{fig:gd1_apar} shows the distribution
+of stripping times for members of the simulated stream as well as the
+mean and dispersion calculated from
+\equationname~(\ref{eq:pdtapar}). Similar to the distribution of
+$\Delta \opar$ as a function of $\Delta \apar$, the distribution of
+$\Delta t$ has a kink at the angle offset that can only be reached by
+stars that have to have been stripped at large frequency offsets (that
+is, larger than average).
+
+We can now also calculate $p(\Delta \aperp|\Delta \apar)$
+\begin{equation}
+  p(\Delta \aperpii|\Delta \apar) = \int \dd \Delta t \,\dd \Delta
+  \operpii \,p(\Delta \aperpii|\Delta \operpii,\Delta t)\,p(\Delta \operpii|\Delta t,\Delta \apar)\,p(\Delta t|\Delta \apar)\,,
+\end{equation}
+where $ii=2,3$ for the first and second perpendicular direction. For
+the fiducial model, this simplifies to
+\begin{equation}\label{eq:paperpapar}
+  p(\Delta \aperpii|\Delta \apar) = \int \dd \Delta t \,\mathcal{N}\left(\Delta \aperpii|0,\sigma_{\theta}^2+(\Delta t\,\sigma_{\Omega,ii})^2\right)\,p(\Delta t|\Delta \apar)\,,
+\end{equation}
+where $\sigma_{\Omega,ii}$ is the model frequency dispersion in the
+perpendicular direction (corresponding to the middle and smallest
+eigenvalue of $V_\veco$). In the fiducial model, this distribution is
+the only one of those considered in this section that requires an
+explicit integration. However, we can approximate the distribution by
+using the mean $\Delta t(\Delta \apar)$ rather than integrating over
+$\Delta t$. The lower left panel of \figurename~\ref{fig:gd1_apar}
+shows the distribution of perpendicular angle offsets and the
+grayscale bands show the dispersion calculated using
+\equationname~(\ref{eq:paperpapar}); the white dashed line shows the
+approximate estimate obtained using the mean stripping time rather
+than integrating over it. This simple estimate agrees well with the
+exact calculation.
+
+We can then estimate the dispersion around the stream track, which is
+useful for getting a sense of the width of the model stream and to
+find appropriate integration intervals when marginalizing the stream
+PDF as discussed in \sectionname~\ref{sec:pdf}. At a given angle
+offset $\Delta \apar$ we calculate the dispersion in $\Delta \opar$
+from \equationname~(\ref{eq:poparapar}) and we substitute this
+appropriately in $V_\veco$. We can also calculate the dispersion in
+$\Delta \aperp$ using \equationname~(\ref{eq:paperpapar}). How to
+estimate the dispersion in $\Delta \apar$ near a given $\Delta \apar$
+is more difficult, and I simply use a dispersion of $1$, as the stream
+typically spreads out over $\approx 1$ radian. This choice could be
+improved, as it has the effect of overestimating the true dispersion
+(see \sectionname~\ref{sec:pdf} below). We could calculate the
+correlation between $\Delta \operpii$ and $\Delta aperpii$ using
+similar equations as those given in the previous paragraphs, but for a
+simple estimate we can approximate these as $0.5$; we set the
+correlation between $\Delta \opar$ and $\Delta \apar$ to zero. While
+these estimates are not perfect, they give a reasonable estimate of
+the dispersion along the stream track. This provides an adequate
+starting point for more precise calculations of the dispersion in
+\sectionname~\ref{sec:pdf}.
+
+\subsection{In position--velocity coordinates}\label{sec:trackxv}
+
+%BOVY: reference point
 
@@
 
+\begin{figure}
+  \includegraphics[width=0.8\textwidth,clip=]{track_aa.ps}\\
+  \includegraphics[width=0.8\textwidth,clip=]{track_xz.ps}\\
+  \includegraphics[width=0.8\textwidth,clip=]{track_ld.ps}
+  \caption{}\label{fig:track}
+%python illustrate_track.py ../tex/track_aa.ps ../tex/track_xz.ps ../tex/track_ld.ps
+\end{figure}
+
 \begin{figure}[t!]

### 2013-12-28 tex/ms.tex
streams: description of the stream track

@@
 \subsection{In position--velocity coordinates}\label{sec:trackxv}
+
+Having estimated the track of the stream and the dispersion around
+this in frequency--angle space, we can propagate the track to
+position--velocity coordinates by inverting the transformation
+$(\vecx,\vecv) \rightarrow (\veco,\veca)$. We do this by linearizing
+the transformation along the orbit of the progenitor in the vicinity
+of the stream and inverting this linearized transformation. That this
+works relies on the stream being cold for a few different
+reasons. First, the track of a cold stream, while in general offset
+from the orbit of the progenitor, does not stray too far from it, such
+that the transformation to frequency--angle coordinates can be
+linearized over the range of the stream--orbit offset. Second, cold
+streams are essentially one-dimensional, that is, they only spread
+significantly over a single direction. This means that non-linearity
+in the $(\vecx,\vecv) \rightarrow (\veco,\veca)$ transformation only
+affects a single direction, which roughly follows the orbit of the
+progenitor. Therefore, we can linearize the $(\vecx,\vecv) \rightarrow
+(\veco,\veca)$ transformation along a one-dimensional grid of points
+on the progenitor's orbit, rather than on a full six-dimensional
+grid. Third, the coldness of the stream means that any significant
+mass of the full stream PDF is close enough to the stream track that
+all frequency--angle calculations can be performed using the linear
+approximation along the progenitor orbit.
+
+\figurename~\ref{fig:track} illustrates how I linearize the
+$(\vecx,\vecv) \rightarrow (\veco,\veca)$ transformation along the
+progenitor orbit and use it to propagate the track of the stream to
+position--velocity coordinates and to observable quantities
+$(l,b,D,\vlos,\pmll,\pmbb)$. The top panel shows the orbit of the
+progenitor and the mean stream track in frequency--angle
+coordinates. Only the parallel direction is shown here; the mean
+stream track coincides with the progenitor orbit in the perpendicular
+direction. At a small number of points along the progenitor orbit,
+chosen here to span $1.5$ radians, we calculate the Jacobian $\partial
+(\veco,\veca) / \partial (\vecx,\vecv)$. In this figure this is done
+at four points, but for all other figures and calculations in this
+paper eleven points are used. Each Jacobian calculation requires seven
+frequency--angle calcuations, such that the total number of such
+computations is less than $100$ to adequately model the stream (each
+computation involves a single orbit integration, see
+\appendixname~\ref{sec:aa}).
+
+We then calculate the stream track in Galactocentric
+position--velocity coordinates at the parallel angles for which we
+have linearized the frequency--angle transformation. This is shown in
+the middle panel of \figurename~\ref{fig:track}. Similarly, we
+transform the approximated variance described at the end of the
+previous section to position--velocity coordinates under the linear
+approximation. We can then interpolate the track of the stream between
+this small number of points at which it is calculated; this gives the
+full track of the stream. In \sectionname~\ref{sec:pdf}, we require an
+estimate of the dispersion around the track at any angle along the
+stream (for calculating the marginalized PDF). We can obtain this from
+interpolating the estimated variances at the calculated track
+points. This interpolation can be performed practically as follows. We
+decompose the variance matrix at each calculated track point into its
+eigendecomposition and order the eigenvalues by size. We then
+interpolate the values of each of the six eigenvalues using spline
+interpolation. The direction of each eigenvector is interpolated using
+\emph{slerp} \citep{Shoemake85a}, which is a type of spherical linear
+interpolation. The variance matrix at interpolated track points can
+then be constructed by from the interpolated eigenvalues and
+eigenvectors. All other $(\vecx,\vecv) \leftrightarrow (\veco,\veca)$
+calculations can then be performed by using the closest interpolated
+track point in $(\vecx,\vecv)$ or $(\veco,\veca)$ and using the
+calculated Jacobian from the closest calculated track point.
+
+Finally, we can calculate the stream track in observable quantities
+$(l,b,D,\vlos,\pmll,\pmbb)$ by converting the interpolated track in
+$(\vecx,\vecv)$ to these observable coordinates. The dispersion at
+calculated track points is calculated from that in $(\veco,\veca)$
+using the appropriate Jacobians and the six-dimensional dispersion can
+again be interpolated using the eigendecomposition. This is
+illustrated in the bottom panel of \figurename~\ref{fig:track}.
+
+The average stream location computed in the manner described in this
+section is an approximate track. The mean stream location, for
+example, in distance from the Sun at a given Galactic longitude, can
+be exactly calculated by marginalizing over the full six-dimensional
+stream PDF in $(l,b,D,\vlos,\pmll,\pmbb)$ over the unobserved
+dimensions $(b,\vlos,\pmll,\pmbb)$. In general this will give a
+slightly different stream location, because the stream DF is not
+exactly Gaussian. However, it is demonstrated in
+\sectionname~\ref{sec:pdf}, where I calculate the full stream PDF,
+that the approximate mean stream location of this section is
+sufficiently close to the true average position for all practical
+purposes. The same should hold for any cold stream.
+
+\figurename~\ref{fig:gd1_xz} shows the simulated stream particles as
+well as the stream track calculated in this section in Galactocentric
+$X$ and $Z$ coordinates. It is clear that the average stream location
+tracks the position of the simulated stream. \figurename
+s~\ref{fig:gd1_lbd} and \ref{fig:gd1_lbv} show the same, but in
+observed coordinates $(l,b,D,\vlos,\pmll,\pmbb)$. These figures show
+that the simple fiducial stream model does a good job of predicting
+where the stream is located as a function of Galactic longitude.
 
@@
   Sanders,~J. \& Binney,~J. 2013b, \mnras, 433, 1826
+\bibitem[Shoemake(1985)]{Shoemake85}
+  Shoemake,~K. 1985, ACM SIGGRAPH Computer Graphics 3, 45
 \bibitem[Teuben(1995)]{Teuben95a}

### 2013-12-28 tex/ms.tex
streams: illustration caption

@@
   Sanders,~J. \& Binney,~J. 2013b, \mnras, 433, 1826
-\bibitem[Shoemake(1985)]{Shoemake85}
+\bibitem[Shoemake(1985)]{Shoemake85a}
   Shoemake,~K. 1985, ACM SIGGRAPH Computer Graphics 3, 45
@@
   \includegraphics[width=0.8\textwidth,clip=]{track_ld.ps}
-  \caption{}\label{fig:track}
+  \caption{Illustration of the determination of the average stream
+    track described in \sectionname~\ref{sec:track}. The small black
+    dots in each panel are the points along the progenitor track at
+    which the Jacobian $\partial (\veco,\veca) / \partial
+    (\vecx,\vecv)$ is calculated. The large black dots are the
+    coresponding points on the stream where the track is
+    calculated. The dashed line in each panel is the progenitor's
+    orbit, while the solid line and dash-dotted lines are the mean
+    track and $2\sigma$ dispersion.}\label{fig:track}
 %python illustrate_track.py ../tex/track_aa.ps ../tex/track_xz.ps ../tex/track_ld.ps

### 2013-12-28 py/illustrate_track.py
streams: punctuation

@@ -43,15 +43,15 @@ def illustrate_track(plotfilename1,plotfilename2,plotfilename3):
     bovy_plot.bovy_plot(sdf._thetasTrack+progAngle,sdf._progenitor_Omega_along_dOmega*bovy_conversion.freq_in_Gyr(220.,8.)*numpy.ones(len(sdf._thetasTrack)),
                         'k--',lw=1.5,overplot=True)
     bovy_plot.bovy_plot(sdf._thetasTrack+progAngle,sdf._progenitor_Omega_along_dOmega*bovy_conversion.freq_in_Gyr(220.,8.)*numpy.ones(len(sdf._thetasTrack)),
-                        'ko',ms=8.,overplot=True)
+                        'ko',ms=6.,overplot=True)
     bovy_plot.bovy_text(1.05+progAngle,22.475,r'$\mathrm{progenitor\ orbit}$',size=16.)
     bovy_plot.bovy_text(progAngle+0.05,22.50,r'$\mathrm{current\ progenitor\ position}$',size=16.)
     bovy_plot.bovy_plot([progAngle+0.05,progAngle],[22.50,sdf._progenitor_Omega_along_dOmega*bovy_conversion.freq_in_Gyr(220.,8.)],'k:',overplot=True)
-    bovy_plot.bovy_text(-0.6,22.33,r'$\mathrm{at\ a\ small\ number\ of\ points\ along\ the}$'+'\n'+r'$\mathrm{progenitor\ orbit, we\ calculate}\ \frac{\partial(\mathbf{\Omega},\boldsymbol{\theta})}{\partial (\mathbf{x},\mathbf{v})}$',size=16.)
+    bovy_plot.bovy_text(-0.6,22.33,r'$\mathrm{At\ a\ small\ number\ of\ points\ along\ the}$'+'\n'+r'$\mathrm{progenitor\ orbit, we\ calculate}\ \frac{\partial(\mathbf{\Omega},\boldsymbol{\theta})}{\partial (\mathbf{x},\mathbf{v})}.$',size=16.)
     bovy_plot.bovy_plot([progAngle+sdf._thetasTrack[1],-0.6],
                         [sdf._progenitor_Omega_along_dOmega*bovy_conversion.freq_in_Gyr(220.,8.),22.38],
                         'k:',overplot=True)
[…]

### 2013-12-28 tex/ms.tex
streams: re-arrange figures
streams: tweak
streams: spelling

@@
 
+\begin{figure}[t!]
+  \includegraphics[width=0.5\textwidth,clip=]{gd1_evol_xz.ps}
+  \caption{Mock tidal stream simulated as described in
+    \sectionname~\ref{sec:sim} after $5.011\Gyr$ of evolution in
+    Galactocentric $X$ and $Z$ coordinates. The grey dot shows the
+    current position of the progenitor cluster and the dashed line
+    shows its orbit. The solid line gives the stream track, calculated
+    using the method described in \sectionname~\ref{sec:track}. The
+    inset shows a zoomed-in version of part of the leading stream that
+    more clearly shows the offset between the stream track and the
+    progenitor orbit.}\label{fig:gd1_xz}
+%python plot_stream.py ../tex/gd1_evol_xz.ps
+\end{figure}
+
+\begin{figure}[t!]
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aajrjz.ps}
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaoroz.ps}
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaaraz.ps}\\
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aajrjp.ps}
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaorop.ps}
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaarap.ps}
+  \caption{Simulated tidal stream from \sectionname~\ref{sec:sim}
+    after $5.011\Gyr$ of evolution in action--frequency--angle
+    coordinates. The top row displays projections of the actions
+    (left), frequencies (middle), and angles (right) in the $R$ and
+    $Z$ directions; the bottom panels shows the same for $R$ and
+    $\phi$. While the action distribution has a quite complicated
+    form, the frequency distribution is close to one-dimensional and
+    the leading and trailing parts of the stream clearly separate. In
+    the right panels, the angles of the progenitor have each been
+    shifted to $\pi$ and the dashed lines show the direction of the
+    progenitor's frequency vector. The misalignment between the
+    progenitor's frequency vector and the mean frequency vector of the
+    debris is $0.18^\circ$.}\label{fig:gd1_jt}
+%python plot_stream.py ../tex/gd1_evol_aajrjz.ps
+%python plot_stream.py ../tex/gd1_evol_aaoroz.ps
+%python plot_stream.py ../tex/gd1_evol_aaaraz.ps
+%python plot_stream.py ../tex/gd1_evol_aajrjp.ps
+%python plot_stream.py ../tex/gd1_evol_aaorop.ps
+%python plot_stream.py ../tex/gd1_evol_aaarap.ps
+\end{figure}
+
 The final position of the stream in $(X,Z)$ is shown in
@@
 isotropic Gaussian distribution.
+
+\begin{figure}[t!]
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_timeshist.ps}
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_timesdO.ps}
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_timesda.ps}
+  \caption{Properties of the simulated tidal stream of
+    \sectionname~\ref{sec:sim} as a function of time. Stripping times
+    are calculated by linear regression of the angle difference
+    between a stream particle and the progenitor versus the equivalent
+    frequency difference. The distribution of stripping times is shown
+    in the left panel.  While the stripping process happens in bursts
+    with a period of roughly $400\Myr$ (the radial period of the
+    progenitor), it can be approximated as constant (more
+    coarsely-binned solid histogram). The magnitude of the frequency
+    offset between stream particles and the progenitor is shown in the
+    middle panel. While particles stripped at pericenter have larger
+    frequency offsets than those between pericenter passages, there is
+    no secular trend and the typical frequency offset (averaged over a
+    radial period) can be considered constant with time. The right
+    panel shows the magnitude of the angle offset between stream
+    particles and the progenitor as a function of stripping
+    time. There is a spread in the angle offset at all stripping times
+    because of the dispersion in frequency offsets, but typically
+    stars that have been stripped earlier reach larger angle
+    offsets. That is, self-sorting happens only to a small
+    degree.}\label{fig:gd1_times}
+%python plot_stream.py ../tex/gd1_evol_timeshist.ps
+%python plot_stream.py ../tex/gd1_evol_timesdO.ps
+%python plot_stream.py ../tex/gd1_evol_timesda.ps
+\end{figure}
 
@@
 
+\begin{figure}[t!]
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aadohist.ps}
+  \caption{Distribution of the magnitude of frequency offsets between
+    stream particles and the progenitor along the direction of the
+    mean frequency offset. The dashed line is a Gaussian fit to this
+    distribution, while the solid line is a fit of our preferred form
+    for this distribution, given in \equationname~(\ref{eq:do}). Both
+    forms provide adequate fits to the
+    distribution. }\label{fig:gd1_dohist}
+%python plot_stream.py ../tex/gd1_evol_aadohist.ps
+\end{figure}
+
 We then further constrain the mean frequency offset to lie along the
@@
 
+\begin{figure}[t!]
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaaparopar.ps}
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaaparoperp.ps}\\
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaaparaperp.ps}
+  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaapartime.ps}
+  \caption{Stream properties as a function of angle along the
+    stream. The top, left panel shows the magnitude of frequency
+    offsets along the stream. The gray line shows the mean predicted
+    frequency offset calculated using the framework of
+    \sectionname\sectionname~\ref{sec:modeloa} and \ref{sec:track},
+    while the grayscale bands show the $68\,\%$ and $95\,\%$ limits of
+    the predicted distribution. The top, right panel shows the
+    magnitude of frequency offsets perpendicular to those in the left
+    panel. The predicted distribution is constant as a function of
+    angle and therefore not shown. The bottom, left panel shows the
+    angle offset perpendicular to the stream and the grayscale bands
+    show $68\,\%$ and $95\,\%$ limits of the predicted distribution of
+    angle offsets; the white dashed line shows the $68\,\%$ limit
+    calculated using a simplified method. The bottom, right panel
+    shows the stripping time. The gray line and grayscale bands again
+    show the mean and $68\,\%$ and $95\,\%$ confidence limits of the
+    predicted distribution. While the ``burstiness'' related to the
+    energetic stripping at pericenter that is apparent in the
+    simulated data's distributions is (by design) absent in the model,
+    the overall scale and the mean trends with angle are well
+    represented by the model.}\label{fig:gd1_apar}
+%python plot_stream.py ../tex/gd1_evol_aaaparopar.ps
+%python plot_stream.py ../tex/gd1_evol_aaaparoperp.ps
+%python plot_stream.py ../tex/gd1_evol_aaaparaperp.ps
+%python plot_stream.py ../tex/gd1_evol_aaapartime.ps
+\end{figure}
+
 Here, I discuss how to calculate the stream's properties as a function
@@
 \end{equation}
-\equationname~(\ref{eq:poparapar}) shows why we chose to model
+\Eqnname~(\ref{eq:poparapar}) shows why we chose to model
 $p(\Delta \opar)$ as a Gaussian multiplied with $|\Delta
@@
 approximation along the progenitor orbit.
+
+\begin{figure}
+  \includegraphics[width=0.8\textwidth,clip=]{track_aa.ps}\\
+  \includegraphics[width=0.8\textwidth,clip=]{track_xz.ps}\\
+  \includegraphics[width=0.8\textwidth,clip=]{track_ld.ps}
+  \caption{Illustration of the determination of the average stream
+    track described in \sectionname~\ref{sec:track}. The small black
+    dots in each panel are the points along the progenitor track at
+    which the Jacobian $\partial (\veco,\veca) / \partial
+    (\vecx,\vecv)$ is calculated. The large black dots are the
+    corresponding points on the stream where the track is
+    calculated. The dashed line in each panel is the progenitor's
+    orbit, while the solid line and dash-dotted lines are the mean
+    track and $2\sigma$ dispersion.}\label{fig:track}
+%python illustrate_track.py ../tex/track_aa.ps ../tex/track_xz.ps ../tex/track_ld.ps
+\end{figure}
 
@@
 
+\begin{figure}[t!]
+  \includegraphics[width=\textwidth,clip=]{gd1_evol_lb.ps}
+  \includegraphics[width=\textwidth,clip=]{gd1_evol_ld.ps}
+  \caption{Simulated stream after $5.011\Gyr$ as a function of Galactic
+    longitude. The top panel shows the stream in Galactic latitude and
+    the bottom panel shows the distance from the Sun. The progenitor
+    is shown as the large gray dot and its orbit as the dashed
+    lines. The predicted track of the stream is shown as the solid
+    line. Insets show various zoomed-in parts of the stream to show
+    the stream, progenitor orbit, and predicted stream track up
+    close.}\label{fig:gd1_lbd}
+%python plot_stream.py ../tex/gd1_evol_lb.ps
+%python plot_stream.py ../tex/gd1_evol_ld.ps
+\end{figure}
+
+\begin{figure}[tp!!]
+  \includegraphics[width=\textwidth,clip=]{gd1_evol_lvlos.ps}
+  \includegraphics[width=\textwidth,clip=]{gd1_evol_lpmll.ps}
+  \includegraphics[width=\textwidth,clip=]{gd1_evol_lpmbb.ps}
+  \caption{Simulated stream velocities as a function of Galactic
+    longitude. Same as \figurename~\ref{fig:gd1_lbd}, but for the
+    stream in heliocentric line-of-sight velocity $V_{\mathrm{los}}$
+    (top panel), proper motion in the direction of Galactic longitude
+    (middle panel), and in the direction of Galactic latitude (bottom
+    panel). Insets again show various interesting parts of the stream
+    up close.}\label{fig:gd1_lbv}
+%python plot_stream.py ../tex/gd1_evol_lvlos.ps
+%python plot_stream.py ../tex/gd1_evol_lpmll.ps
+%python plot_stream.py ../tex/gd1_evol_lpmbb.ps
+\end{figure}
+
 The average stream location computed in the manner described in this
@@
 \section{Mock stream data}\label{sec:mock}
+
+\begin{figure}[t!]
+ \includegraphics[width=0.48\textwidth,clip=]{gd1_evol_2Gya_xz.ps}
+ \includegraphics[width=0.48\textwidth,clip=]{gd1_evol_2Gya_xz_sim.ps}
+  \caption{Mock stream data generated from the model. The left panel
+    shows the same $N$-body stream as shown in previous figures, but
+    $1.96\Gyr$ earlier in its evolution. Symbols, lines, and the
+    stream model are the same as those in
+    \figurename~\ref{fig:gd1_xz}, except that the model progenitor
+    phase-space location from \figurename~\ref{fig:gd1_xz} has been
+    integrated backward in time for $1.96\Gyr$ (no other changes were
+    made to the model's parameters). The right panel shows mock stream
+    data drawn from the model using the procedures discussed in
+    \sectionname~\ref{sec:mock}. Overall, the spatial distribution of
+    the mock data is very similar to that of the $N$-body
+    stream.}\label{fig:gd1_sim}
+%python plot_stream_2Gya.py ../tex/gd1_evol_2Gya_xz.ps
+%python plot_stream_2Gya.py ../tex/gd1_evol_2Gya_xz_sim.ps
+\end{figure}
 
@@
 \section{An efficient, general method for calculating action-angle coordinates using orbit integration}\label{sec:aa}
+
+\begin{figure}[t!]
+ \includegraphics[width=0.48\textwidth,clip=]{aAI_jr.ps}
+ \includegraphics[width=0.48\textwidth,clip=]{aAI_araz.ps}
+  \caption{Action--angle calculations using an orbit-integration-based
+    method. The right panel shows the cumulative mean of the radial
+    actions calculated in an auxiliary isochrone potential calculated
+    along the orbit of the progenitor used in the main body of this
+    paper. The color-coding corresponds to the radial angle
+    (calculated in the isochrone potential) along the orbit. The time
+    is in units of $35.56\Myr$ and the radial action is normalized by
+    its value at $t=200$. The left panel shows the radial and vertical
+    angles calculated in the auxiliary isochrone potential, with
+    color-coding corresponding to the azimuthal angle. The wiggles in
+    the angles are due to the isochrone potential not being the
+    correct potential, but these wiggles are fit and removed to
+    calculate the true angles.}\label{fig:aAI}
+%python plot_aAI.py ../tex/aAI_jr.ps
+%python plot_aAI.py ../tex/aAI_araz.ps
+\end{figure}
 
@@
 
-\begin{figure}[t!]
-  \includegraphics[width=0.5\textwidth,clip=]{gd1_evol_xz.ps}
-  \caption{Mock tidal stream simulated as described in
-    \sectionname~\ref{sec:sim} after $5.011\Gyr$ of evolution in
-    Galactocentric $X$ and $Z$ coordinates. The grey dot shows the
-    current position of the progenitor cluster and the dashed line
-    shows its orbit. The solid line gives the stream track, calculated
-    using the method described in \sectionname~\ref{sec:track}. The
-    inset shows a zoomed-in version of part of the leading stream that
-    more clearly shows the offset between the stream track and the
-    progenitor orbit.}\label{fig:gd1_xz}
-%python plot_stream.py ../tex/gd1_evol_xz.ps
-\end{figure}
-
-\begin{figure}[t!]
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aajrjz.ps}
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaoroz.ps}
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaaraz.ps}\\
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aajrjp.ps}
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaorop.ps}
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaarap.ps}
-  \caption{Simulated tidal stream from \sectionname~\ref{sec:sim}
-    after $5.011\Gyr$ of evolution in action--frequency--angle
-    coordinates. The top row displays projections of the actions
-    (left), frequencies (middle), and angles (right) in the $R$ and
-    $Z$ directions; the bottom panels shows the same for $R$ and
-    $\phi$. While the action distribution has a quite complicated
-    form, the frequency distribution is close to one-dimensional and
-    the leading and trailing parts of the stream clearly separate. In
-    the right panels, the angles of the progenitor have each been
-    shifted to $\pi$ and the dashed lines show the direction of the
-    progenitor's frequency vector. The misalignment between the
-    progenitor's frequency vector and the mean frequency vector of the
-    debris is $0.18^\circ$.}\label{fig:gd1_jt}
-%python plot_stream.py ../tex/gd1_evol_aajrjz.ps
-%python plot_stream.py ../tex/gd1_evol_aaoroz.ps
-%python plot_stream.py ../tex/gd1_evol_aaaraz.ps
-%python plot_stream.py ../tex/gd1_evol_aajrjp.ps
-%python plot_stream.py ../tex/gd1_evol_aaorop.ps
-%python plot_stream.py ../tex/gd1_evol_aaarap.ps
-\end{figure}
-
-\begin{figure}[t!]
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_timeshist.ps}
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_timesdO.ps}
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_timesda.ps}
-  \caption{Properties of the simulated tidal stream of
-    \sectionname~\ref{sec:sim} as a function of time. Stripping times
-    are calculated by linear regression of the angle difference
-    between a stream particle and the progenitor versus the equivalent
-    frequency difference. The distribution of stripping times is shown
-    in the left panel.  While the stripping process happens in bursts
-    with a period of roughly $400\Myr$ (the radial period of the
-    progenitor), it can be approximated as constant (more
-    coarsely-binned solid histogram). The magnitude of the frequency
-    offset between stream particles and the progenitor is shown in the
-    middle panel. While particles stripped at pericenter have larger
-    frequency offsets than those between pericenter passages, there is
-    no secular trend and the typical frequency offset (averaged over a
-    radial period) can be considered constant with time. The right
-    panel shows the magnitude of the angle offset between stream
-    particles and the progenitor as a function of stripping
-    time. There is a spread in the angle offset at all stripping times
-    because of the dispersion in frequency offsets, but typically
-    stars that have been stripped earlier reach larger angle
-    offsets. That is, self-sorting happens only to a small
-    degree.}\label{fig:gd1_times}
-%python plot_stream.py ../tex/gd1_evol_timeshist.ps
-%python plot_stream.py ../tex/gd1_evol_timesdO.ps
-%python plot_stream.py ../tex/gd1_evol_timesda.ps
-\end{figure}
-
-\begin{figure}[t!]
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aadohist.ps}
-  \caption{Distribution of the magnitude of frequency offsets between
-    stream particles and the progenitor along the direction of the
-    mean frequency offset. The dashed line is a Gaussian fit to this
-    distribution, while the solid line is a fit of our preferred form
-    for this distribution, given in \equationname~(\ref{eq:do}). Both
-    forms provide adequate fits to the
-    distribution. }\label{fig:gd1_dohist}
-%python plot_stream.py ../tex/gd1_evol_aadohist.ps
-\end{figure}
-
-\begin{figure}[t!]
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaaparopar.ps}
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaaparoperp.ps}\\
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaaparaperp.ps}
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaapartime.ps}
-  \caption{Stream properties as a function of angle along the
-    stream. The top, left panel shows the magnitude of frequency
-    offsets along the stream. The gray line shows the mean predicted
-    frequency offset calculated using the framework of
-    \sectionname\sectionname~\ref{sec:modeloa} and \ref{sec:track},
-    while the grayscale bands show the $68\,\%$ and $95\,\%$ limits of
-    the predicted distribution. The top, right panel shows the
-    magnitude of frequency offsets perpendicular to those in the left
-    panel. The predicted distribution is constant as a function of
-    angle and therefore not shown. The bottom, left panel shows the
-    angle offset perpendicular to the stream and the grayscale bands
-    show $68\,\%$ and $95\,\%$ limits of the predicted distribution of
-    angle offsets; the white dashed line shows the $68\,\%$ limit
-    calculated using a simplified method. The bottom, right panel
-    shows the stripping time. The gray line and grayscale bands again
-    show the mean and $68\,\%$ and $95\,\%$ confidence limits of the
-    predicted distribution. While the ``burstiness'' related to the
-    energetic stripping at pericenter that is apparent in the
-    simulated data's distributions is (by design) absent in the model,
-    the overall scale and the mean trends with angle are well
-    represented by the model.}\label{fig:gd1_apar}
-%python plot_stream.py ../tex/gd1_evol_aaaparopar.ps
-%python plot_stream.py ../tex/gd1_evol_aaaparoperp.ps
-%python plot_stream.py ../tex/gd1_evol_aaaparaperp.ps
-%python plot_stream.py ../tex/gd1_evol_aaapartime.ps
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=0.8\textwidth,clip=]{track_aa.ps}\\
-  \includegraphics[width=0.8\textwidth,clip=]{track_xz.ps}\\
-  \includegraphics[width=0.8\textwidth,clip=]{track_ld.ps}
-  \caption{Illustration of the determination of the average stream
-    track described in \sectionname~\ref{sec:track}. The small black
-    dots in each panel are the points along the progenitor track at
-    which the Jacobian $\partial (\veco,\veca) / \partial
-    (\vecx,\vecv)$ is calculated. The large black dots are the
-    coresponding points on the stream where the track is
-    calculated. The dashed line in each panel is the progenitor's
-    orbit, while the solid line and dash-dotted lines are the mean
-    track and $2\sigma$ dispersion.}\label{fig:track}
-%python illustrate_track.py ../tex/track_aa.ps ../tex/track_xz.ps ../tex/track_ld.ps
-\end{figure}
-
-\begin{figure}[t!]
-  \includegraphics[width=\textwidth,clip=]{gd1_evol_lb.ps}
-  \includegraphics[width=\textwidth,clip=]{gd1_evol_ld.ps}
-  \caption{Simulated stream after $5.011\Gyr$ as a function of Galactic
-    longitude. The top panel shows the stream in Galactic latitude and
-    the bottom panel shows the distance from the Sun. The progenitor
-    is shown as the large gray dot and its orbit as the dashed
-    lines. The predicted track of the stream is shown as the solid
-    line. Insets show various zoomed-in parts of the stream to show
-    the stream, progenitor orbit, and predicted stream track up
-    close.}\label{fig:gd1_lbd}
-%python plot_stream.py ../tex/gd1_evol_lb.ps
-%python plot_stream.py ../tex/gd1_evol_ld.ps
-\end{figure}
-\begin{figure}[t!]
-  \includegraphics[width=\textwidth,clip=]{gd1_evol_lvlos.ps}
-  \includegraphics[width=\textwidth,clip=]{gd1_evol_lpmll.ps}
-  \includegraphics[width=\textwidth,clip=]{gd1_evol_lpmbb.ps}
-  \caption{Simulated stream velocities as a function of Galactic
-    longitude. Same as \figurename~\ref{fig:gd1_lbd}, but for the
-    stream in heliocentric line-of-sight velocity $V_{\mathrm{los}}$
-    (top panel), proper motion in the direction of Galactic longitude
-    (middle panel), and in the direction of Galactic latitude (bottom
-    panel). Insets again show various interesting parts of the stream
-    up close.}\label{fig:gd1_lbv}
-%python plot_stream.py ../tex/gd1_evol_lvlos.ps
-%python plot_stream.py ../tex/gd1_evol_lpmll.ps
-%python plot_stream.py ../tex/gd1_evol_lpmbb.ps
-\end{figure}
-
-\begin{figure}[t!]
- \includegraphics[width=0.48\textwidth,clip=]{gd1_evol_2Gya_xz.ps}
- \includegraphics[width=0.48\textwidth,clip=]{gd1_evol_2Gya_xz_sim.ps}
-  \caption{Mock stream data generated from the model. The left panel
-    shows the same $N$-body stream as shown in previous figures, but
-    $1.96\Gyr$ earlier in its evolution. Symbols, lines, and the
-    stream model are the same as those in
-    \figurename~\ref{fig:gd1_xz}, except that the model progenitor
-    phase-space location from \figurename~\ref{fig:gd1_xz} has been
-    integrated backward in time for $1.96\Gyr$ (no other changes were
-    made to the model's parameters). The right panel shows mock stream
-    data drawn from the model using the procedures discussed in
-    \sectionname~\ref{sec:mock}. Overall, the spatial distribution of
-    the mock data is very similar to that of the $N$-body
-    stream.}\label{fig:gd1_sim}
-%python plot_stream_2Gya.py ../tex/gd1_evol_2Gya_xz.ps
-%python plot_stream_2Gya.py ../tex/gd1_evol_2Gya_xz_sim.ps
-\end{figure}
-
-\begin{figure}[t!]
- \includegraphics[width=0.48\textwidth,clip=]{aAI_jr.ps}
- \includegraphics[width=0.48\textwidth,clip=]{aAI_araz.ps}
-  \caption{Action--angle calculations using an orbit-integration-based
-    method. The right panel shows the cumulative mean of the radial
-    actions calculated in an auxiliary isochrone potential calculated
-    along the orbit of the progenitor used in the main body of this
-    paper. The color-coding corresponds to the radial angle
-    (calculated in the isochrone potential) along the orbit. The time
-    is in units of $35.56\Myr$ and the radial action is normalized by
-    its value at $t=200$. The left panel shows the radial and vertical
-    angles calculated in the auxiliary isochrone potential, with
-    color-coding corresponding to the azimuthal angle. The wiggles in
-    the angles are due to the isochrone potential not being the
-    correct potential, but these wiggles are fit and removed to
-    calculate the true angles.}\label{fig:aAI}
-%python plot_aAI.py ../tex/aAI_jr.ps
-%python plot_aAI.py ../tex/aAI_araz.ps
-\end{figure}
-
 \end{document}

### 2013-12-30 tex/ms.tex
streams: mock section

@@
 
-\subsection{The fiducial stream model}
+\subsection{The fiducial stream model}\label{sec:fidmodel}
 
@@
 
-%BOVY: reference point
-
 \section{Mock stream data}\label{sec:mock}
@@
     phase-space location from \figurename~\ref{fig:gd1_xz} has been
-    integrated backward in time for $1.96\Gyr$ (no other changes were
+    integrated backward in time for $1.96\Gyr$ and the model
+    disruption time was reduced by $1.96\Gyr$ (no other changes were
     made to the model's parameters). The right panel shows mock stream
@@
     the mock data is very similar to that of the $N$-body
-    stream.}\label{fig:gd1_sim}
+    stream.}\label{fig:gd1_mock}
 %python plot_stream_2Gya.py ../tex/gd1_evol_2Gya_xz.ps
@@
 \end{figure}
+
+Using the generative model described in
+\sectionname~\ref{sec:modeloa}, it is straightforward how to draw mock
+stream data in frequency--angle space. The three ingredients---time,
+frequency, and angle distributions of the debris---allow for mock data
+to be generated in four steps. First, a stripping time $t_s$ is
+sampled from the distribution of stripping times. Second, a frequency
+offset $\Delta \veco$ with respect to the progenitor is drawn from the
+frequency distribution at $t_s$. Third, an initial angle offset is
+drawn. Fourth, the initial angle offset is incremented by $\Delta
+\veco \,ts$. This generates the final frequency--angle coordinates of
+the mock stream member $(\Delta \veco,\Delta \veca+\Delta \veco\,t_s)$.
+
+For the fiducial stream model of \sectionname~\ref{sec:fidmodel} the
+procedure for producing mock data is simplified to sampling a small
+number of easy-to-sample distributions. First, $t_s$ is drawn
+uniformly between $0$ and $t_d$. Second, $\Delta \operp$ is drawn from
+the Gaussian distribution of $\Delta \operp$. The distribution of
+$\Delta \opar$ is not Gaussian, as it is a Gaussian in $|\Delta
+\opar|$ multiplied with $|\Delta \opar|$ over the range $\Delta \opar
+> 0$. This distribution is log-concave, that is, the derivative of its
+logarithm is negative everywhere, and therefore it can be efficienly
+sampled using adaptive-rejection sampling \citep{Gilks92a}. The sign
+of $\Delta \opar$ is determined by whether we are generating a leading
+or trailing stream. Third, initial angle offsets are drawn from the
+Gaussian distribution of such offsets.
+
+To transform the mock stream in frequency--angle coordinates to
+position--velocity space, we use the approximate procedure for this
+transformation along the stream track described at the end of the
+previous section. The mock stream in $(\vecx,\vecv)$ can then be
+further transformed to observable coordinates
+$(l,b,D,\vlos,\pmll,\pmbb)$ or any other similar quantities as
+desired.
+
+\figurename~\ref{fig:gd1_mock} shows mock data generated for the model
+of the simulated stream described in
+\sectionname~\ref{sec:fidmodel}. The mock data in this Figure has been
+generated $1.96\Gyr$ before the final snapshot of the simulation. The
+model from \sectionname~\ref{sec:fidmodel}, which was for the final
+snapshot, was adjusted to this earlier time by backward orbit
+integration of the model progenitor and by revising the disruption
+time downward by $1.96\Gyr$. No other changes were made to the
+model. The right panel of \figurename~\ref{fig:gd1_mock} shows the
+mock data generated from the model in $(X,Z)$, while the left panel
+shows the simulated stream data at this time. Overall, the
+distribution of the mock and simulated data are similar.
+
+%BOVY: reference point
 
@@
   Eyre,~A. \& Binney,~J. 2011, \mnras, 413, 1852
+\bibitem[Gilks \& Wild(1992)]{Gilks92a}
+  Gilks,~W.~R. \& Wild,~P. 1992, Applied Statistics, 41, 337
 \bibitem[Helmi \& White(1999)]{Helmi99a}

### 2013-12-30 tex/ms.tex
streams: cut-out version of l figures

@@
 \begin{figure}[t!]
-  \includegraphics[width=\textwidth,clip=]{gd1_evol_lb.ps}
-  \includegraphics[width=\textwidth,clip=]{gd1_evol_ld.ps}
+  \includegraphics[width=0.8\textwidth,clip=]{gd1_evol_lb.eps}
+  \includegraphics[width=0.8\textwidth,clip=]{gd1_evol_ld.eps}
   \caption{Simulated stream after $5.011\Gyr$ as a function of Galactic
@@
 \begin{figure}[tp!!]
-  \includegraphics[width=\textwidth,clip=]{gd1_evol_lvlos.ps}
-  \includegraphics[width=\textwidth,clip=]{gd1_evol_lpmll.ps}
-  \includegraphics[width=\textwidth,clip=]{gd1_evol_lpmbb.ps}
+  \includegraphics[width=0.8\textwidth,clip=]{gd1_evol_lvlos.eps}
+  \includegraphics[width=0.8\textwidth,clip=]{gd1_evol_lpmll.eps}
+  \includegraphics[width=0.8\textwidth,clip=]{gd1_evol_lpmbb.eps}
   \caption{Simulated stream velocities as a function of Galactic

### 2013-12-30 tex/ms.tex
streams: also cut out the illustration
streams: tweaks to the mock section

@@
 \begin{figure}
-  \includegraphics[width=0.8\textwidth,clip=]{track_aa.ps}\\
-  \includegraphics[width=0.8\textwidth,clip=]{track_xz.ps}\\
-  \includegraphics[width=0.8\textwidth,clip=]{track_ld.ps}
+  \includegraphics[width=0.8\textwidth,clip=]{track_aa.eps}\\
+  \includegraphics[width=0.8\textwidth,clip=]{track_xz.eps}\\
+  \includegraphics[width=0.8\textwidth,clip=]{track_ld.eps}
   \caption{Illustration of the determination of the average stream
@@
 drawn. Fourth, the initial angle offset is incremented by $\Delta
-\veco \,ts$. This generates the final frequency--angle coordinates of
-the mock stream member $(\Delta \veco,\Delta \veca+\Delta \veco\,t_s)$.
+\veco \,t_s$. This generates the final frequency--angle coordinates of
+the mock stream member $(\Delta \veco,\Delta \veca+\Delta \veco\,t_s)
++ (\veco_p,\veca_p)$.
 
@@
 or trailing stream. Third, initial angle offsets are drawn from the
-Gaussian distribution of such offsets.
+Gaussian distribution of such offsets. The fourth step is as above.
 
@@
 \sectionname~\ref{sec:fidmodel}. The mock data in this Figure has been
-generated $1.96\Gyr$ before the final snapshot of the simulation. The
-model from \sectionname~\ref{sec:fidmodel}, which was for the final
+generated $1.96\Gyr$ before the final snapshot of the simulation (that
+is, $1.96\Gyr$ before \figurename~\ref{fig:gd1_xz}). The model
+from \sectionname~\ref{sec:fidmodel}, which was for the final
 snapshot, was adjusted to this earlier time by backward orbit

### 2013-12-30 tex/ms.tex
streams: start on PDF section

@@
 
+\section{The full stream PDF}\label{sec:pdf}
+
+Data on tidal streams typically come in one of two flavors: data on
+individual stream members, often with multiple missing phase--space
+components, or data describing the mean position and width of the
+stream. For the proper analysis of these kinds of data it is useful to
+be able to evaluate the PDF for the stream model described in this
+paper and to marginalize over or condition it on certain dimensions.
+
+The generative model described in \sectionname~\ref{sec:modeloa}
+corresponds to the stream PDF in frequency $\veco$, angle $\veca$, and
+stripping time $t_s$
+\begin{equation}
+  p(\veco,\veca,t_s) = p(t_s)\,p(\veco|t_s)\,p(\veca|\veco,t_s)\,,
+\end{equation}
+where the three factors on the right are specified through the three
+ingredients of \sectionname~\ref{sec:modeloa}. The PDF in
+position--velocity space is then given by
+\begin{equation}
+  p(\vecx,\vecv,t_s) = p(\veco,\veca,t_s)\,\left|\frac{\partial
+    \veco}{\partial \vecj}\right|\,.
+\end{equation}
+The PDF in observable quantities $(l,b,D,\vlos,\pmll,\pmbb)$ can be
+obtained from $p(\vecx,\vecv,t_s)$ by multiplying by the Jacobian
+$\left|\partial(\vecx,\vecv)/\partial(l,b,D,\vlos,\pmll,\pmbb)\right|$.
+
+The stripping time $t_s$ cannot be observed and should therefore by
+marginalized over when evaluating the PDF. One of the great advantages
+of modeling a stream in frequency--angle coordinates is that this
+marginalization can be done analytically for many choices of $p(t_s)$
+if $p(\veco|t_s) \equiv p(\veco)$. Specifically, for the fiducial
+model $p(t_s)$ is uniform up to a maximum disruption time $t_d$ (see
+\equationname~[\ref{eq:pt}]) such that we can write
+\begin{equation}
+\begin{split}
+  p(\veco,\veca) & = \int \dd t_s\,p(\veco,\veca,t_s)\,,\\
+  & = p(\veco)\,\int_0^{t_d} \dd t_s\,
+  \frac{1}{(2\,\pi)^{3/2}\,\sigma_\theta^3}\,\exp\left(-\frac{1}{2\,\sigma_\theta^2}\left[\sum_{R,\phi,Z}\left(\Delta \veca_i-\Delta \veco_i\,t_s\right)^2\right]\right)\,,
+\end{split}
+\end{equation}
+where the sum in the argument of the exponential is over the three
+components of $\Delta \veca_i$ and $\Delta \veco_i$. This integral can
+be done analytically, resulting in
+\begin{equation}\label{eq:pdf}
+\begin{split}
+  p(\veco,\veca) & = 
+  p(\veco) \,  \frac{\mathrm{erf}(a_0) + \mathrm{erf}(a_d)}{4\,\pi\,\sigma_\theta^2\,|\Delta \veco|}\,\exp\left(-\frac{1}{2\,\sigma_\theta^2}\left[\sum_{R,\phi,Z}\Delta \veca^2_i-\frac{\left(\sum_{R,\phi,Z} \Delta \veco_i\Delta \veca_i\right)^2}{|\Delta \veco|}\right]\right)\,.
+\end{split}
+\end{equation}
+In this expression, $\Delta X^2 \equiv (X_1-X_2)^2$. The expression in
+square brackets is positive by using the Cauchy-Schwarz
+inequality. The quantities $a_0$ and $a_d$ in this expression are
+defined by the following expressions
+\begin{align}
+  \tilde{t}_s & = \frac{\sum \Delta \veco_i\Delta \veca_i}{|\Delta \veco_i|^2}\,,\\
+  a_0 & = \frac{|\Delta \veco_i|}{\sqrt{2}\,\sigma_\theta}\,\tilde{t}_s\,,\\
+  a_d & = \frac{|\Delta \veco_i|}{\sqrt{2}\,\sigma_\theta}\,\left(t_d - \tilde{t}_s\right)\,.
+\end{align}
+The time $\tilde{t}_s$ is the best estimate of the stripping time for
+a given $\Delta \veco$ and $\Delta \veca$. The quantity $a_0$ then
+serves to suppress the PDF in case the stripping time is smaller than
+zero, \ie, a stream member appears to have been stripping \emph{in the
+  future}; in this case $\mathrm{erf}(a_0) \approx -1$ and the PDF
+goes to zero. Similarly, $a_d$ suppresses the PDF when the stripping
+time is larger than $t_d$, \ie, the stream member seems to have been
+stripped before disruption began. Both of these get a tolerance
+corresponding to the initial angle spread. For example, if
+$\tilde{t}_s$ is negative, but so small such that $|\Delta
+\veco|\,\tilde{t}_s \sim \sigma_\theta$, the PDF is only mildly
+suppressed. For stars well within the stream $\mathrm{erf}(a_0) +
+\mathrm{erf}(a_d) \approx 2$.
+
+The probability in \eqnname~(\ref{eq:pdf}) incorporates and elucidates
+commonly used methods for fitting tidal streams. On the one hand are
+approaches that search to minimize the spread in energy, integrals of
+the motion, or actions (\eg, \citealt{Binney08a,Penarrubia12a},
+Sanderson \etal, 2013, in preparation). These approaches ignore the
+correlations between actions and angles in the stream (the exponential
+in \eqnname~(\ref{eq:pdf}), which describes this correlation through
+the frequencies) and only use $p(\veco)$; finding the gravitational
+potential that minimizes the spread in frequencies (or equivalently,
+actions) optimizes $p(\veco)$. Recently, a new method was proposed
+that only uses the correlation between the actions and the angles,
+expressed by the exponential in \eqnname~(\ref{eq:pdf})
+\citep{Sanders13b}; this method does not use the fact that the spread
+in actions in a tidal stream is small (expressed by $p(\veco)$), but
+only uses the fact that the angles and actions/frequencies are highly
+correlated. Approaches that require stream members to re-unite with
+the progenitor when integrating their orbits backward in time
+\citep{Johnston99a,PriceWhelan13a}) use the full PDF (but expressed in
+position--velocity space), but cannot be marginalized over unobserved
+or noisy phase--space dimensions as easily.
+
 %BOVY: reference point
-
-\section{The full stream PDF}\label{sec:pdf}
 
@@
 that of the actual progenitor.
-
-The probability in \eqnname~(\ref{eq:aadfsimple}) incorporates and
-elucidates commonly used methods for fitting tidal streams. On the one
-hand are approaches that search to minimize the spread in energy,
-integrals of the motion, or actions (\eg,
-\citealt{Binney08a,Penarrubia12a}, Sanderson \etal, 2013, in
-preparation). These approaches ignore the correlations between actions
-and angles in the stream (the exponential in
-\eqnname~(\ref{eq:aadfsimple}), which describes this correlation
-through the frequencies) and only use $f_{\vecj}(\Delta
-\vecj_i|\paramsdiff)$; finding the gravitational potential that
-minimizes the spread in actions optimizes $f_{\vecj}(\Delta
-\vecj_i|\paramsdiff)$. Recently, a method was proposed that only uses
-the correlation between the actions and the angles, expressed by the
-exponential in \eqnname~(\ref{eq:aadfsimple}) \citep{Sanders13b}; this
-method does not use the fact that the spread in actions in a tidal
-stream is small (expressed by $f_{\vecj}(\Delta
-\vecj_i|\paramsdiff)$).
 

### 2013-12-30 tex/ms.tex
streams: rm some old material

@@
 
-\subsection{Action--angle stream modeling}
-
-Similar as for the real--space approach, we can write down a
-generative model for the positions and velocities ($\vecx_i,\vecv_i$)
-of stars $i$ in a tidal stream in the action--angle formalism by
-specifying: (a) the gravitational potential $\Phi$, (b) the current
-action-angle coordinates of the progenitor $(\vecj_p,\veca_p)(t=0)$,
-(c) the time $\Delta t_i$ at which each star was removed from the
-progenitor, and (d) the distribution $f(\Delta \vecj,\Delta
-\veca|\paramsdiff)$, characterized by some parameters $\paramsdiff$,
-of initial phase-space offsets from which the actual offsets $(\Delta
-\vecj_i,\Delta \veca_i)$ are drawn. In practice, $f$ can again be
-taken to be a Gaussian, although more general models could take the
-expected `bowtie' structure of the action distribution into account
-(\citealt{Eyre11a}; see also \figurename~\ref{fig:gd1_jt}).
-
-The probability for observing a star today at $(\vecx,\vecv)$ given
-items (a) through (d) is then
-\begin{align}\label{eq:aadft}
-  p(\vecx_i,\vecv_i | \Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff) 
-  & = p(\vecj_i,\veca_i | \Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff) \,,\nonumber\\
-  & = p(\vecj_i-\vecj_p,\veca_i-\veca_p-\Delta \veco_i\Delta t_i | \Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff) \,,\\
-  & = f(\Delta \vecj_i,\Delta \veca_i(-\Delta t_i) | \Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff) \,,\nonumber
-\end{align}
-where the first equality holds because the Jacobian of the
-$(\vecx,\vecv)\rightarrow(\vecj,\veca)$ transformation is unity and
-the second equality is a simple linear transformation. The final
-probability can again be evaluated using the parameters
-\paramsdiff\ following (d).
-
-As in the real--space formalism, we should integrate
-\eqnname~(\ref{eq:aadft}) over the time $\Delta t_i$ at which the star
-is removed from the progenitor (cf.~\eqnname~[\ref{eq:realdf}]):
-\begin{equation}\label{eq:aadf}
-  p(\vecx_i,\vecv_i | \Phi,\vecj_p,\veca_p,\paramsdiff) = \int \dd
-  \Delta t_i\, f(\Delta \vecj_i,\Delta \veca_i(-\Delta t_i) |
-  \Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff) \,.
-\end{equation}
-Without further assumptions the action-angle formalism proceeds in the
-same way as the real--space formalism, evaluating
-\eqnname~(\ref{eq:aadft}) for a grid $j$ of times $\Delta t_i^j$ and
-summing $f(\Delta \vecj_i,\Delta \veca_i(-\Delta t^j_i) |
-\Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff)$ over $j$.
-
-Even without further assumptions about the shape of $f(\Delta
-\vecj,\Delta \veca|\paramsdiff)$, the action-angle formalism already
-allows for a simplification over the real--space method. Because the
-progenitor orbit is specified in action-angle coordinates, it can be
-changed \emph{without} expensive orbit integrations of all possible
-source orbits. As the progenitor's current phase-space position is a
-six-dimensional parameter, this is a major simplification of the
-dynamical model when the source orbit is unknown. Note that this
-simplification does require us to be able to compute $\veco(\vecj)$,
-to calculate the new source frequencies $\veco_p$ when changing the
-source actions $\vecj_p$. This mapping can be constructed using a
-small number of (action,frequency) calculations covering the range of
-possible source actions and is relatively easy to construct due to the
-approximately one-dimensional structure of a stream (see discussion at
-the end of \sectionname~\ref{sec:dynamics}).
-
-We can simplify \eqnname~(\ref{eq:aadf}) by assuming (a) that
-$f(\Delta \vecj,\Delta \veca|\paramsdiff)$ is separable into
-$f_{\veca}(\Delta \veca|\paramsdiff)\,f_{\vecj}(\Delta
-\vecj|\paramsdiff)$ and (b) that the distribution $f_{\veca}(\Delta
-\veca|\paramsdiff)$ is Gaussian. In what follows, we will assume that
-this Gaussian is isotropic with dispersion $\sigma_\theta$, but all of
-the results are the same after appropriate diagonalization and
-rescaling of the angles for more general covariance matrices. In this
-case, \eqnname~(\ref{eq:aadf}) can be written as
-\begin{equation}
-\begin{split}
-  p(\vecx_i,\vecv_i & | \Phi,\vecj_p,\veca_p,\paramsdiff) 
-   = \\
-   & f_{\vecj}(\Delta \vecj_i|\paramsdiff) \int \dd \Delta t_i\,
-  \frac{1}{(2\,\pi)^{3/2}\,\sigma_\theta^3}\,\exp\left(-\frac{1}{2\,\sigma_\theta^2}\left[\sum_{R,\phi,Z}\left(\Delta \veca_i-\Delta \veco_i\Delta t_i\right)^2\right]\right)\,,
-\end{split}
-\end{equation}
-where the sum in the argument of the exponential is over the three
-components of $\Delta \veca_i$ and $\Delta \veco_i$. This integral can
-be done analytically, resulting in
-\begin{equation}\label{eq:aadfsimple}
-\begin{split}
-  p(\vecx_i,\vecv_i & | \Phi,\vecj_p,\veca_p,\paramsdiff) 
-   = \\
-   & f_{\vecj}(\Delta \vecj_i|\paramsdiff) \,  \frac{1}{2\,\pi\,\sigma_\theta^2\,\sqrt{\sum_{R,\phi,Z}\Delta \veco_i^2}}\,\exp\left(-\frac{1}{2\,\sigma_\theta^2}\left[\sum_{R,\phi,Z}\Delta \veca^2_i-\frac{\left(\sum_{R,\phi,Z} \Delta \veco_i\Delta \veca_i\right)^2}{\sum_{R,\phi,Z}\Delta \veco_i^2}\right]\right)\,.
-\end{split}
-\end{equation}
-In this expression, $\Delta X^2 \equiv (X_1-X_2)^2$. The expression in
-square brackets is positive by using the Cauchy-Schwarz inequality.
-
-The quantity in the square brackets in \eqnname~(\ref{eq:aadfsimple})
-expresses the fact that the present-day difference in angles between a
-star in a tidal stream and the source must be linearly related by a
-scalar (the time since the removal of the star) to the difference in
-frequencies, within a tolerance set by the width of the distribution
-of initial angle differences ($\sigma_\theta$). The ratio
-$\sum_{R,\phi,Z} \Delta \veco_i\Delta \veca_i/\sum_{R,\phi,Z}\Delta
-\veco_i^2$ is the best-guess for $\Delta t_i$ when performing a linear
-regression between $\Delta \veca_i$ and $\Delta \veco_i$. The
-expression in square brackets is minimized when the correlation
-between $\Delta \veca_i$ and $\Delta \veco_i$ is the largest, \ie,
-when a single time difference $\Delta t_i$ explains all three of the
-angle differences.
-
-A detail that is ignored in \equationname~(\ref{eq:aadfsimple}) is
-that the uniform prior on $\Delta t$ should only have support for
-$\Delta t > 0$, \ie, for times in the past, which introduces an
-additional factor of $[1+\mathrm{erf}(a_0)]$, with
-\[ a_0 = \frac{\sqrt{\sum_{R,\phi,Z}\Delta
-  \veco_i^2}}{\sqrt{2}\,\sigma_\theta}\,\frac{\sum_{R,\phi,Z} \Delta
-  \veco_i\Delta \veca_i}{\sum_{R,\phi,Z}\Delta \veco_i^2}\,.\] Because
-the second factor on the right-hand side is the best guess for $\Delta
-t_i$, if this is large and negative (\ie, the star was stripped
-\emph{in the future}), the probability of the star being stripped from
-the progenitor is strongly suppressed. Only when it appears that the
-star was only recently stripped with $\Delta t_i$ negative, but
-$\approx \sigma_\theta/\sqrt{\sum_{R,\phi,Z}\Delta \veco_i^2}$, is the
-suppression for negative stripping times less.
-
-The discussion so far has been entirely general and applies to any
-sort of stellar debris that originates from a single orbit, without
-necessarily generating a one-dimensional structure. In most
-observations of tidal streams, however, they appear as one-dimensional
-objects rather than the six-dimensional objects that we have
-considered so far. The reason for this is both physical and
-observational: (a) narrow tidal streams form because a star separates
-from the progenitor over time mostly along a single direction in
-angle-space: the Hessian matrix of the Hamiltonian, $\partial^2 H /
-\partial \vecj \partial \vecj$, has a single large eigenvalue, which
-determines the direction along which the stream is disrupted in
-angle-space (because $\partial \veco / \partial \vecj$ is equal to the
-Hessian); (b) because many tidal streams are dynamically cold with
-velocity dispersions $\lesssim10\kms$, observations are unable to
-resolve the angle differences along the two eigenvectors of the
-Hessian corresponding to the two smallest eigenvalues or the spread in
-actions among stream members. If the latter can be resolved, the
-stream is four-dimensional. In what follows we will be mainly
-concerned with the stream only spreading along a single direction in
-angle space.
-
-If the stream is one-dimensional in angle space then the time since
-the removal from the progenitor becomes unobservable if the progenitor
-orbit is unknown: above, determining the time depended on explaining
-the three-dimensional angle difference $\Delta \veca$ with the
-three-dimensional frequency difference $\Delta \veco$ using a single
-time difference; if two of the angle and frequency differences are
-unable to be resolved, then the frequency difference and the time
-difference become degenerate (aka, you can draw a straight line
-connecting any two points). To make use of this fact to simplify the
-dynamical model, we write the frequency difference as
-\begin{equation}
-  \Delta \veco = \Delta \Omega_1\,\unitvec{e}_1\,,
-\end{equation}
-where $\unitvec{e}_1$ is a unit vector. Then we write the angle
-difference $\Delta \veca$ 
-\begin{equation}
-  \Delta \veca = \Delta \theta_1\,\unitvec{e}_1 + \Delta \veca_\perp\,\unitvec{e}_1^\perp\,,
-\end{equation}
-where$\unitvec{e}_1^\perp$ is the orthogonal complement of
-$\unitvec{e}_1$ (the two directions perpendicular to $\unitvec{e}_1$)
-and $\Delta \veca_\perp$ are the components of $\Delta \veca$ in that
-direction. We can then write \eqnname~(\ref{eq:aadfsimple}) as
-\begin{equation}\label{eq:aadfsimpler}
-\begin{split}
-  p(\vecx_i,\vecv_i & | \Phi,\vecj_p,\veca_p,\paramsdiff) 
-   = \\
-   & f_{\vecj}(\Delta \vecj_i|\paramsdiff) \,  \frac{1}{2\,\pi\,\sigma_\theta^2\,\sqrt{\sum_{R,\phi,Z}\Delta \veco_i^2}}\,\exp\left(-\frac{|\Delta\veca_{i,\perp}|^2}{2\,\sigma_\theta^2}\right)\,.
-\end{split}
-\end{equation}
-\Eqnname~(\ref{eq:aadfsimpler}) is equivalent to
-\eqnname~(\ref{eq:aadfsimple}), but it expresses the dynamical model
-in a slightly different way. Whereas the angle part of
-\eqnname~(\ref{eq:aadfsimple}) allows the correct potential and
-progenitor to be inferred by asking whether a single time difference
-explains the three angle differences given the three frequencies,
-\eqnname~(\ref{eq:aadfsimpler}) `rewinds' the star's orbit along the
-direction of $\Delta \veco$ and asks whether the angle difference
-perpendicular to $\Delta \veco$ is consistent with the initial angle
-spread (because it has not evolved since the removal of the star). In
-re-writing this, we have not yet used the approximation that the
-stream only forms along a single direction in $\Delta \veco$.
-
-\section{Two equivalent generative models of tidal streams}\label{sec:method}
-
-In this Section, we describe two mathematically equivalent generative
-models for a tidal stream. The first is based on orbit integrations in
-real (position-velocity) space: a tidal stream is generated from small
-(phase-space) offsets between a star and a source\footnote{The word
-  `progenitor' is typically used to indicate the actual parent body of
-  a star in a tidal stream. For the dynamical model only a reference
-  orbit from which the tidal stream can be thought of to originate is
-  necessary; this reference orbit does \emph{not} have to coincide
-  with the actual progenitor (this point is most clear in the
-  description of the dynamics in action-angle coordinates). To avoid
-  confusion between the hypothetical progenitor and the actual parent
-  body, we refer to the reference orbit as the `source' rather than
-  the `progenitor'.} that are evolved in time using orbit
-integrations. The second method relies on the equivalent description
-of this process in action-angle coordinates. While these two methods
-are equivalent, the latter provides a conceptually simpler description
-of the dynamical model and, when fitting a tidal stream, it allows for
-fast marginalization over nuisance parameters. In particular, the
-marginalization over the time at which a star was removed from the
-source is analytic and the integration over possible source orbits can
-be performed much faster (by avoiding orbit integrations of all
-possible source orbits).
-
-\subsection{Real--space model}\label{sec:realmethod}
-
-The description of the stripping of a star from a satellite in real
-(\vecx,\vecv) space assumes that at a time $\Delta t$ ago the star was
-offset by a small amount ($\Delta \vecx,\Delta \vecv$) from the
-position of the source $(\vecx^s,\vecv^s)(-\Delta t)$ at that
-time. Then the star orbited under the influence of the same
-gravitational potential $\Phi$ as the source until the present day
-$(t=0)$, when it was observed: $(\vecx,\vecv)(t=0)$ = $\vech_{\Delta
-  t}(\vecx^s(-\Delta t)+\Delta \vecx,\vecv^s(-\Delta t)+\Delta
-\vecv)$, where $\vech$ denotes the Hamiltonian flow from $t=-\Delta t$
-to $t=0$ (see \citealt{binneytremaine}); this is simply the orbit of
-the star between these times. In phase-space this orbit can be
-calculated by solving Hamilton's equations: $\dot{\vecx} = \vecv;
-\dot{\vecv} = - \dd \Phi / \dd \vecx$.
-
-A generative model for the positions and velocities
-($\vecx_i,\vecv_i$) of stars $i$ in a tidal stream is then fully
-specified by: (a) the gravitational potential $\Phi$, (b) the current
-phase-space position of the source $(\vecx_p,\vecv_p)(t=0)$, which can
-be integrated backwards in time by reversing the Hamiltonian flow
-defined by $\Phi$, (c) the time $\Delta t_i$ at which each star was
-removed from the source, and (d) the distribution $f(\Delta
-\vecx,\Delta \vecv|\paramsdiff)$, characterized by some parameters
-$\paramsdiff$, of initial phase-space offsets from which the actual
-offsets $(\Delta \vecx_i,\Delta \vecv_i)$ are drawn. In practice, $f$
-can be taken to be a Gaussian, characterized by a width in position
-$\sigma_x$ and in velocity $\sigma_v$, corresponding to the tidal
-radius and escape velocity of the progenitor, respectively. However,
-more complicated models for the initial phase-space distribution of
-the debris can be used here as well (see discussion below). In the
-case of a Gaussian $f$, the source can be thought of as the mean of
-this Gaussian, with the additional requirement that this mean is an
-orbit in the potential $\Phi$.
-
-The probability for observing a star today at $(\vecx,\vecv)$ given
-items (a) through (d) is then
-\begin{align}\label{eq:realdft}
-  p(\vecx_i,\vecv_i | \Phi,\vecx_p,\vecv_p,\Delta t_i,\paramsdiff) 
-  & = p(\vecx_i(-\Delta t),\vecv_i(-\Delta t) | \Phi,\vecx_p,\vecv_p,\Delta t_i,\paramsdiff) \,,\nonumber\\
-  & = p([\vecx_i-\vecx_p](-\Delta t),[\vecv_i-\vecv_p](-\Delta t) | \Phi,\vecx_p,\vecv_p,\Delta t_i,\paramsdiff) \,\nonumber\\
-  & = p(\Delta\vecx_i(-\Delta t),\Delta\vecv_i(-\Delta t) | \Phi,\vecx_p,\vecv_p,\Delta t_i,\paramsdiff)\,,\\
-  & = f(\Delta\vecx_i(-\Delta t),\Delta\vecv_i(-\Delta t) | \Phi,\vecx_p,\vecv_p,\Delta t_i,\paramsdiff)\, ,\nonumber
-\end{align}
-where the first equality holds because the Hamiltonian flow is unique
-and conserves phase-space volume by Liousville's theorem and the
-second one because the Jacobian of the transformation
-$(\vecx_i,\vecv_i)(-\Delta t) \rightarrow (\vecx_i,\vecv_i)(-\Delta t)
-- (\vecx_p,\vecv_p)(-\Delta t)$ is one. The final probability can be
-evaluated using the parameters $\paramsdiff$ following (d).
-
-The time $\Delta t_i$ at which a star was removed is unobservable and
-the probability in \eqnname~(\ref{eq:realdft}) should therefore be
-integrated over this time using a prior $p(\Delta t_i|\ldots)$, where
-$\ldots$ indicates that this prior could depend on other factors such
-as the potential and the progenitor orbit\footnote{For example, one
-  could insist that stars are removed around pericenter passages. Note
-  that this assumption would require the source orbit to be the actual
-  progenitor orbit.}. For simplicity, we will assume a uniform prior
-for the times $\Delta t$. The probability for a single star then becomes
-\begin{align}\label{eq:realdf}
-  p(\vecx_i,\vecv_i | \Phi,\vecx_p,\vecv_p,\paramsdiff) & = \int \dd
-  \Delta t_i \,p(\vecx_i,\vecv_i | \Phi,\vecx_p,\vecv_p,\Delta t_i,
-  \paramsdiff)\,.
-\end{align}
-If the progenitor's current phase-space position is unknown, then this
-probability should be further marginalized over $(\vecx_p,\vecv_p)$;
-for multiple stars, this marginalization is done after multiplying the
-individual-star likelihoods in \eqnname~(\ref{eq:realdf}).
-
-In practice, the likelihood in \eqnname~(\ref{eq:realdf}) can be
-calculated as follows: for a given potential $\Phi$ and source
-$(\vecx_p,\vecv_p)$, we (a) integrate the orbits of all stars
-$(\vecx_i,\vecv_i)$ and (b) that of the source backward in time for
-many Gyr and record the phase-space positions over time on a grid $j$
-in $\Delta^j_i$; (c) we evaluate $f(\Delta \vecx_i,\Delta
-\vecv_i)(\Delta t^j_i)$ for all stars $i$ and times $j$; (d) $f(\Delta
-\vecx_i,\Delta \vecv_i)(\Delta t^j_i)$ is then summed over $j$ and the
-logarithm of that is summed over $i$. The most computationally
-expensive part of this calculation is the integration of the various
-orbits. The evaluation of the likelihood for different parameters
-\paramsdiff\ of $f$ can be efficiently computed by caching the orbit
-integrations of (a) and (b); different source orbits
-$(\vecx_p,\vecv_p)$ can also make use of the stored orbits in (a).
-
-\subsection{Action-angle-space model}\label{sec:aamethod}
-
-An alternative description of orbits in a gravitational potential
-$\Phi$ is provided by the action-angle formalism (see
-\citealt{binneytremaine} for an extensive description). The
-Hamiltonian flow in the action-angle formalism is trivial: actions are
-conserved and angles increase linearly with time. In
-\appendixname~\ref{sec:aa} we summarize a method to calculate
-action-angle coordinates from an orbit integration with little extra
-work in any (static) potential for any kind of orbit. Therefore, it is
-practical to use action-angle coordinates to describe the dynamical
-model for a tidal stream in these coordinates and make use of the many
-conceptual advantages provided by this formalism.
-
-In the action-angle formalism, a star is removed from the source with
-a small offset $(\Delta \vecj,\Delta \veca)(-\Delta t)$ from the
-source's actions $\vecj_p$ and angles $\veca_p$ a time $\Delta t$ ago,
-similar to the small offset $(\Delta \vecx,\Delta \vecv)(-\Delta t)$
-in the real--space model. The offset in the actions corresponds to an
-offset in the frequencies $\Delta \veco$; ultimately, this offset is
-responsible for the spreading of the stream over long stretches on the
-sky. Because both the source's angles and the star's angles increase
-linearly with time, albeit at different frequencies, the difference in
-the angles also increases linearly in time. Therefore at time $t=0$
-the difference in angles is
-\begin{equation}
-\Delta \veca(t=0) = \Delta \veca(-\Delta t) + \Delta \veco \Delta t\,,
-\end{equation}
-while the difference in actions is constant
-\begin{equation}
-\Delta \vecj(t=0) = \Delta \vecj(-\Delta t)\,.
-\end{equation}
-
-With these preliminaries, we can write down a generative model for the
-positions and velocities ($\vecx_i,\vecv_i$) of stars $i$ in a tidal
-stream in the action-angle formalism by specifying: (a) the
-gravitational potential $\Phi$, (b) the current action-angle
-coordinates of the source $(\vecj_p,\veca_p)(t=0)$, (c) the time
-$\Delta t_i$ at which each star was removed from the source, and (d)
-the distribution $f(\Delta \vecj,\Delta \veca|\paramsdiff)$,
-characterized by some parameters $\paramsdiff$, of initial phase-space
-offsets from which the actual offsets $(\Delta \vecj_i,\Delta
-\veca_i)$ are drawn. In practice, $f$ can again be taken to be a
-Gaussian, although more general models could take the expected
-`bowtie' structure of the action distribution into account
-\citep{Eyre11a}.
-
-The probability for observing a star today at $(\vecx,\vecv)$ given
-items (a) through (d) is then
-\begin{align}\label{eq:aadft}
-  p(\vecx_i,\vecv_i | \Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff) 
-  & = p(\vecj_i,\veca_i | \Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff) \,,\nonumber\\
-  & = p(\vecj_i-\vecj_p,\veca_i-\veca_p-\Delta \veco_i\Delta t_i | \Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff) \,,\\
-  & = f(\Delta \vecj_i,\Delta \veca_i(-\Delta t_i) | \Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff) \,,\nonumber
-\end{align}
-where the first equality holds because the Jacobian of the
-$(\vecx,\vecv)\rightarrow(\vecj,\veca)$ transformation is unity and
-the second equality is a simple linear transformation. The final
-probability can again be evaluated using the parameters
-\paramsdiff\ following (d).
-
-As in the real--space formalism, we should integrate
-\eqnname~(\ref{eq:aadft}) over the time $\Delta t_i$ at which the star
-is removed from the source (cf.~\eqnname~[\ref{eq:realdf}]):
-\begin{equation}\label{eq:aadf}
-  p(\vecx_i,\vecv_i | \Phi,\vecj_p,\veca_p,\paramsdiff) = \int \dd
-  \Delta t_i\, f(\Delta \vecj_i,\Delta \veca_i(-\Delta t_i) |
-  \Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff) \,.
-\end{equation}
-Without further assumptions the action-angle formalism proceeds in the
-same way as the real--space formalism, evaluating
-\eqnname~(\ref{eq:aadft}) for a grid $j$ of times $\Delta t_i^j$ and
-summing $f(\Delta \vecj_i,\Delta \veca_i(-\Delta t^j_i) |
-\Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff)$ over $j$.
-
-Even without further assumptions about the shape of $f(\Delta
-\vecj,\Delta \veca|\paramsdiff)$, the action-angle formalism already
-allows for a simplification over the real--space method. Because the
-source orbit is specified in action-angle coordinates, it can be
-changed \emph{without} expensive orbit integrations of all possible
-source orbits. As the source's current phase-space position is a
-six-dimensional parameter, this is a major simplification of the
-dynamical model when the source orbit is unknown. Note that this
-simplification does require us to be able to compute $\veco(\vecj)$,
-to calculate the new source frequencies $\veco_p$ when changing the
-source actions $\vecj_p$. This mapping can be constructed using a
-small number of (action,frequency) calculations covering the range of
-possible source actions.
-
-We can simplify \eqnname~(\ref{eq:aadf}) by assuming (a) that
-$f(\Delta \vecj,\Delta \veca|\paramsdiff)$ is separable into
-$f_{\veca}(\Delta \veca|\paramsdiff)\,f_{\vecj}(\Delta
-\vecj|\paramsdiff)$ and (b) that the distribution $f_{\veca}(\Delta
-\veca|\paramsdiff)$ is Gaussian. In what follows, we will assume that
-this Gaussian is isotropic with dispersion $\sigma_\theta$, but all of
-the results are the same after appropriate diagonalization and
-rescaling of the angles for more general covariance matrices. In this
-case, \eqnname~(\ref{eq:aadf}) can be written as
-\begin{equation}
-\begin{split}
-  p(\vecx_i,\vecv_i & | \Phi,\vecj_p,\veca_p,\paramsdiff) 
-   = \\
-   & f_{\vecj}(\Delta \vecj_i|\paramsdiff) \int \dd \Delta t_i\,
-  \frac{1}{(2\,\pi)^{3/2}\,\sigma_\theta^3}\,\exp\left(-\frac{1}{2\,\sigma_\theta^2}\left[\sum_{R,\phi,Z}\left(\Delta \veca_i-\Delta \veco_i\Delta t_i\right)^2\right]\right)\,,
-\end{split}
-\end{equation}
-where the sum in the argument of the exponential is over the three
-components of $\Delta \veca_i$ and $\Delta \veco_i$. This integral can
-be done analytically, resulting in
-\begin{equation}\label{eq:aadfsimple}
-\begin{split}
-  p(\vecx_i,\vecv_i & | \Phi,\vecj_p,\veca_p,\paramsdiff) 
-   = \\
-   & f_{\vecj}(\Delta \vecj_i|\paramsdiff) \,  \frac{1}{2\,\pi\,\sigma_\theta^2\,\sqrt{\sum_{R,\phi,Z}\Delta \veco_i^2}}\,\exp\left(-\frac{1}{2\,\sigma_\theta^2}\left[\sum_{R,\phi,Z}\Delta \veca^2_i-\frac{\left(\sum_{R,\phi,Z} \Delta \veco_i\Delta \veca_i\right)^2}{\sum_{R,\phi,Z}\Delta \veco_i^2}\right]\right)\,.
-\end{split}
-\end{equation}
-In this expression, $\Delta X^2 \equiv (X_1-X_2)^2$. The expression in
-square brackets is positive by using the Cauchy-Schwarz inequality.
-
-The quantity in the square brackets in \eqnname~(\ref{eq:aadfsimple})
-expresses the fact that the present-day difference in angles between a
-star in a tidal stream and the source must be linearly related by a
-scalar (the time since the removal of the star) to the difference in
-frequencies, within a tolerance set by the width of the distribution
-of initial angle differences ($\sigma_\theta$). The ratio
-$\sum_{R,\phi,Z} \Delta \veco_i\Delta \veca_i/\sum_{R,\phi,Z}\Delta
-\veco_i^2$ is the best-guess for $\Delta t_i$ when performing a linear
-regression between $\Delta \veca_i$ and $\Delta \veco_i$. The
-expression in square brackets is minimized when the correlation
-between $\Delta \veca_i$ and $\Delta \veco_i$ is the largest, \ie,
-when a single time difference $\Delta t_i$ explains all three of the
-angle differences.
-
-The discussion so far has been entirely general and applies to any
-sort of stellar debris that originates from a single orbit, without
-necessarily generating a one-dimensional structure. In most
-observations of tidal streams, however, they appear as one-dimensional
-objects rather than the six-dimensional objects that we have
-considered so far. The reason for this is both physical and
-observational: (a) narrow tidal streams form because a star separates
-from the progenitor over time mostly along a single direction in
-angle-space: the Hessian matrix of the Hamiltonian, $\partial^2 H /
-\partial \vecj \partial \vecj$, has a single large eigenvalue, which
-determines the direction along which the stream is disrupted in
-angle-space (because $\partial \veco / \partial \vecj$ is equal to the
-Hessian); (b) because many tidal streams are dynamically cold with
-velocity dispersions $\lesssim10\kms$, observations are unable to
-resolve the angle differences along the two eigenvectors of the
-Hessian corresponding to the two smallest eigenvalues or the spread in
-actions among stream members. If the latter can be resolved, the
-stream is four-dimensional. In what follows we will be mainly
-concerned with the stream only spreading along a single direction in
-angle space.
-
-If the stream is one-dimensional in angle space then the time since
-the removal from the progenitor becomes unobservable if the progenitor
-orbit is unknown: above, determining the time depended on explaining
-the three-dimensional angle difference $\Delta \veca$ with the
-three-dimensional frequency difference $\Delta \veco$ using a single
-time difference; if two of the angle and frequency differences are
-unable to be resolved, then the frequency difference and the time
-difference become degenerate (aka, you can draw a straight line
-connecting any two points). To make use of this fact to simplify the
-dynamical model, we write the frequency difference as
-\begin{equation}
-  \Delta \veco = \Delta \Omega_1\,\unitvec{e}_1\,,
-\end{equation}
-where $\unitvec{e}_1$ is a unit vector. Then we write the angle
-difference $\Delta \veca$ 
-\begin{equation}
-  \Delta \veca = \Delta \theta_1\,\unitvec{e}_1 + \Delta \veca_\perp\,\unitvec{e}_1^\perp\,,
-\end{equation}
-where$\unitvec{e}_1^\perp$ is the orthogonal complement of
-$\unitvec{e}_1$ (the two directions perpendicular to $\unitvec{e}_1$)
-and $\Delta \veca_\perp$ are the components of $\Delta \veca$ in that
-direction. We can then write \eqnname~(\ref{eq:aadfsimple}) as
-\begin{equation}\label{eq:aadfsimpler}
-\begin{split}
-  p(\vecx_i,\vecv_i & | \Phi,\vecj_p,\veca_p,\paramsdiff) 
-   = \\
-   & f_{\vecj}(\Delta \vecj_i|\paramsdiff) \,  \frac{1}{2\,\pi\,\sigma_\theta^2\,\sqrt{\sum_{R,\phi,Z}\Delta \veco_i^2}}\,\exp\left(-\frac{|\Delta\veca_{i,\perp}|^2}{2\,\sigma_\theta^2}\right)\,.
-\end{split}
-\end{equation}
-\Eqnname~(\ref{eq:aadfsimpler}) is equivalent to
-\eqnname~(\ref{eq:aadfsimple}), but it expresses the dynamical model
-in a slightly different way. Whereas the angle part of
-\eqnname~(\ref{eq:aadfsimple}) allows the correct potential and
-progenitor to be inferred by asking whether a single time difference
-explains the three angle differences given the three frequencies,
-\eqnname~(\ref{eq:aadfsimpler}) `rewinds' the star's orbit along the
-direction of $\Delta \veco$ and asks whether the angle difference
-perpendicular to $\Delta \veco$ is consistent with the initial angle
-spread (because it has not evolved since the removal of the star). In
-re-writing this, we have not yet used the approximation that the
-stream only forms along a single direction in $\Delta \veco$.
-
-This view of the dynamics of the stream in action-angle coordinates
-also shows that the orbit of the source does not have to coincide with
-that of the actual progenitor.
-
 %BOVY: reference point
@@
 
-Recently, Fox \& Binney, 2014, in preparation, have suggested a new
+Recently, Fox \& Binney (2014, in preparation) have suggested a new
 method for calculating $\vecj(\vecx,\vecv)$ inspired by torus

### 2013-12-30 tex/ms.tex
streams: tweak to action--angle description

@@
 over the auxiliary angle along the path of the orbit, such that the
-oscillating cosine behavior cancels and the target action is obtained
-as the auxiliary-angle-averaged auxilary action. That is,
+mapping defined by the $S_{\vecn}$ is onto orbital tori of the target
+Hamiltonian, the oscillating cosine behavior cancels, and the target
+action is obtained as the auxiliary-angle-averaged auxilary
+action. That is,
 \begin{equation}\label{eq:jjtint}

### 2013-12-30 tex/ms.tex
streams: tweaks to the appendix

@@
 \end{equation}
-where the $\vecn > 0$ sum is over the integer two-dimensional
-half-plane that excludes the origin (two- rather than
-three-dimensional, as the azimuthal angle is perfectly correlated with
-the vertical angle in the auxiliary isochrone potential;
-\citealt{binneytremaine}), and the $S_{\vecn}$ are functions of the
-target actions that are to be determined. The canonical transformation
-corresponding to this generating function is
+where the $\vecn > 0$ sum is over the integer three-dimensional
+half-space that excludes the origin and the $S_{\vecn}$ are functions
+of the target actions that are to be determined. The canonical
+transformation corresponding to this generating function is
 \begin{equation}\label{eq:jjt}
@@
 \end{equation}
+For axisymmetric potentials $J_\phi \equiv L_Z$, such that all
+$S_\vecn = 0$ when $n_\phi \neq 0 $ in the above equations. For
+axisymmetric potentials we therefore only need to consider
+$\theta^A_R$ and $\theta^A_Z$.
 
@@
 factor $\vecn$: the condition $\vecn > 0$ means that when calculating
-$J_R$ there are terms with $\vecn = (n_R,0)$, which would not cancel
+$J_R$ there are terms with $\vecn = (n_R,0,0)$, which would not cancel
 according to the argument below when integrating over $\theta_Z$. For
@@
 \begin{equation}
-  \int \dd \theta_R^A\cos(\vecn\cdot\veca^A) = \mathrm{Re}\left[\int \dd \theta_R^A\,\exp(-i\vecn\cdot\veca^A)\right] = \mathrm{Re}\left[\exp(-i\,n_Z\,\theta^A_Z)\int \dd \theta_R^A\,\exp(-i\,n_R\,\theta^A_R)\right]\,.
+\begin{split}
+  \int \dd \theta_R^A\cos(\vecn\cdot\veca^A) & = \mathrm{Re}\left[\int \dd \theta_R^A\,\exp(-i\vecn\cdot\veca^A)\right]\,, \\
+  & = \mathrm{Re}\left[\exp(-i\,n_\phi\,\theta^A_\phi-i\,n_Z\,\theta^A_Z)\int \dd \theta_R^A\,\exp(-i\,n_R\,\theta^A_R)\right]\,.
+\end{split}
 \end{equation}
@@
 $\sin(\vecn\cdot\veca^A)$ is small, such that a small number of
-expansion terms is sufficient (we limit the terms to $n_R < 4$ and
-$n_Z < 4$).
+expansion terms is sufficient (we limit the terms in this axisymmetric
+case to $n_R < 4$ and $n_Z < 4$).
 

### 2013-12-30 tex/ms.tex
streams: some description of PDF marginalization

@@
 position--velocity space is then given by
-\begin{equation}
+\begin{equation}\label{eq:pxvt}
   p(\vecx,\vecv,t_s) = p(\veco,\veca,t_s)\,\left|\frac{\partial
@@
 or noisy phase--space dimensions as easily.
+
+We can marginalize \equationname~(\ref{eq:pxvt}) over $t_s$ and find
+that
+\begin{equation}\label{eq:pxv}
+  p(\vecx,\vecv) = p(\veco,\veca)\,\left|\frac{\partial
+    \veco}{\partial \vecj}\right|\,.
+\end{equation}
+When evaluating this PDF as a function of $(\vecx,\vecv)$ or as a
+function of observable quantities $(l,b,D,\vlos,\pmll,\pmbb)$ we
+calculate frequencies and angles using the approximate, linear
+transformation near the stream track as described at the end of
+\sectionname~\ref{sec:trackxv}. Thus, $p(\vecx,\vecv)$ can be
+evaluated very quickly by making efficient use of array operations. In
+this approximation the Jacobian $|\partial \veco / \partial \vecj|$ is
+constant everywhere near the track (for a given potential).
+
+We can marginalize the PDF over unobserved directions, convolve it
+with uncertainty distributions, or condition it on certain directions
+using the standard rules of probability theory; I will refer to all of
+these as marginalizations in what follows, as they all involve
+integrations over the PDF. In practice, the perform these
+marginalizations it is useful to estimate the extent of the PDF to
+estimate appropriate intervals for efficient numerical
+integration. For this we can use the average stream track and
+dispersion around it as estimated in
+\sectionname~\ref{sec:track}. This is most easily explained through an
+example. To determine $p(Z|X)$, the distribution of Galactocentric $Z$
+at a given Galactocentric $X$, we need to calculate
+\begin{equation}
+  p(Z|X) = \frac{p(Z,X)}{p(X)} = \frac{\int \dd Y\,\dd \vecv\, p(\vecx,\vecv)}{\int \dd Y\,\dd Z\,\dd \vecv\, p(\vecx,\vecv)}\,.
+\end{equation}
+Thus, we need to marginalize over $(Y,\vecv)$ in the numerator and
+$(Y,Z,\vecv)$ in the denominator. Focusing on the numerator, we find
+the closest point on the stream track calculated as in
+\sectionname~\ref{sec:trackxv} and calculate its six-dimensional
+Gaussian approximation. Then we condition this six-dimensional
+Gaussian on $(X,Z)$ using the standard rules of Gaussian conditioning
+(\eg, Appendix B of arXiv version 1 of \citealt{Bovy11a}) to obtain
+the approximate PDF $p(Y,\vecv|X,Z)$. We then evaluate $\int \dd
+Y\,\dd \vecv\, p(\vecx,\vecv)$ by numerical integration over the, \eg,
+$3\sigma$ range of this Gaussian; this numerical integration can be
+performed efficiently using Gaussian quadrature.
 
@@
   Binney,~J. 2012, \mnras, 426, 1324
+\bibitem[Bovy \etal(2011)]{Bovy11a}
+  Bovy,~J., Hogg,~D.~W., \& Roweis,~S.~T. 2011, Ann. Appl. Stat. 5, 1657, arXiv:0905.2979v1
 \bibitem[Carlberg(2009)]{Carlberg09a}

### 2013-12-30 py/plot_pdfs_xz.py
streams: plot the PDFs in XZ

@@ -1,0 +1,101 @@
+import sys
+import numpy
+from scipy.misc import logsumexp
+from galpy.util import bovy_plot, bovy_conversion
+from galpy import potential
+from galpy.orbit import Orbit
+from galpy.actionAngle_src.actionAngleIsochroneApprox\
+    import actionAngleIsochroneApprox
+from galpy.df_src.streamdf import streamdf
+from matplotlib import pyplot
+_STREAMSNAPDIR= '../sim/snaps'
+_STREAMSNAPAADIR= '../sim/snaps_aai'
+_NTRACKCHUNKS= 11
+_SIGV=0.365
+_NZS= 101
+_NGL= 5
+def plot_pdfs_xz(plotfilename1,plotfilename2,plotfilename3):
+    lp= potential.LogarithmicHaloPotential(q=0.9,normalize=1.)
+    aAI= actionAngleIsochroneApprox(b=0.8,pot=lp)
+    obs= numpy.array([1.56148083,0.35081535,-1.15481504,
+                      0.88719443,-0.47713334,0.12019596])
+    sdf= streamdf(_SIGV/220.,progenitor=Orbit(obs),pot=lp,aA=aAI,
+                  leading=True,nTrackChunks=_NTRACKCHUNKS,
+                  tdisrupt=4.5/bovy_conversion.time_in_Gyr(220.,8.))
+    bovy_plot.bovy_print()
+    pyplot.figure()
+    sdf.plotTrack(d1='x',d2='z',interp=True,color='k',spread=2,
+                  overplot=True,lw=1.5,scaleToPhysical=True)
+    sdf.plotProgenitor(d1='x',d2='z',color='k',
+                       overplot=True,lw=1.5,scaleToPhysical=True,
+                       ls='--')
+    bovy_plot._add_axislabels(r'$X\,(\mathrm{kpc})$',
+                               r'$Z\,(\mathrm{kpc})$')
+    pyplot.xlim(12.,14.6)
+    pyplot.ylim(-3.,7.5)
+    x1= 13.5
+    x2= 14.2
+    pyplot.plot([x1,x1],[4.5,6.2],'k-',lw=3.)
+    pyplot.plot([x2,x2],[-1.,5.],'k-',lw=3.)
+    bovy_plot._add_ticks()
+    bovy_plot.bovy_end_print(plotfilename1)
+    if True:
+        #First PDF
+        cindx= sdf._find_closest_trackpoint(x1/8.,None,5.5/8.,None,None,None,
[…]

### 2013-12-30 py/plot_pdfs_xz.py
streams: better smoothness in figure

@@ -12,7 +12,7 @@ _STREAMSNAPDIR= '../sim/snaps'
 _STREAMSNAPAADIR= '../sim/snaps_aai'
 _NTRACKCHUNKS= 11
 _SIGV=0.365
-_NZS= 101
+_NZS= 301
 _NGL= 5
 def plot_pdfs_xz(plotfilename1,plotfilename2,plotfilename3):
     lp= potential.LogarithmicHaloPotential(q=0.9,normalize=1.)

### 2013-12-30 tex/ms.tex
streams: add PDF figures

@@
 or noisy phase--space dimensions as easily.
+
+\begin{figure}[t!]
+ \includegraphics[width=0.32\textwidth,clip=]{gd1_pdf_xz_0.ps}
+ \includegraphics[width=0.32\textwidth,clip=]{gd1_pdf_xz_1.ps}
+ \includegraphics[width=0.32\textwidth,clip=]{gd1_pdf_xz_2.ps}
+ \caption{Marginalized, conditional $p(Z|X)$ PDFs. The left panel
+   shows the model leading stream (see \figurename~\ref{fig:gd1_xz}),
+   with the stream track shown as the solid line, the estimated
+   $2\sigma$ width of the stream as the dash-dotted line, and the
+   progenitor orbit as the dashed line. The solid line in the middle
+   and right panels shows $p(Z|X)$ calculated at the $X$ indicated in
+   the left panel; the dash-dotted line is a simple Gaussian
+   estimate. The upper branch near $Z \approx 5\kpc$ of the stream is
+   chosen for the middle panel. The right panel shows an $X$ close to
+   where the stream turns around in $X$; the PDF has multiple peaks in
+   this case.}\label{fig:gd1_pdf_xz}
+%python plot_pdfs_xz.py ../tex/gd1_pdf_xz_0.ps ../tex/gd1_pdf_xz_1.ps ../tex/gd1_pdf_xz_2.ps
+\end{figure}
 

### 2013-12-30 tex/ms.tex
streams: finish PDF section

@@
 
+\figurename~\ref{fig:gd1_pdf_xz} shows such PDFs of $p(Z|X)$ for a few
+$X$ on the leading arm of the stream as well as the estimated Gaussian
+PDFs determined using the procedure in
+\sectionname~\ref{sec:trackxv}. It is clear that the average stream
+location as calculated in \sectionname~\ref{sec:trackxv} is very close
+to the actual average location of the stream in $(\vecx,\vecv)$ and
+that the estimated dispersion is close to the true model
+dispersion. The fact that the estimated Gaussian in the right panel is
+only at the highest peak is by choice; there is an estimated Gaussian
+dispersion at the lower peak as well.
+
+Similar procedures can be followed to evaluate other projections and
+marginalizations of the PDF, \eg, $p(b|l)$ for the observed track of
+the stream on the sky, to convolve the PDF with observational
+uncertainties, and to calculate moments of the PDF, \eg, the density
+and velocity dispersion along the stream.
+
 %BOVY: reference point
@@
 
-\acknowledgements It is a pleasure to thank \dots for helpful comments
-and assistance. J.B. was supported by NASA through Hubble Fellowship
-grant HST-HF-51285.01 from the Space Telescope Science Institute,
-which is operated by the Association of Universities for Research in
-Astronomy, Incorporated, under NASA contract NAS5-26555. J.B.
-acknowledges support from SFB 881 (A3) funded by the German Research
-Foundation DFG.
+\acknowledgements It is a pleasure to thank David W. Hogg, Hans-Walter
+Rix, and Jason Sanders for helpful discussions. J.B. was supported by
+NASA through Hubble Fellowship grant HST-HF-51285.01 from the Space
+Telescope Science Institute, which is operated by the Association of
+Universities for Research in Astronomy, Incorporated, under NASA
+contract NAS5-26555. J.B.  acknowledges support from SFB 881 (A3)
+funded by the German Research Foundation DFG.
 

### 2013-12-31 tex/ms.tex
streams: text tweaks after first perusal

@@
 \newcommand{\operpii}{\ensuremath{\Omega_{\perp,ii}}}
+\newcommand{\ts}{\ensuremath{t_s}}
 
@@
   streams. The framework consists of simple models for the initial
-  action--angle distribution of tidal debris as a function of time,
-  which can be straightforwardly evolved forward in time. Taking
-  advantage of the essentially one-dimensional nature of cold tidal
-  streams, the transformation to position--velocity coordinates can be
-  linearized and interpolated near a small number of points along the
-  progenitor orbit, thus allowing for efficient computations of a
-  stream's properties in observable quantities. I illustrate how to
-  calculate the stream's track in different coordinate systems, how to
-  quickly estimate the dispersion perpendicular to the track, and how
-  to draw mock stream data. As a generative model, this framework
-  allows one to compute the full probability distribution function and
-  marginalize over or condition it on certain phase--space dimensions
-  as well as convolve it with observational uncertainties. This will
-  be instrumental in proper data analysis of stream data. In addition
-  to providing a computationally-efficient practical tool for modeling
-  the dynamics of tidal streams, the action--angle nature of the
-  framework helps elucidate in exactly what manner streams do not
-  follow single orbits, how the observed width of the stream relates
-  to the velocity dispersion or mass of the progenitor, and how the
-  progenitor of ``orphan'' streams could be located.
+  action--angle distribution of tidal debris, which can be
+  straightforwardly evolved forward in time. Taking advantage of the
+  essentially one-dimensional nature of cold tidal streams, the
+  transformation to position--velocity coordinates can be linearized
+  and interpolated near a small number of points along the progenitor
+  orbit, thus allowing for efficient computations of a stream's
+  properties in observable quantities. I illustrate how to calculate
+  the stream's average location (its ``track'') in different
+  coordinate systems, how to quickly estimate the dispersion around
+  its track, and how to draw mock stream data. As a generative model,
+  this framework allows one to compute the full probability
+  distribution function and marginalize over or condition it on
+  certain phase--space dimensions as well as convolve it with
+  observational uncertainties. This will be instrumental in proper
+  data analysis of stream data. In addition to providing a
+  computationally-efficient practical tool for modeling the dynamics
+  of tidal streams, the action--angle nature of the framework helps
+  elucidate in exactly what manner streams do not follow single
+  orbits, how the observed width of the stream relates to the velocity
+  dispersion or mass of the progenitor, and how the progenitors of
+  ``orphan'' streams could be located.
 
@@
 and the action distribution of the tidal debris \citep{Eyre11a},
-allowing simple physical models of the stream to be used and
+allowing simple physical models of the stream to be used and to be
 constrained by observational data.
@@
 In this paper, I present a new method for modeling cold tidal streams
-that fundamentally lives in frequency--angle space, because after
-specifying the time at which a star was stripped, the stream
+that fundamentally lives in frequency--angle space, because the stream
 distribution function is essentially one-dimensional in this
@@
 In real space, a tidal stream forms as stars are stripped from a
-progenitor some time $\Delta t$ in the past and then evolve (largely)
-independently in a galaxy's host potential, moving away both in
-position and velocity from the progenitor object as time goes
-on. Thus, a star was offset by a small amount $(\Delta \vecx,\Delta
-\vecv)$ from the position of the progenitor $(\vecx^p,\vecv^p)(-\Delta
-t)$ at that time. Then the star orbited under the influence of the
-same gravitational potential $\Phi$ as the progenitor until the
-present day $(t=0)$, when it was observed: $(\vecx,\vecv)(t=0)$ =
-$\vech_{\Delta t}(\vecx^p(-\Delta t)+\Delta \vecx,\vecv^p(-\Delta
-t)+\Delta \vecv)$, where $\vech$ denotes the Hamiltonian flow from
-$t=-\Delta t$ to $t=0$ (see \citealt{binneytremaine}); this is simply
-the orbit of the star between these times. In phase-space this orbit
-can be calculated by solving Hamilton's equations: $\dot{\vecx} =
-\vecv; \dot{\vecv} = - \dd \Phi / \dd \vecx$.
+progenitor cluster or satellite galaxy some time $\ts$ in the
+past and then evolve (largely) independently in a galaxy's host
+potential, moving away both in position and velocity from the
+progenitor object as time goes on. Thus, a star was offset by a small
+amount $(\Delta \vecx,\Delta \vecv)$ from the position of the
+progenitor $(\vecx^p,\vecv^p)(-\ts)$ at that time. Then the star
+orbited under the influence of the same gravitational potential $\Phi$
+as the progenitor until the present day $(t=0)$, when it was observed:
+$(\vecx,\vecv)(t=0)$ = $\vech_{\ts}(\vecx^p(-ts)+\Delta
+\vecx,\vecv^p(-\ts)+\Delta \vecv)$, where $\vech$ denotes the
+Hamiltonian flow from $t=-\ts$ to $t=0$ (see
+\citealt{binneytremaine}); this is simply the orbit of the star
+between these times. In phase-space this orbit can be calculated by
+solving Hamilton's equations: $\dot{\vecx} = \vecv; \dot{\vecv} = -
+\dd \Phi / \dd \vecx$.
 
@@
 simpler. A star received an offset ($\Delta \vecj,\Delta \veca$) at
-time $-\Delta t$ from the action--angle coordinates of the progenitor
+time $-\ts$ from the action--angle coordinates of the progenitor
 $(\vecj^p,\veca^p)$, similar to the small offset ($\Delta \vecx,\Delta
@@
 \begin{equation}
-\Delta \veca(t=0) = \Delta \veca(-\Delta t) + \Delta \veco \Delta t\,,
+\Delta \veca(t=0) = \Delta \veca(-\ts) + \Delta \veco \,\ts\,,
 \end{equation}
@@
 \begin{equation}
-\Delta \vecj(t=0) = \Delta \vecj(-\Delta t)\,.
-\end{equation}
-As the initial offset $\Delta \veca(-\Delta t)$ in angles is small
-compared to the growth, we have that $\Delta\veco\Delta t$, $\Delta
-\veca(t=0) \approx \Delta \veco \Delta t$.
+\Delta \vecj(t=0) = \Delta \vecj(-\ts)\,.
+\end{equation}
+As the initial offset $\Delta \veca(-\ts)$ in angles is small compared
+to the growth $\Delta \veco \,\ts$, we have that $\Delta \veca(t=0)
+\approx \Delta \veco \,\ts$.
 
@@
 \vecj) \Delta \vecj$, in the limit of small action offsets, because
-$\veco = \partial H / \partial \vecj$. For an approximately
-one-dimensional stream to form, the Hessian matrix $(\partial^2 H /
-\partial \vecj \partial \vecj)\large|_{\vecj^p}$ has to be dominated
-by a single large eigenvalue; in this case the angle difference of
-stars stripped from the stream at any time will fall along
-approximately the same direction $\unitvec\veco$, and the stream will
-essentially be one dimensional. In detail, because the distribution of
-initial offsets $\Delta \vecj$ is not isotropic, the direction
-$\unitvec \veco$ is not just determined by the potential (through $H$)
-and the progenitor orbit, but also by the distribution of $\Delta
-\vecj$, which is set through a combination of the internal dynamics of
-the progenitor and the stripping process (see discussion in
-\citealt{Sanders13b}). For dynamical modeling we can be agnostic about
-the relative importance of all of these contributors to $\unitvec
-\veco$ by flexible modeling of the distribution of $\Delta \vecj$ or
-$\Delta \veco$.
+$\veco = \partial H / \partial \vecj$, where $H$ is the
+Hamiltonian. For an approximately one-dimensional stream to form, the
+Hessian matrix $(\partial^2 H / \partial \vecj \partial
+\vecj)\large|_{\vecj^p}$ has to be dominated by a single large
+eigenvalue; in this case the angle difference of stars stripped from
+the stream at any time will fall along approximately the same
+direction $\unitvec\veco$, and the stream will be essentially one
+dimensional. In detail, because the distribution of initial offsets
+$\Delta \vecj$ is not isotropic, the direction $\unitvec \veco$ is not
+just determined by the potential (through $H$) and the progenitor
+orbit, but also by the distribution of $\Delta \vecj$, which is set
+through a combination of the internal dynamics of the progenitor and
+the stripping process (see discussion in \citealt{Sanders13b}). For
+dynamical modeling we can be agnostic about the relative importance of
+all of these contributors to $\unitvec \veco$ by flexible modeling of
+the distribution of $\Delta \vecj$ or $\Delta \veco$.
 
@@
 of the GD-1 stream at the end of the simulation
-\citep{Koposov10a}. The progenitor's orbit has an eccentricity of
-$0.31$, a pericenter of $13.7\kpc$, a current $r=14.4\kpc$, such that
-the cluster is close to pericenter at the end of the simulation. The
-actions of the progenitor are $(J_R,L_Z,J_Z) =
+\citep{Koposov10a}. The progenitor cluster's orbit has an eccentricity
+of $0.31$, a pericenter of $13.7\kpc$, and a current $r=14.4\kpc$,
+such that the cluster is close to pericenter at the end of the
+simulation. The actions of the progenitor are $(J_R,L_Z,J_Z) =
 (288.5,-3173.7,897.6)\kpc\kms$ and its orbital frequencies are
@@
 \partial \vecj)\large|_{\vecj^p}$ must be strongly dominated by a
-single large eigenvalue leads us to propose that frequency--angle
+single large eigenvalue leads me to propose that frequency--angle
 space is a better coordinate system than action--angle space for
@@
 and the overlapping leading- and trailing-arm distributions in
-($J_R,J_Z$) are hard to model with a simple distribution function. The
-distribution of frequencies, however, is close-to one dimensional, its
-direction of largest variance can be well-modeled as a Gaussian
-(although below we will model it slightly differently), and its
-parameters can all be easily estimated from the velocity dispersion of
-the progenitor, the orbit of the progenitor, and the gravitational
-potential, as discussed below. Therefore, we will model tidal streams
-in frequency--angle space. Because the leading and trailing arm are
-well-separated in frequency space, we model each arm individually.
+($J_R,J_Z$) are difficult to model with a simple distribution
+function. The distribution of frequencies, however, is close-to one
+dimensional, its direction of largest variance can be well-modeled as
+a Gaussian (although below we will model it slightly differently), and
+its parameters can all be easily estimated from the velocity
+dispersion of the progenitor, the orbit of the progenitor, and the
+gravitational potential, as discussed below. Therefore, we will model
+tidal streams in frequency--angle space. Because the leading and
+trailing arm are well-separated in frequency space, we model each arm
+individually.
 
@@
 \begin{equation}
-  \Delta t = \frac{\Delta \veco \cdot \Delta \veca}{|\Delta \veco|^2}\,.
+  \ts = \frac{\Delta \veco \cdot \Delta \veca}{|\Delta \veco|^2}\,.
 \end{equation}
 The dotted line shows a finely-binned histogram of these stripping
-times, which indicates that stripping happens in bursts with a radial
-period of about $400\Myr$, the radial period of the progenitor's
-orbit. Thus, as expected, stripping happens primarily at pericenter
-passages, with a smaller number of particles lost between pericenter
+times, which indicates that stripping happens in bursts with a period
+of about $400\Myr$, the radial period of the progenitor's orbit. Thus,
+as expected, stripping happens primarily at pericenter passages, with
+a smaller number of particles lost between pericenter
 passages. Realistic models of a stream, especially those that require
@@
 small extent but the narrow distribution of frequency offsets limits
-the ability of particles stripped later to reach large distances from
+the ability of particles removed later to reach large distances from
 the progenitor.
@@
   \begin{split}
-    p(\Delta t) & \propto 1\,,\qquad 0 < \Delta t < t_d\\
+    p(\ts) & \propto 1\,,\qquad 0 < \ts < t_d\\
     & = 0\,,\qquad \mathrm{otherwise}\,.
@@
 \sigma_{\Omega,1}$, where $\sigma_{\Omega,1}$ is the square root of
-the largest eigenvalue of $V_\veco$ and $\Delta \Omega^m$ is the mean
-frequency offset. The sign of this parameter sets whether we are
-modeling a leading or a trailing arm. \figurename~\ref{fig:gd1_dohist}
-shows the distribution frequency offsets in the direction of the
-median frequency offset in the stream. The dashed line is a Gaussian
-fit to this distribution, while the solid line is a fit of the form
+the largest eigenvalue of $V_\veco$ and $\Delta \Omega^m$ is the
+one-dimensional mean frequency offset. The sign of this parameter sets
+whether we are modeling a leading or a trailing
+arm. \figurename~\ref{fig:gd1_dohist} shows the distribution of
+frequency offsets in the direction of the median frequency offset in
+the stream. The dashed line is a Gaussian fit to this distribution,
+while the solid line is a fit of the form
 \begin{equation}\label{eq:do}
@@
 as an isotropic Gaussian, has a charactaristic spread $\sigma_{\theta}$
-which we set here to $0.00365$ ($=\sigv/[100\kms]$), and we likewise
+which we set here to $0.003$ ($=\sigv/[122\kms]$), and we likewise
 expect this spread to scale with the velocity dispersion of the
@@
 The fiducial model for a tidal stream in frequency--angle coordinates
-he proposed here is therefore described by essentially two free
+proposed here is therefore described by essentially two free
 parameters in addition to the progenitor's phase-space position: the
@@
 determining the track of the stream. More sophisticated models can
-leave $\mu_\Omega$ and $\sigma_{\theta}$ as free paramters as well. In
-what follows, we will fix these to the values given in the previous
-paragraph. It must be stressed, however, that all of these parameters
-here have been fixed ``by-eye'', and better fits might be possible
-using more quantitative fit procedures. But such procedures are not
-pursued here further.
+leave $\mu_\Omega$ and $\sigma_{\theta}$ as free parameters as
+well. In what follows, we will fix these to the values given in the
+previous paragraph. It must be stressed, however, that all of these
+parameters here have been fixed ``by-eye'', and better fits might be
+possible using more quantitative fit procedures.
 
@@
 
-Here, I discuss how to calculate the stream's properties as a function
-of the angle along the stream. In the generative model described
-above, the direction of the angle along the stream lies along the
-principal eigenvector of $V_\veco$, the variance matrix of the model
-$\Delta \veco$ distribution. We denote the angle in this direction as
-$\apar$ and the frequency in this direction $\opar$; the
-two-dimensional angle- and frequency-space perpendicular to this
-direction is denoted as $\aperp$ and $\operp$.
+In the generative model described above, the direction of the angle
+along the stream lies along the principal eigenvector of $V_\veco$,
+the variance matrix of the model $\Delta \veco$ distribution. We
+denote the angle in this direction as $\apar$ and the frequency in
+this direction $\opar$; the two-dimensional angle- and frequency-space
+perpendicular to this direction is denoted as $\aperp$ and $\operp$.
 
 First, we compute the distribution of $\Delta \opar$, the frequency
-offset between stream members and the progenitor, given the angle
-offset $\Delta \apar$ as follows
+offset between stream members and the progenitor given the angle
+offset $\Delta \apar$ as follows (considering $\Delta \opar$ to be
+positive)
 \begin{equation}\label{eq:poparapargeneral}
 \begin{split}
-  p(\Delta \opar|\Delta \apar) & = \int \dd \Delta t\,p(\Delta \opar,\Delta t|\Delta \apar)\, \\
-  & = \int \dd \Delta t \,p(\Delta \apar | \Delta \opar,\Delta t)\,p(\Delta \opar,\Delta t)\,,\\
-  & \approx \int \dd \Delta t \,\delta(\Delta \apar - \Delta \opar\,\Delta t)\,p(\Delta \opar,\Delta t)\,,\\
-  & = \frac{1}{\left|\Delta \opar\right|}\,p(\Delta \opar,\frac{\Delta \apar}{\Delta \opar})\,,
+  p(\Delta \opar|\Delta \apar) & = \int \dd \ts\,p(\Delta \opar,\ts|\Delta \apar)\, \\
+  & \propto \int \dd \ts \,p(\Delta \apar | \Delta \opar,\ts)\,p(\Delta \opar,\ts)\,,\\
+  & \approx \int \dd \ts \,\delta(\Delta \apar - \Delta \opar\,\ts)\,p(\Delta \opar,\ts)\,,\\
+  & = \frac{1}{\Delta \opar}\,p(\Delta \opar,\frac{\Delta \apar}{\Delta \opar})\,,
 \end{split}
@@
 \begin{split}
-  p(\Delta \opar|\Delta \apar) \propto &\ \frac{1}{\left|\Delta \opar\right|}\,p(\Delta \opar)\,,\qquad 0 < \frac{\Delta \apar}{\Delta \opar} < t_d\,.\\
+  p(\Delta \opar|\Delta \apar) \propto &\ \frac{1}{\Delta \opar}\,p(\Delta \opar)\,,\qquad 0 < \frac{\Delta \apar}{\Delta \opar} < t_d\,.\\
   & 0\,,\qquad \mathrm{otherwise}\,.
@@
 average stream track as a function of $\Delta \apar$ is entirely
-specified by the mean $\Delta \opar(\Delta \apar)$ and zero offsets in
-the perpendicular direction. This track can be rotated into the
-$(R,\phi,Z)$ coordinate system using the eigenvectors of $V_\veco$.
+specified by the mean $\Delta \opar(\Delta \apar)$ in the parallel
+direction and zero offsets in the perpendicular directions. This track
+can be rotated into the $(R,\phi,Z)$ coordinate system using the
+eigenvectors of $V_\veco$.
 
@@
 \apar)$, however, does depend on $\Delta \apar$, because $\Delta
-\aperp = \Delta \aperp^{\mathrm{init}} + \Delta \operp \Delta t$ and
-the distribution of $\Delta t$ depends on $\Delta \apar$. Therefore,
-we first calculate $p(\Delta t | \Delta \apar)$.
-
-We can calculate $p(\Delta t | \Delta \apar)$ as follows
+\aperp = \Delta \aperp^{\mathrm{init}} + \Delta \operp \ts$ and
+the distribution of $\ts$ depends on $\Delta \apar$. Therefore,
+we first calculate $p(\ts | \Delta \apar)$.
+
+We can calculate $p(\ts | \Delta \apar)$ as follows
 \begin{equation}\label{eq:pdtapar}
 \begin{split}
-  p(\Delta t | \Delta \apar) & = \int \dd \Delta \opar\,p(\Delta t,\Delta \opar|\Delta \apar)\,,\\
-  & = \int \dd \Delta \opar\,p(\Delta t|\Delta \opar,\Delta \apar)\,p(\Delta \opar|\Delta \apar)\,,\\
-  & = p\left(\Delta \opar = \frac{\Delta \apar}{\Delta t}\right)\,\left|\frac{\Delta \apar}{(\Delta t)^2}\right|\,,
+  p(\ts | \Delta \apar) & = \int \dd \Delta \opar\,p(\ts,\Delta \opar|\Delta \apar)\,,\\
+  & = \int \dd \Delta \opar\,p(\ts|\Delta \opar,\Delta \apar)\,p(\Delta \opar|\Delta \apar)\,,\\
+  & = p\left(\Delta \opar = \frac{\Delta \apar}{\ts}\right)\,\left|\frac{\Delta \apar}{(\ts)^2}\right|\,,
 \end{split}
@@
 factor in this equation is given by the expression given in
-\equationname~(\ref{eq:poparapargeneral}) (in general) or
+\equationname~(\ref{eq:poparapargeneral}) (in general) or in
 \equationname~(\ref{eq:poparapar}) for the fiducial model. The lower
@@
 $\Delta \opar$ as a function of $\Delta \apar$, the distribution of
-$\Delta t$ has a kink at the angle offset that can only be reached by
-stars that have to have been stripped at large frequency offsets (that
-is, larger than average).
-
-We can now also calculate $p(\Delta \aperp|\Delta \apar)$
+$\ts$ has a kink at the angle offset that can only be reached by stars
+that have to have been stripped at large frequency offsets (that is,
+larger than average).
+
+We can now also calculate $p(\Delta \aperp|\Delta \apar)$:
 \begin{equation}
-  p(\Delta \aperpii|\Delta \apar) = \int \dd \Delta t \,\dd \Delta
-  \operpii \,p(\Delta \aperpii|\Delta \operpii,\Delta t)\,p(\Delta \operpii|\Delta t,\Delta \apar)\,p(\Delta t|\Delta \apar)\,,
+  p(\Delta \aperpii|\Delta \apar) = \int \dd \ts \,\dd \Delta
+  \operpii \,p(\Delta \aperpii|\Delta \operpii,\ts)\,p(\Delta \operpii|\ts,\Delta \apar)\,p(\ts|\Delta \apar)\,,
 \end{equation}
@@
 \begin{equation}\label{eq:paperpapar}
-  p(\Delta \aperpii|\Delta \apar) = \int \dd \Delta t \,\mathcal{N}\left(\Delta \aperpii|0,\sigma_{\theta}^2+(\Delta t\,\sigma_{\Omega,ii})^2\right)\,p(\Delta t|\Delta \apar)\,,
+  p(\Delta \aperpii|\Delta \apar) = \int \dd \ts \,\mathcal{N}\left(\Delta \aperpii|0,\sigma_{\theta}^2+(\ts\,\sigma_{\Omega,ii})^2\right)\,p(\ts|\Delta \apar)\,,
 \end{equation}
@@
 perpendicular direction (corresponding to the middle and smallest
-eigenvalue of $V_\veco$). In the fiducial model, this distribution is
-the only one of those considered in this section that requires an
-explicit integration. However, we can approximate the distribution by
-using the mean $\Delta t(\Delta \apar)$ rather than integrating over
-$\Delta t$. The lower left panel of \figurename~\ref{fig:gd1_apar}
-shows the distribution of perpendicular angle offsets and the
-grayscale bands show the dispersion calculated using
-\equationname~(\ref{eq:paperpapar}); the white dashed line shows the
-approximate estimate obtained using the mean stripping time rather
+eigenvalue of $V_\veco$). We have \emph{not} approximated the initial
+angle-offset distribution as a delta function here, because its width
+is a significant fraction of that of the the final
+perpendicular-angle-offset distribution. In the fiducial model, this
+distribution is the only one of those considered in this section that
+requires an explicit integration. However, we can approximate the
+distribution by using the mean $\ts(\Delta \apar)$ rather than
+integrating over $\ts$. The lower left panel of
+\figurename~\ref{fig:gd1_apar} shows the distribution of perpendicular
+angle offsets and the grayscale bands show the dispersion calculated
+using \equationname~(\ref{eq:paperpapar}); the white dashed line shows
+the approximate estimate obtained using the mean stripping time rather
 than integrating over it. This simple estimate agrees well with the
@@
 is more difficult, and I simply use a dispersion of $1$, as the stream
-typically spreads out over $\approx 1$ radian. This choice could be
-improved, as it has the effect of overestimating the true dispersion
-(see \sectionname~\ref{sec:pdf} below). We could calculate the
-correlation between $\Delta \operpii$ and $\Delta aperpii$ using
+typically spreads out over $\approx 1$ radian. We could calculate the
+correlation between $\Delta \operpii$ and $\Delta \aperpii$ using
 similar equations as those given in the previous paragraphs, but for a
@@
 chosen here to span $1.5$ radians, we calculate the Jacobian $\partial
-(\veco,\veca) / \partial (\vecx,\vecv)$. In this figure this is done
+(\veco,\veca) / \partial (\vecx,\vecv)$. In this Figure this is done
 at four points, but for all other figures and calculations in this
@@
 interpolation. The variance matrix at interpolated track points can
-then be constructed by from the interpolated eigenvalues and
+then be constructed from the interpolated eigenvalues and
 eigenvectors. All other $(\vecx,\vecv) \leftrightarrow (\veco,\veca)$
@@
 Using the generative model described in
-\sectionname~\ref{sec:modeloa}, it is straightforward how to draw mock
+\sectionname~\ref{sec:modeloa}, it is straightforward to draw mock
 stream data in frequency--angle space. The three ingredients---time,
@@
 
-The stripping time $t_s$ cannot be observed and should therefore by
+The stripping time $t_s$ cannot be observed and should therefore be
 marginalized over when evaluating the PDF. One of the great advantages
 of modeling a stream in frequency--angle coordinates is that this
-marginalization can be done analytically for many choices of $p(t_s)$
-if $p(\veco|t_s) \equiv p(\veco)$. Specifically, for the fiducial
-model $p(t_s)$ is uniform up to a maximum disruption time $t_d$ (see
-\equationname~[\ref{eq:pt}]) such that we can write
+marginalization can be calculated analytically for many choices of
+$p(t_s)$ if $p(\veco|t_s) \equiv p(\veco)$. Specifically, for the
+fiducial model $p(t_s)$ is uniform up to a maximum disruption time
+$t_d$ (see \equationname~[\ref{eq:pt}]) such that we can write
 \begin{equation}
@@
   p(\veco,\veca) & = 
-  p(\veco) \,  \frac{\mathrm{erf}(a_0) + \mathrm{erf}(a_d)}{4\,\pi\,\sigma_\theta^2\,|\Delta \veco|}\,\exp\left(-\frac{1}{2\,\sigma_\theta^2}\left[\sum_{R,\phi,Z}\Delta \veca^2_i-\frac{\left(\sum_{R,\phi,Z} \Delta \veco_i\Delta \veca_i\right)^2}{|\Delta \veco|}\right]\right)\,.
+  p(\veco) \,  \frac{\mathrm{erf}(a_0) + \mathrm{erf}(a_d)}{4\,\pi\,\sigma_\theta^2\,|\Delta \veco|}\,\exp\left(-\frac{1}{2\,\sigma_\theta^2}\left[\sum_{R,\phi,Z}\Delta \veca^2_i-\frac{\left(\sum_{R,\phi,Z} \Delta \veco_i\Delta \veca_i\right)^2}{|\Delta \veco|^2}\right]\right)\,.
 \end{split}
@@
 \begin{align}
-  \tilde{t}_s & = \frac{\sum \Delta \veco_i\Delta \veca_i}{|\Delta \veco_i|^2}\,,\\
-  a_0 & = \frac{|\Delta \veco_i|}{\sqrt{2}\,\sigma_\theta}\,\tilde{t}_s\,,\\
-  a_d & = \frac{|\Delta \veco_i|}{\sqrt{2}\,\sigma_\theta}\,\left(t_d - \tilde{t}_s\right)\,.
+  \tilde{t}_s & = \frac{\sum \Delta \veco_i\Delta \veca_i}{|\Delta \veco|^2}\,,\\
+  a_0 & = \frac{|\Delta \veco|}{\sqrt{2}\,\sigma_\theta}\,\tilde{t}_s\,,\\
+  a_d & = \frac{|\Delta \veco|}{\sqrt{2}\,\sigma_\theta}\,\left(t_d - \tilde{t}_s\right)\,.
 \end{align}
@@
 serves to suppress the PDF in case the stripping time is smaller than
-zero, \ie, a stream member appears to have been stripping \emph{in the
+zero, \ie, a stream member appears to have been removed \emph{in the
   future}; in this case $\mathrm{erf}(a_0) \approx -1$ and the PDF
@@
 the motion, or actions (\eg, \citealt{Binney08a,Penarrubia12a},
-Sanderson \etal, 2013, in preparation). These approaches ignore the
+Sanderson \etal, 2014, in preparation). These approaches ignore the
 correlations between actions and angles in the stream (the exponential
@@
 the progenitor when integrating their orbits backward in time
-\citep{Johnston99a,PriceWhelan13a}) use the full PDF (but expressed in
+\citep{Johnston99a,PriceWhelan13a} use the full PDF (but expressed in
 position--velocity space), but cannot be marginalized over unobserved
@@
    and right panels shows $p(Z|X)$ calculated at the $X$ indicated in
-   the left panel; the dash-dotted line is a simple Gaussian
-   estimate. The upper branch near $Z \approx 5\kpc$ of the stream is
-   chosen for the middle panel. The right panel shows an $X$ close to
-   where the stream turns around in $X$; the PDF has multiple peaks in
-   this case.}\label{fig:gd1_pdf_xz}
+   the left panel by thick vertical lines; the dash-dotted line is a
+   simple Gaussian estimate. The upper branch near $Z \approx 5\kpc$
+   of the stream is chosen for the middle panel. The right panel shows
+   an $X$ close to where the stream turns around in $X$; the PDF has
+   multiple peaks in this case.}\label{fig:gd1_pdf_xz}
 %python plot_pdfs_xz.py ../tex/gd1_pdf_xz_0.ps ../tex/gd1_pdf_xz_1.ps ../tex/gd1_pdf_xz_2.ps
@@
 these as marginalizations in what follows, as they all involve
-integrations over the PDF. In practice, the perform these
+integrations over the PDF. In practice, to perform these
 marginalizations it is useful to estimate the extent of the PDF to
-estimate appropriate intervals for efficient numerical
-integration. For this we can use the average stream track and
-dispersion around it as estimated in
-\sectionname~\ref{sec:track}. This is most easily explained through an
-example. To determine $p(Z|X)$, the distribution of Galactocentric $Z$
-at a given Galactocentric $X$, we need to calculate
+define appropriate intervals for efficient numerical integration. For
+this we can use the average stream track and dispersion around it as
+estimated in \sectionname~\ref{sec:track}. This is most easily
+explained through an example. To determine $p(Z|X)$, the distribution
+of Galactocentric $Z$ at a given Galactocentric $X$, we need to
+calculate
 \begin{equation}
@@
 only at the highest peak is by choice; there is an estimated Gaussian
-dispersion at the lower peak as well.
+dispersion at the lower peak as well, but it is not shown.
 
@@
 
-As in torus modeling, the Fox \& Binney method uses a auxiliary
+As in torus modeling, the Fox \& Binney method uses an auxiliary
 isochrone potential $\Phi^A$ that should be close to the target
@@
   \int \dd \theta_R^A\cos(\vecn\cdot\veca^A) & = \mathrm{Re}\left[\int \dd \theta_R^A\,\exp(-i\vecn\cdot\veca^A)\right]\,, \\
-  & = \mathrm{Re}\left[\exp(-i\,n_\phi\,\theta^A_\phi-i\,n_Z\,\theta^A_Z)\int \dd \theta_R^A\,\exp(-i\,n_R\,\theta^A_R)\right]\,.
+  & \approx \mathrm{Re}\left[\exp(-i\,n_\phi\,\theta^A_\phi-i\,n_Z\,\theta^A_Z)\int \dd \theta_R^A\,\exp(-i\,n_R\,\theta^A_R)\right]\,.
 \end{split}
@@
 to using it is that a sufficiently close isochrone potential has to be
-supplied. Sufficiently close means specifically that both the
-auxiliary radial and vertical angle have to go through the full range
-of $0$ to $2\pi$ along the orbit in the target potential. Whether they
-do or not is easy to check and I have found that with limited trial
-and error a good enough isochrone potential can be determined quickly
-for Milky-Way-like axisymmetric potentials, although it remains to be
-checked whether this holds for triaxial potentials as well. For
-analyzing stream data, the necessity of this trial-and-error
-determinition of the auxiliary potential does not pose a practical
-problem, as all of the stream stars are on very similar
-orbits. Therefore, the same auxiliary potential can be used for all
-action--angle calculations for a given potential and for different
-potentials as well, as long as they are not too different. More
-widespread use of this technique may require automated methods for
-determining a good auxiliary potential. In the unlikely circumstance
-that no good auxiliary isochrone potential can be found, the method
-discussed here can also be used with other auxiliary potentials for
-which the actions and angles can be calculated, such as other
-spherical potentials or the family of St\"{a}ckel potentials. This
-will, however, add a non-negligible amount of computation for the
-necessary numerical integrations.
+supplied. Sufficiently close means specifically that the auxiliary
+angles have to go through the full range of $0$ to $2\pi$ along the
+orbit in the target potential. Whether they do or not is easy to check
+and I have found that with limited trial and error a good enough
+isochrone potential can be determined quickly for Milky-Way-like
+axisymmetric potentials, although it remains to be checked whether
+this holds for triaxial potentials as well. For analyzing stream data,
+the necessity of this trial-and-error determination of the auxiliary
+potential does not pose a practical problem, as all of the stream
+stars are on very similar orbits. Therefore, the same auxiliary
+potential can be used for all action--angle calculations for a given
+potential and for different potentials as well, as long as they are
+not too different. More widespread use of this technique may require
+automated methods for determining a good auxiliary potential. In the
+unlikely circumstance that no good auxiliary isochrone potential can
+be found, the method discussed here can also be used with other
+auxiliary potentials for which the actions and angles can be
+calculated, such as other spherical potentials or the family of
+St\"{a}ckel potentials. This will, however, add a non-negligible
+amount of computation for the necessary numerical integrations.
 

### 2013-12-31 py/plot_pdfs_x.py
streams: plot density as a function of x for the trailing tail

@@ -1,0 +1,65 @@
+import sys
+import os, os.path
+import numpy
+from scipy.misc import logsumexp
+from galpy.util import bovy_plot, bovy_conversion, multi, bovy_coords
+import multiprocessing
+from galpy import potential
+from galpy.orbit import Orbit
+from galpy.actionAngle_src.actionAngleIsochroneApprox\
+    import actionAngleIsochroneApprox
+from galpy.df_src.streamdf import streamdf
+_STREAMSNAPDIR= '../sim/snaps'
+_STREAMSNAPAADIR= '../sim/snaps_aai'
+_NTRACKCHUNKS= 11
+_SIGV=0.365
+_NLS= 301
+_NGL= 5
+def plot_pdfs_x(plotfilename):
+    lp= potential.LogarithmicHaloPotential(q=0.9,normalize=1.)
+    aAI= actionAngleIsochroneApprox(b=0.8,pot=lp)
+    obs= numpy.array([1.56148083,0.35081535,-1.15481504,
+                      0.88719443,-0.47713334,0.12019596])
+    sdft= streamdf(_SIGV/220.,progenitor=Orbit(obs),pot=lp,aA=aAI,
+                   leading=False,nTrackChunks=_NTRACKCHUNKS,
+                   vsun=[0.,30.24*8.,0.],
+                   tdisrupt=4.5/bovy_conversion.time_in_Gyr(220.,8.),
+                   multi=_NTRACKCHUNKS)
+    #Calculate the density as a function of l, p(l)
+    txs= numpy.linspace(3.,12.4,_NLS)
+    tlogps= multi.parallel_map((lambda x: sdft.callMarg([txs[x]/8.,None,None,None,None,None],
+                                                        interp=True,ngl=_NGL,
+                                                        nsigma=3)),
+                              range(_NLS),
+                              numcores=numpy.amin([_NLS,
+                                                   multiprocessing.cpu_count()]))
+    tlogps= numpy.array(tlogps)
+    tlogps[numpy.isnan(tlogps)]= -100000000000000000.
+    tps= numpy.exp(tlogps-logsumexp(tlogps))
+    tps/= numpy.nansum(tps)*(txs[1]-txs[0])
+    bovy_plot.bovy_print()
+    bovy_plot.bovy_plot(txs,tps,'k-',lw=1.5,
+                        xlabel=r'$X\,(\mathrm{kpc})$',
+                        ylabel=r'$p(X)$',
[…]

### 2013-12-31 py/plot_stream.py

@@ -746,7 +746,7 @@ def plot_stream_aa(plotfilename):
         xrange=[0.,1.3]
         yrange=[0.,5.]
         xlabel= r'$\Large|\Delta \mathbf{\theta}_\parallel\Large|$'
-        ylabel= r'$\Delta t\,(\mathrm{Gyr})$'
+        ylabel= r'$t_s\,(\mathrm{Gyr})$'
         fmt= 'k.'
     bovy_plot.bovy_print()
     bovy_plot.bovy_plot(plotx,ploty,fmt,
@@ -867,7 +867,7 @@ def plot_stream_times(plotfilename):
     if 'hist' in plotfilename:
         bovy_plot.bovy_print()
         bovy_plot.bovy_hist(dts,range=[0.,6.],bins=13,
-                            xlabel=r'$\Delta t\ (\mathrm{Gyr})$',
+                            xlabel=r'$t_s\ (\mathrm{Gyr})$',
                             histtype='step',color='k',normed=True,lw=2.)
         bovy_plot.bovy_hist(dts,range=[0.,6.],bins=61,normed=True,
                             overplot=True,ls='dotted',histtype='step',
@@ -905,7 +905,7 @@ def plot_stream_times(plotfilename):
                             'k.',
                             xrange=[0.,6.],
                             yrange=[0.1,0.35],
-                            xlabel=r'$\Delta t\ (\mathrm{Gyr})$',
+                            xlabel=r'$t_s\ (\mathrm{Gyr})$',
                             ylabel=r'$|\Delta \mathbf{\Omega}|\ (\mathrm{Gyr}^{-1})$')
     elif 'da' in plotfilename:
         bovy_plot.bovy_print()
@@ -914,7 +914,7 @@ def plot_stream_times(plotfilename):
                             'k.',
                             xrange=[0.,6.],
                             yrange=[0.,1.5],
-                            xlabel=r'$\Delta t\ (\mathrm{Gyr})$',
+                            xlabel=r'$t_s\ (\mathrm{Gyr})$',
                             ylabel=r'$|\Delta \boldsymbol\theta|$')
     bovy_plot.bovy_end_print(plotfilename)
     return None

### 2013-12-31 py/plot_pdfs_l.py
streams: plot the PDF in l

@@ -1,0 +1,78 @@
+import sys
+import os, os.path
+import numpy
+from scipy import interpolate
+from scipy.misc import logsumexp
+from galpy.util import bovy_plot, bovy_conversion, multi, bovy_coords
+import multiprocessing
+from galpy import potential
+from galpy.orbit import Orbit
+from galpy.actionAngle_src.actionAngleIsochroneApprox\
+    import actionAngleIsochroneApprox
+from galpy.df_src.streamdf import streamdf
+_STREAMSNAPDIR= '../sim/snaps'
+_STREAMSNAPAADIR= '../sim/snaps_aai'
+_NTRACKCHUNKS= 11
+_SIGV=0.365
+_NLS= 1001
+def plot_pdfs_l(plotfilename):
+    lp= potential.LogarithmicHaloPotential(q=0.9,normalize=1.)
+    aAI= actionAngleIsochroneApprox(b=0.8,pot=lp)
+    obs= numpy.array([1.56148083,0.35081535,-1.15481504,
+                      0.88719443,-0.47713334,0.12019596])
+    sdf= streamdf(_SIGV/220.,progenitor=Orbit(obs),pot=lp,aA=aAI,
+                  leading=True,nTrackChunks=_NTRACKCHUNKS,
+                  vsun=[0.,30.24*8.,0.],
+                  tdisrupt=4.5/bovy_conversion.time_in_Gyr(220.,8.),
+                  multi=_NTRACKCHUNKS)
+    sdft= streamdf(_SIGV/220.,progenitor=Orbit(obs),pot=lp,aA=aAI,
+                   leading=False,nTrackChunks=_NTRACKCHUNKS,
+                   vsun=[0.,30.24*8.,0.],
+                   tdisrupt=4.5/bovy_conversion.time_in_Gyr(220.,8.),
+                   multi=_NTRACKCHUNKS)
+    #Calculate the density as a function of l, p(l)
+    #Sample from sdf
+    llbd= sdf.sample(n=40000,lb=True)
+    tlbd= sdft.sample(n=50000,lb=True)
+    b,e= numpy.histogram(llbd[0],bins=101,normed=True)
+    t= ((numpy.roll(e,1)-e)/2.+e)[1:]
+    lspl= interpolate.UnivariateSpline(t,numpy.log(b),k=3,s=1.)
+    lls= numpy.linspace(t[0],t[-1],_NLS)
[…]

### 2013-12-31 tex/ms.tex
streams: bib tweak
streams: add figure of density distribution and description

@@
 \citep[\eg,][]{Johnston99a,Koposov10a} and its small-scale
-fluctuations \citep{Carlberg09a,Carlberg12a}. However, two factors
+fluctuations \citep{Carlberg12a}. However, two factors
 have hampered the practical use of tidal streams in obtaining
@@
 
+\begin{figure}[t!]
+ \includegraphics[width=0.8\textwidth,clip=]{gd1_pdf_l.ps}
+ \caption{Density distribution as a function of Galactic
+   longitude. The smooth solid line shows the fiducial model's
+   prediction, while the histogram shows the density along the
+   simulated stream. The simple fiducial stream model can accurately
+   predict the density distribution along the
+   stream.}\label{fig:gd1_pdf_l}
+%python plot_pdfs_l.py ../tex/gd1_pdf_l.ps
+\end{figure}
+
 \figurename~\ref{fig:gd1_pdf_xz} shows such PDFs of $p(Z|X)$ for a few
@@
 uncertainties, and to calculate moments of the PDF, \eg, the density
-and velocity dispersion along the stream.
+and velocity dispersion along the stream. Another example is shown in
+\figurename~\ref{fig:gd1_pdf_l}. This Figure shows the model's density
+as a function of Galactic longitude and compares it to that of the
+simulated stream in the final snapshot (\ie, that of
+\figurename~s~\ref{fig:gd1_lbd} and \ref{fig:gd1_lbv}). The model
+density here is calculated using a Monte Carlo mock stream sample,
+drawn as described in \sectionname~\ref{sec:mock}, rather than through
+direct numerical integration.
 
@@
   Bovy,~J., Hogg,~D.~W., \& Roweis,~S.~T. 2011, Ann. Appl. Stat. 5, 1657, arXiv:0905.2979v1
-\bibitem[Carlberg(2009)]{Carlberg09a}
-  Carlberg,~R.~G. 2009, \apj, 705, L223
 \bibitem[Carlberg(2012)]{Carlberg12a}

### 2013-12-31 tex/ms.tex
streams: typo
streams: first section of the discussion

@@
 simulated stream in the final snapshot (\ie, that of
-\figurename~s~\ref{fig:gd1_lbd} and \ref{fig:gd1_lbv}). The model
+\figurename s~\ref{fig:gd1_lbd} and \ref{fig:gd1_lbv}). The model
 density here is calculated using a Monte Carlo mock stream sample,
@@
 
+\section{Discussion}\label{sec:discussion}
+
+\subsection{Dynamical fitting}
+
+There are multiple existing methods for fitting data on tidal streams
+with models for the Milky Way potential, however, none of these allow
+for the likelihood of a model to be evaluated using a well-defined
+generative model. The only exceptions to this are orbit-fitting
+methods, but these use a faulty generative model (see below), and the
+method of \citet{Varghese11a}. The latter is in many ways similar to
+the method proposed here, in that it calculates the stream track and
+uses this as the basis of the stream inference. They calculate the
+track directly in position--velocity space by orbit integration of
+stream members released at the Lagrange points. However, they do not
+relate the observed width of the stream to the velocity dispersion of
+the progenitor---and thus the location of the Lagrange points and the
+position of the track---and it is therefore only weakly constrained
+and has to be assumed. In general, it is straightforward to write down
+generative models of tidal streams in position--velocity space by
+substituting our initial offsets $(\Delta \veco,\Delta \veca)$ by
+similar offsets $(\Delta \vecx,\Delta \vecv)$ and integrating both the
+progenitor and offset stream members forward in time. However,
+evaluating such models when fitting observational data requires large
+numbers of orbit integrations to marginalize over stripping time
+(which cannot be performed analytically in this case) and missing
+phase--space dimensions. Thus, any such method will likely be orders
+of magnitude slower than the method put forth here.
+
+The forward model of this paper is superior to all other
+stream-fitting methods in that it allows easy marginalization over
+missing or noisily measured phase--space coordinates. The obvious
+exception to this is orbit fitting, which as a purely one-dimensional
+model is easy to integrate over. However, orbit fitting likely
+produces biased results as streams do not delineate single orbits over
+their full angle widths (see below). The fiducial model I propose in
+this paper allows for a straightforward replacement of orbit fitting,
+at the expense of adding two extra parameters, $\sigv$ and $t_d$, in
+addition to the six-dimensional position of the progenitor. The orbit
+in orbit fitting then gets replaced by the stream track, calculated as
+in \sectionname~\ref{sec:track}. To constrain $\sigv$ and $t_d$ it is
+necessary to also use the observed width and length of the
+stream. While cold streams are typically too narrow to be resolved in
+$D$ or any of the velocity components, the observed width of the
+stream's sky projection can be measured and used as a
+constraint. Future high-resolution spectroscopic data,
+$\mu\mathrm{as}$-level astrometric data, or percent-level distances
+for standard candles may allow the width and structure of tidal
+streams to be resolved and provide further constraints on the
+generative model. The switch from orbit fitting to stream-track
+fitting comes at the expense of $\approx100$ orbit integrations for
+each model rather than a single integration. If the progenitor is
+unknown we also need to marginalize over whether we are seeing a
+leading or a trailing stream.
+
+The dynamical framework proposed here can be used in situations where
+fitting a stream track is insufficient, for exampe, when high-quality
+data in some dimensions on individual stars are available. Many of the
+approaches in the literature do not require a model progenitor
+position. This includes methods that attempt to minimize the spread of
+orbital energies or actions in the stream (\eg,
+\citealt{Binney08a,Penarrubia12a}, Sanderson \etal, 2014, in
+preparation) as well as the method of \citet{Sanders13b}, which
+constrains the potential by requiring angle differences in the stream
+to lie along the same direction as frequency differences (see
+\equationname~\ref{eq:pdf} and subsequent discussion). These methods
+require good measurements of the phase--space coordinates of stream
+members, or at the very least, \emph{some} measurement of all of the
+six phase--space coordinates (although the latter can be relaxed when
+excellent measurements of the line-of-sight velocity along the stream
+are available; \citealt{Binney08a}). If such high-quality data are
+present, these methods can provide good initial guesses for the PDF of
+potential parameters that fit a stream, which can subsequently be used
+in a more thorough exploration of the PDF and structure of the stream
+using the framework of \sectionname~\ref{sec:pdf}. That these methods
+work also goes to show that the progenitor parameters---phase--space
+position, $\sigv$ and $t_d$---are not that important for the
+constraints on the gravitational potential.
+
 %BOVY: reference point
 
-\section{Discussion}\label{sec:discussion}
+\subsection{Do streams follow orbits?}
+
+\subsection{Constraining progenitor properties with stream kinematics}
+
+\subsection{Baseline models for stream-gap finding}
 
@@
 \bibitem[Tremaine(1999)]{Tremaine99a}
-  Tremaine,~S. 1999, \mnras, 307, 877
+  Tremaine,~S. 1999, \mnras, 307, 877\
+\bibitem[Varghese \etal(2011)]{Varghese11a}
+  Varghese,~A., Ibata,~R., \& Lewis,~G.~F. 2011, \mnras, 417, 198
 \bibitem[Yoon \etal(2011)]{Yoon11a}

### 2013-12-31 tex/ms.tex
streams: do streams follow orbit

@@
 
+\subsection{Do streams follow orbits?}
+
+The answer to this question is clearly no. The real question therefore
+is how bad the approximation that streams follow orbits really is. The
+first question is what is meant by the approximation of streams as a
+single orbit. \citet{Sanders13a} define this as meaning that the
+frequency differences along the orbit lie in the direction of the
+progenitor frequency $\veco_p$. However, this does not faithfully
+represent how orbit fitting works in practice
+\citep[\eg,][]{Koposov10a}. Orbit fitting works by fitting a
+\emph{single} orbit to a stream, such that there are no frequency
+differences along the orbit which can be compared with
+$\veco_p$. Orbit fitting is more correctly represented by stating that
+$t_d$ is infinite in the framework of this paper and using the stream
+track to fit data. In this case, the mean $\Delta \veco$ is constant
+along the stream and the stream track calculated as in
+\sectionname~\ref{sec:track} will be a single orbit.
+
+In our stream model the distribution of $\Delta \veco$ is constant
+with time, such that stream members on average are removed to the same
+orbit. However, the finite disruption time means that for stream
+members to have reached the largest angle offsets from the progenitor,
+they have to have been removed onto an orbit that is separated from
+the progenitor's by a larger $\Delta \veco$ than average. In our model
+this is explicitly expressed by the distribution $p(\Delta
+\opar|\Delta \apar)$, calculated in \equationname~(\ref{eq:poparapar})
+and shown in the top left panel of
+\figurename~\ref{fig:gd1_apar}. Thus, the average orbit changes as a
+function of positon along the stream, whether expressed as $\Delta
+\apar$, $X$, or $l$, and a single-orbit approximation is necessarily
+wrong over the full length of the stream.
+
+It is clear from \figurename~\ref{fig:gd1_apar} that a large part of
+the stream near the progenitor, approximately $\Delta \apar < \Delta
+\Omega^m\,t_d$, is on average on the same orbit, such that it could be
+fit with a single orbit. However, for constraining the gravitational
+potential with a tidal stream, it is essential to use as large an arc
+of the stream as possible, such that limiting the stream to a small
+region around the progenitor to be fit with a single orbit is not
+optimal.
+
+Because the orbit-fitting definition used by \citet{Sanders13a} does
+not simulate real orbit fitting, the conclusions of their analysis as
+regards biases in the recovered potential from the faulty single-orbit
+assumption should be treated with caution. While orbit fitting is
+wrong, how biased inferences based on it are is therefore still an
+open question. However, orbit-fitting-based constraints on the Milky
+Way halo's potential based on streams such as GD-1 or the Orphan
+stream must be treated with suspicion. The fact that the progenitors
+of these streams have not been located means that the observed stream
+segments are likely far from the progenitor, where the average orbit
+changes as a function of angle along the stream (\ie, right of the
+dashed line in \figurename~\ref{fig:gd1_apar}).
+
+\subsection{Constraining progenitor properties with stream kinematics}
+
 %BOVY: reference point
-
-\subsection{Do streams follow orbits?}
-
-\subsection{Constraining progenitor properties with stream kinematics}
 

### 2014-01-01 tex/ms.tex
streams: discussion of stream properties

@@
 
+The generative model for a tidal stream proposed here also connects
+the dynamics of the stream, in particular its average track and the
+dispersion around this track, to the properties of the
+progenitor. Primarily, we can constrain $\sigv$, which is proportional
+to the velocity dispersion or mass$^{1/3}$ of the progenitor (see
+\citealt{Sanders13a}), from the observed width of the stream, even if
+the stream's sky projection is the only projection for which we can
+resolve the stream. Further $N$-body simulations are necessary to
+characterize the exact relation between the \sigv\ parameter and the
+velocity dispersion of the progenitor.
+
+Our calculation of the mean stream orbit as a function of angle along
+the stream also points the way to determining the location of an
+``orphan'' stream's progenitor (an orphan stream is a stream for which
+the progenitor is not know). This requires high-quality data that
+allows the average orbit as a function of position along the stream to
+be determined. If this can be done, then the observation of a steady
+change in the mean orbit followed by a plateau indicates that the
+progenitor lies in the direction of the plateau. The change in the
+mean orbit along the stream is a function of the mean orbit offset
+($\Delta \Omega^m$ in the direction of $\Delta \opar$), the dispersion
+in orbit offsets ($\sigma^2_{\Omega,1}$ in $\Delta \opar$), and the
+disruption time $t_d$. The first two of these can be measured from the
+width of the stream and the plateau in the mean orbit as a function of
+stream position. The disruption time $t_d$ can then be measured from
+the change in mean orbit at the end of the stream. With all of the
+parameters of the forward model measured, we can extrapolate the
+stream backward to the progenitor position. Similarly, we can
+extrapolate a leading stream to its trailing counterpart, if this is
+unknown (and vice versa).
+
+While this requires high-quality data, this should be achievable in
+the near future. We can average or ``stack'' noisily measured
+phase--space coordinates in small segments of the stream to determine
+high quality average phase--space positions and use these instead of
+data on individual stream stars. These considerations show that
+further observations of members of the GD-1 and Orphan streams would
+be highly informative for determining the location or fate of their
+progenitors.
+
+\subsection{Baseline models for stream-gap finding}
+
 %BOVY: reference point
-
-\subsection{Baseline models for stream-gap finding}
 

### 2014-01-01 tex/ms.tex
streams: note

@@
 \subsection{Baseline models for stream-gap finding}
+
+NOTES: stream feathering not included in our model, because of the
+linearization of the dynamics (feathering means it's not linear over
+the width of the stream. When important, can calculate actions and
+angles directly (data good enough -> no need for that much
+marginalization?) or, when samping mock data, use torus machinery to
+calculate (X,V) better.
 

### 2014-01-01 tex/ms.tex
streams: typo

@@
 ($\Delta \Omega^m$ in the direction of $\Delta \opar$), the dispersion
-in orbit offsets ($\sigma^2_{\Omega,1}$ in $\Delta \opar$), and the
+in orbit offsets ($\sigma_{\Omega,1}$ in $\Delta \opar$), and the
 disruption time $t_d$. The first two of these can be measured from the

### 2014-01-02 tex/ms.tex
streams: Sanders initial
streams: gaps section

@@
+%BOVY:NOTES:
+% - GO through citations and be a little more reluctant to cite Sanders
+% - do more justice to Eyre & Binney work
+% - discuss hotter streams somewhere
+% - say more about why dynammical large-scale modeling doesn't care about the small-scale structure
 \documentclass[12pt,preprint]{aastex}
@@
 
-NOTES: stream feathering not included in our model, because of the
-linearization of the dynamics (feathering means it's not linear over
-the width of the stream. When important, can calculate actions and
-angles directly (data good enough -> no need for that much
-marginalization?) or, when samping mock data, use torus machinery to
-calculate (X,V) better.
+Besides allowing for sensitive measurements of the Milky Way's
+large-scale gravitational potential through the wide arcs traced by
+tidal streams, their structure on smaller scales is also sensitive to
+more subtle features of the halo's density distribution. In
+particular, the perturbations from dark--matter subhalos orbiting
+within the Milky Way's halo can have the effect of removing stars from
+small segments of a tidal stream \citep{Yoon11a,Carlberg12a}. Thus,
+the small-scale density structure of tidal streams can be used to
+constrain the subhalo mass function at the high-mass end ($M \gtrsim
+10^5\msun$). This holds a great promise for testing the basic
+predictions of dark--mater clustering in the $\Lambda$CDM framework
+and for shedding light on galaxy formation in the smallest galaxies in
+the Universe.
+
+Tidal streams, even in the absence of perturbations from orbiting
+dark--matter subhalos, are not entirely smooth as the effects of the
+preferential stripping at pericenter and orbital dynamics can create
+over- and underdensities along the observed stream track (see
+\figurename~\ref{fig:gd1_times}; \citealt{Kuepper10a}). These effects
+can lead to spurious detections of gaps in streams that provide a
+source of noise in measurements of gaps due to substructure in the
+halo \citep{Ngan14a}.
+
+The generative stream models proposed in this paper can be useful for
+generating tailor-made background models for stream gap-finding
+algorithms. The generative model allows for the density along the
+stream to be predicted quickly for a given progenitor orbit and
+distribution of stripping times. The example given in
+\figurename~\ref{fig:gd1_pdf_l} using the simple fiducial model shows
+that this works well. However, the fiducial model assumes a uniform
+distribution of stripping times rather than bursts at pericenter
+passages, and therefore it cannot model the under- and over-densities
+along the stream (\eg, ``feathering'') in detail. The
+mock-stream-generation algorithm given in \sectionname~\ref{sec:mock}
+works almost as simply for more complicated models of the stripping
+process (\eg, a bursty distribution of stripping times and more
+energetic stripping at pericenter) and it is therefore straightforward
+to produce more realistic models of the small-scale density structure
+of observed tidal streams. These background models will allow more
+sensitive detections of substructure due to halo substructure.
 
@@
   Koposov,~S.~E., Rix,~H.-W., \& Hogg,~D.~W. 2010, \apj, 712, 260
+\bibitem[K\"{u}pper \etal(2010)]{Kuepper10a}
+  K\"{u}pper,~A.~H.~W., Kroupa,~P., Baumgardt,~H., \& Heggie,~D.~C. 2010, \mnras, 401, 105
 \bibitem[McGill \& Binney(1990)]{McGill90a}
@@
 \bibitem[Sanders \& Binney(2013a)]{Sanders13a}
-  Sanders,~J. \& Binney,~J. 2013a, \mnras, 433, 1813
+  Sanders,~J.~L. \& Binney,~J. 2013a, \mnras, 433, 1813
 \bibitem[Sanders \& Binney(2013b)]{Sanders13b}
-  Sanders,~J. \& Binney,~J. 2013b, \mnras, 433, 1826
+  Sanders,~J.~L. \& Binney,~J. 2013b, \mnras, 433, 1826
 \bibitem[Shoemake(1985)]{Shoemake85a}

### 2014-01-02 tex/ms.tex
streams: conclusion

@@
 % - say more about why dynammical large-scale modeling doesn't care about the small-scale structure
+% - what about spherical potentials? dO/dJ has zero eigenvalue
+% - We/I distinction
 \documentclass[12pt,preprint]{aastex}
@@
 
+\section{Conclusion}\label{sec:conclusion}
+
+I have presented a new method for modeling the dynamics of tidal
+streams, making extensive use of action--angle variables. This new
+framework consists of simple models for the disruption of the stream
+in frequency--angle space, coupled with a novel method for calculating
+action--angle variables for any orbit in any potential. The model is a
+generative model, meaning that it can be used to sample mock stream
+data and evaluate the likelihood of different models for observed
+tidal stream data. I have described fast methods for calculating the
+mean location of the stream in various coordinate systems, for
+estimating the dispersion of the stream and relating it to the
+velocity dispersion of mass of the progenitor, and for marginalizing
+the stream likelihood over noisily or entirely unobserved dimensions
+of phase space.
+
+The framework allows for the proper analysis of data on tidal streams
+by taking into account all the relevant dynamical effects, most
+notably the stream--single-orbit offset. As such, it should be
+immediately useful for the proper analysis of data on the GD-1
+\citep{Koposov10a}, Orphan \citep[\eg,][]{Sesar13a}, Pal 5 and othter
+streams. The ability to quickly marginalize the model over unobserved
+dimensions will also prove highly useful when Gaia data for these
+streams becomes available in the near future. The novel framework also
+provides a straightforward way to generate mock stream data that is
+useful for detecting anomalies in the structure of tidal streams, such
+as those that can be caused by fly-bys of dark--mater subhalos.
+
+Besides being a practical tool for proper stream fitting, the
+framework presented here also clarifies the relation between the
+stream, the potential, and properties of the progenitor. I have
+discussed how we can relate the properties of a stream to those of its
+progenitor, potentially allowing for the unknown progenitor of streams
+such as GD-1 or the Orphan stream to be located. Further
+investigations of the dynamics of cluster disruption and tidal-stream
+generation for different progenitor orbits, a larger and more
+realistic set of potentials, and different progenitor structures will
+be necessary to perfect the models in frequency--angle space proposed
+here and to parameterize the mapping between model parameters and
+actual progenitor properties.
+
+The full modeling and action--angle framework presented here is
+available as part of the \emph{galpy} Galactic Dynamics
+code\footnote{Available at
+  \protect{\url{http://github.com/jobovy/galpy}}~.}. A tutorial on how
+to use this code is given at
+\[
+\protect{\mathrm{\url{http://jobovy.github.io/galpy/}}}\,.
+\]
+BOVY:WRITE AND LINK TO TUTORIAL.
+
 %BOVY: reference point
 
-\section{Conclusion}\label{sec:conclusion}
-
-%BOVY: reference point
-
 \acknowledgements It is a pleasure to thank David W. Hogg, Hans-Walter
-Rix, and Jason Sanders for helpful discussions. J.B. was supported by
-NASA through Hubble Fellowship grant HST-HF-51285.01 from the Space
-Telescope Science Institute, which is operated by the Association of
-Universities for Research in Astronomy, Incorporated, under NASA
-contract NAS5-26555. J.B.  acknowledges support from SFB 881 (A3)
-funded by the German Research Foundation DFG.
+Rix, Jason Sanders, and Scott Tremaine for helpful
+discussions. J.B. was supported by NASA through Hubble Fellowship
+grant HST-HF-51285.01 from the Space Telescope Science Institute,
+which is operated by the Association of Universities for Research in
+Astronomy, Incorporated, under NASA contract NAS5-26555. J.B.
+acknowledges support from SFB 881 (A3) funded by the German Research
+Foundation DFG.
 
@@
   Sanders,~J.~L. \& Binney,~J. 2013b, \mnras, 433, 1826
+\bibitem[Sesar\etal(2013)]{Sesar13a}
+  Sesar,~B., \etal\ 2013, \apj, 776, 26
 \bibitem[Shoemake(1985)]{Shoemake85a}

### 2014-01-02 tex/ms.tex
streams: add reference to discussion section

@@
 \end{equation}
-\Eqnname~(\ref{eq:poparapar}) shows why we chose to model
-$p(\Delta \opar)$ as a Gaussian multiplied with $|\Delta
-\opar|$. $p(\Delta \opar|\Delta \apar)$ is then a Gaussian with mean
-$\Delta \Omega^m$ and variance $\sigma^2_{\Omega,1}$, for $\Delta
-\opar > \Delta \apar / t_d$, and zero otherwise. The mean and variance
-of such a Gaussian is straightforward to calculate in terms of error
-functions. The top left panel of \figurename~\ref{fig:gd1_apar} shows
-the distribution of $|\Delta \opar|$ versus $|\Delta \apar|$ (absolute
-value in order to show the trailing and leading stream together) for
-the simulated stream, as well as the mean and dispersion calculated
-from the fiducial model. The dashed line shows $\Delta \apar = \Delta
-\Omega^m \,t_d$, approximately the angle where the finite disruption
-time starts to influence the mean $\Delta \opar$ of the stream. The
-finite disruption time, $\approx 4.5\Gyr$ in this case, of a tidal
-stream means that for stream members to have reached very large angle
+\Eqnname~(\ref{eq:poparapar}) shows why we chose to model $p(\Delta
+\opar)$ as a Gaussian multiplied with $|\Delta \opar|$. $p(\Delta
+\opar|\Delta \apar)$ is then a Gaussian with mean $\Delta \Omega^m$
+and variance $\sigma^2_{\Omega,1}$, for $\Delta \opar > \Delta \apar /
+t_d$, and zero otherwise. The mean and variance of such a Gaussian is
+straightforward to calculate in terms of error functions. The top left
+panel of \figurename~\ref{fig:gd1_apar} shows the distribution of
+$|\Delta \opar|$ versus $|\Delta \apar|$ (absolute value in order to
+show the trailing and leading stream together) for the simulated
+stream, as well as the mean and dispersion calculated from the
+fiducial model. The dashed line shows $\Delta \apar = \Delta \Omega^m
+\,t_d$, approximately the angle where the finite disruption time
+starts to influence the mean $\Delta \opar$ of the stream. The finite
+disruption time, $\approx 4.5\Gyr$ in this case, of a tidal stream
+means that for stream members to have reached very large angle
 differences with respect to the progenitor, they must have been
@@
 with distance from the progenitor. It is in this sense that streams do
-not follow a single orbit. This is discussed further in BOVY:REF.
+not follow a single orbit. This is discussed further in
+\sectionname~\ref{sec:discussorbit}.
 
@@
 
-\subsection{Do streams follow orbits?}
+\subsection{Do streams follow orbits?}\label{sec:discussorbit}
 

### 2014-01-02 tex/ms.tex
streams: first full draft pre-spell-check

@@
 %BOVY:NOTES:
-% - GO through citations and be a little more reluctant to cite Sanders
-% - do more justice to Eyre & Binney work
-% - discuss hotter streams somewhere
-% - say more about why dynammical large-scale modeling doesn't care about the small-scale structure
-% - what about spherical potentials? dO/dJ has zero eigenvalue
 % - We/I distinction
@@
 efficiently. Until now this has required accurate phase--space data
-\citep{Sanders13b} and specialized algorithms to calculate
-action--angle coordinates that break down for the eccentric orbits on
-which tidal-stream progenitors are typically found (that is, radial
-and/or vertical actions of similar size as the angular momentum;
-\citealt{Sanders12a}). However, even with Gaia, future stream data
-will typically have large uncertainties compared to the intrinsic
-dispersion of the stream and in particular the line-of-sight
-velocities of the faintest stream members will not be observed for
-many stars. Additionally, streams are superimposed on a non-negligible
-background of field stars and background contamination cannot be
-easily taken into account in any of the current stream-fitting
-methods.
+and specialized algorithms to calculate action--angle coordinates that
+break down for the eccentric orbits on which tidal-stream progenitors
+are typically found (that is, radial and/or vertical actions of
+similar size as the angular momentum; \citealt{Sanders12a}). However,
+even with Gaia, future stream data will typically have large
+uncertainties compared to the intrinsic dispersion of the stream and
+in particular the line-of-sight velocities of the faintest stream
+members will not be observed for many stars. Additionally, streams are
+superimposed on a non-negligible background of field stars and
+background contamination cannot be easily taken into account in any of
+the current stream-fitting methods.
 
@@
 through a combination of the internal dynamics of the progenitor and
-the stripping process (see discussion in \citealt{Sanders13b}). For
-dynamical modeling we can be agnostic about the relative importance of
-all of these contributors to $\unitvec \veco$ by flexible modeling of
-the distribution of $\Delta \vecj$ or $\Delta \veco$.
+the stripping process. For dynamical modeling we can be agnostic about
+the relative importance of all of these contributors to $\unitvec
+\veco$ by flexible modeling of the distribution of $\Delta \vecj$ or
+$\Delta \veco$.
 
@@
 dynamical evolution; for the simulated stream used here this happens
-after approximately $20\Myr$). For that reason, we will assume that
-the initial angle offsets are independent of both stripping time and
+after approximately $20\Myr$). For that reason, I will assume that the
+initial angle offsets are independent of both stripping time and
 frequency offset. In what follows, we will model them using a simple
@@
 
+We can model streams in frequency space because the mapping $\vecj
+\rightarrow \veco$ is of full rank for axisymmetric and triaxial
+potentials. However, for spherical potentials this mapping only has
+rank two, because $\Omega_\phi = \mathrm{sgn}(L_Z)\,\Omega_Z$, such
+that the determinant $|\partial \veco / \partial \vecj|$ is zero. For
+spherical potentials, tidal streams only spread in a five-dimensional
+subspace and in modeling stream in spherical potentials we need to
+keep the direction perpendicular to this subspace constant.
+
 \subsection{The fiducial stream model}\label{sec:fidmodel}
@@
 simple models of the structure of tidal streams when considering them
-in frequency--angle space. In this section, we propose a very simple
-fiducial model that we will use in the remainder of this paper. This
+in frequency--angle space. In this section, I propose a very simple
+fiducial model that I will use in the remainder of this paper. This
 model combines simple analytic forms for ease of computation that are
@@
 model that needs to be determined for each stream. For the model of
-the simulated stream used here we set $t_d = 4.5\Gyr$.
+the simulated stream used here I set $t_d = 4.5\Gyr$.
 
@@
 differences at pericentric passages than between them. The fact that
-there is no long-term trend in the typical frequency offset means that
-this is a good model on time-scales larger than the radial
+there is only a small long-term trend in the typical frequency offset
+means that this is a good model on time-scales larger than the radial
 period. While this means that the small-scale structure of the stream
@@
 radii of the progenitor orbit, respectively, and $Z_{\mathrm{max}}$ is
-the maximum $Z$ reached on this orbit. We assume that the all
+the maximum $Z$ reached on this orbit. We assume that all of the
 correlations between the actions are zero (see
@@
 
-We then further constrain the mean frequency offset to lie along the
+I then further constrain the mean frequency offset to lie along the
 principal eigenvector of $V_\veco$. This mean frequency offset can be
@@
 distribution. The fit is equally good (because the dispersion is much
-smaller than the mean) and we choose the second form because it
+smaller than the mean) and I choose the second form because it
 simplifies some of the stream-track calculations below. The best-fit
@@
 
-Because we expect the typical $\Delta\veco$ to scale with the
-velocity dispersion of the progenitor (see also \citealt{Sanders13a}),
-I conjecture that a constant $\mu_\Omega$ can be used for modeling
-cold tidal streams of any (small-ish) velocity dispersion. Further
-modeling of simulated streams, however, is necessary to check this and
-to find the optimizal value of $\mu_\Omega$. Fixing $\mu_\Omega$
-leaves a single free parameter \sigv\ for modeling the frequency
-offset distribution of a tidal stream. In modeling to observed data,
-this parameter needs to be fit and we expect it to be proportional to
-the velocity dispersion of the progenitor, although this
-proportionality needs to be checked more carefully (but see Figure 3
-in \citealt{Sanders13a}, which shows that the size of the frequency
-distribution scales as mass$^{1/3}$, or approximately as the velocity
-dispersion). Similarly, the initial angle distribution, modeled here
-as an isotropic Gaussian, has a charactaristic spread $\sigma_{\theta}$
-which we set here to $0.003$ ($=\sigv/[122\kms]$), and we likewise
-expect this spread to scale with the velocity dispersion of the
-progenitor.
+Because we expect the typical $\Delta\veco$ to scale with the velocity
+dispersion of the progenitor, I conjecture that a constant
+$\mu_\Omega$ can be used for modeling cold tidal streams of any
+(small-ish) velocity dispersion. Further modeling of simulated
+streams, however, is necessary to check this and to find the optimizal
+value of $\mu_\Omega$. Fixing $\mu_\Omega$ leaves a single free
+parameter \sigv\ for modeling the frequency offset distribution of a
+tidal stream. In modeling to observed data, this parameter needs to be
+fit and we expect it to be proportional to the velocity dispersion of
+the progenitor, although this proportionality needs to be checked more
+carefully (but see Figure 3 in \citealt{Sanders13a}, which shows that
+the size of the frequency distribution scales as mass$^{1/3}$, or
+approximately as the velocity dispersion). Similarly, the initial
+angle distribution, modeled here as an isotropic Gaussian, has a
+charactaristic spread $\sigma_{\theta}$ which I set here to $0.003$
+($=\sigv/[122\kms]$), and we likewise expect this spread to scale with
+the velocity dispersion of the progenitor.
 
@@
 leave $\mu_\Omega$ and $\sigma_{\theta}$ as free parameters as
-well. In what follows, we will fix these to the values given in the
+well. In what follows, I will fix these to the values given in the
 previous paragraph. It must be stressed, however, that all of these
@@
 \end{equation}
-where in the penultimate step we have approximated the initial angle
+where in the penultimate step I have approximated the initial angle
 distribution as a delta function (as the initial angle offset is small
@@
 \end{equation}
-\Eqnname~(\ref{eq:poparapar}) shows why we chose to model $p(\Delta
+\Eqnname~(\ref{eq:poparapar}) shows why I chose to model $p(\Delta
 \opar)$ as a Gaussian multiplied with $|\Delta \opar|$. $p(\Delta
@@
 \end{equation}
-where in the last step we have again approximated the initial
+where in the last step I have again approximated the initial
 distribution of parallel angle offsets as a delta function. The first
@@
 perpendicular direction (corresponding to the middle and smallest
-eigenvalue of $V_\veco$). We have \emph{not} approximated the initial
+eigenvalue of $V_\veco$). I have \emph{not} approximated the initial
 angle-offset distribution as a delta function here, because its width
@@
 similar equations as those given in the previous paragraphs, but for a
-simple estimate we can approximate these as $0.5$; we set the
+simple estimate we can approximate these as $0.5$; I set the
 correlation between $\Delta \opar$ and $\Delta \apar$ to zero. While
@@
 position--velocity coordinates by inverting the transformation
-$(\vecx,\vecv) \rightarrow (\veco,\veca)$. We do this by linearizing
+$(\vecx,\vecv) \rightarrow (\veco,\veca)$. I do this by linearizing
 the transformation along the orbit of the progenitor in the vicinity
@@
 found by simple linear algebra for any desired number of expansion
-terms $\sin(\vecn\cdot\veca^A)$. For extra stability, we integrate the
+terms $\sin(\vecn\cdot\veca^A)$. For extra stability, I integrate the
 orbit both forward and backward in time, such that the desired angle
@@
 $\sin(\vecn\cdot\veca^A)$ is small, such that a small number of
-expansion terms is sufficient (we limit the terms in this axisymmetric
+expansion terms is sufficient (I limit the terms in this axisymmetric
 case to $n_R < 4$ and $n_Z < 4$).

### 2014-01-02 tex/ms.tex
streams: first draft pre-reading it over

@@
-%BOVY:NOTES:
-% - We/I distinction
 \documentclass[12pt,preprint]{aastex}
@@
 angle along the stream, both in action--angle coordinates and in
-poition--velocity coordinates. \sectionname~\ref{sec:mock} discusses
+position--velocity coordinates. \sectionname~\ref{sec:mock} discusses
 how to generate mock stream data using the generative model of a tidal
@@
     \sectionname~\ref{sec:sim} after $5.011\Gyr$ of evolution in
-    Galactocentric $X$ and $Z$ coordinates. The grey dot shows the
+    Galactocentric $X$ and $Z$ coordinates. The gray dot shows the
     current position of the progenitor cluster and the dashed line
@@
 a simple, analytic form (see also the discussion in
-\citealt{Eyre11a}). Especially the ``bowtie'' structure in $(J_R,L_Z)$
+\citealt{Eyre11a}). Especially the ``bow tie'' structure in $(J_R,L_Z)$
 and the overlapping leading- and trailing-arm distributions in
@@
 frequency offsets between stream particles and the progenitor as a
-function of strippng time, corresponding to ingredient (b) above. The
+function of stripping time, corresponding to ingredient (b) above. The
 same burstiness that is apparent in the distribution of stripping
@@
 (small-ish) velocity dispersion. Further modeling of simulated
-streams, however, is necessary to check this and to find the optimizal
+streams, however, is necessary to check this and to find the optimal
 value of $\mu_\Omega$. Fixing $\mu_\Omega$ leaves a single free
@@
 angle distribution, modeled here as an isotropic Gaussian, has a
-charactaristic spread $\sigma_{\theta}$ which I set here to $0.003$
+characteristic spread $\sigma_{\theta}$ which I set here to $0.003$
 ($=\sigv/[122\kms]$), and we likewise expect this spread to scale with
@@
 paper eleven points are used. Each Jacobian calculation requires seven
-frequency--angle calcuations, such that the total number of such
+frequency--angle calculations, such that the total number of such
 computations is less than $100$ to adequately model the stream (each
@@
 > 0$. This distribution is log-concave, that is, the derivative of its
-logarithm is negative everywhere, and therefore it can be efficienly
+logarithm is negative everywhere, and therefore it can be efficiently
 sampled using adaptive-rejection sampling \citep{Gilks92a}. The sign
@@
 of the simulated stream described in
-\sectionname~\ref{sec:fidmodel}. The mock data in this Figure has been
-generated $1.96\Gyr$ before the final snapshot of the simulation (that
-is, $1.96\Gyr$ before \figurename~\ref{fig:gd1_xz}). The model
+\sectionname~\ref{sec:fidmodel}. The mock data in this Figure have
+been generated $1.96\Gyr$ before the final snapshot of the simulation
+(that is, $1.96\Gyr$ before \figurename~\ref{fig:gd1_xz}). The model
 from \sectionname~\ref{sec:fidmodel}, which was for the final
@@
 The dynamical framework proposed here can be used in situations where
-fitting a stream track is insufficient, for exampe, when high-quality
+fitting a stream track is insufficient, for example, when high-quality
 data in some dimensions on individual stars are available. Many of the
@@
 \figurename~\ref{fig:gd1_apar}. Thus, the average orbit changes as a
-function of positon along the stream, whether expressed as $\Delta
+function of position along the stream, whether expressed as $\Delta
 \apar$, $X$, or $l$, and a single-orbit approximation is necessarily
@@
 immediately useful for the proper analysis of data on the GD-1
-\citep{Koposov10a}, Orphan \citep[\eg,][]{Sesar13a}, Pal 5 and othter
+\citep{Koposov10a}, Orphan \citep[\eg,][]{Sesar13a}, Pal 5 and other
 streams. The ability to quickly marginalize the model over unobserved
 dimensions will also prove highly useful when Gaia data for these
-streams becomes available in the near future. The novel framework also
+streams become available in the near future. The novel framework also
 provides a straightforward way to generate mock stream data that is
@@
 Hamiltonian, the oscillating cosine behavior cancels, and the target
-action is obtained as the auxiliary-angle-averaged auxilary
+action is obtained as the auxiliary-angle-averaged auxiliary
 action. That is,

### 2014-01-02 tex/ms.tex
streams: tweaks after reading; first draft

@@
 efficiently. Until now this has required accurate phase--space data
-and specialized algorithms to calculate action--angle coordinates that
-break down for the eccentric orbits on which tidal-stream progenitors
-are typically found (that is, radial and/or vertical actions of
-similar size as the angular momentum; \citealt{Sanders12a}). However,
-even with Gaia, future stream data will typically have large
-uncertainties compared to the intrinsic dispersion of the stream and
-in particular the line-of-sight velocities of the faintest stream
-members will not be observed for many stars. Additionally, streams are
-superimposed on a non-negligible background of field stars and
-background contamination cannot be easily taken into account in any of
-the current stream-fitting methods.
+and specialized algorithms that break down for the eccentric orbits on
+which tidal-stream progenitors are typically found (that is, radial
+and/or vertical actions of similar magnitude as the angular momentum;
+\citealt{Sanders12a}). However, even with Gaia, future stream data
+will typically have large uncertainties compared to the intrinsic
+dispersion of the stream and in particular the line-of-sight
+velocities of the faintest stream members will not be observed for
+many stars. Additionally, streams are superimposed on a non-negligible
+background of field stars and background contamination cannot be
+easily taken into account in any of the current stream-fitting
+methods.
 
@@
 are independent of stripping time. While this model oversimplifies the
-real stripping process, it is argued that at least for constraining
-the gravitational potential with stream data, this simple model is
-likely unbiased. For detailed mock stream data, more elaborate models
-of the stripping process as a function of time might be necessary and
-I discuss how these could easily be incorporated into the proposed
+real stripping process, I argue that at least for constraining the
+gravitational potential with stream data, this simple model is likely
+unbiased. For detailed mock stream data, more elaborate models of the
+stripping process as a function of time might be necessary and I
+discuss how these could easily be incorporated into the proposed
 framework.
@@
 Appendix~\ref{sec:aa}. This transformation is linearized at a small
-number of points along the progenitor orbit that span the width of the
-stream. For cold streams, the separation between the stream track and
-the progenitor orbit as well as the internal differences within the
-stream are small, such that this linear approximation is accurate
+number of points along the progenitor orbit that span the length of
+the stream. For cold streams, the separation between the stream track
+and the progenitor orbit as well as the internal differences within
+the stream are small, such that this linear approximation is accurate
 enough. The linear approximation allows for fast evaluation of the
@@
 
-The dynamics of tidal streams both in real (position--velocity) and
-action--angle space has been described by various previous authors
+The dynamics of tidal streams both in position--velocity and
+action--angle space has been described previously by various authors
 \citep[\eg,][]{Helmi99a,Tremaine99a,Johnston99a,Sanders13b}. I provide
@@
 
-In real space, a tidal stream forms as stars are stripped from a
-progenitor cluster or satellite galaxy some time $\ts$ in the
-past and then evolve (largely) independently in a galaxy's host
+In position--velocity space, a tidal stream forms as stars are
+stripped from a progenitor cluster or satellite galaxy some time $\ts$
+in the past and then evolve (largely) independently in a galaxy's host
 potential, moving away both in position and velocity from the
@@
 $(\vecj^p,\veca^p)$, similar to the small offset ($\Delta \vecx,\Delta
-\vecv)$ in the real--space description. The offset in the actions
-corresponds to an offset in the frequencies $\Delta \veco$;
-ultimately, this frequency offset is responsible for the spreading of
-the stream over long stretches on the sky. Because both the
-progenitor's angles and the star's angles increase linearly with time,
-albeit at different frequencies, the difference in the angles also
-increases linearly in time. Therefore at time $t=0$ the difference in
-angles is
+\vecv)$. The offset in the actions corresponds to an offset in the
+frequencies $\Delta \veco$; ultimately, this frequency offset is
+responsible for the spreading of the stream over long stretches on the
+sky. Because both the progenitor's angles and the star's angles
+increase linearly with time, albeit at different frequencies, the
+difference in the angles also increases linearly in time. Therefore at
+time $t=0$ the difference in angles is
 \begin{equation}
@@
 just determined by the potential (through $H$) and the progenitor
-orbit, but also by the distribution of $\Delta \vecj$, which is set
-through a combination of the internal dynamics of the progenitor and
-the stripping process. For dynamical modeling we can be agnostic about
-the relative importance of all of these contributors to $\unitvec
-\veco$ by flexible modeling of the distribution of $\Delta \vecj$ or
-$\Delta \veco$.
+orbit $\vecj^p$, but also by the distribution of $\Delta \vecj$, which
+is set through a combination of the internal dynamics of the
+progenitor and the stripping process. For dynamical modeling we can be
+agnostic about the relative importance of all of these contributors to
+$\unitvec \veco$ by flexible modeling of the distribution of $\Delta
+\vecj$ or $\Delta \veco$.
 
@@
 ratio of the central potential to the velocity dispersion squared of
-$2.0$ is sampled using $10^4$ particles. The cluster is evolved in an
-external logarithmic host potential with $V_c = 220\kms$ and a
-flattening $q=0.9$ ($\Phi = V_c^2 \ln [R^2 + z^2/q^2]$) using the
-gyrfalcon code \citep{Dehnen00a,Dehnen02a} with a softening of
-$1.5\pc$ in the NEMO toolkit \citep{Teuben95a}. The cluster is evolved
-for $5.011\Gyr$ ($=5.125\kpc/[\kms]$) from the initial condition
-$(X,Y,Z) = (-11.63337239,-20.76235661,-10.631736273934635)\kpc$ and
+$2.0$ is sampled using $10^4$ particles. The cluster is evolved with
+self-gravity in an external logarithmic host potential with $V_c =
+220\kms$ and a flattening $q=0.9$ ($\Phi = V_c^2 \ln [R^2 + z^2/q^2]$)
+using the gyrfalcon code \citep{Dehnen00a,Dehnen02a} with a softening
+of $1.5\pc$ in the NEMO toolkit \citep{Teuben95a}. The cluster is
+evolved for $5.011\Gyr$ ($=5.125\kpc/[\kms]$) from the initial
+condition $(X,Y,Z) =
+(-11.63337239,-20.76235661,-10.631736273934635)\kpc$ and
 $(V_X,V_Y,V_Z) = (-128.8281653,42.88727925,79.172383882274971)\kms$,
@@
 initial angle offsets are independent of both stripping time and
-frequency offset. In what follows, we will model them using a simple
+frequency offset. In what follows, I will model them using a simple
 isotropic Gaussian distribution.
@@
     frequency offsets than those between pericenter passages, there is
-    no secular trend and the typical frequency offset (averaged over a
-    radial period) can be considered constant with time. The right
-    panel shows the magnitude of the angle offset between stream
+    almost no secular trend and the typical frequency offset (averaged
+    over a radial period) can be considered constant with time. The
+    right panel shows the magnitude of the angle offset between stream
     particles and the progenitor as a function of stripping
@@
 offset (at the end of the simulation) versus the stripping time. There
-is a spread in the angle offset reach for any given stripping time,
+is a spread in the angle offset reached for any given stripping time,
 but it is nevertheless typically the case that particles that have
@@
 spherical potentials, tidal streams only spread in a five-dimensional
-subspace and in modeling stream in spherical potentials we need to
+subspace and in modeling streams in spherical potentials we need to
 keep the direction perpendicular to this subspace constant.
@@
 stripping times, indicating that the uniform distribution is a good
-approximation over intervals larger than a radial period ($\approx
+approximation over intervals longer than a radial period ($\approx
 400\Myr$). The \emph{disruption time} $t_d$ is a free parameter of the
@@
     distribution, while the solid line is a fit of our preferred form
-    for this distribution, given in \equationname~(\ref{eq:do}). Both
+    for this distribution, given in \equationname~(\ref{eq:do}) (the
+    dashed line is almost entirely covered by the solid line). Both
     forms provide adequate fits to the
@@
 the largest eigenvalue of $V_\veco$ and $\Delta \Omega^m$ is the
-one-dimensional mean frequency offset. The sign of this parameter sets
+one-dimensional mean frequency offset. The sign of $\mu_\Omega$ sets
 whether we are modeling a leading or a trailing
@@
 well. In what follows, I will fix these to the values given in the
-previous paragraph. It must be stressed, however, that all of these
+previous paragraphs. It must be stressed, however, that all of these
 parameters here have been fixed ``by-eye'', and better fits might be
@@
 
-In this section, I discuss how to calculate the track of a model tidal
+In this Section, I discuss how to calculate the track of a model tidal
 stream in the generative model described in
@@
     while the grayscale bands show the $68\,\%$ and $95\,\%$ limits of
-    the predicted distribution. The top, right panel shows the
-    magnitude of frequency offsets perpendicular to those in the left
-    panel. The predicted distribution is constant as a function of
-    angle and therefore not shown. The bottom, left panel shows the
-    angle offset perpendicular to the stream and the grayscale bands
-    show $68\,\%$ and $95\,\%$ limits of the predicted distribution of
-    angle offsets; the white dashed line shows the $68\,\%$ limit
-    calculated using a simplified method. The bottom, right panel
-    shows the stripping time. The gray line and grayscale bands again
-    show the mean and $68\,\%$ and $95\,\%$ confidence limits of the
-    predicted distribution. While the ``burstiness'' related to the
-    energetic stripping at pericenter that is apparent in the
-    simulated data's distributions is (by design) absent in the model,
-    the overall scale and the mean trends with angle are well
+    the predicted distribution. The dashed line shows the angle $\apar
+    = \Delta \Omega^m\,t_d$, angles beyond which can only be reached
+    by having a larger $\Delta \opar$ than average. The top, right
+    panel shows the magnitude of frequency offsets perpendicular to
+    those in the left panel. The predicted distribution is constant as
+    a function of angle and therefore not shown. The bottom, left
+    panel shows the angle offset perpendicular to the stream and the
+    grayscale bands show $68\,\%$ and $95\,\%$ limits of the predicted
+    distribution of angle offsets; the white dashed line shows the
+    $68\,\%$ limit calculated using a simplified method. The bottom,
+    right panel shows the stripping time. The gray line and grayscale
+    bands again show the mean and $68\,\%$ and $95\,\%$ confidence
+    limits of the predicted distribution. While the ``burstiness''
+    related to the energetic stripping at pericenter that is apparent
+    in the simulated data's distributions is (by design) absent in the
+    model, the overall scale and the mean trends with angle are well
     represented by the model.}\label{fig:gd1_apar}
@@
 angle offsets. Therefore, even though the average frequency offset
-does not change over time (see \figurename~\ref{fig:gd1_times}), the
-average frequency offset, and therefore the average orbit, does change
-with distance from the progenitor. It is in this sense that streams do
-not follow a single orbit. This is discussed further in
+does not change much over time (see \figurename~\ref{fig:gd1_times}),
+the average frequency offset, and therefore the average orbit, does
+change with distance from the progenitor. It is in this sense that
+streams do not follow a single orbit. This is discussed further in
 \sectionname~\ref{sec:discussorbit}.
@@
   & = \int \dd \Delta \opar\,p(\ts|\Delta \opar,\Delta \apar)\,p(\Delta \opar|\Delta \apar)\,,\\
-  & = p\left(\Delta \opar = \frac{\Delta \apar}{\ts}\right)\,\left|\frac{\Delta \apar}{(\ts)^2}\right|\,,
+  & \approx p\left(\Delta \opar = \frac{\Delta \apar}{\ts}\right)\,\left|\frac{\Delta \apar}{(\ts)^2}\right|\,,
 \end{split}
@@
 eigendecomposition and order the eigenvalues by size. We then
-interpolate the values of each of the six eigenvalues using spline
+interpolate each of the six eigenvalues using spline
 interpolation. The direction of each eigenvector is interpolated using
@@
 $(l,b,D,\vlos,\pmll,\pmbb)$ by converting the interpolated track in
-$(\vecx,\vecv)$ to these observable coordinates. The dispersion at
-calculated track points is calculated from that in $(\veco,\veca)$
-using the appropriate Jacobians and the six-dimensional dispersion can
-again be interpolated using the eigendecomposition. This is
-illustrated in the bottom panel of \figurename~\ref{fig:track}.
+$(\vecx,\vecv)$ to these coordinates. The dispersion at calculated
+track points is calculated from that in $(\veco,\veca)$ using the
+appropriate Jacobians and the six-dimensional dispersion can again be
+interpolated using the eigendecomposition. This is illustrated in the
+bottom panel of \figurename~\ref{fig:track}.
 
@@
   \includegraphics[width=0.8\textwidth,clip=]{gd1_evol_ld.eps}
-  \caption{Simulated stream after $5.011\Gyr$ as a function of Galactic
-    longitude. The top panel shows the stream in Galactic latitude and
-    the bottom panel shows the distance from the Sun. The progenitor
-    is shown as the large gray dot and its orbit as the dashed
-    lines. The predicted track of the stream is shown as the solid
-    line. Insets show various zoomed-in parts of the stream to show
-    the stream, progenitor orbit, and predicted stream track up
-    close.}\label{fig:gd1_lbd}
+  \caption{Simulated stream after $5.011\Gyr$ as a function of
+    Galactic longitude. The top panel shows the stream in Galactic
+    latitude and the bottom panel shows it in the distance from the
+    Sun. The progenitor is shown as the large gray dot and its orbit
+    as the dashed lines. The predicted track of the stream is shown as
+    the solid line. Insets show various zoomed-in parts of the stream
+    to show the stream, progenitor orbit, and predicted stream track
+    up close.}\label{fig:gd1_lbd}
 %python plot_stream.py ../tex/gd1_evol_lb.ps
@@
 Sanderson \etal, 2014, in preparation). These approaches ignore the
-correlations between actions and angles in the stream (the exponential
-in \eqnname~(\ref{eq:pdf}), which describes this correlation through
-the frequencies) and only use $p(\veco)$; finding the gravitational
-potential that minimizes the spread in frequencies (or equivalently,
-actions) optimizes $p(\veco)$. Recently, a new method was proposed
-that only uses the correlation between the actions and the angles,
-expressed by the exponential in \eqnname~(\ref{eq:pdf})
-\citep{Sanders13b}; this method does not use the fact that the spread
-in actions in a tidal stream is small (expressed by $p(\veco)$), but
-only uses the fact that the angles and actions/frequencies are highly
-correlated. Approaches that require stream members to re-unite with
-the progenitor when integrating their orbits backward in time
-\citep{Johnston99a,PriceWhelan13a} use the full PDF (but expressed in
+correlations between actions and angles in the stream and only use
+$p(\veco)$; finding the gravitational potential that minimizes the
+spread in frequencies (or equivalently, actions) optimizes
+$p(\veco)$. Recently, a new method was proposed that only uses the
+correlation between the actions and the angles, expressed by the
+exponential in \eqnname~(\ref{eq:pdf}) \citep{Sanders13b}; this method
+does not use the fact that the spread in actions in a tidal stream is
+small (expressed by $p(\veco)$), but only uses the fact that the
+angles and actions/frequencies are highly correlated. Approaches that
+require stream members to re-unite with the progenitor when
+integrating their orbits backward in time
+\citep{Johnston99a,PriceWhelan13a} use the full PDF (expressed in
 position--velocity space), but cannot be marginalized over unobserved
@@
 Similar procedures can be followed to evaluate other projections and
-marginalizations of the PDF, \eg, $p(b|l)$ for the observed track of
-the stream on the sky, to convolve the PDF with observational
+marginalizations of the PDF (\eg, $p(b|l)$ for the observed track of
+the stream on the sky), to convolve the PDF with observational
 uncertainties, and to calculate moments of the PDF, \eg, the density
@@
 as a function of Galactic longitude and compares it to that of the
-simulated stream in the final snapshot (\ie, that of
-\figurename s~\ref{fig:gd1_lbd} and \ref{fig:gd1_lbv}). The model
-density here is calculated using a Monte Carlo mock stream sample,
-drawn as described in \sectionname~\ref{sec:mock}, rather than through
-direct numerical integration.
+simulated stream in the final snapshot (\ie, that of \figurename
+s~\ref{fig:gd1_lbd} and \ref{fig:gd1_lbv}). The model density here is
+calculated using a Monte Carlo mock stream sample, drawn as described
+in \sectionname~\ref{sec:mock}, rather than through direct numerical
+integration.
 
@@
 There are multiple existing methods for fitting data on tidal streams
-with models for the Milky Way potential, however, none of these allow
+with models for the Milky Way potential. However, none of these allow
 for the likelihood of a model to be evaluated using a well-defined
@@
 the progenitor---and thus the location of the Lagrange points and the
-position of the track---and it is therefore only weakly constrained
-and has to be assumed. In general, it is straightforward to write down
+position of the track---which is therefore only weakly constrained and
+has to be assumed. In general, it is straightforward to write down
 generative models of tidal streams in position--velocity space by
@@
 segments are likely far from the progenitor, where the average orbit
-changes as a function of angle along the stream (\ie, right of the
+changes as a function of angle along the stream (\ie, rightward of the
 dashed line in \figurename~\ref{fig:gd1_apar}).
@@
 ``orphan'' stream's progenitor (an orphan stream is a stream for which
-the progenitor is not know). This requires high-quality data that
+the progenitor is not known). This requires high-quality data that
 allows the average orbit as a function of position along the stream to
@@
 distribution of stripping times rather than bursts at pericenter
-passages, and therefore it cannot model the under- and over-densities
-along the stream (\eg, ``feathering'') in detail. The
+passages, and therefore it cannot fully model the under- and
+over-densities along the stream (\eg, ``feathering'') in detail. The
 mock-stream-generation algorithm given in \sectionname~\ref{sec:mock}
@@
 streams, making extensive use of action--angle variables. This new
-framework consists of simple models for the disruption of the stream
-in frequency--angle space, coupled with a novel method for calculating
-action--angle variables for any orbit in any potential. The model is a
-generative model, meaning that it can be used to sample mock stream
-data and evaluate the likelihood of different models for observed
-tidal stream data. I have described fast methods for calculating the
-mean location of the stream in various coordinate systems, for
-estimating the dispersion of the stream and relating it to the
-velocity dispersion of mass of the progenitor, and for marginalizing
-the stream likelihood over noisily or entirely unobserved dimensions
-of phase space.
+framework consists of simple models for the disruption of a star
+cluster in frequency--angle space coupled with a novel method for
+calculating action--angle variables for any orbit in any
+potential. The model is a generative model, meaning that it can be
+used to sample mock stream data and evaluate the likelihood of
+different models for observed tidal stream data. I have described fast
+methods for calculating the mean location of the stream in various
+coordinate systems, for estimating the dispersion of the stream and
+relating it to the velocity dispersion or mass of the progenitor, and
+for marginalizing the stream likelihood over noisily or entirely
+unobserved dimensions of phase space.
 
@@
 immediately useful for the proper analysis of data on the GD-1
-\citep{Koposov10a}, Orphan \citep[\eg,][]{Sesar13a}, Pal 5 and other
+\citep{Koposov10a}, Orphan \citep[\eg,][]{Sesar13a}, Pal 5, and other
 streams. The ability to quickly marginalize the model over unobserved
 dimensions will also prove highly useful when Gaia data for these
-streams become available in the near future. The novel framework also
+streams become available in the near future. The new framework also
 provides a straightforward way to generate mock stream data that is

### 2014-01-04 sim/make_sim_movie.py
sim: plot action distribution of just the debris

@@ -222,7 +222,9 @@ def make_sim_movie(proj='xz',comov=False,skippng=False,
         print "'ffmpeg' failed"
     return None                           
 
-def make_sim_movie_aa(proj='aaarazap',comov=False,skippng=False,
+def make_sim_movie_aa(proj='aaarazap',comov=False,
+                      debris=False,
+                      skippng=False,
                       includeorbit=True,aas=False):
     #Directories
     if aas:
@@ -262,15 +264,34 @@ def make_sim_movie_aa(proj='aaarazap',comov=False,skippng=False,
         elif 'aas' in snapaadir:
             xrange=[1.1,1.5]
             yrange=[3.85,4.3]
+    if debris:
+        basefilename+= 'debris_'
+        moviefilename+= '_debris'
     if not skippng:
         nx= 10000
         nt= len(glob.glob(os.path.join(snapaadir,
                                        'gd1_evol_hitres_aa_*.dat')))
+        if debris:
+            #Load final snapshot first, determine which stars are debris
+            data= numpy.loadtxt(os.path.join(snapaadir,
+                                             'gd1_evol_hitres_aa_%s.dat' % str(1312).zfill(5)),
+                                delimiter=',')
+            thetar= data[:,6]
+            thetar= (numpy.pi+(thetar-numpy.median(thetar))) % (2.*numpy.pi)
+            debrisIndx= numpy.fabs(thetar-numpy.pi) > (5.*numpy.median(numpy.fabs(thetar-numpy.median(thetar))))
         for ii in range(nt):
             #Read data
             data= numpy.loadtxt(os.path.join(snapaadir,
                                        'gd1_evol_hitres_aa_%s.dat' % str(ii).zfill(5)),
                                 delimiter=',')
+            if debris:
+                thetar= data[:,6]
+                thetar= (numpy.pi+(thetar-numpy.median(thetar))) % (2.*numpy.pi)
+                indx= numpy.fabs(thetar-numpy.pi) > (5.*numpy.median(numpy.fabs(thetar-numpy.median(thetar))))
+                indx*= debrisIndx
+                data= data[indx,:]
+                if numpy.sum(indx) == 0:
+                    data= -1000.+numpy.random.uniform(size=(2,9))*50.
             if proj.lower() == 'aaarazap':
                 plotx= data[:,6]
                 ploty= data[:,8]
[…]

### 2014-01-04 tex/ms.tex
streams: tweak about L_Z

@@
 a simple, analytic form (see also the discussion in
-\citealt{Eyre11a}). Especially the ``bow tie'' structure in $(J_R,L_Z)$
-and the overlapping leading- and trailing-arm distributions in
-($J_R,J_Z$) are difficult to model with a simple distribution
-function. The distribution of frequencies, however, is close-to one
-dimensional, its direction of largest variance can be well-modeled as
-a Gaussian (although below we will model it slightly differently), and
-its parameters can all be easily estimated from the velocity
-dispersion of the progenitor, the orbit of the progenitor, and the
-gravitational potential, as discussed below. Therefore, we will model
-tidal streams in frequency--angle space. Because the leading and
-trailing arm are well-separated in frequency space, we model each arm
-individually.
+\citealt{Eyre11a}). Especially the ``bow tie'' structure in
+$(J_R,L_Z)$ and the overlapping leading- and trailing-arm
+distributions in ($J_R,J_Z$) are difficult to model with a simple
+distribution function\footnote{For triaxial potentials, the azimuthal
+  action $J_\phi \neq L_Z$ is the correct action to use instead of
+  $L_Z$. Even though the modeling in this paper also applies to
+  triaxial potentials, we denote $J_\phi$ as $L_Z$, because the
+  simulation uses an axisymmetric potential.}. The distribution of
+frequencies, however, is close-to one dimensional, its direction of
+largest variance can be well-modeled as a Gaussian (although below we
+will model it slightly differently), and its parameters can all be
+easily estimated from the velocity dispersion of the progenitor, the
+orbit of the progenitor, and the gravitational potential, as discussed
+below. Therefore, we will model tidal streams in frequency--angle
+space. Because the leading and trailing arm are well-separated in
+frequency space, we model each arm individually.
 

### 2014-01-06 tex/ms.tex
streams: edit

@@
 characterize the exact relation between the \sigv\ parameter and the
-velocity dispersion of the progenitor.
+velocity dispersion of the progenitor and to determine the importance
+of the cluster's internal structure, primarily its concentration, in
+determining \sigv.
 

### 2014-01-06 sim/make_sim_movie.py
streams: higher mass sim

@@ -355,7 +355,10 @@ def make_sim_movie_aa(proj='aaarazap',comov=False,
     return None
 
 if __name__ == '__main__':
-    if 'aa' in sys.argv[1].lower():
+    if 'oparapar' in sys.argv[1].lower():
+        make_sim_movie_oparapar(proj=sys.argv[1],comov=len(sys.argv) > 2,
+                                debris=len(sys.argv) > 3)
+    elif 'aa' in sys.argv[1].lower():
         make_sim_movie_aa(proj=sys.argv[1],comov=len(sys.argv) > 2,
                           debris=len(sys.argv) > 3)
     else:

### 2014-01-06 tex/ms.tex
streams: tweak to appendix

@@
 desired (that is, a convergence criterion can be applied and
-convergence can be attained). I describe this method briefly here and
-then show how to extend it to also calculate the frequencies and
-angles: $(\vecx,\vecv) \rightarrow (\veco,\veca)$.
+convergence can be attained). I describe the Fox \& Binney method
+briefly here and then show how to extend it to also calculate the
+frequencies and angles: $(\vecx,\vecv) \rightarrow (\veco,\veca)$.
 
@@
 \end{equation}
-As long as the auxiliary angle $\theta_R$ goes through the full range
+As long as the auxiliary angle $\theta^A_R$ goes through the full range
 $0$ to $2\pi$ along the orbit, the oscillatory behavior integrates to
@@
 \end{equation}
-This a linear system of equations for each point along the orbit with
-unknowns $\veca(t=0)$, $\veco(\vecj)$, and the functions
+This is a linear system of equations for each point along the orbit
+with unknowns $\veca(t=0)$, $\veco(\vecj)$, and the functions
 $\frac{\partial S_{\vecn}(\vecj)}{\partial \vecj}$. For a finely
@@
 
-The method described here is completely general, but the only obstacle
+The method described here is completely general and the only obstacle
 to using it is that a sufficiently close isochrone potential has to be
@@
 isochrone potential can be determined quickly for Milky-Way-like
-axisymmetric potentials, although it remains to be checked whether
-this holds for triaxial potentials as well. For analyzing stream data,
-the necessity of this trial-and-error determination of the auxiliary
-potential does not pose a practical problem, as all of the stream
-stars are on very similar orbits. Therefore, the same auxiliary
-potential can be used for all action--angle calculations for a given
-potential and for different potentials as well, as long as they are
-not too different. More widespread use of this technique may require
-automated methods for determining a good auxiliary potential. In the
-unlikely circumstance that no good auxiliary isochrone potential can
-be found, the method discussed here can also be used with other
-auxiliary potentials for which the actions and angles can be
-calculated, such as other spherical potentials or the family of
-St\"{a}ckel potentials. This will, however, add a non-negligible
-amount of computation for the necessary numerical integrations.
+axisymmetric potentials in regions where the flattening is mild (\ie,
+in the halo). Whether this method works near the plane of the disk
+where the potential is strongly flattened remains to be checked,
+although the fact that disk orbits are close to circular makes the
+details of the auxiliary potential important. Similarly, the method
+should be tested in more detail for triaxial potentials as well. For
+analyzing stream data, the necessity of this trial-and-error
+determination of the auxiliary potential does not pose a practical
+problem, as all of the stream stars are on very similar
+orbits. Therefore, the same auxiliary potential can be used for all
+action--angle calculations for a given potential and for different
+potentials as well, as long as they are not too different. More
+widespread use of this technique may require automated methods for
+determining a good auxiliary potential. In the unlikely circumstance
+that no good auxiliary isochrone potential can be found, the method
+discussed here can also be used with other auxiliary potentials for
+which the actions and angles can be calculated, such as other
+spherical potentials or the family of St\"{a}ckel potentials. This
+will, however, add a non-negligible amount of computation for the
+necessary numerical integrations.
 

### 2014-01-06 tex/ms.tex
streams: some of Scott's comments

@@
 \citep[\eg,][]{Johnston99a,Koposov10a} and its small-scale
-fluctuations \citep{Carlberg12a}. However, two factors
-have hampered the practical use of tidal streams in obtaining
-constraints on the MW gravitational potential. First, streams do not
-trace a single orbit, which has led to confusion over how to best fit
-streams and over how problematic the single-orbit approximation really
-is \citep{Eyre11a,Sanders13a}. Second, approaches that go beyond the
+fluctuations \citep{Carlberg12a}. However, two factors have hampered
+the practical use of tidal streams in obtaining constraints on the MW
+gravitational potential. First, streams do not trace a single orbit,
+which has led to confusion over how to best fit streams and over how
+problematic the single-orbit approximation really is
+\citep{Eyre11a,Sanders13a}. Second, approaches that go beyond the
 single-orbit assumption encounter practical difficulties of
@@
 along a tidal stream in order to constrain the number of encounters
-with dark satellites and their masses \citep{Ngan14a}. Fast,
-generative models of tidal streams would greatly benefit both of these
-applications.
+with dark satellites and their masses \citep{Ngan14a}. Simple,
+analytic models of how tidal streams are generated---\emph{generative}
+models---would greatly benefit both of these applications.
 
@@
 and/or vertical actions of similar magnitude as the angular momentum;
-\citealt{Sanders12a}). However, even with Gaia, future stream data
-will typically have large uncertainties compared to the intrinsic
+\citealt{Sanders12a}). However, even with Gaia \citep{Perryman01a},
+all of the phase--space coordinates except for the sky position will
+typically have non-negligible uncertainties compared to the intrinsic
 dispersion of the stream and in particular the line-of-sight
@@
 offsets between stream members and the progenitor, and a uniform
-distribution of stripping times; the frequency and angle distributions
-are independent of stripping time. While this model oversimplifies the
-real stripping process, I argue that at least for constraining the
-gravitational potential with stream data, this simple model is likely
-unbiased. For detailed mock stream data, more elaborate models of the
-stripping process as a function of time might be necessary and I
-discuss how these could easily be incorporated into the proposed
-framework.
+distribution of stripping times; the frequency and angle offset
+distributions are independent of stripping time. While this model
+oversimplifies the real stripping process, I argue that at least for
+constraining the gravitational potential with stream data, this simple
+model is likely unbiased. For detailed mock stream data, more
+elaborate models of the stripping process as a function of time might
+be necessary and I discuss how these could easily be incorporated into
+the proposed framework.
 
@@
 \sectionname~\ref{sec:discussion} and I conclude in
-\sectionname~\ref{sec:conclusion}. In \appendixname~\ref{sec:aa} a
-novel method for calculating actions, frequencies, and angles for any
-orbit in any static gravitational potential is presented.
+\sectionname~\ref{sec:conclusion}. In \appendixname~\ref{sec:aa} I
+review the method of Fox \& Binney (2014, in preparation) for
+calculating actions in any static gravitational potential and I show
+how it can be extended to compute frequencies and angles.
 
@@
 In position--velocity space, a tidal stream forms as stars are
-stripped from a progenitor cluster or satellite galaxy some time $\ts$
-in the past and then evolve (largely) independently in a galaxy's host
-potential, moving away both in position and velocity from the
+stripped from a progenitor cluster or satellite galaxy some time $\ts
+> 0$ in the past and then evolve (largely) independently in a galaxy's
+host potential, moving away both in position and velocity from the
 progenitor object as time goes on. Thus, a star was offset by a small
@@
 as the progenitor until the present day $(t=0)$, when it was observed:
-$(\vecx,\vecv)(t=0)$ = $\vech_{\ts}(\vecx^p(-ts)+\Delta
+$(\vecx,\vecv)(t=0)$ = $\vech_{\ts}(\vecx^p(-\ts)+\Delta
 \vecx,\vecv^p(-\ts)+\Delta \vecv)$, where $\vech$ denotes the
@@
 that used in \citet{Sanders13b}. The stream is constructed by running
-an $N$-body simulation of a King cluster on an orbit similar to that
-of the GD-1 stream \citep{Koposov10a}. Specifically, a King cluster
-with a mass of $2\times10^4\msun$, a tidal radius of $0.07\kpc$, and a
-ratio of the central potential to the velocity dispersion squared of
-$2.0$ is sampled using $10^4$ particles. The cluster is evolved with
-self-gravity in an external logarithmic host potential with $V_c =
-220\kms$ and a flattening $q=0.9$ ($\Phi = V_c^2 \ln [R^2 + z^2/q^2]$)
-using the gyrfalcon code \citep{Dehnen00a,Dehnen02a} with a softening
-of $1.5\pc$ in the NEMO toolkit \citep{Teuben95a}. The cluster is
-evolved for $5.011\Gyr$ ($=5.125\kpc/[\kms]$) from the initial
-condition $(X,Y,Z) =
+an $N$-body simulation of a King cluster \citep{King66a} on an orbit
+similar to that of the GD-1 stream \citep{Koposov10a}. Specifically, a
+King cluster with a mass of $2\times10^4\msun$, a tidal radius of
+$0.07\kpc$, and a ratio of the central potential to the velocity
+dispersion squared of $2.0$ is sampled using $10^4$ particles. The
+cluster is evolved with self-gravity in an external logarithmic host
+potential with $V_c = 220\kms$ and a flattening $q=0.9$ ($\Phi = V_c^2
+\ln [R^2 + z^2/q^2]$) using the gyrfalcon code
+\citep{Dehnen00a,Dehnen02a} with a softening of $1.5\pc$ in the NEMO
+toolkit \citep{Teuben95a}. The cluster is evolved for $5.011\Gyr$
+($=5.125\kpc/[\kms]$) from the initial condition $(X,Y,Z) =
 (-11.63337239,-20.76235661,-10.631736273934635)\kpc$ and
@@
     using the method described in \sectionname~\ref{sec:track}. The
-    inset shows a zoomed-in version of part of the leading stream that
-    more clearly shows the offset between the stream track and the
-    progenitor orbit.}\label{fig:gd1_xz}
+    inset shows a zoomed-in version of part of the leading stream
+    (stretched in $X$ and $Z$) that more clearly shows the offset
+    between the stream track and the progenitor
+    orbit.}\label{fig:gd1_xz}
 %python plot_stream.py ../tex/gd1_evol_xz.ps
@@
 function of stream angle, need to model this non-uniform stripping
-process.
+process. In what follows, we will use a uniform distribution of
+stripping times. This is appropriate for modeling the large-scale
+structure of streams, such as what is used when constraining the
+gravitational potential.
 
@@
 smaller frequency differences and thus to erase correlations between
-particle position along the stream and stripping time---happens to a
-small extent but the narrow distribution of frequency offsets limits
-the ability of particles removed later to reach large distances from
-the progenitor.
+particle position along the stream and stripping time and sort the
+stream in frequency offset---happens to a small extent but the narrow
+distribution of frequency offsets limits the ability of particles
+removed later to reach large distances from the progenitor.
 
@@
 spherical potentials, tidal streams only spread in a five-dimensional
-subspace and in modeling streams in spherical potentials we need to
+subspace and when modeling streams in spherical potentials we need to
 keep the direction perpendicular to this subspace constant.
@@
 deviations
-\begin{equation}
-  \sigma_{J_R} = \frac{1}{\pi}\,\sigv\,\left(r_{\mathrm{apo}}-r_{\mathrm{peri}}\right)\,,\quad 
-  \sigma_{L_Z} = \sigv\,r_{\mathrm{peri}}\,,\quad 
-  \sigma_{J_Z} = \frac{2}{\pi}\,\sigv\,Z_{\mathrm{max}}\,,
-\end{equation}
+\begin{align}
+  \sigma_{J_R} & = \frac{1}{2\,\pi}\oint\dd r\,p_r \approx \frac{1}{\pi}\,\sigv\,\left(r_{\mathrm{apo}}-r_{\mathrm{peri}}\right)\,,\\
+  \sigma_{L_Z} & \approx  \sigv\,r_{\mathrm{peri}}\,,\\
+  \sigma_{J_Z} & = \frac{1}{\pi}\oint\dd z\,p_z \approx \frac{2}{\pi}\,\sigv\,Z_{\mathrm{max}}\,,
+\end{align}
 where $\sigv$ is the velocity dispersion of the progenitor,
@@
   Johnston,~K.~V., Zhao,~H., Spergel,~D.~N., \& Hernquist,~L. 1999, \apj, 512, L109
+\bibitem[King(1966)]{King66a}
+  King,~I.~R. 1966, \aj, 71, 64
 \bibitem[Koposov \etal(2010)]{Koposov10a}
@@
   Pe{\~n}arrubia, J., Koposov,~S.~E., \& Walker,~M.~G. 2012, \apj, 760, 2
+\bibitem[Perryman \etal(2001)]{Perryman01a}
+  Perryman,~M.~A.~C., \etal\ 2001, \aap, 369, 339
 \bibitem[Price-Whelan \& Johnston(2013)]{PriceWhelan13a}

### 2014-01-06 tex/ms.tex
streams: Scott's comments

@@
 approximate the \emph{action} distribution as a Gaussian with standard
-deviations
+deviations given by the approximate spread in the actions $(\delta
+J_R,\delta L_Z,\delta J_Z)$ in the cluster
 \begin{align}
-  \sigma_{J_R} & = \frac{1}{2\,\pi}\oint\dd r\,p_r \approx \frac{1}{\pi}\,\sigv\,\left(r_{\mathrm{apo}}-r_{\mathrm{peri}}\right)\,,\\
-  \sigma_{L_Z} & \approx  \sigv\,r_{\mathrm{peri}}\,,\\
-  \sigma_{J_Z} & = \frac{1}{\pi}\oint\dd z\,p_z \approx \frac{2}{\pi}\,\sigv\,Z_{\mathrm{max}}\,,
+  \sigma_{J_R} & \approx \delta \left(\frac{1}{2\,\pi}\oint\dd r\,p_r\right) \approx \frac{1}{\pi}\,\sigv\,\left(r_{\mathrm{apo}}-r_{\mathrm{peri}}\right)\,,\\
+  \sigma_{L_Z} & \approx  \delta L_Z \approx \sigv\,r_{\mathrm{peri}}\,,\\
+  \sigma_{J_Z} & \approx \delta\left(\frac{1}{\pi}\oint\dd z\,p_z \right)\approx \frac{2}{\pi}\,\sigv\,Z_{\mathrm{max}}\,,
 \end{align}
@@
 eigenvector of $V_\veco$ and the progenitor's frequency vector is
-$0.50^\circ$ while the value measured from the simulation is
-$0.18^\circ$. For a model with an isotropic action distribution, the
-misalignment would be $1.28^\circ$.
+$0.50^\circ$ which is close to the value measured from the simulation
+is $0.18^\circ$. For a model with an isotropic action distribution,
+the misalignment would be $1.28^\circ$, much larger than the measured
+value.
 
@@
 
-The forward model of this paper is superior to all other
-stream-fitting methods in that it allows easy marginalization over
-missing or noisily measured phase--space coordinates. The obvious
-exception to this is orbit fitting, which as a purely one-dimensional
-model is easy to integrate over. However, orbit fitting likely
-produces biased results as streams do not delineate single orbits over
-their full angle widths (see below). The fiducial model I propose in
-this paper allows for a straightforward replacement of orbit fitting,
-at the expense of adding two extra parameters, $\sigv$ and $t_d$, in
-addition to the six-dimensional position of the progenitor. The orbit
-in orbit fitting then gets replaced by the stream track, calculated as
-in \sectionname~\ref{sec:track}. To constrain $\sigv$ and $t_d$ it is
+The forward model of this paper is superior to other stream-fitting
+methods in that it allows easy marginalization over missing or noisily
+measured phase--space coordinates. The obvious exception to this is
+orbit fitting, which as a purely one-dimensional model is easy to
+integrate over. However, orbit fitting likely produces biased results
+as streams do not delineate single orbits over their full angle widths
+(see below). The fiducial model I propose in this paper allows for a
+straightforward replacement of orbit fitting, at the expense of adding
+two extra parameters, $\sigv$ and $t_d$, in addition to the
+six-dimensional position of the progenitor. The orbit in orbit fitting
+then gets replaced by the stream track, calculated as in
+\sectionname~\ref{sec:track}. To constrain $\sigv$ and $t_d$ it is
 necessary to also use the observed width and length of the
@@
 
-While this requires high-quality data, this should be achievable in
-the near future. We can average or ``stack'' noisily measured
-phase--space coordinates in small segments of the stream to determine
-high quality average phase--space positions and use these instead of
-data on individual stream stars. These considerations show that
-further observations of members of the GD-1 and Orphan streams would
-be highly informative for determining the location or fate of their
-progenitors.
+While determining the mean orbit as a function of position along the
+stream requires high-quality data, this should be achievable in the
+near future. We can average or ``stack'' noisily measured phase--space
+coordinates in small segments of the stream to determine high quality
+average phase--space positions and use these instead of data on
+individual stream stars. These considerations show that further
+observations of members of the GD-1 and Orphan streams would be highly
+informative for determining the location or fate of their progenitors.
 

### 2014-01-07 tex/ms.tex
streams: re-format to ApJ

@@
-\documentclass[12pt,preprint]{aastex}
-%\documentclass{emulateapj}
+%\documentclass[12pt,preprint]{aastex}
+\documentclass{emulateapj}
 \usepackage{amssymb,amsmath}
@@
 
-%\submitted{}
+\submitted{}
 
@@
 \affil{Institute for Advanced Study, Einstein Drive, Princeton, NJ 08540, USA}
-\email{bovy@ias.edu}
+\email{Email: bovy@ias.edu}
 \altaffiltext{\theaddress}{\label{Hubble}\stepcounter{address} Hubble fellow}
@@
 toolkit \citep{Teuben95a}. The cluster is evolved for $5.011\Gyr$
-($=5.125\kpc/[\kms]$) from the initial condition $(X,Y,Z) =
-(-11.63337239,-20.76235661,-10.631736273934635)\kpc$ and
-$(V_X,V_Y,V_Z) = (-128.8281653,42.88727925,79.172383882274971)\kms$,
+($=5.125\kpc/[\kms]$) from the initial condition $(X,Y,Z) =$
+$(-11.63337239$ $,-20.76235661,$ $-10.631736273934635)\kpc$ and
+$(V_X,V_Y,V_Z) = $ $(-128.8281653$ $,42.88727925,$ $79.172383882274971)\kms$,
 chosen such that the cluster ends up at roughly the observed location
@@
 
-\begin{figure}[t!]
+\begin{figure*}[t!]
   \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aajrjz.ps}
@@
 %python plot_stream.py ../tex/gd1_evol_aaarap.ps
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[t!]
+\begin{figure*}[t!]
   \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_timeshist.ps}
@@
 %python plot_stream.py ../tex/gd1_evol_timesda.ps
-\end{figure}
+\end{figure*}
 
@@
 \begin{figure}[t!]
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aadohist.ps}
+  \includegraphics[width=0.49\textwidth,clip=]{gd1_evol_aadohist.ps}
   \caption{Distribution of the magnitude of frequency offsets between
@@
 
-\begin{figure}[t!]
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaaparopar.ps}
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaaparoperp.ps}\\
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaaparaperp.ps}
-  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaapartime.ps}
+\begin{figure*}[t!]
+\begin{center}
+  \includegraphics[width=0.36\textwidth,clip=]{gd1_evol_aaaparopar.ps}
+  \includegraphics[width=0.36\textwidth,clip=]{gd1_evol_aaaparoperp.ps}\\
+  \includegraphics[width=0.36\textwidth,clip=]{gd1_evol_aaaparaperp.ps}
+  \includegraphics[width=0.36\textwidth,clip=]{gd1_evol_aaapartime.ps}
   \caption{Stream properties as a function of angle along the
@@
 %python plot_stream.py ../tex/gd1_evol_aaapartime.ps
-\end{figure}
+\end{center}
+\end{figure*}
 
@@
 \begin{equation}
-  p(\Delta \aperpii|\Delta \apar) = \int \dd \ts \,\dd \Delta
-  \operpii \,p(\Delta \aperpii|\Delta \operpii,\ts)\,p(\Delta \operpii|\ts,\Delta \apar)\,p(\ts|\Delta \apar)\,,
+\begin{split}
+  p(\Delta \aperpii|\Delta \apar) & = \int \dd \ts \,\dd \Delta
+  \operpii \,p(\Delta \aperpii|\Delta \operpii,\ts)\\
+  & \qquad \times p(\Delta \operpii|\ts,\Delta \apar)\,p(\ts|\Delta \apar)\,,
+\end{split}
 \end{equation}
@@
 \begin{equation}\label{eq:paperpapar}
-  p(\Delta \aperpii|\Delta \apar) = \int \dd \ts \,\mathcal{N}\left(\Delta \aperpii|0,\sigma_{\theta}^2+(\ts\,\sigma_{\Omega,ii})^2\right)\,p(\ts|\Delta \apar)\,,
+\begin{split}
+  p(\Delta \aperpii|\Delta \apar) & = \int \dd \ts \,\mathcal{N}\left(\Delta \aperpii|0,\sigma_{\theta}^2+(\ts\,\sigma_{\Omega,ii})^2\right)\qquad\\
+& \qquad \times p(\ts|\Delta \apar)\,,
+\end{split}
 \end{equation}
@@
 
-\begin{figure}
+\begin{figure*}
+\begin{center}
   \includegraphics[width=0.8\textwidth,clip=]{track_aa.eps}\\
@@
 %python illustrate_track.py ../tex/track_aa.ps ../tex/track_xz.ps ../tex/track_ld.ps
-\end{figure}
+\end{center}
+\end{figure*}
+
+\begin{figure*}[t!]
+\begin{center}
+  \includegraphics[width=0.8\textwidth,clip=]{gd1_evol_lb.eps}
+  \includegraphics[width=0.8\textwidth,clip=]{gd1_evol_ld.eps}
+  \caption{Simulated stream after $5.011\Gyr$ as a function of
+    Galactic longitude. The top panel shows the stream in Galactic
+    latitude and the bottom panel shows it in the distance from the
+    Sun. The progenitor is shown as the large gray dot and its orbit
+    as the dashed lines. The predicted track of the stream is shown as
+    the solid line. Insets show various zoomed-in parts of the stream
+    to show the stream, progenitor orbit, and predicted stream track
+    up close.}\label{fig:gd1_lbd}
+%python plot_stream.py ../tex/gd1_evol_lb.ps
+%python plot_stream.py ../tex/gd1_evol_ld.ps
+\end{center}
+\end{figure*}
+
+\begin{figure*}[tp!!]
+\begin{center}
+  \includegraphics[width=0.8\textwidth,clip=]{gd1_evol_lvlos.eps}
+  \includegraphics[width=0.8\textwidth,clip=]{gd1_evol_lpmll.eps}
+  \includegraphics[width=0.8\textwidth,clip=]{gd1_evol_lpmbb.eps}
+  \caption{Simulated stream velocities as a function of Galactic
+    longitude. Same as \figurename~\ref{fig:gd1_lbd}, but for the
+    stream in heliocentric line-of-sight velocity $V_{\mathrm{los}}$
+    (top panel), proper motion in the direction of Galactic longitude
+    (middle panel), and in the direction of Galactic latitude (bottom
+    panel). Insets again show various interesting parts of the stream
+    up close.}\label{fig:gd1_lbv}
+%python plot_stream.py ../tex/gd1_evol_lvlos.ps
+%python plot_stream.py ../tex/gd1_evol_lpmll.ps
+%python plot_stream.py ../tex/gd1_evol_lpmbb.ps
+\end{center}
+\end{figure*}
 
@@
 
-\begin{figure}[t!]
-  \includegraphics[width=0.8\textwidth,clip=]{gd1_evol_lb.eps}
-  \includegraphics[width=0.8\textwidth,clip=]{gd1_evol_ld.eps}
-  \caption{Simulated stream after $5.011\Gyr$ as a function of
-    Galactic longitude. The top panel shows the stream in Galactic
-    latitude and the bottom panel shows it in the distance from the
-    Sun. The progenitor is shown as the large gray dot and its orbit
-    as the dashed lines. The predicted track of the stream is shown as
-    the solid line. Insets show various zoomed-in parts of the stream
-    to show the stream, progenitor orbit, and predicted stream track
-    up close.}\label{fig:gd1_lbd}
-%python plot_stream.py ../tex/gd1_evol_lb.ps
-%python plot_stream.py ../tex/gd1_evol_ld.ps
-\end{figure}
-
-\begin{figure}[tp!!]
-  \includegraphics[width=0.8\textwidth,clip=]{gd1_evol_lvlos.eps}
-  \includegraphics[width=0.8\textwidth,clip=]{gd1_evol_lpmll.eps}
-  \includegraphics[width=0.8\textwidth,clip=]{gd1_evol_lpmbb.eps}
-  \caption{Simulated stream velocities as a function of Galactic
-    longitude. Same as \figurename~\ref{fig:gd1_lbd}, but for the
-    stream in heliocentric line-of-sight velocity $V_{\mathrm{los}}$
-    (top panel), proper motion in the direction of Galactic longitude
-    (middle panel), and in the direction of Galactic latitude (bottom
-    panel). Insets again show various interesting parts of the stream
-    up close.}\label{fig:gd1_lbv}
-%python plot_stream.py ../tex/gd1_evol_lvlos.ps
-%python plot_stream.py ../tex/gd1_evol_lpmll.ps
-%python plot_stream.py ../tex/gd1_evol_lpmbb.ps
-\end{figure}
-
 The average stream location computed in the manner described in this
@@
 \begin{figure}[t!]
- \includegraphics[width=0.48\textwidth,clip=]{gd1_evol_2Gya_xz.ps}
+ \includegraphics[width=0.48\textwidth,clip=]{gd1_evol_2Gya_xz.ps}\\
  \includegraphics[width=0.48\textwidth,clip=]{gd1_evol_2Gya_xz_sim.ps}
@@
 
+\begin{figure*}[t!]
+ \includegraphics[width=0.32\textwidth,clip=]{gd1_pdf_xz_0.ps}
+ \includegraphics[width=0.32\textwidth,clip=]{gd1_pdf_xz_1.ps}
+ \includegraphics[width=0.32\textwidth,clip=]{gd1_pdf_xz_2.ps}
+ \caption{Marginalized, conditional $p(Z|X)$ PDFs. The left panel
+   shows the model leading stream (see \figurename~\ref{fig:gd1_xz}),
+   with the stream track shown as the solid line, the estimated
+   $2\sigma$ width of the stream as the dash-dotted line, and the
+   progenitor orbit as the dashed line. The solid line in the middle
+   and right panels shows $p(Z|X)$ calculated at the $X$ indicated in
+   the left panel by thick vertical lines; the dash-dotted line is a
+   simple Gaussian estimate. The upper branch near $Z \approx 5\kpc$
+   of the stream is chosen for the middle panel. The right panel shows
+   an $X$ close to where the stream turns around in $X$; the PDF has
+   multiple peaks in this case.}\label{fig:gd1_pdf_xz}
+%python plot_pdfs_xz.py ../tex/gd1_pdf_xz_0.ps ../tex/gd1_pdf_xz_1.ps ../tex/gd1_pdf_xz_2.ps
+\end{figure*}
+
 The stripping time $t_s$ cannot be observed and should therefore be
@@
 \begin{split}
-  p(\veco,\veca) & = \int \dd t_s\,p(\veco,\veca,t_s)\,,\\
+  p(& \veco,\veca)\\
+  & = \int \dd t_s\,p(\veco,\veca,t_s)\,,\\
   & = p(\veco)\,\int_0^{t_d} \dd t_s\,
-  \frac{1}{(2\,\pi)^{3/2}\,\sigma_\theta^3}\,\exp\left(-\frac{1}{2\,\sigma_\theta^2}\left[\sum_{R,\phi,Z}\left(\Delta \veca_i-\Delta \veco_i\,t_s\right)^2\right]\right)\,,
+  \frac{1}{(2\,\pi)^{3/2}\,\sigma_\theta^3}\\
+  & \qquad \qquad \times\exp\left(-\frac{1}{2\,\sigma_\theta^2}\left[\sum_{R,\phi,Z}\left(\Delta \veca_i-\Delta \veco_i\,t_s\right)^2\right]\right)\,,
 \end{split}
@@
 \begin{split}
-  p(\veco,\veca) & = 
-  p(\veco) \,  \frac{\mathrm{erf}(a_0) + \mathrm{erf}(a_d)}{4\,\pi\,\sigma_\theta^2\,|\Delta \veco|}\,\exp\left(-\frac{1}{2\,\sigma_\theta^2}\left[\sum_{R,\phi,Z}\Delta \veca^2_i-\frac{\left(\sum_{R,\phi,Z} \Delta \veco_i\Delta \veca_i\right)^2}{|\Delta \veco|^2}\right]\right)\,.
+  p(& \veco,\veca)  = 
+  p(\veco) \,  \frac{\mathrm{erf}(a_0) + \mathrm{erf}(a_d)}{4\,\pi\,\sigma_\theta^2\,|\Delta \veco|}\\
+  & \ \times \exp\left(-\frac{1}{2\,\sigma_\theta^2}\left[\sum_{R,\phi,Z}\Delta \veca^2_i-\frac{\left(\sum_{R,\phi,Z} \Delta \veco_i\Delta \veca_i\right)^2}{|\Delta \veco|^2}\right]\right)\,.
 \end{split}
@@
 
-\begin{figure}[t!]
- \includegraphics[width=0.32\textwidth,clip=]{gd1_pdf_xz_0.ps}
- \includegraphics[width=0.32\textwidth,clip=]{gd1_pdf_xz_1.ps}
- \includegraphics[width=0.32\textwidth,clip=]{gd1_pdf_xz_2.ps}
- \caption{Marginalized, conditional $p(Z|X)$ PDFs. The left panel
-   shows the model leading stream (see \figurename~\ref{fig:gd1_xz}),
-   with the stream track shown as the solid line, the estimated
-   $2\sigma$ width of the stream as the dash-dotted line, and the
-   progenitor orbit as the dashed line. The solid line in the middle
-   and right panels shows $p(Z|X)$ calculated at the $X$ indicated in
-   the left panel by thick vertical lines; the dash-dotted line is a
-   simple Gaussian estimate. The upper branch near $Z \approx 5\kpc$
-   of the stream is chosen for the middle panel. The right panel shows
-   an $X$ close to where the stream turns around in $X$; the PDF has
-   multiple peaks in this case.}\label{fig:gd1_pdf_xz}
-%python plot_pdfs_xz.py ../tex/gd1_pdf_xz_0.ps ../tex/gd1_pdf_xz_1.ps ../tex/gd1_pdf_xz_2.ps
-\end{figure}
-
 We can marginalize \equationname~(\ref{eq:pxvt}) over $t_s$ and find
@@
 
-\begin{figure}[t!]
+\begin{figure*}[t!]
+\begin{center}
  \includegraphics[width=0.8\textwidth,clip=]{gd1_pdf_l.ps}
@@
 %python plot_pdfs_l.py ../tex/gd1_pdf_l.ps
-\end{figure}
+\end{center}
+\end{figure*}
 

### 2014-01-07 py/plot_stream.py
streams: scales in insets

@@ -168,11 +168,13 @@ def plot_stream_lb(plotfilename):
         sdf= streamdf(_SIGV/220.,progenitor=Orbit(obs),pot=lp,aA=aAI,
                       leading=True,nTrackChunks=_NTRACKCHUNKS,
                       vsun=[0.,30.24*8.,0.],
-                      tdisrupt=4.5/bovy_conversion.time_in_Gyr(220.,8.))
+                      tdisrupt=4.5/bovy_conversion.time_in_Gyr(220.,8.),
+                      multi=_NTRACKCHUNKS)
         sdft= streamdf(_SIGV/220.,progenitor=Orbit(obs),pot=lp,aA=aAI,
                        leading=False,nTrackChunks=_NTRACKCHUNKS,
                        vsun=[0.,30.24*8.,0.],
-                       tdisrupt=4.5/bovy_conversion.time_in_Gyr(220.,8.))
+                       tdisrupt=4.5/bovy_conversion.time_in_Gyr(220.,8.),
+                       multi=_NTRACKCHUNKS)
     #Plot
     bovy_plot.bovy_print(fig_width=8.25,fig_height=3.5)
     if 'ld' in plotfilename:
@@ -250,6 +252,11 @@ def plot_stream_lb(plotfilename):
                                 overplot=True)
             sdf.plotTrack(d1=d1,d2=d2,interp=True,color='k',spread=0,
                            overplot=True,lw=1.)
+            #Plot approximate scale
+            bovy_plot.bovy_plot([240.,240.],[250.,275.],'k-',lw=2.,
+                                overplot=True)
+            bovy_plot.bovy_text(241.,255.,r'$25\,\mathrm{km\,s}^{-1}$',
+                                size=16.)
             nullfmt   = NullFormatter()         # no labels
             insetAxes.xaxis.set_major_formatter(nullfmt)
             insetAxes.yaxis.set_major_formatter(nullfmt)
@@ -272,6 +279,11 @@ def plot_stream_lb(plotfilename):
                                 overplot=True)
             sdf.plotTrack(d1=d1,d2=d2,interp=True,color='k',spread=0,
                            overplot=True,lw=1.)
+            #Plot approximate scale
+            bovy_plot.bovy_plot([168.5,168.5],[10.75,11.25],'k-',lw=2.,
+                                overplot=True)
+            bovy_plot.bovy_text(169.8,10.875,r'$0.5\,\mathrm{mas\,yr}^{-1}$',
[…]

### 2014-01-07 py/plot_stream.py
streams: tweaks to labeling

@@ -406,7 +406,7 @@ def plot_stream_lb(plotfilename):
             #Plot approximate scale
             bovy_plot.bovy_plot([74.,74.],[11.95,12.45],'k-',lw=2.,
                                 overplot=True)
-            bovy_plot.bovy_text(76.,11.9,r'$0.5\,\mathrm{kpc}$',
+            bovy_plot.bovy_text(76.,12.01,r'$0.5\,\mathrm{kpc}$',
                                 size=16.)
             nullfmt   = NullFormatter()         # no labels
             insetAxes.xaxis.set_major_formatter(nullfmt)
@@ -433,7 +433,7 @@ def plot_stream_lb(plotfilename):
             #Plot approximate scale
             bovy_plot.bovy_plot([115.,115.],[48.5,49.5],'k-',lw=2.,
                                 overplot=True)
-            bovy_plot.bovy_text(117.2,48.4,r'$1^\circ$',
+            bovy_plot.bovy_text(117.2,48.5,r'$1^\circ$',
                                 size=16.)
             nullfmt   = NullFormatter()         # no labels
             insetAxes.xaxis.set_major_formatter(nullfmt)

### 2014-01-07 tex/ms.tex
streams: cut outs of some figures

@@
 \begin{figure}[t!]
-  \includegraphics[width=0.5\textwidth,clip=]{gd1_evol_xz.ps}
+  \includegraphics[width=0.48\textwidth,clip=]{gd1_evol_xz.eps}
   \caption{Mock tidal stream simulated as described in
@@
 \begin{figure}[t!]
-  \includegraphics[width=0.49\textwidth,clip=]{gd1_evol_aadohist.ps}
+  \includegraphics[width=0.48\textwidth,clip=]{gd1_evol_aadohist.eps}
   \caption{Distribution of the magnitude of frequency offsets between
@@
 \begin{figure}[t!]
- \includegraphics[width=0.48\textwidth,clip=]{gd1_evol_2Gya_xz.ps}\\
- \includegraphics[width=0.48\textwidth,clip=]{gd1_evol_2Gya_xz_sim.ps}
+ \includegraphics[width=0.48\textwidth,clip=]{gd1_evol_2Gya_xz.eps}\\
+ \includegraphics[width=0.48\textwidth,clip=]{gd1_evol_2Gya_xz_sim.eps}
   \caption{Mock stream data generated from the model. The left panel
@@
 \begin{figure*}[t!]
- \includegraphics[width=0.32\textwidth,clip=]{gd1_pdf_xz_0.ps}
- \includegraphics[width=0.32\textwidth,clip=]{gd1_pdf_xz_1.ps}
- \includegraphics[width=0.32\textwidth,clip=]{gd1_pdf_xz_2.ps}
+ \includegraphics[width=0.32\textwidth,clip=]{gd1_pdf_xz_0.eps}
+ \includegraphics[width=0.32\textwidth,clip=]{gd1_pdf_xz_1.eps}
+ \includegraphics[width=0.32\textwidth,clip=]{gd1_pdf_xz_2.eps}
  \caption{Marginalized, conditional $p(Z|X)$ PDFs. The left panel
@@
 \begin{figure}[t!]
- \includegraphics[width=0.48\textwidth,clip=]{aAI_jr.ps}
- \includegraphics[width=0.48\textwidth,clip=]{aAI_araz.ps}
+ \includegraphics[width=0.48\textwidth,clip=]{aAI_jr.eps}
+ \includegraphics[width=0.48\textwidth,clip=]{aAI_araz.eps}
   \caption{Action--angle calculations using an orbit-integration-based

### 2014-01-07 sim/make_sim_movie.py
streams: make apar opar movie

@@ -1,10 +1,11 @@
 import sys
 import os, os.path
+import copy
 import glob
 import subprocess
 import numpy
 from numpy import linalg
-from galpy.util import bovy_plot, bovy_coords
+from galpy.util import bovy_plot, bovy_coords, bovy_conversion
 from galpy import potential
 from galpy.orbit import Orbit
 def make_sim_movie(proj='xz',comov=False,skippng=False,
@@ -354,10 +355,138 @@ def make_sim_movie_aa(proj='aaarazap',comov=False,
         print "'ffmpeg' failed"
     return None
 
+def make_sim_movie_oparapar(proj='aaarazap',
+                            skippng=True,#False,
+                            aas=False):
+    #Directories
+    if aas:
+        snapaadir= 'snaps_aas/'
+        aastr= 'aas'
+    else:
+        snapaadir= 'snaps_aai/'
+        aastr= 'aai'
+    savedirpng= './movies/gd1_aa/pngs/'
+    basefilename= 'gd1_evol_%s_' % aastr
+    moviefilename= 'gd1_evol_%s' % aastr
+    if True:
+        basefilename+= 'oparapar_'
+        moviefilename+= '_oparapar'
+        xlabel=r'$|\theta_\parallel|$'
+        ylabel=r'$|\Omega_\parallel|$'
+        xrange=[0.,1.3]
+        yrange=[0.1,0.3]
+        zrange=[0.,4.5]
+    if not skippng:
+        nt= len(glob.glob(os.path.join(snapaadir,
+                                       'gd1_evol_hitres_aa_*.dat')))
+        #Load final snapshot first, determine which stars are debris and when they were stripped
+        data= numpy.loadtxt(os.path.join(snapaadir,
+                                         'gd1_evol_hitres_aa_%s.dat' % str(1312).zfill(5)),
+                            delimiter=',')
+        thetar= data[:,6]
+        thetar= (numpy.pi+(thetar-numpy.median(thetar))) % (2.*numpy.pi)
+        debrisIndx= numpy.fabs(thetar-numpy.pi) > (5.*numpy.median(numpy.fabs(thetar-numpy.median(thetar))))
+        thetar= thetar[debrisIndx]
+        #Calculate times at which stars were stripped, angles
+        thetap= data[:,7]
+        thetap= (numpy.pi+(thetap-numpy.median(thetap))) % (2.*numpy.pi)
+        thetap= thetap[debrisIndx]
+        thetaz= data[:,8]
[…]

### 2014-01-07 sim/make_sim_movie.py
streams: final tweaks for opar apar movie

@@ -356,7 +356,7 @@ def make_sim_movie_aa(proj='aaarazap',comov=False,
     return None
 
 def make_sim_movie_oparapar(proj='aaarazap',
-                            skippng=True,#False,
+                            skippng=False,
                             aas=False):
     #Directories
     if aas:
@@ -371,21 +371,21 @@ def make_sim_movie_oparapar(proj='aaarazap',
     if True:
         basefilename+= 'oparapar_'
         moviefilename+= '_oparapar'
-        xlabel=r'$|\theta_\parallel|$'
-        ylabel=r'$|\Omega_\parallel|$'
+        xlabel=r'$|\Delta\theta_\parallel|$'
+        ylabel=r'$|\Delta\Omega_\parallel|$'
         xrange=[0.,1.3]
         yrange=[0.1,0.3]
         zrange=[0.,4.5]
     if not skippng:
         nt= len(glob.glob(os.path.join(snapaadir,
                                        'gd1_evol_hitres_aa_*.dat')))
-        #Load final snapshot first, determine which stars are debris and when they were stripped
+        #Load one of the final snapshots first, determine which stars are debris and when they were stripped
         data= numpy.loadtxt(os.path.join(snapaadir,
-                                         'gd1_evol_hitres_aa_%s.dat' % str(1312).zfill(5)),
+                                         'gd1_evol_hitres_aa_%s.dat' % str(1287).zfill(5)),
                             delimiter=',')
         thetar= data[:,6]
         thetar= (numpy.pi+(thetar-numpy.median(thetar))) % (2.*numpy.pi)
-        debrisIndx= numpy.fabs(thetar-numpy.pi) > (5.*numpy.median(numpy.fabs(thetar-numpy.median(thetar))))
+        debrisIndx= numpy.fabs(thetar-numpy.pi) > (6.*numpy.median(numpy.fabs(thetar-numpy.median(thetar))))
         thetar= thetar[debrisIndx]
         #Calculate times at which stars were stripped, angles
         thetap= data[:,7]
@@ -418,19 +418,19 @@ def make_sim_movie_oparapar(proj='aaarazap',
             pass
         dOdir= numpy.median(dO4dir,axis=1)
         dOdir/= numpy.sqrt(numpy.sum(dOdir**2.))
-        for ii in range(115,nt):#Skip the first 100, because nothing happens anyway
+        for ii in range(80,nt):#Skip the first 100, because nothing happens anyway
             #Read data
[…]

### 2014-01-07 tex/ms.tex
streams: move a figure, remove a galpy link
streams: add link

@@
 
+Ingredients (a) and (b) require more careful modeling. The right panel
+of \figurename~\ref{fig:gd1_times} shows the distribution of times at
+which stream particles are stripped from the progenitor. These
+stripping times are calculated from the final snapshot by linear
+regression of a particle's angle offset $\Delta \veca$ from the
+progenitor, versus the frequency offset $\Delta \veco$ as 
+\begin{equation}
+  \ts = \frac{\Delta \veco \cdot \Delta \veca}{|\Delta \veco|^2}\,.
+\end{equation}
+The dotted line shows a finely-binned histogram of these stripping
+times, which indicates that stripping happens in bursts with a period
+of about $400\Myr$, the radial period of the progenitor's orbit. Thus,
+as expected, stripping happens primarily at pericenter passages, with
+a smaller number of particles lost between pericenter
+passages. Realistic models of a stream, especially those that require
+good models of the surface-density structure of the stream as a
+function of stream angle, need to model this non-uniform stripping
+process. In what follows, we will use a uniform distribution of
+stripping times. This is appropriate for modeling the large-scale
+structure of streams, such as what is used when constraining the
+gravitational potential.
+
+The middle panel of \figurename~\ref{fig:gd1_times} shows the
+frequency offsets between stream particles and the progenitor as a
+function of stripping time, corresponding to ingredient (b) above. The
+same burstiness that is apparent in the distribution of stripping
+times shows up here. Particles stripped at pericenter typically have
+larger frequency offsets than those lost at larger radii. This can be
+described as saying that particles are \emph{removed} at pericenter,
+but only \emph{peeled off} between pericenter passages. Again,
+realistic models of tidal streams need to take this detailed structure
+into account.
+
+The right panel of \figurename~\ref{fig:gd1_times} shows the angle
+offset (at the end of the simulation) versus the stripping time. There
+is a spread in the angle offset reached for any given stripping time,
+but it is nevertheless typically the case that particles that have
+been stripped earlier are found at larger angle offsets
+today. Self-sorting---the tendency of particles stripped later with
+larger frequency offsets to overtake particles stripped earlier at
+smaller frequency differences and thus to erase correlations between
+particle position along the stream and stripping time and sort the
+stream in frequency offset---happens to a small extent but the narrow
+distribution of frequency offsets limits the ability of particles
+removed later to reach large distances from the progenitor.
+
 \begin{figure*}[t!]
@@
 \end{figure*}
-
-Ingredients (a) and (b) require more careful modeling. The right panel
-of \figurename~\ref{fig:gd1_times} shows the distribution of times at
-which stream particles are stripped from the progenitor. These
-stripping times are calculated from the final snapshot by linear
-regression of a particle's angle offset $\Delta \veca$ from the
-progenitor, versus the frequency offset $\Delta \veco$ as 
-\begin{equation}
-  \ts = \frac{\Delta \veco \cdot \Delta \veca}{|\Delta \veco|^2}\,.
-\end{equation}
-The dotted line shows a finely-binned histogram of these stripping
-times, which indicates that stripping happens in bursts with a period
-of about $400\Myr$, the radial period of the progenitor's orbit. Thus,
-as expected, stripping happens primarily at pericenter passages, with
-a smaller number of particles lost between pericenter
-passages. Realistic models of a stream, especially those that require
-good models of the surface-density structure of the stream as a
-function of stream angle, need to model this non-uniform stripping
-process. In what follows, we will use a uniform distribution of
-stripping times. This is appropriate for modeling the large-scale
-structure of streams, such as what is used when constraining the
-gravitational potential.
-
-The middle panel of \figurename~\ref{fig:gd1_times} shows the
-frequency offsets between stream particles and the progenitor as a
-function of stripping time, corresponding to ingredient (b) above. The
-same burstiness that is apparent in the distribution of stripping
-times shows up here. Particles stripped at pericenter typically have
-larger frequency offsets than those lost at larger radii. This can be
-described as saying that particles are \emph{removed} at pericenter,
-but only \emph{peeled off} between pericenter passages. Again,
-realistic models of tidal streams need to take this detailed structure
-into account.
-
-The right panel of \figurename~\ref{fig:gd1_times} shows the angle
-offset (at the end of the simulation) versus the stripping time. There
-is a spread in the angle offset reached for any given stripping time,
-but it is nevertheless typically the case that particles that have
-been stripped earlier are found at larger angle offsets
-today. Self-sorting---the tendency of particles stripped later with
-larger frequency offsets to overtake particles stripped earlier at
-smaller frequency differences and thus to erase correlations between
-particle position along the stream and stripping time and sort the
-stream in frequency offset---happens to a small extent but the narrow
-distribution of frequency offsets limits the ability of particles
-removed later to reach large distances from the progenitor.
 
@@
 \[
-\protect{\mathrm{\url{http://jobovy.github.io/galpy/}}}\,.
+\protect{\mathrm{\url{http://jobovy.github.io/galpy/streamdf.html}}}\,.
 \]
-BOVY:WRITE AND LINK TO TUTORIAL.
 
@@
 The action--angle method described in this Appendix is implemented in
-the \emph{galpy} Galactic dynamics code\footnote{Available at
-  \protect{\url{http://github.com/jobovy/galpy}}~.}.
+the \emph{galpy} Galactic dynamics code.
 

### 2014-01-07 py/plot_stream_hisigv.py
streams: higher sigv figure

@@ -1,0 +1,88 @@
+import sys
+import os, os.path
+import copy
+import numpy
+from scipy import optimize, special
+from galpy.util import bovy_plot, bovy_coords, bovy_conversion
+from galpy import potential
+from galpy.orbit import Orbit
+from galpy.actionAngle_src.actionAngleIsochroneApprox\
+    import actionAngleIsochroneApprox
+from galpy.df_src.streamdf import streamdf
+from matplotlib import pyplot
+from matplotlib.ticker import NullFormatter
+_STREAMSNAPDIR= '../sim/snaps-hisigv'
+_STREAMSNAPAADIR= '../sim/snaps_aai'
+_NTRACKCHUNKS= 64
+_SIGV=0.365*10. #x 10 because stream is 10x hotter
+def plot_stream_xz(plotfilename):
+    #Read stream
+    data= numpy.loadtxt(os.path.join(_STREAMSNAPDIR,'gd1-hisigv_evol_00041.dat'),
+                        delimiter=',')
+    includeorbit= True
+    if includeorbit:
+        npts= 201
+        pot= potential.LogarithmicHaloPotential(normalize=1.,q=0.9)
+        pts= numpy.linspace(0.,17.,npts)
+        #Calculate progenitor orbit around this point
+        pox= numpy.median(data[:,1])
+        poy= numpy.median(data[:,3])
+        poz= numpy.median(data[:,2])
+        povx= numpy.median(data[:,4])
+        povy= numpy.median(data[:,6])
+        povz= numpy.median(data[:,5])
+        pR,pphi,pZ= bovy_coords.rect_to_cyl(pox,poy,poz)
+        pvR,pvT,pvZ= bovy_coords.rect_to_cyl_vec(povx,povy,povz,pR,
+                                                 pphi,pZ,cyl=True)
+        ppo= Orbit([pR/8.,pvR/220.,pvT/220.,pZ/8.,pvZ/220.,pphi])
+        pno= Orbit([pR/8.,-pvR/220.,-pvT/220.,pZ/8.,-pvZ/220.,pphi])
+        ppo.integrate(pts,pot)
+        pno.integrate(pts,pot)
+        pvec= numpy.zeros((3,npts*2-1))
+        pvec[0,:npts-1]= pno.x(pts)[::-1][:-1]
+        pvec[1,:npts-1]= pno.z(pts)[::-1][:-1]
+        pvec[2,:npts-1]= pno.y(pts)[::-1][:-1]
[…]

### 2014-01-07 tex/ms.tex
streams: description of new figure

@@
 \section{Discussion}\label{sec:discussion}
+
+\begin{figure}[t!]
+  \includegraphics[width=0.48\textwidth,clip=]{gd1-hisigv_evol_xz.eps}
+  \caption{Same as \figurename~\ref{fig:gd1_xz}, but for a simulation
+    of a progenitor with a mass $M^p$ that is $1,000$ times larger and
+    therefore a velocity dispersion $\sigma_v^p$ that is ten times
+    larger. The model used to predict the stream track (solid line) is
+    the same as that used in \figurename~\ref{fig:gd1_xz}, except that
+    the model velocity dispersion \sigv\ has been increased by a
+    factor of ten. The stream spreads over a much larger volume for a
+    heavier progenitor, but the fiducial model still accurately
+    predicts its location in phase space.}\label{fig:gd1-hisigv_xz}
+%python plot_stream_hisigv.py ../tex/gd1-hisigv_evol_xz.ps
+\end{figure}
+
+\subsection{At what progenitor mass does the framework break down?}
+
+I have illustrated the new framework for modeling tidal streams using
+a cold, narrow stream with a progenitor velocity dispersion of only
+$\sigma_v^p = 1.4\kms$ (and a progenitor mass of $M^p = 2 \times
+10^4\msun$). For a tidal stream that is as cold as that, the offset
+between the progenitor's orbit and the stream track as well as the
+dispersion within the stream are small (we expect these relative
+offsets to be approximately $\sigv/V_c$). Therefore, we can linearize
+the $(\vecx,\vecv) \leftrightarrow (\veco,\veca)$ transformation over
+the relevant range of $(\vecx,\vecv)$.  The stream dispersion and the
+stream--progenitor-orbit offset scale approximately as mass$^{1/3}$,
+such that for streams arising from the tidal disruption of small dwarf
+galaxies, we expect the stream--progenitor-orbit offset to be around
+$10\,\%$. In the fiducial model and for the simulated stream used in
+this paper, the internal frequency dispersion in the stream is six
+times smaller than the stream--progenitor-orbit offset. Therefore,
+bridging the latter with a linearized frequency--angle transformation
+is the more stringent constraint.
+
+To test whether the framework presented applies to streams arising
+from the tidal disruption of small dwarf galaxies, I have run the same
+simulation as that described in \sectionname~\ref{sec:sim}, but for a
+progenitor mass of $M^p=2\times 10^7\msun$ and a velocity dispersion
+of $\sigma_v^p=14\kms$. The tidal radius of this progenitor is
+$0.69\kpc$ and I use a softening parameter of $14\pc$ in this
+case. The final snapshot of this simulation in Galactocentric $X$ and
+$Z$ coordinates is shown in \figurename~\ref{fig:gd1-hisigv_xz}. I
+predict the stream track for this simulated stream by taking the same
+fiducial model as decribed in \sectionname~\ref{sec:fidmodel}, except
+that the model velocity dispersion \sigv\ is increased by a factor of
+ten. This track is shown in \figurename~\ref{fig:gd1-hisigv_xz}.
+
+It is clear from \figurename~\ref{fig:gd1-hisigv_xz} that the fiducial
+model as well as the action--angle approximations used in this paper
+still work for heavier progenitors with masses $\approx
+10^7\msun$. Similar experiments with progenitors with masses $\approx
+10^8\msun$ show that the framework presented here can model the tidal
+streams produced by such heavy progenitor's as well, although the
+stream--progenitor-orbit offset becomes so large that the algorithm
+for calculating the stream track in $(\vecx,\vecv)$ described in
+\sectionname~\ref{sec:trackxv} may need to be improved. When the
+stream--progenitor-orbit becomes so large that the assumption of a
+linear $(\vecx,\vecv) \leftrightarrow (\veco,\veca)$ transformation
+becomes invalid, we can improve the calculation of the stream track by
+iterating the procedure described in \sectionname~\ref{sec:trackxv}
+and illustrated in \figurename~\ref{fig:track}. After calculating the
+stream track, we can calculate the $(\vecx,\vecv) \rightarrow
+(\veco,\veca)$ transformation at the estimated track points and use
+this mapping to refine the estimated trackpoints. A single iteration
+of this procedure will likely suffice. With this procedure, streams
+originating from progenitors up to masses of $10^9\msun$ should be
+able to be modeled with the framework presented in this paper.
+
+Above $M^p = 10^9\msun$ the effects of dynamical friction, the dwarf
+galaxy's gravity, and non-linearity in the $\vecj \rightarrow \veco$
+transformation are expected to start affecting the structure and
+location of the stream \citep[\eg,][]{Sanders13a}. Therefore, streams
+originating from the heaviest dwarf satellites will have to be modeled
+with more intricate means.
 

### 2014-01-07 tex/ms.tex
streams: some re-arranging

@@
 
+The framework presented in this paper differs from other approaches
+that model streams in action--angle coordinates in a few respects. As
+in \citet{Helmi99a}, the model is a generative one in that it gives an
+analytic prescription for the initial action--angle distribution of
+tidal debris that is then evolved in time to generate a tidal
+stream. The difference between these two generative models is that I
+propose simple, few-parameter models for the initial action--angle
+distributions, that I work out the properties of streams in observable
+coordinates in more detail, and that I present and use a general
+action--angle transformation that allows stream modeling in general
+potentials. The average location of a stream in phase-space can be
+computed in the framework presented here and it could be used to
+replace orbit fitting; \citet{Varghese11a} discussed a similar method
+to model observed streams. The method in this paper differs from that
+of \citet{Varghese11a} in that the model is fundamentally located in
+action--angle space rather than position--velocity space and in that
+the width and length of the stream are related to the offsets between
+the stream, the progenitor's orbit, and the average orbit as a
+function of stream position. Thus, the present framework does not have
+to assume a progenitor velocity dispersion, but can constrain it from
+the stream data as well. Compared to the recent methods of
+\citet{Penarrubia12a}, \citet{Sanders13b}, and \citet{PriceWhelan13a},
+which allow for fitting the Milky Way's gravitational potential using
+observations of the positions and velocities of individual stream
+stars, the method presented here is a generative model and it can be
+used with noisy or entirely missing phase-space dimensions.
+
 This paper is outlined as follows. In \sectionname~\ref{sec:dynamics}
@@
 isotropic Gaussian distribution.
-
-Ingredients (a) and (b) require more careful modeling. The right panel
-of \figurename~\ref{fig:gd1_times} shows the distribution of times at
-which stream particles are stripped from the progenitor. These
-stripping times are calculated from the final snapshot by linear
-regression of a particle's angle offset $\Delta \veca$ from the
-progenitor, versus the frequency offset $\Delta \veco$ as 
-\begin{equation}
-  \ts = \frac{\Delta \veco \cdot \Delta \veca}{|\Delta \veco|^2}\,.
-\end{equation}
-The dotted line shows a finely-binned histogram of these stripping
-times, which indicates that stripping happens in bursts with a period
-of about $400\Myr$, the radial period of the progenitor's orbit. Thus,
-as expected, stripping happens primarily at pericenter passages, with
-a smaller number of particles lost between pericenter
-passages. Realistic models of a stream, especially those that require
-good models of the surface-density structure of the stream as a
-function of stream angle, need to model this non-uniform stripping
-process. In what follows, we will use a uniform distribution of
-stripping times. This is appropriate for modeling the large-scale
-structure of streams, such as what is used when constraining the
-gravitational potential.
-
-The middle panel of \figurename~\ref{fig:gd1_times} shows the
-frequency offsets between stream particles and the progenitor as a
-function of stripping time, corresponding to ingredient (b) above. The
-same burstiness that is apparent in the distribution of stripping
-times shows up here. Particles stripped at pericenter typically have
-larger frequency offsets than those lost at larger radii. This can be
-described as saying that particles are \emph{removed} at pericenter,
-but only \emph{peeled off} between pericenter passages. Again,
-realistic models of tidal streams need to take this detailed structure
-into account.
-
-The right panel of \figurename~\ref{fig:gd1_times} shows the angle
-offset (at the end of the simulation) versus the stripping time. There
-is a spread in the angle offset reached for any given stripping time,
-but it is nevertheless typically the case that particles that have
-been stripped earlier are found at larger angle offsets
-today. Self-sorting---the tendency of particles stripped later with
-larger frequency offsets to overtake particles stripped earlier at
-smaller frequency differences and thus to erase correlations between
-particle position along the stream and stripping time and sort the
-stream in frequency offset---happens to a small extent but the narrow
-distribution of frequency offsets limits the ability of particles
-removed later to reach large distances from the progenitor.
 
@@
 \end{figure*}
+
+Ingredients (a) and (b) require more careful modeling. The right panel
+of \figurename~\ref{fig:gd1_times} shows the distribution of times at
+which stream particles are stripped from the progenitor. These
+stripping times are calculated from the final snapshot by linear
+regression of a particle's angle offset $\Delta \veca$ from the
+progenitor, versus the frequency offset $\Delta \veco$ as 
+\begin{equation}
+  \ts = \frac{\Delta \veco \cdot \Delta \veca}{|\Delta \veco|^2}\,.
+\end{equation}
+The dotted line shows a finely-binned histogram of these stripping
+times, which indicates that stripping happens in bursts with a period
+of about $400\Myr$, the radial period of the progenitor's orbit. Thus,
+as expected, stripping happens primarily at pericenter passages, with
+a smaller number of particles lost between pericenter
+passages. Realistic models of a stream, especially those that require
+good models of the surface-density structure of the stream as a
+function of stream angle, need to model this non-uniform stripping
+process. In what follows, we will use a uniform distribution of
+stripping times. This is appropriate for modeling the large-scale
+structure of streams, such as what is used when constraining the
+gravitational potential.
+
+The middle panel of \figurename~\ref{fig:gd1_times} shows the
+frequency offsets between stream particles and the progenitor as a
+function of stripping time, corresponding to ingredient (b) above. The
+same burstiness that is apparent in the distribution of stripping
+times shows up here. Particles stripped at pericenter typically have
+larger frequency offsets than those lost at larger radii. This can be
+described as saying that particles are \emph{removed} at pericenter,
+but only \emph{peeled off} between pericenter passages. Again,
+realistic models of tidal streams need to take this detailed structure
+into account.
+
+The right panel of \figurename~\ref{fig:gd1_times} shows the angle
+offset (at the end of the simulation) versus the stripping time. There
+is a spread in the angle offset reached for any given stripping time,
+but it is nevertheless typically the case that particles that have
+been stripped earlier are found at larger angle offsets
+today. Self-sorting---the tendency of particles stripped later with
+larger frequency offsets to overtake particles stripped earlier at
+smaller frequency differences and thus to erase correlations between
+particle position along the stream and stripping time and sort the
+stream in frequency offset---happens to a small extent but the narrow
+distribution of frequency offsets limits the ability of particles
+removed later to reach large distances from the progenitor.
 
@@
 streams, however, is necessary to check this and to find the optimal
-value of $\mu_\Omega$. Fixing $\mu_\Omega$ leaves a single free
+value of $\mu_\Omega$. In particular, if the progenitor has internal
+rotation the action distribution of the debris and $\mu_\Omega$ may be
+significantly different. Fixing $\mu_\Omega$ leaves a single free
 parameter \sigv\ for modeling the frequency offset distribution of a
@@
 
+\begin{figure*}
+\begin{center}
+  \includegraphics[width=0.8\textwidth,clip=]{track_aa.eps}\\
+  \includegraphics[width=0.8\textwidth,clip=]{track_xz.eps}\\
+  \includegraphics[width=0.8\textwidth,clip=]{track_ld.eps}
+  \caption{Illustration of the determination of the average stream
+    track described in \sectionname~\ref{sec:track}. The small black
+    dots in each panel are the points along the progenitor track at
+    which the Jacobian $\partial (\veco,\veca) / \partial
+    (\vecx,\vecv)$ is calculated. The large black dots are the
+    corresponding points on the stream where the track is
+    calculated. The dashed line in each panel is the progenitor's
+    orbit, while the solid line and dash-dotted lines are the mean
+    track and $2\sigma$ dispersion.}\label{fig:track}
+%python illustrate_track.py ../tex/track_aa.ps ../tex/track_xz.ps ../tex/track_ld.ps
+\end{center}
+\end{figure*}
+
 We can then estimate the dispersion around the stream track, which is
@@
 \subsection{In position--velocity coordinates}\label{sec:trackxv}
+
+\begin{figure*}[th!!]
+\begin{center}
+  \includegraphics[width=0.8\textwidth,clip=]{gd1_evol_lb.eps}
+  \includegraphics[width=0.8\textwidth,clip=]{gd1_evol_ld.eps}
+  \caption{Simulated stream after $5.011\Gyr$ as a function of
+    Galactic longitude. The top panel shows the stream in Galactic
+    latitude and the bottom panel shows it in the distance from the
+    Sun. The progenitor is shown as the large gray dot and its orbit
+    as the dashed lines. The predicted track of the stream is shown as
+    the solid line. Insets show various zoomed-in parts of the stream
+    to show the stream, progenitor orbit, and predicted stream track
+    up close.}\label{fig:gd1_lbd}
+%python plot_stream.py ../tex/gd1_evol_lb.ps
+%python plot_stream.py ../tex/gd1_evol_ld.ps
+\end{center}
+\end{figure*}
 
@@
 
-\begin{figure*}
-\begin{center}
-  \includegraphics[width=0.8\textwidth,clip=]{track_aa.eps}\\
-  \includegraphics[width=0.8\textwidth,clip=]{track_xz.eps}\\
-  \includegraphics[width=0.8\textwidth,clip=]{track_ld.eps}
-  \caption{Illustration of the determination of the average stream
-    track described in \sectionname~\ref{sec:track}. The small black
-    dots in each panel are the points along the progenitor track at
-    which the Jacobian $\partial (\veco,\veca) / \partial
-    (\vecx,\vecv)$ is calculated. The large black dots are the
-    corresponding points on the stream where the track is
-    calculated. The dashed line in each panel is the progenitor's
-    orbit, while the solid line and dash-dotted lines are the mean
-    track and $2\sigma$ dispersion.}\label{fig:track}
-%python illustrate_track.py ../tex/track_aa.ps ../tex/track_xz.ps ../tex/track_ld.ps
-\end{center}
-\end{figure*}
-
-\begin{figure*}[t!]
-\begin{center}
-  \includegraphics[width=0.8\textwidth,clip=]{gd1_evol_lb.eps}
-  \includegraphics[width=0.8\textwidth,clip=]{gd1_evol_ld.eps}
-  \caption{Simulated stream after $5.011\Gyr$ as a function of
-    Galactic longitude. The top panel shows the stream in Galactic
-    latitude and the bottom panel shows it in the distance from the
-    Sun. The progenitor is shown as the large gray dot and its orbit
-    as the dashed lines. The predicted track of the stream is shown as
-    the solid line. Insets show various zoomed-in parts of the stream
-    to show the stream, progenitor orbit, and predicted stream track
-    up close.}\label{fig:gd1_lbd}
-%python plot_stream.py ../tex/gd1_evol_lb.ps
-%python plot_stream.py ../tex/gd1_evol_ld.ps
-\end{center}
-\end{figure*}
+\figurename~\ref{fig:track} illustrates how I linearize the
+$(\vecx,\vecv) \rightarrow (\veco,\veca)$ transformation along the
+progenitor orbit and use it to propagate the track of the stream to
+position--velocity coordinates and to observable quantities
+$(l,b,D,\vlos,\pmll,\pmbb)$. The top panel shows the orbit of the
+progenitor and the mean stream track in frequency--angle
+coordinates. Only the parallel direction is shown here; the mean
+stream track coincides with the progenitor orbit in the perpendicular
+direction. At a small number of points along the progenitor orbit,
+chosen here to span $1.5$ radians, we calculate the Jacobian $\partial
+(\veco,\veca) / \partial (\vecx,\vecv)$. In this Figure this is done
+at four points, but for all other figures and calculations in this
+paper eleven points are used. Each Jacobian calculation requires seven
+frequency--angle calculations, such that the total number of such
+computations is less than $100$ to adequately model the stream (each
+computation involves a single orbit integration, see
+\appendixname~\ref{sec:aa}).
 
@@
 \end{figure*}
-
-\figurename~\ref{fig:track} illustrates how I linearize the
-$(\vecx,\vecv) \rightarrow (\veco,\veca)$ transformation along the
-progenitor orbit and use it to propagate the track of the stream to
-position--velocity coordinates and to observable quantities
-$(l,b,D,\vlos,\pmll,\pmbb)$. The top panel shows the orbit of the
-progenitor and the mean stream track in frequency--angle
-coordinates. Only the parallel direction is shown here; the mean
-stream track coincides with the progenitor orbit in the perpendicular
-direction. At a small number of points along the progenitor orbit,
-chosen here to span $1.5$ radians, we calculate the Jacobian $\partial
-(\veco,\veca) / \partial (\vecx,\vecv)$. In this Figure this is done
-at four points, but for all other figures and calculations in this
-paper eleven points are used. Each Jacobian calculation requires seven
-frequency--angle calculations, such that the total number of such
-computations is less than $100$ to adequately model the stream (each
-computation involves a single orbit integration, see
-\appendixname~\ref{sec:aa}).
 
@@
 \apar$, $X$, or $l$, and a single-orbit approximation is necessarily
-wrong over the full length of the stream.
+wrong over the full length of the stream. This is the case regardless
+of the progenitor's orbit.
 

### 2014-01-07 tex/ms.tex
streams: re-arrange, looks okay

@@
 \begin{abstract} 
-  I present a new framework for modeling the dynamics of cold tidal
+  I present a new framework for modeling the dynamics of tidal
   streams. The framework consists of simple models for the initial
@@
   straightforwardly evolved forward in time. Taking advantage of the
-  essentially one-dimensional nature of cold tidal streams, the
+  essentially one-dimensional nature of tidal streams, the
   transformation to position--velocity coordinates can be linearized
@@
 
-In this paper, I present a new method for modeling cold tidal streams
-that fundamentally lives in frequency--angle space, because the stream
+In this paper, I present a new method for modeling tidal streams that
+fundamentally lives in frequency--angle space, because the stream
 distribution function is essentially one-dimensional in this
@@
 number of points along the progenitor orbit that span the length of
-the stream. For cold streams, the separation between the stream track
-and the progenitor orbit as well as the internal differences within
-the stream are small, such that this linear approximation is accurate
-enough. The linear approximation allows for fast evaluation of the
-average stream location (the stream ``track'') in position--velocity
-coordinates and efficient mock data generation. For a uniform
-distribution of stripping times, an observed stream member's
-phase--space probability distribution function (PDF) can be
-analytically marginalized over stripping time, such that using the
-linear approximation any evaluation---including marginalization and
-uncertainty convolution---of the PDF is extremely rapid. This clears
-the way for proper probabilistic inference of the Milky Way halo's
-gravitational potential and progenitor properties using stream data
-for individual stars.
+the stream. For streams originating from progenitors with masses
+$\lesssim 10^8\msun$, the separation between the stream track and the
+progenitor orbit as well as the internal differences within the stream
+are small, such that this linear approximation is accurate enough. The
+linear approximation allows for fast evaluation of the average stream
+location (the stream ``track'') in position--velocity coordinates and
+efficient mock data generation. For a uniform distribution of
+stripping times, an observed stream member's phase--space probability
+distribution function (PDF) can be analytically marginalized over
+stripping time, such that using the linear approximation any
+evaluation---including marginalization and uncertainty
+convolution---of the PDF is extremely rapid. This clears the way for
+proper probabilistic inference of the Milky Way halo's gravitational
+potential and progenitor properties using stream data for individual
+stars.
 
@@
 \begin{figure}[t!]
-  \includegraphics[width=0.48\textwidth,clip=]{gd1_evol_aadohist.eps}
+  \includegraphics[width=0.44\textwidth,clip=]{gd1_evol_aadohist.eps}
   \caption{Distribution of the magnitude of frequency offsets between
@@
 dispersion of the progenitor, I conjecture that a constant
-$\mu_\Omega$ can be used for modeling cold tidal streams of any
-(small-ish) velocity dispersion. Further modeling of simulated
-streams, however, is necessary to check this and to find the optimal
-value of $\mu_\Omega$. In particular, if the progenitor has internal
-rotation the action distribution of the debris and $\mu_\Omega$ may be
+$\mu_\Omega$ can be used for modeling tidal streams of any (small-ish)
+velocity dispersion. Further modeling of simulated streams, however,
+is necessary to check this and to find the optimal value of
+$\mu_\Omega$. In particular, if the progenitor has internal rotation
+the action distribution of the debris and $\mu_\Omega$ may be
 significantly different. Fixing $\mu_\Omega$ leaves a single free
@@
 
-\section{The track of a tidal stream}\label{sec:track}
-
-In this Section, I discuss how to calculate the track of a model tidal
-stream in the generative model described in
-\sectionname~\ref{sec:generative}. The track consists of the mean
-location of the stream as a function of angle along the stream. I also
-describe how to estimate the dispersion of the stream along this
-track. In \sectionname~\ref{sec:trackaa}, I explain how to calculate
-the track of the stream in frequency--angle coordinates as well as how
-to estimate the spread around this track. In
-\sectionname~\ref{sec:trackxv} I discuss how to efficiently project
-the stream track and dispersion into position--velocity coordinates
-$(\vecx,\vecv)$ or observable coordinates $(l,b,D,\vlos,\pmll,\pmbb)$.
-
-\subsection{In frequency--angle coordinates}\label{sec:trackaa}
-
 \begin{figure*}[t!]
 \begin{center}
-  \includegraphics[width=0.36\textwidth,clip=]{gd1_evol_aaaparopar.ps}
-  \includegraphics[width=0.36\textwidth,clip=]{gd1_evol_aaaparoperp.ps}\\
-  \includegraphics[width=0.36\textwidth,clip=]{gd1_evol_aaaparaperp.ps}
-  \includegraphics[width=0.36\textwidth,clip=]{gd1_evol_aaapartime.ps}
+  \includegraphics[width=0.34\textwidth,clip=]{gd1_evol_aaaparopar.eps}
+  \includegraphics[width=0.34\textwidth,clip=]{gd1_evol_aaaparoperp.eps}\\
+  \includegraphics[width=0.34\textwidth,clip=]{gd1_evol_aaaparaperp.eps}
+  \includegraphics[width=0.34\textwidth,clip=]{gd1_evol_aaapartime.eps}
   \caption{Stream properties as a function of angle along the
@@
 
+\section{The track of a tidal stream}\label{sec:track}
+
+In this Section, I discuss how to calculate the track of a model tidal
+stream in the generative model described in
+\sectionname~\ref{sec:generative}. The track consists of the mean
+location of the stream as a function of angle along the stream. I also
+describe how to estimate the dispersion of the stream along this
+track. In \sectionname~\ref{sec:trackaa}, I explain how to calculate
+the track of the stream in frequency--angle coordinates as well as how
+to estimate the spread around this track. In
+\sectionname~\ref{sec:trackxv} I discuss how to efficiently project
+the stream track and dispersion into position--velocity coordinates
+$(\vecx,\vecv)$ or observable coordinates $(l,b,D,\vlos,\pmll,\pmbb)$.
+
+\begin{figure*}[t!p]
+\begin{center}
+  \includegraphics[width=0.8\textwidth,clip=]{track_aa.eps}\\
+  \includegraphics[width=0.8\textwidth,clip=]{track_xz.eps}\\
+  \includegraphics[width=0.8\textwidth,clip=]{track_ld.eps}
+  \caption{Illustration of the determination of the average stream
+    track described in \sectionname~\ref{sec:track}. The small black
+    dots in each panel are the points along the progenitor track at
+    which the Jacobian $\partial (\veco,\veca) / \partial
+    (\vecx,\vecv)$ is calculated. The large black dots are the
+    corresponding points on the stream where the track is
+    calculated. The dashed line in each panel is the progenitor's
+    orbit, while the solid line and dash-dotted lines are the mean
+    track and $2\sigma$ dispersion.}\label{fig:track}
+%python illustrate_track.py ../tex/track_aa.ps ../tex/track_xz.ps ../tex/track_ld.ps
+\end{center}
+\end{figure*}
+
+\subsection{In frequency--angle coordinates}\label{sec:trackaa}
+
 In the generative model described above, the direction of the angle
@@
 
-\begin{figure*}
+\begin{figure*}[t!]
 \begin{center}
-  \includegraphics[width=0.8\textwidth,clip=]{track_aa.eps}\\
-  \includegraphics[width=0.8\textwidth,clip=]{track_xz.eps}\\
-  \includegraphics[width=0.8\textwidth,clip=]{track_ld.eps}
-  \caption{Illustration of the determination of the average stream
-    track described in \sectionname~\ref{sec:track}. The small black
-    dots in each panel are the points along the progenitor track at
-    which the Jacobian $\partial (\veco,\veca) / \partial
-    (\vecx,\vecv)$ is calculated. The large black dots are the
-    corresponding points on the stream where the track is
-    calculated. The dashed line in each panel is the progenitor's
-    orbit, while the solid line and dash-dotted lines are the mean
-    track and $2\sigma$ dispersion.}\label{fig:track}
-%python illustrate_track.py ../tex/track_aa.ps ../tex/track_xz.ps ../tex/track_ld.ps
+  \includegraphics[width=0.8\textwidth,clip=]{gd1_evol_lb.eps}
+  \includegraphics[width=0.8\textwidth,clip=]{gd1_evol_ld.eps}
+  \caption{Simulated stream after $5.011\Gyr$ as a function of
+    Galactic longitude. The top panel shows the stream in Galactic
+    latitude and the bottom panel shows it in the distance from the
+    Sun. The progenitor is shown as the large gray dot and its orbit
+    as the dashed lines. The predicted track of the stream is shown as
+    the solid line. Insets show various zoomed-in parts of the stream
+    to show the stream, progenitor orbit, and predicted stream track
+    up close.}\label{fig:gd1_lbd}
+%python plot_stream.py ../tex/gd1_evol_lb.ps
+%python plot_stream.py ../tex/gd1_evol_ld.ps
 \end{center}
@@
 
-\begin{figure*}[th!!]
-\begin{center}
-  \includegraphics[width=0.8\textwidth,clip=]{gd1_evol_lb.eps}
-  \includegraphics[width=0.8\textwidth,clip=]{gd1_evol_ld.eps}
-  \caption{Simulated stream after $5.011\Gyr$ as a function of
-    Galactic longitude. The top panel shows the stream in Galactic
-    latitude and the bottom panel shows it in the distance from the
-    Sun. The progenitor is shown as the large gray dot and its orbit
-    as the dashed lines. The predicted track of the stream is shown as
-    the solid line. Insets show various zoomed-in parts of the stream
-    to show the stream, progenitor orbit, and predicted stream track
-    up close.}\label{fig:gd1_lbd}
-%python plot_stream.py ../tex/gd1_evol_lb.ps
-%python plot_stream.py ../tex/gd1_evol_ld.ps
-\end{center}
-\end{figure*}
-
 Having estimated the track of the stream and the dispersion around
@@
 of the stream and inverting this linearized transformation. That this
-works relies on the stream being cold for a few different
-reasons. First, the track of a cold stream, while in general offset
-from the orbit of the progenitor, does not stray too far from it, such
-that the transformation to frequency--angle coordinates can be
-linearized over the range of the stream--orbit offset. Second, cold
-streams are essentially one-dimensional, that is, they only spread
-significantly over a single direction. This means that non-linearity
-in the $(\vecx,\vecv) \rightarrow (\veco,\veca)$ transformation only
-affects a single direction, which roughly follows the orbit of the
+works relies on the stream being relatively cold (in that $\sigv/V_c
+\ll 1$) for a few different reasons. First, the track of a cold
+stream, while in general offset from the orbit of the progenitor, does
+not stray too far from it, such that the transformation to
+frequency--angle coordinates can be linearized over the range of the
+stream--orbit offset. Second, cold streams are essentially
+one-dimensional, that is, they only spread significantly over a single
+direction. This means that non-linearity in the $(\vecx,\vecv)
+\rightarrow (\veco,\veca)$ transformation only affects a single
+direction, which roughly follows the orbit of the
 progenitor. Therefore, we can linearize the $(\vecx,\vecv) \rightarrow

### 2014-01-07 py/plot_stream_2Gya.py
streams: smaller mock figures

@@ -1,4 +1,5 @@
 import sys
+import copy
 import os, os.path
 import numpy
 from galpy.util import bovy_plot, bovy_coords, bovy_conversion
@@ -75,11 +76,11 @@ def plot_stream_xz(plotfilename):
         sdf= streamdf(_SIGV/220.,progenitor=Orbit(obs),pot=lp,aA=aAI,
                       leading=True,nTrackChunks=_NTRACKCHUNKS,
                       tdisrupt=tdisrupt/bovy_conversion.time_in_Gyr(220.,8.),
-                      deltaAngleTrack=1.5*3./5.)
+                      deltaAngleTrack=1.5*3./5.,multi=_NTRACKCHUNKS)
         sdft= streamdf(_SIGV/220.,progenitor=Orbit(obs),pot=lp,aA=aAI,
                        leading=False,nTrackChunks=_NTRACKCHUNKS,
                        tdisrupt=tdisrupt/bovy_conversion.time_in_Gyr(220.,8.),
-                       deltaAngleTrack=1.5*3./5.)
+                       deltaAngleTrack=1.5*3./5.,multi=_NTRACKCHUNKS)
     if 'sim' in plotfilename:
         #Replace data with simulated data
         forwardXY= sdf.sample(int(round(numpy.sum(sindx)/2.)),
@@ -99,14 +100,20 @@ def plot_stream_xz(plotfilename):
         data[forwardXY.shape[1]:,4]= backwardXY[3]*220.
         data[forwardXY.shape[1]:,5]= backwardXY[5]*220.
         data[forwardXY.shape[1]:,6]= backwardXY[4]*220.
-        sindx= numpy.ones(data.shape[0],dtype='bool')
+        sindx= numpy.ones(data.shape[0],dtype='bool')      
     #Plot
     bovy_plot.bovy_print()
-    bovy_plot.bovy_plot(data[:,1],data[:,2],'k.',ms=2.,
+    bovy_plot.bovy_plot(data[sindx,1],data[sindx,2],'k,',ms=2.,
                         xlabel=r'$X\,(\mathrm{kpc})$',
                         ylabel=r'$Z\,(\mathrm{kpc})$',
                         xrange=[-12.5,-3.],
                         yrange=[-12.5,-7.])
+    if numpy.sum(True-sindx) > 0:
+        #Also plot progenitor
+        pindx= copy.copy(True-sindx)
+        pindx[0:9900]= False #subsample
[…]

### 2014-01-07 py/plot_stream.py
streams: smaller l figures

@@ -108,6 +108,12 @@ def plot_stream_lb(plotfilename):
     #Read stream
     data= numpy.loadtxt(os.path.join(_STREAMSNAPDIR,'gd1_evol_hitres_01312.dat'),
                         delimiter=',')
+    aadata= numpy.loadtxt(os.path.join(_STREAMSNAPAADIR,
+                                       'gd1_evol_hitres_aa_01312.dat'),
+                          delimiter=',')
+    thetar= aadata[:,6]
+    thetar= (numpy.pi+(thetar-numpy.median(thetar))) % (2.*numpy.pi)
+    sindx= numpy.fabs(thetar-numpy.pi) > (1.5*numpy.median(numpy.fabs(thetar-numpy.median(thetar)))) #stars in the stream
     #Transform to (l,b)
     XYZ= bovy_coords.galcenrect_to_XYZ(data[:,1],data[:,3],data[:,2],Xsun=8.)
     lbd= bovy_coords.XYZ_to_lbd(XYZ[0],XYZ[1],XYZ[2],degree=True)
@@ -198,17 +204,26 @@ def plot_stream_lb(plotfilename):
         yrange=[-10.,60.]
         ylabel=r'$\mathrm{Galactic\ latitude}\,(\mathrm{deg})$'
     if 'vlos' in plotfilename or 'pm' in plotfilename:
-        bovy_plot.bovy_plot(lbd[:,0],vlbd[:,lbindx],'k,',
+        #Stream
+        bovy_plot.bovy_plot(lbd[sindx,0],vlbd[sindx,lbindx],'k,',
                             xlabel=r'$\mathrm{Galactic\ longitude}\,(\mathrm{deg})$',
                             ylabel=ylabel,
                             xrange=[0.,290.],
                             yrange=yrange)
+        #Progenitor
+        pindx= copy.copy(True-sindx)
+        pindx[0:9900]= False
+        bovy_plot.bovy_plot(lbd[pindx,0],vlbd[pindx,lbindx],'k,',overplot=True)
     else:
-        bovy_plot.bovy_plot(lbd[:,0],lbd[:,lbindx],'k,',
+        bovy_plot.bovy_plot(lbd[sindx,0],lbd[sindx,lbindx],'k,',
                             xlabel=r'$\mathrm{Galactic\ longitude}\,(\mathrm{deg})$',
                             ylabel=ylabel,
                             xrange=[0.,290.],
                             yrange=yrange)
+        #Progenitor
[…]

### 2014-01-07 py/plot_stream.py
streams: try reducing file sizes, but give up...

@@ -889,6 +889,8 @@ def plot_stream_aa(plotfilename):
                             (sas),
                             color='0.6',zorder=1)
         pyplot.plot(das,sass,'--',color='w',zorder=7,lw=2.)
+    #ax= pyplot.gca()
+    #ax.set_rasterized(True)
     bovy_plot.bovy_end_print(plotfilename)
 
 def plot_stream_times(plotfilename):

### 2014-01-08 tex/ms.tex
streams: fix Sesar ref
streams: add note about averaging J_Z and L_Z

@@
 action over the auxiliary angle.
+
+In particular, because for any spherical potential $\Phi^A$ we have
+that $J_Z = |L|-L_Z$, independent of the particular form of $\Phi^A$,
+\equationname~(\ref{eq:aAIJ}) implies that as long as there is a
+spherical auxiliary potential in which the auxiliary angle
+$\theta^A_Z$ goes through the full range $0$ to $2\pi$ along the orbit
+integrated in a general potential, $J_Z$ can be obtained by averaging
+$|L|-L_Z$ over the orbit. Similarly, as $J_\phi = L_Z$ for any
+axisymmetric potential, we can approximate $J_\phi$ for any triaxial
+potential by averaging $L_Z$ over the orbit, as long as an auxiliary
+axisymmetric potential exists in which $\theta^A_\phi$ goes through
+the full range $0$ to $2\pi$ along the orbit (\ie, if the orbit is a
+loop orbit).
 
@@
   Sanders,~J.~L. \& Binney,~J. 2013b, \mnras, 433, 1826
-\bibitem[Sesar\etal(2013)]{Sesar13a}
+\bibitem[Sesar \etal(2013)]{Sesar13a}
   Sesar,~B., \etal\ 2013, \apj, 776, 26

### 2014-01-08 tex/ms.tex
streams: fix L-LZ abs

@@
 In particular, because for any spherical potential $\Phi^A$ we have
-that $J_Z = |L|-L_Z$, independent of the particular form of $\Phi^A$,
+that $J_Z = L-|L_Z|$, independent of the particular form of $\Phi^A$,
 \equationname~(\ref{eq:aAIJ}) implies that as long as there is a
@@
 integrated in a general potential, $J_Z$ can be obtained by averaging
-$|L|-L_Z$ over the orbit. Similarly, as $J_\phi = L_Z$ for any
+$L-|L_Z|$ over the orbit. Similarly, as $J_\phi = L_Z$ for any
 axisymmetric potential, we can approximate $J_\phi$ for any triaxial

### 2014-01-09 tex/ms.tex
streams: add readthedocs link as alternative, add Binney to acks

@@
 \usepackage{amssymb,amsmath}
-%\usepackage{caption}
-%\DeclareCaptionLabelSeparator{dotemdash}{.--- }
-%\captionsetup[figure]{labelformat=simple, labelsep=dotemdash}
-%\usepackage{color,hyperref}
+\usepackage{color,hyperref}
 % hypertex insanity
-%\definecolor{linkcolor}{rgb}{0,0,0.25}
-%\hypersetup{
-%  colorlinks=true,        % false: boxed links; true: colored links
-%  linkcolor=linkcolor,    % color of internal links
-%  citecolor=linkcolor,    % color of links to bibliography
-%  filecolor=linkcolor,    % color of file links
-%  urlcolor=linkcolor      % color of external links
-%}
+\definecolor{linkcolor}{rgb}{0,0,0.25}
+\hypersetup{
+  colorlinks=true,        % false: boxed links; true: colored links
+  linkcolor=linkcolor,    % color of internal links
+  citecolor=linkcolor,    % color of links to bibliography
+  filecolor=linkcolor,    % color of file links
+  urlcolor=linkcolor      % color of external links
+}
 \newcounter{address}
 \setcounter{address}{1}
-%\usepackage{sidecap}
 \setlength{\emergencystretch}{2em}%No overflowing references
@@
   \protect{\url{http://github.com/jobovy/galpy}}~.}. A tutorial on how
-to use this code is given at
-\[
-\protect{\mathrm{\url{http://jobovy.github.io/galpy/streamdf.html}}}\,.
-\]
+to use this code is given at\footnote{An alternative link is\\ \protect{\url{http://galpy.readthedocs.org/en/latest/streamdf.html}}~.}
+\begin{center}
+\protect{\url{http://jobovy.github.io/galpy/streamdf.html}}\,.
+\end{center}
 
@@
 
-\acknowledgements It is a pleasure to thank David W. Hogg, Hans-Walter
-Rix, Jason Sanders, and Scott Tremaine for helpful
-discussions. J.B. was supported by NASA through Hubble Fellowship
-grant HST-HF-51285.01 from the Space Telescope Science Institute,
-which is operated by the Association of Universities for Research in
-Astronomy, Incorporated, under NASA contract NAS5-26555. J.B.
-acknowledges support from SFB 881 (A3) funded by the German Research
-Foundation DFG.
+\acknowledgements It is a pleasure to thank James Binney, David
+W. Hogg, Hans-Walter Rix, Jason Sanders, and Scott Tremaine for
+helpful discussions. J.B. was supported by NASA through Hubble
+Fellowship grant HST-HF-51285.01 from the Space Telescope Science
+Institute, which is operated by the Association of Universities for
+Research in Astronomy, Incorporated, under NASA contract
+NAS5-26555. J.B.  acknowledges support from SFB 881 (A3) funded by the
+German Research Foundation DFG.
 
@@
 \section{An efficient, general method for calculating action-angle coordinates using orbit integration}\label{sec:aa}
-
-\begin{figure}[t!]
- \includegraphics[width=0.48\textwidth,clip=]{aAI_jr.eps}
- \includegraphics[width=0.48\textwidth,clip=]{aAI_araz.eps}
-  \caption{Action--angle calculations using an orbit-integration-based
-    method. The right panel shows the cumulative mean of the radial
-    actions calculated in an auxiliary isochrone potential calculated
-    along the orbit of the progenitor used in the main body of this
-    paper. The color-coding corresponds to the radial angle
-    (calculated in the isochrone potential) along the orbit. The time
-    is in units of $35.56\Myr$ and the radial action is normalized by
-    its value at $t=200$. The left panel shows the radial and vertical
-    angles calculated in the auxiliary isochrone potential, with
-    color-coding corresponding to the azimuthal angle. The wiggles in
-    the angles are due to the isochrone potential not being the
-    correct potential, but these wiggles are fit and removed to
-    calculate the true angles.}\label{fig:aAI}
-%python plot_aAI.py ../tex/aAI_jr.ps
-%python plot_aAI.py ../tex/aAI_araz.ps
-\end{figure}
 
@@
 orbits \citep[\eg,][]{Sanders13a}.
+
+\begin{figure}[t!]
+ \includegraphics[width=0.48\textwidth,clip=]{aAI_jr.eps}
+ \includegraphics[width=0.48\textwidth,clip=]{aAI_araz.eps}
+  \caption{Action--angle calculations using an orbit-integration-based
+    method. The right panel shows the cumulative mean of the radial
+    actions calculated in an auxiliary isochrone potential calculated
+    along the orbit of the progenitor used in the main body of this
+    paper. The color-coding corresponds to the radial angle
+    (calculated in the isochrone potential) along the orbit. The time
+    is in units of $35.56\Myr$ and the radial action is normalized by
+    its value at $t=200$. The left panel shows the radial and vertical
+    angles calculated in the auxiliary isochrone potential, with
+    color-coding corresponding to the azimuthal angle. The wiggles in
+    the angles are due to the isochrone potential not being the
+    correct potential, but these wiggles are fit and removed to
+    calculate the true angles.}\label{fig:aAI}
+%python plot_aAI.py ../tex/aAI_jr.ps
+%python plot_aAI.py ../tex/aAI_araz.ps
+\end{figure}
 

### 2014-01-09 tex/ms.tex
streams: tweaks to aa description after actually seeing the Fox work

@@
 \sectionname~\ref{sec:conclusion}. In \appendixname~\ref{sec:aa} I
-review the method of Fox \& Binney (2014, in preparation) for
-calculating actions in any static gravitational potential and I show
-how it can be extended to compute frequencies and angles.
+review the method of \citet{Fox12a} for calculating actions in any
+static gravitational potential and I show how it can be simplified and
+extended to compute frequencies and angles.
 
@@
 
-Recently, Fox \& Binney (2014, in preparation) have suggested a new
-method for calculating $\vecj(\vecx,\vecv)$ inspired by torus
-modeling. This method requires only a single orbit integration and
-allows the actions to be calculated accurately and as precisely as
-desired (that is, a convergence criterion can be applied and
-convergence can be attained). I describe the Fox \& Binney method
-briefly here and then show how to extend it to also calculate the
+Recently, \citet{Fox12a} has suggested a new method for calculating
+$\vecj(\vecx,\vecv)$ inspired by torus modeling. This method requires
+only a single orbit integration and allows the actions to be
+calculated accurately and as precisely as desired (that is, a
+convergence criterion can be applied and convergence can be
+attained). I describe the Fox method briefly here, discuss how it can
+be simplified, and then show how to extend it to also calculate the
 frequencies and angles: $(\vecx,\vecv) \rightarrow (\veco,\veca)$.
 
-As in torus modeling, the Fox \& Binney method uses an auxiliary
+As in torus modeling, the method of \citet{Fox12a} uses an auxiliary
 isochrone potential $\Phi^A$ that should be close to the target
@@
 calculated (a precise definition of ``close'' will be given
-below). Actions and angles $(\vecj^A,\veca^A)$ calculated in the
-auxiliary potential can be related to those in the target potential
-$(\vecj,\veca)$ by a generating function that can be written as
-\citep{McGill90a}
+below). The discussion here focuses on loop orbits; for box orbits an
+auxiliary harmonic oscillator potential can be used. Actions and
+angles $(\vecj^A,\veca^A)$ calculated in the auxiliary potential can
+be related to those in the target potential $(\vecj,\veca)$ by a
+generating function that can be written as \citep{McGill90a}
 \begin{equation}
@@
 
-The Fox \& Binney method consists of averaging \eqnname~(\ref{eq:jjt})
-over the auxiliary angle along the path of the orbit, such that the
+The Fox method consists of averaging \eqnname~(\ref{eq:jjt}) over the
+three-dimensional torus $\veca^a$ covered by the orbit, such that the
 mapping defined by the $S_{\vecn}$ is onto orbital tori of the target
 Hamiltonian, the oscillating cosine behavior cancels, and the target
-action is obtained as the auxiliary-angle-averaged auxiliary
-action. That is,
+actions are obtained as the auxiliary-angles-averaged auxiliary
+actions. Performing these three-dimensional integrals is difficult,
+because it is necessary to approximate the three-dimensional volume
+$\dd \veca^A$ from a one-dimensional set of points along the
+integrated orbit. Here I show that the three-dimensional integrals
+over $\veca^A$ can be simplified for non-resonant orbits to
+one-dimensional integrals over one of the auxiliary angles along the
+path of the orbit. This significantly simplifies the calculations,
+especially of $J_\phi$ and $J_Z$ (see below). That is, we integrate
+over \eqnname~(\ref{eq:jjt}) as
 \begin{equation}\label{eq:jjtint}
@@
 that $J_Z = L-|L_Z|$, independent of the particular form of $\Phi^A$,
-\equationname~(\ref{eq:aAIJ}) implies that as long as there is a
-spherical auxiliary potential in which the auxiliary angle
+\equationname~(\ref{eq:aAIJ}) implies that as long as there is
+\emph{some} spherical auxiliary potential in which the auxiliary angle
 $\theta^A_Z$ goes through the full range $0$ to $2\pi$ along the orbit
@@
 although the fact that disk orbits are close to circular makes the
-details of the auxiliary potential important. Similarly, the method
+details of the auxiliary potential unimportant. Similarly, the method
 should be tested in more detail for triaxial potentials as well. For
@@
   Eyre,~A. \& Binney,~J. 2011, \mnras, 413, 1852
+\bibitem[Fox(2012)]{Fox12a}
+  Fox,~M. 2012, M.~Phys.~thesis, Univ.~Oxford
 \bibitem[Gilks \& Wild(1992)]{Gilks92a}
